---
title: "sierra_exploratory"
format: html
editor: source
---
### Load Packages

```{r}

if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(tidyverse, here, DBI, RPostgres, mgcv, lme4, lmerTest, plm)

```

### Connect to Database

```{r}
tryCatch({
    drv <- dbDriver("Postgres")
    print("Connecting to Database…")
    connection <- dbConnect(drv,
                 dbname = Sys.getenv("aws_dbname"),
                 host = Sys.getenv("aws_host"),
                 port = Sys.getenv("aws_port"),
                 user = Sys.getenv("aws_user"),
                 password = Sys.getenv("aws_password"),
                 timezone=NULL)
    print("Database Connected!")
    },
    error=function(cond) {
            print("Unable to connect to Database.")
    })

#search path
dbExecute(connection, "set search_path to survey_data")
```

### query data of interest

```{r}


q <- "select l.location, r.region, s.site, s.wilderness, v.date, s2.detection_type, c.species_capture, 
      c.svl_mm, c.body_mass_g, c.life_stage, qbr.target_quant_per_swab, qbr.detected, qbr.bd_swab_id 
      from location l 
      join region r on l.location_id = r.location_id 
      join site s on r.region_id = s.region_id 
      join visit v on s.site_id = v.site_id 
      join survey s2 on v.visit_id = s2.visit_id 
      join capture c on s2.survey_id = c.survey_id 
      join qpcr_bd_results qbr on c.bd_swab_id = qbr.bd_swab_id
      where c.svl_mm is not null
      and c.body_mass_g is not null
      and c.life_stage = 'adult'
      and qbr.detected = '1';"

raw_data <- dbGetQuery(connection, q) 

```

### Wrangle data - Site

#### Initial wrangle

```{r}

data <- raw_data %>% 
  filter(species_capture == "rana_muscosa") %>% 
  mutate(target_quant_per_swab = log10(target_quant_per_swab + 1)) %>%
  mutate(r = if_else(species_capture == "rana_muscosa" & svl_mm < 40, "0", "1")) %>%
  filter(!r == "0") %>%
  select(!r) %>%
  ungroup() %>% 
   mutate(body_mass_g = log(body_mass_g),
          svl_mm = log(svl_mm)) %>% 
   mutate(svl_mm = as.numeric(scale(svl_mm)),
          body_mass_g = as.numeric(scale(body_mass_g)),
          site = as.factor(site),
          wilderness = as.factor(wilderness))


site_count <- data %>% 
  group_by(site) %>% 
  summarise(n = n()) %>% 
  filter(n > 100) %>% 
  select(site)

# 100 counts of more at site
final_data <- data %>% 
  filter(site %in% c(site_count$site))


```

### Models

#### Explore Cor
```{r}

# initial explore - not looking good....
hist(final_data$svl_mm)

hist(final_data$body_mass_g)

cor(final_data$svl_mm, final_data$target_quant_per_swab)

cor(final_data$body_mass_g, final_data$target_quant_per_swab)
```

#### `plm()` check if fixed or random effect for site is better

```{r}
# run plm target_quant_per_swab ~ svl_mm to determine if fixed or random better

fixed <- plm(target_quant_per_swab ~ svl_mm, index = c("site"), model = "within", data=final_data) #fixed model

random <- plm(target_quant_per_swab ~ svl_mm, index = c("site"), model = "random", data=final_data) #random

phtest(fixed,random)

# random site better because p>0.05


# run plm target_quant_per_swab ~ body_mass_g to determine if fixed or random better

fixed <- plm(target_quant_per_swab ~ body_mass_g, index = c("site"), model = "within", data=final_data) #fixed model

random <- plm(target_quant_per_swab ~ body_mass_g, index = c("site"), model = "random", data=final_data) #random

phtest(fixed,random)

# random site better because p>0.05



# run plm target_quant_per_swab ~ svl_mm*body_mass_g to determine if fixed or random better

fixed <- plm(target_quant_per_swab ~ svl_mm*body_mass_g, index = c("site"), model = "within", data=final_data) #fixed model

random <- plm(target_quant_per_swab ~ svl_mm*body_mass_g, index = c("site"), model = "random", data=final_data) #random

phtest(fixed,random)

# fixed site better because p<0.05

```

#### `lm()` to see sig and trend of site level fixed effects
```{r}
# svl_mm
sit_random <- lm(target_quant_per_swab ~ svl_mm + site - 1, data = final_data)

summary(sit_random)

# Seems longer frogs are more infected...


# body_mass_g
sit_random <- lm(target_quant_per_swab ~ body_mass_g + site - 1, data = final_data)

summary(sit_random)

# Seems bigger frogs are more infected...


# svl_mm * body_mass_g
sit_random <- lm(target_quant_per_swab ~ svl_mm*body_mass_g + site - 1, data = final_data)

summary(sit_random)

```

### Wrangle data - Wilderness

#### Initial wrangle

```{r}

data <- raw_data %>% 
  filter(species_capture == "rana_muscosa") %>% 
  mutate(target_quant_per_swab = log10(target_quant_per_swab + 1)) %>%
  mutate(r = if_else(species_capture == "rana_muscosa" & svl_mm < 40, "0", "1")) %>%
  filter(!r == "0") %>%
  select(!r) %>%
  ungroup() %>% 
   mutate(body_mass_g = log(body_mass_g),
          svl_mm = log(svl_mm)) %>% 
   mutate(svl_mm = as.numeric(scale(svl_mm)),
          body_mass_g = as.numeric(scale(body_mass_g)),
          site = as.factor(site),
          wilderness = as.factor(wilderness))


wilderness_count <- data %>% 
  group_by(wilderness) %>% 
  summarise(n = n()) %>% 
  filter(n > 100) %>% 
  select(wilderness)

# 100 counts of more at site
final_data <- data %>% 
  filter(wilderness %in% c(wilderness_count$wilderness))


```

### Models

#### Explore Cor
```{r}

# initial explore - not looking good....
hist(final_data$svl_mm)

hist(final_data$body_mass_g)

cor(final_data$svl_mm, final_data$target_quant_per_swab)

cor(final_data$body_mass_g, final_data$target_quant_per_swab)
```

#### `plm()` check if fixed or random effect for wilderness is better

```{r}
# run plm target_quant_per_swab ~ svl_mm to determine if fixed or random better

fixed <- plm(target_quant_per_swab ~ svl_mm, index = c("wilderness"), model = "within", data=final_data) #fixed model

random <- plm(target_quant_per_swab ~ svl_mm, index = c("wilderness"), model = "random", data=final_data) #random

phtest(fixed,random)

# random site better because p>0.05



fixed <- plm(target_quant_per_swab ~ body_mass_g, index = c("wilderness"), model = "within", data=final_data) #fixed model

random <- plm(target_quant_per_swab ~ body_mass_g, index = c("wilderness"), model = "random", data=final_data) #random

phtest(fixed,random)

# random site better because p>0.05



fixed <- plm(target_quant_per_swab ~ svl_mm*body_mass_g, index = c("wilderness"), model = "within", data=final_data) #fixed model

random <- plm(target_quant_per_swab ~ svl_mm*body_mass_g, index = c("wilderness"), model = "random", data=final_data) #random

phtest(fixed,random)

# random site better because p>0.05

```

#### `lm(target_quant_per_swab ~ svl_mm + factor(wilderness) - 1)` to see sig and trend of site level fixed effects
```{r}
wild_random <- lm(target_quant_per_swab ~ svl_mm + wilderness - 1, data = final_data)

summary(wild_random)

# Seems bigger frogs are more infected...

```