---
title: "site_70550"
format: html
editor: source
---

### Load Packages

```{r}

if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(tidyverse, here, DBI, RPostgres, mgcv, lme4, lmerTest, plm)

```

### Connect to Database (optiion 1)

```{r}
tryCatch({
    drv <- dbDriver("Postgres")
    print("Connecting to Database…")
    connection <- dbConnect(drv,
                 dbname = Sys.getenv("aws_dbname"),
                 host = Sys.getenv("aws_host"),
                 port = Sys.getenv("aws_port"),
                 user = Sys.getenv("aws_user"),
                 password = Sys.getenv("aws_password"),
                 timezone=NULL)
    print("Database Connected!")
    },
    error=function(cond) {
            print("Unable to connect to Database.")
    })

#search path
dbExecute(connection, "set search_path to survey_data")
```

### Connect to the database (option 2)
```{r}
tryCatch({
    drv <- dbDriver("Postgres")
    print("Connecting to Database…")
    connection <- dbConnect(drv,
                 dbname = "ribbitr",
                 host = "ribbitr.c6p56tuocn5n.us-west-1.rds.amazonaws.com",
                 port = "5432",
                 user = rstudioapi::askForPassword("aws_user"),
                 password = rstudioapi::askForPassword("aws_password"),
                 timezone=NULL)
    print("Database Connected!")
    },
    error=function(cond) {
            print("Unable to connect to Database.")
    })

#search path
dbExecute(connection, "set search_path to survey_data")
```

### query data of interest

```{r}

# This query returns all organisms that are infected and have a SVL and body mass measurement 

q <- "select l.location, r.region, s.site, s.wilderness, v.date, s2.detection_type, c.species_capture, 
      c.svl_mm, c.body_mass_g, c.life_stage, qbr.target_quant_per_swab, qbr.detected, qbr.bd_swab_id 
      from location l 
      join region r on l.location_id = r.location_id 
      join site s on r.region_id = s.region_id 
      join visit v on s.site_id = v.site_id 
      join survey s2 on v.visit_id = s2.visit_id 
      join capture c on s2.survey_id = c.survey_id 
      join qpcr_bd_results qbr on c.bd_swab_id = qbr.bd_swab_id
      where c.svl_mm is not null
      and c.body_mass_g is not null;"
      #and qbr.detected = '1';"  # this restriction could bias results. it also explains why log10(Bd load) data are normally distributed (which was unexpected)

raw_data <- dbGetQuery(connection, q) 

```

### Wrangle Adult data - Site 70550

#### Initial wrangle

```{r}

# Here we filter only for adult RAMU, log10 +1 transform the bd load, and filter for adult life stage only. We then center and scale the variables. 

adult_data <- raw_data %>% 
  filter(species_capture == "rana_muscosa",
         life_stage == "adult") %>% 
  mutate(log_target_quant_per_swab = log10(target_quant_per_swab + 1)) %>%  # adding 1 is unnecessary because there are no 0 values in load data
  ungroup() %>% 
   mutate(log_body_mass_g = log(body_mass_g),
          log_svl_mm = log(svl_mm)) %>% 
#   mutate(svl_mm = as.numeric(scale(svl_mm)),
#          body_mass_g = as.numeric(scale(body_mass_g)),
   mutate(site = as.factor(site),
          wilderness = as.factor(wilderness))


# briefly examine how many sites we have available with greater then 100 observations. 

adult_site_count <- adult_data %>% 
  group_by(site) %>% 
  summarise(n = n()) %>% 
  filter(n > 100) %>% 
  select(site)


# Site 70550 data

final_adult_data <- adult_data %>% 
  filter(site == "70550")


```

### Models

#### Explore 
```{r}

# initial explore
hist(final_adult_data$svl_mm)

hist(final_adult_data$log_svl_mm)

hist(final_adult_data$body_mass_g)

hist(final_adult_data$log_body_mass_g)

hist(final_adult_data$target_quant_per_swab)

hist(final_adult_data$log_target_quant_per_swab)

cor(final_adult_data$svl_mm, final_adult_data$log_target_quant_per_swab)

cor(final_adult_data$log_svl_mm, final_adult_data$log_target_quant_per_swab)

cor(final_adult_data$body_mass_g, final_adult_data$log_target_quant_per_swab)

cor(final_adult_data$log_body_mass_g, final_adult_data$log_target_quant_per_swab)
```

#### `lm()` to see sig and trend for adults at site 70550
```{r}
# svl_mm
adult_svl_mod <- lm(log_target_quant_per_swab ~ svl_mm , data = final_adult_data)

summary(adult_svl_mod)

# Seems longer frogs are less infected


# body_mass_g
adult_bm_mod <- lm(target_quant_per_swab ~ log_body_mass_g, data = final_adult_data)

summary(adult_bm_mod)

# Seems bigger frogs are less infected


# svl_mm * body_mass_g
adult_svl.bm_mod <- lm(target_quant_per_swab ~ svl_mm*log_body_mass_g, data = final_adult_data)

summary(adult_svl.bm_mod)

```

#### Visualize Adult `lm(target_quant_per_swab ~ svl_mm)`
```{r}

pred_svl_adult <- seq(min(final_adult_data$svl_mm), max(final_adult_data$svl_mm), len=100)

newdata_adult <- data.frame(svl_mm = pred_svl_adult)

mod_pred_adult = predict(adult_svl_mod, newdata = newdata_adult, se.fit = T, type = 'response')
newdata_adult$pred <- mod_pred_adult$fit
newdata_adult$upper_se <- mod_pred_adult$fit + (2 * mod_pred_adult$se.fit)
newdata_adult$lower_se <- mod_pred_adult$fit - (2 * mod_pred_adult$se.fit)

ggplot(newdata_adult) + 
  geom_point(data=final_adult_data, aes(x=svl_mm, y=log_target_quant_per_swab), size=0.5, alpha = 0.2) +
  geom_line(data=newdata_adult, aes(x=svl_mm, y=pred), linewidth = 1) +
  geom_ribbon(aes(ymin = lower_se, ymax= upper_se, x = svl_mm), alpha = 0.2, fill = "blue") +
  xlab("svl_mm") + ylab("Infection Intensity") + 
  theme_classic() +
  ggtitle(paste("Adult RAMU, Site 70550, Slope:", 
                round(summary(adult_svl_mod)$coefficients[2,1], 3), 
                ", p-value:", 
                round(summary(adult_svl_mod)$coefficients[2,4], 3), "*"))

```


### Wrangle SubAdult data - Site

#### Initial wrangle

```{r}
# Here we filter only for subadult RAMU, log10 +1 transform the bd load, and filter for adult life stage only. We then center and scale the variables. 
sub_data <- raw_data %>% 
  filter(species_capture == "rana_muscosa",
         life_stage == "subadult") %>% 
  mutate(log_target_quant_per_swab = log10(target_quant_per_swab)) %>%
  ungroup() %>% 
   mutate(log_body_mass_g = log(body_mass_g),
          log_svl_mm = log(svl_mm)) %>% 
#   mutate(svl_mm = as.numeric(scale(svl_mm)),
#          body_mass_g = as.numeric(scale(body_mass_g)),
    mutate(site = as.factor(site),
          wilderness = as.factor(wilderness))


# briefly examine how many sites we have available with greater then 50 observations. 
sub_site_count <- sub_data %>% 
  group_by(site) %>% 
  summarise(n = n()) %>% 
  filter(n > 50) %>% 
  select(site)


# Site 70550 data
final_sub_data <- sub_data %>% 
  filter(site == "70550")


```

### Models

#### Explore Cor
```{r}

# initial explore 
hist(final_sub_data$svl_mm)

hist(final_sub_data$log_svl_mm)

hist(final_sub_data$body_mass_g)

hist(final_sub_data$log_body_mass_g)

hist(final_sub_data$target_quant_per_swab)

hist(final_sub_data$log_target_quant_per_swab)

cor(final_sub_data$svl_mm, final_sub_data$log_target_quant_per_swab)

cor(final_sub_data$log_svl_mm, final_sub_data$log_target_quant_per_swab)

cor(final_sub_data$body_mass_g, final_sub_data$target_quant_per_swab)

cor(final_sub_data$log_body_mass_g, final_sub_data$log_target_quant_per_swab)
```

#### `lm()` to see sig and trend for subadults at site 70550
```{r}
# svl_mm
sub_svl_mod <- lm(log_target_quant_per_swab ~ svl_mm, data = final_sub_data)

summary(sub_svl_mod)

# Seems longer frogs are less infected


# body_mass_g
sub_bm_mod <- lm(log_target_quant_per_swab ~ log_body_mass_g, data = final_sub_data)

summary(sub_bm_mod)

# Seems bigger frogs are less infected...


# svl_mm * body_mass_g
sub_svl.bm_mod <- lm(target_quant_per_swab ~ svl_mm*body_mass_g, data = final_sub_data)

summary(sub_svl.bm_mod)
```


#### Visualize Subadult `lm(target_quant_per_swab ~ svl_mm)`
```{r}

pred_svl_sub <- seq(min(final_sub_data$svl_mm), max(final_sub_data$svl_mm), len=100)

newdata_sub <- data.frame(svl_mm = pred_svl_sub)

mod_pred_adult = predict(sub_svl_mod, newdata = newdata_sub, se.fit = T, type = 'response')
newdata_sub$pred <- mod_pred_adult$fit
newdata_sub$upper_se <- mod_pred_adult$fit + (2 * mod_pred_adult$se.fit)
newdata_sub$lower_se <- mod_pred_adult$fit - (2 * mod_pred_adult$se.fit)

ggplot(newdata_sub) + 
  geom_point(data=final_sub_data, aes(x=svl_mm, y=log_target_quant_per_swab), size=0.5, alpha = 0.2) +
  geom_line(data=newdata_sub, aes(x=svl_mm, y=pred), linewidth = 1) +
  geom_ribbon(aes(ymin = lower_se, ymax= upper_se, x = svl_mm), alpha = 0.2, fill = "blue") +
  xlab("svl_mm") + ylab("Infection Intensity") + 
  theme_classic() +
  ggtitle(paste("Sub-Adult RAMU, Site 70550, Slope:", 
                round(summary(sub_svl_mod)$coefficients[2,1], 3), 
                ", p-value:", 
                round(summary(sub_svl_mod)$coefficients[2,4], 3), "**")) 

```

### All post-metamorphic frogs

```{r}

# Here we filter for adult/subadult RAMU, log10 +1 transform the bd load. We then center and scale the variables. 

postmeta_data <- raw_data %>% 
  filter(species_capture == "rana_muscosa") %>% 
  mutate(log_target_quant_per_swab = log10(target_quant_per_swab+1),
         log_body_mass_g = log(body_mass_g),
         log_svl_mm = log(svl_mm)) %>% 
#   mutate(svl_mm = as.numeric(scale(svl_mm)),
#          body_mass_g = as.numeric(scale(body_mass_g)),
  mutate(site = as.factor(site),
          wilderness = as.factor(wilderness))


# briefly examine how many sites we have available with greater then 100 observations. 

postmeta_site_count <- postmeta_data %>% 
  group_by(site) %>% 
  summarise(n = n()) %>% 
  filter(n > 100) %>% 
  select(site)


# Site 70550 data

final_postmeta_data <- postmeta_data %>% 
  filter(site == "70550")


```

### Models

#### Explore 
```{r}

# initial explore
hist(final_postmeta_data$svl_mm)

hist(final_postmeta_data$log_svl_mm)

hist(final_postmeta_data$body_mass_g)

hist(final_postmeta_data$log_body_mass_g)

hist(final_postmeta_data$target_quant_per_swab)

hist(final_postmeta_data$log_target_quant_per_swab)

cor(final_postmeta_data$svl_mm, final_postmeta_data$log_target_quant_per_swab)

cor(final_postmeta_data$log_svl_mm, final_postmeta_data$log_target_quant_per_swab)

cor(final_postmeta_data$body_mass_g, final_postmeta_data$log_target_quant_per_swab)

cor(final_postmeta_data$log_body_mass_g, final_postmeta_data$log_target_quant_per_swab)

```

#### `lm()` to see sig and trend for subadults at site 70550
```{r}
# svl_mm
postmet_svl_mod <- lm(log_target_quant_per_swab ~ svl_mm, data = final_postmeta_data)

summary(postmet_svl_mod)

# Seems longer frogs are less infected


# body_mass_g
sub_bm_mod <- lm(log_target_quant_per_swab ~ log_body_mass_g, data = final_sub_data)

summary(sub_bm_mod)

# Seems bigger frogs are less infected...


# svl_mm * body_mass_g
sub_svl.bm_mod <- lm(target_quant_per_swab ~ svl_mm*body_mass_g, data = final_sub_data)

summary(sub_svl.bm_mod)
```


#### Visualize postmeta `lm(target_quant_per_swab ~ svl_mm)`
```{r}

pred_svl_postmet <- seq(min(final_postmeta_data$svl_mm), max(final_postmeta_data$svl_mm), len=100)

newdata_postmet <- data.frame(svl_mm = pred_svl_postmet)

mod_pred_adult = predict(postmet_svl_mod, newdata = newdata_postmet, se.fit = T, type = 'response')
newdata_postmet$pred <- mod_pred_adult$fit
newdata_postmet$upper_se <- mod_pred_adult$fit + (2 * mod_pred_adult$se.fit)
newdata_postmet$lower_se <- mod_pred_adult$fit - (2 * mod_pred_adult$se.fit)

ggplot(newdata_postmet) + 
  geom_point(data=final_postmeta_data, aes(x=svl_mm, y=log_target_quant_per_swab), size=0.5, alpha = 0.2) +
  geom_line(data=newdata_postmet, aes(x=svl_mm, y=pred), linewidth = 1) +
  geom_ribbon(aes(ymin = lower_se, ymax= upper_se, x = svl_mm), alpha = 0.2, fill = "blue") +
  xlab("svl_mm") + ylab("Infection Intensity") + 
  theme_classic() +
  ggtitle(paste("Post Meta-Metamorphic RAMU, Site 70550, Slope:", 
                round(summary(postmet_svl_mod)$coefficients[2,1], 3), 
                ", p-value:", 
                round(summary(postmet_svl_mod)$coefficients[2,4], 3), "***")) 

```


#### Plot Bd load data vs SVL

```{r}
final_postmeta_data %>% 
  ggplot(aes(svl_mm, target_quant_per_swab)) +
    scale_y_log10() +
    geom_point() +
    geom_smooth() + 
    geom_smooth(method = lm, color = "green")
```
Exclusion only of Bd-negative frogs could bias results. 70550 is a Bd-positive site, and excluding zero load values could change the Bd load vs SVL relationship. This 
definitely needs additional discussion. 

Let's discuss log10-transformation and scaling of variables. There are some issues there worth digging into.

Linear relationship underestimates loads in subadults and the biggest adults. Smoother shows that Bd loads decrease linear from SVL 30 to 50 mm, and then either
remain constant or increase slightly from 50 to 70 mm. However, the magnitude of difference between subadults and adults is less than I expected based on earlier work. 
Interesting! Here are some points to consider for future modeling:

Small subadults (30-35 mm) show odd bifurcation of Bd load values: One group has relatively constant Bd loads with SVL, the other shows very high loads at small SVL and 
rapid decreases with SVL. Weird. Is this an artifact or something else?

Hierarchical structure that may affect independence of data:
  - Same adults swabbed over time. This may also be the case for subadults, but we have no way of knowing. Difficult to account for across subadults and adults. 
  - Sampling occurred on distinct dates. Include year and/or day-of-year as a fixed effect, date as a random (group-level) effect?
  
Conduct similar analysis using data from epizootic population(s)
  - Difficult because we rarely have many samples from such populations due to high frog mortality and resulting small population size. 
  - 10100, 10101, 10102, 11858, 12590 during epizootic (2015) might have potential (Knapp et al. 2022, PeerJ), but I don't think we swabbed many subadults. 
