---
title: "Size vs. Bd Prevalence & Load"
author: Cob Staines
date: 2025-05-06
format: html
df-print: kable
bibliography: references.bib
csl: ecology.csl
---

# Motivation

Patterns of greater Bd load in recently metamorphosed individuals have been documented in some cases[@adams_chytridiomycosis-induced_2020] [@humphries_immune_2022] [@humphries_chytridiomycosis_2024] [@humphries_chytridiomycosis_2025] (and the opposite pattern in others[@bradley_host_2019]), with possible implications for:

- population dynamics (Q2)
- assessing Bd resistance/tolerance mechanisms (Q3)
- resilience trajectory assessment (Q1)

... with the goal of complementing ongoing RIBBiTR research efforts :)

```{r}
#| eval: true
#| output: false
#| code-fold: true
#| code-summary: "Show setup code"

# setup
librarian::shelf(tidyverse, dbplyr, here, janitor, RPostgres, DBI, RIBBiTR-BII/ribbitrrr)

## Connect to DB
dbcon <- hopToDB("ribbitr")

# table pointers
db_bd = tbl(dbcon, Id("survey_data", "bd_qpcr_results"))
db_sample = tbl(dbcon, Id("survey_data", "sample"))
db_capture = tbl(dbcon, Id("survey_data", "capture"))
db_survey = tbl(dbcon, Id("survey_data", "survey"))
db_visit = tbl(dbcon, Id("survey_data", "visit"))
db_site = tbl(dbcon, Id("survey_data", "site"))
db_region = tbl(dbcon, Id("survey_data", "region"))
db_country = tbl(dbcon, Id("survey_data", "country"))

bd_sample = db_sample %>%
  right_join(db_capture, by = "capture_id") %>%
  left_join(db_survey, by = "survey_id") %>%
  left_join(db_visit, by = "visit_id") %>%
  left_join(db_site, by = "site_id") %>%
  left_join(db_region, by = "region_id") %>%
  left_join(db_country, by = "country_id") %>%
  filter(sample_type == "bd")
```
## A case study:

- Region: Sierra Nevada
- Site: 70550
- Temporal range: 2006 - 2024
- Species: Rana sierrae
- Data: Capture surveys (SVL), Bd swabs (Bd ITS1 copies)
- Sample size: 3874
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show code"

data_70550 = bd_sample %>%
  inner_join(db_bd, by = "sample_id") %>%
  filter(site == "70550",
         !is.na(svl_mm)) %>%
  group_by(sample_id,
           capture_id,
           svl_mm,
           life_stage,
           site,
           date) %>%
  summarise(bd_its1_copies_per_swab = mean(bd_its1_copies_per_swab),
            detected = any(detected),
            .groups = "drop") %>%
  ungroup() %>%
  mutate(svl_mm_noise = svl_mm + runif(n(), min = -0.5, max = 0.5)) %>%
  collect()

stats_70550 = data_70550 %>%
  summarise(min_date = min(date),
            max_date = max(date),
            count = n())

ggplot(data_70550, aes(svl_mm_noise, log(bd_its1_copies_per_swab))) + 
  geom_point() +
  xlab("Snout-vent length [mm]") +
  ylab("ln( ITS1 copies / swab )")
```

```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show code"
#| 
ggplot(data_70550, aes(svl_mm_noise, log(bd_its1_copies_per_swab), color = life_stage)) + 
  geom_point() +
  xlab("Snout-vent length [mm]") +
  ylab("ln( Bd_load + 1)")
```


# Working hypotheses:

a. Younger individuals are more likely to be infected with Bd than older individuals in a given population
b. Younger individuals are more likely to carry higher Bd loads than older individuals in a given population

**Broad-enough for cross-RIBBiTR analysis!**

## A first look at 70550
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show code"

  (prev_70550 = data_70550 %>%
  group_by(life_stage) %>%
  summarise(n = n(),
            prevalence = sum(bd_its1_copies_per_swab > 0) / n()))

ggplot(data_70550, aes(, log(bd_its1_copies_per_swab + 1), color = life_stage)) + 
  geom_boxplot() +
  xlab("Snout-vent length [mm]") +
  ylab("ln( Bd_load  + 1)")

```

## Temporal variability
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show code"

# prevalence & bd load through time
grouped_70550 = data_70550 %>%
  mutate(year = year(date),
         year_group_force = cut(year, breaks = seq(min(year) - 1, max(year) + 5, by = 5))) %>%
  group_by(year_group_force) %>%
  mutate(min_year = min(year),
         max_year = max(year),
         year_group = as.factor(paste0("[", min_year, " - ", max_year, "]"))) %>%
  ungroup()

timeline_70550 = grouped_70550%>%
  group_by(year, life_stage) %>%
  summarise(n = n(),
            prevalence = sum(bd_its1_copies_per_swab > 0) / n(),
            load_mean = mean(bd_its1_copies_per_swab),
            load_infected_mean = mean(bd_its1_copies_per_swab[bd_its1_copies_per_swab > 0]),
            load_infected_median = median(bd_its1_copies_per_swab[bd_its1_copies_per_swab > 0])) %>%
  ungroup()

ggplot(timeline_70550, aes(year, log(load_infected_median), color = life_stage)) +
  geom_point() +
  geom_line() +
  ylab("ln( median Bd_load, infected only )")

ggplot(timeline_70550, aes(year, prevalence, color = life_stage)) +
  geom_point() +
  geom_line()
```

##  Infected only, grouped by...

::: {.panel-tabset}
## lifestage
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show code"

grouped_70550_pos = grouped_70550 %>%
  filter(bd_its1_copies_per_swab > 0)

ggplot(grouped_70550_pos, aes(x = date, y = log(bd_its1_copies_per_swab), color = life_stage)) +
  geom_point() +
  geom_smooth(method = "lm") +
  ylab("ln( Bd_load )")
```

## year
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show code"

ggplot(grouped_70550_pos, aes(x = svl_mm, y = log(bd_its1_copies_per_swab))) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(. ~ year_group) +
  ylab("ln( Bd_load )") +
  xlab("Snout-vent length [mm]")
```

## life-stage & year
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show code"

ggplot(grouped_70550_pos, aes(year_group, log(bd_its1_copies_per_swab), color = life_stage)) +
  geom_boxplot() +
  ylab("ln( Bd_load )") +
  xlab("year group")
```

:::

#### Variables of interest

- **Bd load**
- **age/size** (svl_mm or body_mass_g)
- species
- population/site
- time (date/year/season)
- climate (time, space)
- sampling (methods, bias)
- sex
- enzootic/epizootic
- others...?


# RIBBiTR data inventory

#### Preliminary selection criteria:

- Bd swab + body size (SVL or weight) data for at least 60 individuals in a given population
- populations sampled over at least 3 years
- species found in at least valid 2 populations

```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show code"

# define some populations

pop_84313 = c("84313",
              "84325",
              "84326",
              "84327",
              "84329")

pop_84309 = c("84309",
              "84317",
              "84962",
              "84307",
              "84963",
              "82644",
              "82643",
              "84312",
              "84310",
              "84306",
              "84307",
              "84303",
              "87495",
              "82645",
              "84292",
              "84294",
              "82646",
              "84277")

pop_84235 = c("84235",
              "84266",
              "84255",
              "84237",
              "84226")

pop_84218 = c("84218",
              "82682")

pop_72973 = c("72973",
              "74976")

pop_72808 = c("72808",
              "72336")

pop_70505 = c("70505",
              "72092",
              "72654",
              "71971",
              "72604",
              "71899",
              "72894",
              "70494",
              "73011",
              "74124",
              "74132",
              "70037")

pop_70567 = c("70567",
              "70062",
              "72620",
              "72580",
              "70284")

pop_70413 = c("70413",
              "71968",
              "72442",
              "70175",
              "72264",
              "72093",
              "70279",
              "72008",
              "70114",
              "72696",
              "72472")

pop_50837 = c("50837",
              "50858",
              "50899",
              "50895",
              "52212")

pop_20198 = c("20198",
              "20199")

pop_10037 = c("10037",
              "12464",
              "12012",
              "10036",
              "12011")

pop_10090 = c("10090",
              "10091",
              "11870")

# identify populations
bd_population = bd_sample %>%
  # filter(!is.na(svl_mm) | !is.na(body_mass_g)) %>%
  filter(!is.na(svl_mm)) %>%
  mutate(population = site,
         population = ifelse(country == "brazil", region, site),
         population = case_when(
           site %in% pop_84313 ~ "pop_84313",
           site %in% pop_84309 ~ "pop_84309",
           site %in% pop_84235 ~ "pop_84235",
           site %in% pop_84218 ~ "pop_84218",
           site %in% pop_72973 ~ "pop_72973",
           site %in% pop_72808 ~ "pop_72808",
           site %in% pop_70505 ~ "pop_70505",
           site %in% pop_70567 ~ "pop_70567",
           site %in% pop_70413 ~ "pop_70413",
           site %in% pop_50837 ~ "pop_50837",
           site %in% pop_20198 ~ "pop_20198",
           site %in% pop_10037 ~ "pop_10037",
           site %in% pop_10090 ~ "pop_10090",
           .default = population
         ))

# determine populations which meet criteria
bd_population_valid = bd_population %>%
  group_by(population, region, country, taxon_capture) %>%
  summarize(sample_count = n(),
            year_first = year(min(date)),
            year_last = year(max(date)),
            year_count = n_distinct(year(date))) %>%
  filter(sample_count >= 60,
         year_count >= 3) %>%
  group_by(taxon_capture) %>%
  mutate(population_count = n()) %>%
  ungroup() %>%
  filter(population_count >= 2) %>%
  collect() %>%
  arrange(desc(sample_count), taxon_capture)

# join with bd data, select
bd_sample_valid = bd_population %>%
    filter(population %in% bd_population_valid$population &
             taxon_capture %in% bd_population_valid$taxon_capture) %>%
  left_join(db_bd %>%
              group_by(sample_id) %>%
              summarise(bd_swab_count = n(),
                        bd_detected = any(detected),
                        bd_mean_its1_copies_per_swab = mean(bd_its1_copies_per_swab),
                        .groups = "drop"),
            by = "sample_id") %>%
  select(sample_id,
         bd_swab_count,
         bd_detected,
         bd_mean_its1_copies_per_swab,
         capture_id,
         taxon_capture,
         life_stage,
         svl_mm,
         body_mass_g,
         sex,
         survey_id,
         date,
         site,
         site_id,
         geographic_area,
         population,
         region,
         country
         ) %>%
  collect() %>%
  mutate(svl_mm_noise = svl_mm + runif(n(), min = -0.5, max = 0.5)) %>%
  filter(!(taxon_capture == "silverstoneia_flotator" & svl_mm > 100))

(bd_population_agg = bd_sample_valid %>%
    group_by(taxon_capture,
             country) %>%
    summarise(regions = str_flatten(sort(unique(region)), collapse = ", "),
              sample_count = n(),
              year_first = year(min(date)),
              year_last = year(max(date)),
              year_count = n_distinct(year(date)),
              pop_count = n_distinct(population)) %>%
    arrange(country, taxon_capture))
```
# Data exploration

```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 10
#| fig-height: 10

ggplot(bd_sample_valid %>%
         arrange(taxon_capture) %>%
         filter(!is.na(bd_mean_its1_copies_per_swab),
                !is.na(svl_mm_noise)), aes(x = svl_mm_noise, y = log(bd_mean_its1_copies_per_swab + 1))) + 
  geom_point(size = 0.5) +
  theme_bw() +
  facet_wrap("taxon_capture",
             ncol = 5,
             scales = "free_x") +
  xlab("Snout-vent length [mm]") +
  ylab("ln( Bd load + 1 )")

```

# SN population selection
```{r}
#| fig-width: 10
#| fig-height: 40
sn_pops = bd_sample_valid %>%
  filter(region == "california") %>%
  group_by(population, taxon_capture) %>%
  count() %>%
  arrange(desc(n))

sn_sample_valid = bd_sample_valid %>%
  filter(region == "california",
         life_stage %in% c("adult", "subadult"))

sn_summary = sn_sample_valid %>%
  group_by(population) %>%
  summarise(n = n(),
            adult_n = sum(life_stage == "adult"),
            subadult_n = sum(life_stage == "subadult"))
  

ggplot(sn_sample_valid, aes(x = date, y = log(bd_mean_its1_copies_per_swab), color = life_stage)) +
  geom_point() +
  ylab("ln( Bd_load )") +
  geom_text(data = sn_summary, 
            aes(x = as.Date("2020-01-01"), y = 23, label = paste0("n = ", n)),
            color = "#000000") +
  geom_text(data = sn_summary, 
            aes(x = as.Date("2020-01-01"), y = 20, label = paste0("adult_n = ", adult_n)),
            color = "#F8766D") +
  geom_text(data = sn_summary, 
            aes(x = as.Date("2020-01-01"), y = 17, label = paste0("subadult_n = ", subadult_n)),
            color = "#00BFC4") +
  facet_wrap("population",
             ncol = 4)

ggplot(sn_sample_valid, aes(x = date, y = log(bd_mean_its1_copies_per_swab), color = life_stage)) +
  geom_point() +
  ylab("ln( Bd_load )") +
  facet_wrap("geographic_area",
             ncol = 4) +
  geom_smooth(method = "lm")

```

# SN population analysis
```{r}

select_pops = c("70550",
                "72996",
                "pop_70413",
                "pop_70505",
                "pop_72808",
                "pop_72973",
                "pop_84218",
                "pop_84235")

sn_pn = bd_sample_valid %>%
  filter(population %in% select_pops,
         !(population == "pop_72808" & date < "2015-01-01"))
  
sn_pn_pos = sn_pn %>%
  filter(bd_mean_its1_copies_per_swab > 0)

sn_pn_summary = sn_pn %>%
  mutate(year = year(date)) %>%
  group_by(population, life_stage) %>%
  summarise(n = n(),
            min_year = min(year),
            max_year = max(year),
            year_count = n_distinct(year)) %>%
  pivot_wider(id_cols = c(population, min_year, max_year, year_count),
              values_from = n,
              names_from = life_stage)

ggplot(sn_pn_pos, aes(x = date, y = log(bd_mean_its1_copies_per_swab), color = life_stage)) +
  geom_point() +
  geom_smooth(method = "lm") +
  ylab("ln( Bd_load )") +
  facet_wrap("population")

```

# climate variables
```{r}
db_ts_temp = tbl(dbcon, Id("microclimate_data", "ts_temperature"))
db_sensor = tbl(dbcon, Id("microclimate_data", "sensor"))
db_logger = tbl(dbcon, Id("microclimate_data", "logger"))

surveys_valid = bd_sample_valid %>%
  select(survey_id, population, site_id, date) %>%
  distinct()

temp_water_5 = db_ts_temp %>%
  left_join(db_sensor, by = "sensor_id") %>%
  left_join(db_logger, by = "logger_id") %>%
  filter(microhabitat == "water",
         sensor_type == "temperature",
         height_cm == 5) %>%
  select(all_of(colnames(db_ts_temp)))  # fix this bug!

temp_water_35 = db_ts_temp %>%
  left_join(db_sensor, by = "sensor_id") %>%
  left_join(db_logger, by = "logger_id") %>%
  filter(microhabitat == "water",
         sensor_type == "temperature",
         height_cm == 35) %>%
  select(all_of(colnames(db_ts_temp)))  # fix this bug!

test_1 = microclimate_presurvey(surveys_valid, dbcon, buffer = days(14), temp_water_5, time_sensitive = FALSE, output_timezone = "America/Los_Angeles")

```