---
title: "Size vs. Bd modeling: Panamá preliminary results"
author: Cob Staines
date: 2025-09-29
format:
  html:
    toc: true
df-print: kable
bibliography: references.bib
csl: ecology.csl
---
# Introduction

## Motivation

Patterns of greater Bd load in recently metamorphosed individuals have been documented in some cases[@adams_chytridiomycosis-induced_2020] [@humphries_immune_2022] [@humphries_chytridiomycosis_2024] [@humphries_chytridiomycosis_2025] (and the opposite pattern in others[@bradley_host_2019]), with possible implications for:

  - resilience trajectory assessment (Q1)
  - population dynamics (Q2)
  - assessing Bd resistance/tolerance mechanisms (Q3)
  
... with the goal of complementing ongoing RIBBiTR research efforts

## Working hypotheses:

  a. Recently-metamorphosed individuals are more likely to be infected with Bd than older individuals across species
  b. Recently-metamorphosed individuals are more likely to carry higher Bd loads than older individuals across species

# Methods
## Setup

```{r}
#| eval: true
#| output: false
#| code-fold: true
#| code-summary: "Show setup code"

# setup
librarian::shelf(tidyverse, dbplyr, here, janitor, RPostgres, DBI, RIBBiTR-BII/ribbitrrr, brms, tidybayes, ggplot2, ggrepel, bayesplot, runner, cmdstanr, bayestestR, ropensci/rotl, ggtree, ape, patchwork, viridis, terra, emmeans, rio)

```
## Data selection
### Classifying by life stage
These analyses were carried out for post-metamorphic individuals only. Individuals were classified as "adult" or "subadult" with a threshold snout-vent-length (SVL [mm]), identified independently for each species from the literature (see data collection code for specific thresholds used and their sources).
  
### Classification by population
"Populations" were defined at the region level within Panamá. In the Sierra Nevada, sites were grouped where evidence supported connectivity at he scale of individuals. Elsewhere in North America, populations were defined at the site level.

### Species selection criteria:
Species were included if supported by least 14 swabbed adults and subadults (each).

### Bd density and Zero-cutoff
I scaled all Bd quantities by the Snout-Vent-Length (mm) to arrive at a Bd density per linear distance of skin swabbed (within species), to correct for linear effects of larger swabbing area in larger frogs. I also set any Bd ITS copy numbers < 10 to zero, to represent uncertainty in qPCR results and support consistency in zero-inflation modeling.

### RACA
```{r}

# csv data
##  RACA data
raca_clean = read.csv(here("data", "JPSL_RACA_Swabbed_Individuals_2026-03-03.csv")) %>%
  clean_names() %>%
  filter(life_stage %in% c("A", "S", "Y")) %>%
  rename(site = site_name,
         body_mass_g = weight_g,
         site_utme = utme_site,
         site_utmn = utmn_site,
         bd_its_copies = bd_swab_results_its) %>%
  mutate(life_stage_simple = case_when(
    svl_mm >= 50 ~ "adult",
    svl_mm < 50 ~ "subadult"),
    date = as.Date(date),
    site = tolower(gsub(" ", "_", site)),
    site_utm_zone = 11,  # verify
    population = site,
    region = "W",
    country = "usa",
    project = "RACA",
    taxon_capture = "rana_cascadae") %>%
  filter(!is.na(bd_its_copies))

raca_population_valid = raca_clean %>%
  group_by(population) %>%
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop")
# %>%
#   filter(n_adult > 0,
#          n_subadult > 0)

raca_sites = raca_clean %>%
  select(site,
         site_utme,
         site_utmn) %>%
  distinct()

# convert to lat lon
raca_points = cbind(raca_sites$site_utme,
                    raca_sites$site_utmn)

v_utm = vect(raca_points, crs="+proj=utm +zone=10 +datum=WGS84  +units=m")
v_ll = project(v_utm, "+proj=longlat +datum=WGS84")
v_lonlat = geom(v_ll)[, c("x", "y")]

raca_sites_ll = raca_sites %>%
  bind_cols(v_lonlat) %>%
  rename(site_longitude = x,
         site_latitude = y) %>%
  select(site,
         site_latitude,
         site_longitude)

raca_valid = raca_clean %>%
  filter(population %in% raca_population_valid$population) %>%
  left_join(raca_sites_ll, by = "site") %>%
  select(taxon_capture,
         date,
         site,
         population,
         bd_its_copies,
         life_stage_simple,
         life_stage,
         sex,
         svl_mm,
         site_latitude,
         site_longitude,
         region,
         country,
         project)

```

### RALU
```{r}
ralu_raw = import_list(here("data", "Data Russell et al 2010 DAO.xlsx"))

# site coordinates
ralu_sites = ralu_raw[[2]] %>%
  clean_names() %>%
  mutate(site = tolower(pond_name))

ralu_points = cbind(ralu_sites$x_nad83_zone11n,
                    ralu_sites$y_nad83_zone11n)

w_utm = vect(ralu_points, crs="+proj=utm +zone=11 +datum=WGS84  +units=m")
w_ll = project(w_utm, "+proj=longlat +datum=WGS84")
w_lonlat = geom(w_ll)[, c("x", "y")]

ralu_sites_ll = ralu_sites %>%
  bind_cols(w_lonlat) %>%
  rename(site_longitude = x,
         site_latitude = y) %>%
  select(site,
         site_latitude,
         site_longitude)

ralu_clean = ralu_raw[[4]] %>%
  clean_names() %>%
  rename(life_stage = age_class) %>%
  mutate(taxon_capture = "rana_luteiventris",
         date = as.Date(bd_sample_date),
         site = tolower(pond_name),
         population = site,
         region = "NW",
         country = "usa",
         life_stage_simple = recode_values(life_stage,
                                           "ADULT" ~ "adult",
                                           "JUVENILE" ~ "subadult",
                                           "METAMORPH" ~ "metamorph"),
         bd_its_copies = actual_bd_load * 60,
         sex = tolower(sex),
         project = "RALU",
         svl_mm = 25  # using smallest age glass SVL, as researchers used standardized swab lengths.
         ) %>%
  left_join(ralu_sites_ll, by = "site") %>%
  filter(!is.na(life_stage_simple),
         !is.na(bd_its_copies),
         life_stage_simple %in% c("adult", "subadult"))

ralu_population_valid = ralu_clean %>%
  group_by(population) %>%
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop")
# %>%
#   filter(n_adult > 0,
#          n_subadult > 0)

ralu_valid = ralu_clean %>%
  select(taxon_capture,
         date,
         site,
         population,
         bd_its_copies,
         life_stage_simple,
         life_stage,
         sex,
         svl_mm,
         site_latitude,
         site_longitude,
         region,
         country,
         project)
```

### USGS
```{r}

## laod USGS data
usgs_raw = read.csv(here("data", "USGSBsalBdDataFull.csv"))

usgs_clean = usgs_raw %>%
  clean_names() %>%
  rename(site_latitude = latitude,
         site_longitude = longitude,
         svl_mm = body_length,
         life_stage = age,
         region = reg,
         bd_its_copies = copy_num) %>%
  mutate(taxon_capture = tolower(gsub(" ", "_", scientific_name)),
         life_stage_simple = tolower(life_stage),
         life_stage_simple = recode_values(life_stage_simple,
                                        "juvenile" ~ "subadult",
                                        default = life_stage_simple),
         date = mdy(collection_date),
         site = paste0("usgs_", site_id),
         population = site,
         project = "usgs",
         country = "usa") %>%
  filter(!is.na(taxon_capture),
         !is.na(bd_its_copies),
         !is.na(svl_mm),
         life_stage_simple %in% c("subadult", "adult"))

usgs_taxon_population_valid = usgs_clean %>%
  group_by(taxon_capture, population) %>%
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop")
# %>%
#   filter(n_adult > 0,
#          n_subadult > 0)

usgs_taxon_valid = usgs_clean %>%
  inner_join(usgs_taxon_population_valid %>%
               select(taxon_capture,
                      population), by = c("taxon_capture", "population")) %>%
  group_by(taxon_capture) %>%
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop") %>%
  filter(n_adult >= 14,
         n_subadult >= 14,
         taxon_capture != "lithobates_chiricahuensis")

usgs_valid = usgs_clean %>%
  inner_join(usgs_taxon_population_valid %>%
               select(taxon_capture,
                      population), by = c("taxon_capture", "population")) %>%
  filter(taxon_capture %in% usgs_taxon_valid$taxon_capture) %>%
  select(taxon_capture,
         svl_mm,
         sex,
         life_stage_simple,
         life_stage,
         date,
         bd_its_copies,
         population,
         site,
         site_latitude,
         site_longitude,
         region,
         country,
         project)

usgs_spp = usgs_valid %>%
  group_by(taxon_capture,
           life_stage_simple) %>%
  count()

```

### RIBBiTR, etc.
```{r}

## SN populations
sn_populations = read.csv(here("data", "sn_site_pop_2_2025-10-31.csv")) %>%
  rename(population = pop_id) %>%
  mutate(population_sn = if_else(!is.na(population), paste0(as.character(population), "_pop"), site),
         site = as.character(site)) %>%
  select(site,
         population_sn)


population_valid_pa = c("altos_de_campana",
                        "caribbean",
                        "el_cope",
                        "el_valle",
                        "santa_fe")

# RIBBiTR data

## Connect to DB
dbcon = hopToDB("ribbitr", hopReg = FALSE)

# table pointers
db_bd = tbl(dbcon, Id("survey_data", "bd_qpcr_results"))
db_sample = tbl(dbcon, Id("survey_data", "sample"))
db_capture = tbl(dbcon, Id("survey_data", "capture"))
db_survey = tbl(dbcon, Id("survey_data", "survey"))
db_visit = tbl(dbcon, Id("survey_data", "visit"))
db_project = tbl(dbcon, Id("survey_data", "project"))
db_site = tbl(dbcon, Id("survey_data", "site"))
db_region = tbl(dbcon, Id("survey_data", "region"))
db_country = tbl(dbcon, Id("survey_data", "country"))


# preliminary data
bd_sample = db_sample %>%
  right_join(db_capture, by = "capture_id") %>%
  left_join(db_survey, by = "survey_id") %>%
  left_join(db_visit, by = "visit_id") %>%
  left_join(db_site, by = "site_id") %>%
  left_join(db_region, by = "region_id") %>%
  left_join(db_country, by = "country_id") %>%
  left_join(db_project, by = "project_id") %>%
  filter(sample_type == "bd",
         !is.na(taxon_capture)) %>%
  inner_join(db_bd %>%
              filter(!is.na(bd_detected)) %>%
               # limit to one qPCR result per sample
              group_by(sample_id) %>%
              slice_sample(n = 1) %>%
              ungroup() %>%
              select(sample_id,
                     bd_detected,
                     bd_its1_copies_per_swab), by = "sample_id")

# define populations
bd_population = bd_sample %>%
  left_join(sn_populations, by = "site", copy = TRUE) %>%
  mutate(population = case_when(
    country %in% c("brazil", "panama") ~ region,
    region == "california" ~ population_sn,
    .default = site
  )) %>%
  filter(country != "panama" | (country == "panama" & population %in% population_valid_pa),
         !is.na(site_latitude))

# define life stage simple
bd_life_stage = bd_population %>%
  filter(life_stage %in% c("adult",
                           "subadult",
                           "juvenile",
                           NA)) %>%
  mutate(life_stage_simple = case_when(
    # us
    # https://animaldiversity.org/accounts/Acris_blanchardi/
    ((taxon_capture == "acris_blanchardi") & (svl_mm >= 16)) ~ "adult",
    ((taxon_capture == "acris_blanchardi") & (svl_mm < 16)) ~ "subadult",
    # Wright and Wright 1949
    ((taxon_capture == "anaxyrus_americanus") & (svl_mm >= 54)) ~ "adult",
    ((taxon_capture == "anaxyrus_americanus") & (svl_mm < 54)) ~ "subadult",
    # Lannoo 2005 (Amphibian Declines: The Conservation Status of United States Species)
    ((taxon_capture == "anaxyrus_canorus") & (svl_mm >= 48)) ~ "adult",
    ((taxon_capture == "anaxyrus_canorus") & (svl_mm < 48)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "anaxyrus_debilis") & (svl_mm >= 37)) ~ "adult",
    ((taxon_capture == "anaxyrus_debilis") & (svl_mm < 37)) ~ "subadult",
    # Conant and Collins 1991
    ((taxon_capture == "anaxyrus_fowleri") & (svl_mm >= 51)) ~ "adult",
    ((taxon_capture == "anaxyrus_fowleri") & (svl_mm < 51)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "hyla_chrysoscelis") & (svl_mm >= 32)) ~ "adult",
    ((taxon_capture == "hyla_chrysoscelis") & (svl_mm < 32)) ~ "subadult",
    # Conant & Collins 1998
    ((taxon_capture == "hyla_cinerea") & (svl_mm >= 32)) ~ "adult",
    ((taxon_capture == "hyla_cinerea") & (svl_mm < 32)) ~ "subadult",
    # Johnson 1987
    ((taxon_capture == "hyla_versicolor") & (svl_mm >= 32)) ~ "adult",
    ((taxon_capture == "hyla_versicolor") & (svl_mm < 32)) ~ "subadult",
    # Minton, 2001 / Lanoo 2005
    ((taxon_capture == "lithobates_blairi") & (svl_mm >= 51)) ~ "adult",
    ((taxon_capture == "lithobates_blairi") & (svl_mm < 51)) ~ "subadult",
    # Platz and Mecham 1963
    ((taxon_capture == "lithobates_chiricahuensis") & (svl_mm >= 57)) ~ "adult",
    ((taxon_capture == "lithobates_chiricahuensis") & (svl_mm < 57)) ~ "subadult",
    # Minton 2001
    ((taxon_capture == "lithobates_palustris") & (svl_mm >= 43)) ~ "adult",
    ((taxon_capture == "lithobates_palustris") & (svl_mm < 43)) ~ "subadult",
    # Conant and Collins, 1991
    ((taxon_capture == "lithobates_sphenocephalus") & (svl_mm >= 51)) ~ "adult",
    ((taxon_capture == "lithobates_sphenocephalus") & (svl_mm < 51)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "pseudacris_crucifer") & (svl_mm >= 20)) ~ "adult",
    ((taxon_capture == "pseudacris_crucifer") & (svl_mm < 20)) ~ "subadult",
    # Hulse, C. and McCoy C. J. and Ellen Censky ,1998.
    ((taxon_capture == "pseudacris_feriarum") & (svl_mm >= 19)) ~ "adult",
    ((taxon_capture == "pseudacris_feriarum") & (svl_mm < 19)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "pseudacris_regilla") & (svl_mm >= 25)) ~ "adult",
    ((taxon_capture == "pseudacris_regilla") & (svl_mm < 25)) ~ "subadult",
    # Howard 1981
    ((taxon_capture == "rana_catesbeiana") & (svl_mm >= 95)) ~ "adult",
    ((taxon_capture == "rana_catesbeiana") & (svl_mm < 95)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "rana_clamitans") & (svl_mm >= 60)) ~ "adult",
    ((taxon_capture == "rana_clamitans") & (svl_mm < 60)) ~ "subadult",
    # Knapp
    ((taxon_capture == "rana_muscosa") & is.na(svl_mm)) ~ life_stage,
    ((taxon_capture == "rana_muscosa") & (svl_mm >= 40)) ~ "adult",
    ((taxon_capture == "rana_muscosa") & (svl_mm < 40)) ~ "subadult",
    # https://www.nrc.gov/docs/ML1428/ML14282A848.pdf
    ((taxon_capture == "rana_pipiens") & (svl_mm >= 51)) ~ "adult",
    ((taxon_capture == "rana_pipiens") & (svl_mm < 51)) ~ "subadult",
    # Knapp
    ((taxon_capture == "rana_sierrae") & is.na(svl_mm)) ~ life_stage,
    ((taxon_capture == "rana_sierrae") & (svl_mm >= 40)) ~ "adult",
    ((taxon_capture == "rana_sierrae") & (svl_mm < 40)) ~ "subadult",
    # Martof 1963
    ((taxon_capture == "rana_sylvatica") & (svl_mm >= 37)) ~ "adult",
    ((taxon_capture == "rana_sylvatica") & (svl_mm < 37)) ~ "subadult",
    # Conant and Collins, 1991
    ((taxon_capture == "scaphiopus_holbrookii") & (svl_mm >= 44)) ~ "adult",
    ((taxon_capture == "scaphiopus_holbrookii") & (svl_mm < 44)) ~ "subadult",
    # Stebbins, 1985
    ((taxon_capture == "spea_multiplicata") & (svl_mm >= 37)) ~ "adult",
    ((taxon_capture == "spea_multiplicata") & (svl_mm < 37)) ~ "subadult",
    
    # brasil
    # Heyer, W. R., and R. B. Cocroft. 1986
    ((taxon_capture == "hylodes_phyllodes") & (svl_mm >= 27.5)) ~ "adult",
    ((taxon_capture == "hylodes_phyllodes") & (svl_mm < 27.5)) ~ "subadult",
    # Kwet, A., and M. Solé. 2005
    ((taxon_capture == "ischnocnema_henselii") & (svl_mm >= 23.0)) ~ "adult",
    ((taxon_capture == "ischnocnema_henselii") & (svl_mm < 23.0)) ~ "subadult",
    # Kwet, A., and M. Solé. 2005
    ((taxon_capture == "brachycephalus_pitanga") & (svl_mm >= 10.8)) ~ "adult",
    ((taxon_capture == "brachycephalus_pitanga") & (svl_mm <10.8)) ~ "subadult",
    ((taxon_capture == "boana_faber") & (svl_mm >= 70)) ~ "adult",
    ((taxon_capture == "boana_faber") & (svl_mm < 70)) ~ "subadult",
   
    # country == "brazil" & life_stage == "adult" ~ "adult",
    # country == "brazil" & life_stage %in% c("subadult", "juvenile") ~ "subadult",
    
    # panama
    # Savage 1972 & Crump 1988
    ((taxon_capture == "atelopus_varius") & (svl_mm >= 25)) ~ "adult",
    ((taxon_capture == "atelopus_varius") & (svl_mm < 25)) ~ "subadult",
    # Lynch and Myers 1983
    ((taxon_capture == "craugastor_crassidigitus") & (svl_mm >= 20.2)) ~ "adult",
    ((taxon_capture == "craugastor_crassidigitus") & (svl_mm < 20.2)) ~ "subadult",
    # Lynch and Myers 1983
    ((taxon_capture == "craugastor_fitzingeri") & (svl_mm >= 23.5)) ~ "adult",
    ((taxon_capture == "craugastor_fitzingeri") & (svl_mm < 23.5)) ~ "subadult",
    # Amphibiaweb
    ((taxon_capture == "colostethus_panamansis") & (svl_mm >= 18.8)) ~ "adult",
    ((taxon_capture == "colostethus_panamansis") & (svl_mm < 18.8)) ~ "subadult",
    # Leenders 2001
    ((taxon_capture == "dendrobates_auratus") & (svl_mm >= 25)) ~ "adult",
    ((taxon_capture == "dendrobates_auratus") & (svl_mm < 25)) ~ "subadult",
    # Amphibiaweb
    ((taxon_capture == "espadarana_prosoblepon") & (svl_mm >= 21)) ~ "adult",
    ((taxon_capture == "espadarana_prosoblepon") & (svl_mm < 21)) ~ "subadult",
    # Basante 2025
    ((taxon_capture == "lithobates_warszewitschii") & (svl_mm >= 35)) ~ "adult",
    ((taxon_capture == "lithobates_warszewitschii") & (svl_mm < 35)) ~ "subadult",
    # Wang 2008
    ((taxon_capture == "pristimantis_ridens") & (svl_mm >= 16)) ~ "adult",
    ((taxon_capture == "pristimantis_ridens") & (svl_mm < 16)) ~ "subadult",
    # Savage 2002
    ((taxon_capture == "rhaebo_haematiticus") & (svl_mm >= 42)) ~ "adult",
    ((taxon_capture == "rhaebo_haematiticus") & (svl_mm < 42)) ~ "subadult",
    # Easteal 1963
    ((taxon_capture == "rhinella_marina") & (svl_mm >= 85)) ~ "adult",
    ((taxon_capture == "rhinella_marina") & (svl_mm < 85)) ~ "subadult",
    # Savage 2002
    ((taxon_capture == "sachatamia_albomaculata") & (svl_mm >= 20.5)) ~ "adult",
    ((taxon_capture == "sachatamia_albomaculata") & (svl_mm < 20.5)) ~ "subadult",
    # Ibunez & Smith 1995
    ((taxon_capture == "silverstoneia_flotator") & (svl_mm >= 15.1)) ~ "adult",
    ((taxon_capture == "silverstoneia_flotator") & (svl_mm < 15.1)) ~ "subadult",
    # Grant & Myers 2013
    ((taxon_capture == "silverstoneia_nubicola") & (svl_mm >= 15.4)) ~ "adult",
    ((taxon_capture == "silverstoneia_nubicola") & (svl_mm < 15.4)) ~ "subadult",
    # Duellman, W. E. 2001
    ((taxon_capture == "smilisca_sila") & (svl_mm >= 31.6)) ~ "adult",
    ((taxon_capture == "smilisca_sila") & (svl_mm < 31.6)) ~ "subadult",
    .default = NA_character_
  )) %>%
  filter(!is.na(life_stage_simple),
         !is.na(svl_mm))

# identify valid species & populations
## drop pops without both adults and juveniles
## drop species without 14 or more adults and juveniles each
pop_taxa_valid = bd_life_stage %>%
  group_by(taxon_capture, population) %>%
  summarize(n_adult = sum(as.integer(life_stage_simple == "adult")),
            n_subadult = sum(as.integer(life_stage_simple == "subadult")),
            .groups = "drop")
# %>%
#   filter(n_adult > 0,
#          n_subadult > 0)

taxa_valid = bd_life_stage %>%
  inner_join(pop_taxa_valid %>%
               select(taxon_capture,
                      population), by = c("taxon_capture", "population")) %>%
  group_by(taxon_capture) %>%
  summarize(n_adult = sum(as.integer(life_stage_simple == "adult"), na.rm = TRUE),
            n_subadult = sum(as.integer(life_stage_simple == "subadult"), na.rm = TRUE),
            .groups = "drop") %>%
  filter(n_adult >= 14,
         n_subadult >= 14)

# identify valid samples
bd_sample_valid = bd_life_stage %>%
  mutate(year = year(date),
         project = project_id) %>%
  inner_join(pop_taxa_valid, by = c("taxon_capture", "population")) %>%
  inner_join(taxa_valid, by = "taxon_capture") %>%
  # limit to 1 bd sample per capture
  group_by(capture_id) %>%
  slice_sample(n = 1) %>%
  ungroup() %>%
  select(sample_id,
         bd_detected,
         bd_its1_copies_per_swab,
         sample_name,
         capture_id,
         taxon_capture,
         life_stage,
         life_stage_simple,
         svl_mm,
         body_mass_g,
         sex,
         date,
         year,
         site,
         site_latitude,
         site_longitude,
         population,
         region,
         country,
         project) %>%
  collect()

# final data selection and cleaning for model
ribbitr_valid = bd_sample_valid %>%
  rename(bd_its_copies = bd_its1_copies_per_swab) %>%
  select(bd_its_copies,
         taxon_capture,
         life_stage,
         life_stage_simple,
         svl_mm,
         body_mass_g,
         sex,
         date,
         year,
         site,
         site_latitude,
         site_longitude,
         population,
         region,
         country,
         project)

```

### Combine datasets
```{r}

column_check = compare_df_cols(raca_valid,
                               ralu_valid,
                               usgs_valid,
                               ribbitr_valid)

bd_model = bind_rows(raca_valid,
                     ralu_valid,
                     usgs_valid,
                     ribbitr_valid) %>%
  mutate(bd_detected_co = bd_its_copies >= 10,
         bd_load_co = if_else(bd_its_copies < 10,
                              0,
                              bd_its_copies),
         bd_load_co_density = bd_load_co / svl_mm)

(bd_model_stats = bd_model %>%
  group_by(taxon_capture, country) %>%
  summarise(regions = paste(sort(unique(region)), collapse = ", "),
            population_n = n_distinct(population),
            site_n = n_distinct(site),
            year_first = year(min(date)),
            year_last = year(max(date)),
            year_count = n_distinct(year(date)),
            adult_n = sum(life_stage_simple == "adult"),
            subadult_n = sum(life_stage_simple == "subadult"),
            .groups = "drop") %>%
  arrange(country,
          taxon_capture))

# write_csv(bd_model_stats, here("data", "size_bd_modeling_unified_stats.csv"))

```

## Environmental data
To support comparing Bd prevalence and loads between sites and species across a range of environmental conditions, I incorporated local (~5 km) daily maximum temperature (CHIRTS) and daily precipitation (CHIRPS) time series. Following Sonn et al. 2019 [sonn2019effects], I created 15-day and 30-day means (for tempurature) and sums (for precipitation). Preliminary model fitting showed stronger relationships between Bd prevalence and 15-day lag products, while Bd load showed stronger relationships with 30-day lag products, similar to Sonn et al. 2019.

### Spatio-temporal coverage
```{r}
#| eval: false
#| warning: false
#| code-fold: true
#| code-summary: "Code to export spatio-temporal coverage"

temporal_report = bd_model %>%
  group_by(year) %>%
  summarise(n_obs = n(),
            n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "adult"),
            n_taxa = n_distinct(taxon_capture),
            n_site = n_distinct(site),
            n_pop = n_distinct(population),
            n_country = n_distinct(country),
            .groups = "drop")



# export site list and coords
temporal_buffer = 60  # days

site_space_time = bd_model %>%
  group_by(site,
           site_latitude,
           site_longitude) %>%
  summarize(start_date = min(date) - days(temporal_buffer),
            end_date = max(date),
            .groups = "drop") %>%
  rename(site_name = site,
         latitude = site_latitude,
         longitude = site_longitude)

# site_space_time_ralu = site_space_time %>%
#   filter(project == "RALU")

site_space_time_before_2017 = site_space_time %>%
  filter(start_date <= "2016-12-31") %>%
  mutate(end_date = if_else(end_date > "2016-12-31", as.Date("2016-12-31"), end_date))

site_space_time_after_2017 = site_space_time %>%
  filter(end_date >= "2017-01-01") %>%
  mutate(start_date = if_else(start_date < "2017-01-01", as.Date("2017-01-01"), start_date))

# write_csv(site_space_time, here("python", "environmental_timeseries_pull", "inputs", paste0("site_coords_dates_all_", today(), ".csv")))
# write_csv(site_space_time_before_2017, here("python", "environmental_timeseries_pull", "inputs", paste0("site_coords_dates_through_2016_", today(), ".csv")))
# write_csv(site_space_time_after_2017, here("python", "environmental_timeseries_pull", "inputs", paste0("site_coords_dates_2017_beyond_", today(), ".csv")))

```

### Join with env. data
```{r}
#| eval: false
#| warning: false
#| code-fold: true
#| code-summary: "Show environmental data mergin code"
#| 
# env_in = read_csv(here("python", "gee_env_pull", "data", "gee_chirps_modis_data_all.csv"))
env_in = read_csv(here("python", "environmental_timeseries_pull", "outputs", "all_sites_combined_chirps_chirts_2026-03-17.csv")) %>%
  rename(site = site_name,
         precipitation_mm = precip_mm,
         temperature_max_c = tmax_c)

sliding_window <- function(df, ts_col_name, window_size, f, result_col_name) {
  ts_sym <- sym(ts_col_name)
  res_sym <- sym(result_col_name)
  
  df %>%
    group_modify(~{
      n <- nrow(.x)
      ts_vals <- pull(.x, !!ts_sym)
      agg_vals <- map_dbl(seq_len(n), function(i) {
        if (i >= window_size) {
          window_vals <- ts_vals[(i - window_size + 1):i]
          f(window_vals)
        } else {
          NA_real_
        }
      })
      .x[[result_col_name]] <- agg_vals
      .x
    })
}

rolling_aggregate <- function(x, date, n_days, fun) {
  runner(
    x   = x,
    idx = date,
    k   = paste0(n_days, " days"),
    f   = fun
  )
}


n_val = function(x) sum(!is.na(x))

na_mean = function(x) {
  mean(x, na.rm = TRUE)
}

na_max = function(x) {
  max(x, na.rm = TRUE)
}

na_sum = function(x) {
  sum(x, na.rm = TRUE)
}

env_agg = env_in %>%
  arrange(site, date) %>%
  filter(!(is.na(precipitation_mm) & is.na(temperature_max_c))) %>%
  group_by(site) %>%
  mutate(precip_mm_15_day_sum = rolling_aggregate(precipitation_mm, date, 15, na_sum),
         precip_mm_30_day_sum = rolling_aggregate(precipitation_mm, date, 30, na_sum),
         temp_c_15_day_mean = rolling_aggregate(temperature_max_c, date, 15, na_mean),
         temp_c_30_day_mean = rolling_aggregate(temperature_max_c, date, 30, na_mean)) %>%
  ungroup()

write_csv(env_agg, here("data", "unified_env_agg.csv"))
env_agg = read_csv(here("data", "unified_env_agg.csv"))


bd_model_env = bd_model %>%
  left_join(env_agg, by = c("site", "date"))

save(bd_model_env, file = here("data", paste0("bd_model_env_unified_", today(), ".RData")))

```

## Load saved data
```{r}
#| eval: true
#| warning: false
#| echo: false
#| code-fold: true
#| code-summary: "Show data loading code"

# load bd_model_env
load(here("data", "bd_model_env_unified_2026-03-17.RData")) 

# checks
sum(is.na(bd_model_env$bd_load_co_density))
sum(is.na(bd_model_env$precip_mm_30_day_sum))
sum(is.na(bd_model_env$temperature_max_c))

bd_model_env_stats = bd_model_env %>%
  group_by(taxon_capture, country) %>%
  summarise(regions = paste(sort(unique(region)), collapse = ", "),
            population_n = n_distinct(population),
            site_n = n_distinct(site),
            year_first = year(min(date)),
            year_last = year(max(date)),
            year_count = n_distinct(year(date)),
            adult_n = sum(life_stage_simple == "adult"),
            adult_n_pos = sum(life_stage_simple == "adult" & bd_detected_co),
            subadult_n = sum(life_stage_simple == "subadult"),
            subadult_n_pos = sum(life_stage_simple == "subadult" & bd_detected_co),
            .groups = "drop") %>%
  arrange(country,
          taxon_capture)

```

# Modeling
I modeled Bd load using a Baysian zero-inflated (or "hurdle") log-normal model [@schrock2025fungi] with fixed effects of species, life stage, and recent precipitation and temperature, and considering population, species, life stage and recent precipitation and temperature as a random effects. The "hurdle" infection probability (hu) and the variability in Bd load (sigma) are allowed to vary with life stage.

## Model setup
```{r}
#| code-fold: true
#| code-summary: "Show model setup code"

# unified, scaled model
brms_scales = bd_model_env %>%
  group_by() %>%
  summarize(precip_mm_15_day_sum_mean = mean(precip_mm_15_day_sum),
            precip_mm_15_day_sum_sd = sd(precip_mm_15_day_sum),
            precip_mm_30_day_sum_mean = mean(precip_mm_30_day_sum),
            precip_mm_30_day_sum_sd = sd(precip_mm_30_day_sum),
            temp_c_15_day_mean_mean = mean(temp_c_15_day_mean),
            temp_c_15_day_mean_sd = sd(temp_c_15_day_mean),
            temp_c_30_day_mean_mean = mean(temp_c_30_day_mean),
            temp_c_30_day_mean_sd = sd(temp_c_30_day_mean),
            precip_mm_15_day_sum2_mean = mean(precip_mm_15_day_sum ^ 2),
            precip_mm_15_day_sum2_sd = sd(precip_mm_15_day_sum ^ 2),
            precip_mm_30_day_sum2_mean = mean(precip_mm_30_day_sum ^ 2),
            precip_mm_30_day_sum2_sd = sd(precip_mm_30_day_sum ^ 2),
            temp_c_15_day_mean2_mean = mean(temp_c_15_day_mean ^ 2),
            temp_c_15_day_mean2_sd = sd(temp_c_15_day_mean ^ 2),
            temp_c_30_day_mean2_mean = mean(temp_c_30_day_mean ^ 2),
            temp_c_30_day_mean2_sd = sd(temp_c_30_day_mean ^ 2))

brms_rescale = function(df, scales) {
  df_out = df %>%
    mutate(precip_mm_15_z = (precip_mm_15_day_sum - scales$precip_mm_15_day_sum_mean)/scales$precip_mm_15_day_sum_sd,
           precip_mm_30_z = (precip_mm_30_day_sum - scales$precip_mm_30_day_sum_mean)/scales$precip_mm_30_day_sum_sd,
           temp_c_15_z = (temp_c_15_day_mean - scales$temp_c_15_day_mean_mean)/scales$temp_c_15_day_mean_sd,
           temp_c_30_z = (temp_c_30_day_mean - scales$temp_c_30_day_mean_mean)/scales$temp_c_30_day_mean_sd,
           precip_mm_15_z2 = (precip_mm_15_day_sum^2 - scales$precip_mm_15_day_sum2_mean)/scales$precip_mm_15_day_sum2_sd,
           precip_mm_30_z2 = (precip_mm_30_day_sum^2 - scales$precip_mm_30_day_sum2_mean)/scales$precip_mm_30_day_sum2_sd,
           temp_c_15_z2 = (temp_c_15_day_mean^2 - scales$temp_c_15_day_mean2_mean)/scales$temp_c_15_day_mean2_sd,
           temp_c_30_z2 = (temp_c_30_day_mean^2 - scales$temp_c_30_day_mean2_mean)/scales$temp_c_30_day_mean2_sd
           )
}

brms_data = bd_model_env %>%
  mutate(
    precip_mm_30_z = as.numeric(scale(precip_mm_30_day_sum, center = TRUE, scale = TRUE)),
    temp_c_30_z = as.numeric(scale(temp_c_30_day_mean, center = TRUE, scale = TRUE)),
    precip_mm_15_z = as.numeric(scale(precip_mm_15_day_sum, center = TRUE, scale = TRUE)),
    temp_c_15_z = as.numeric(scale(temp_c_15_day_mean, center = TRUE, scale = TRUE)),
    precip_mm_30_z2 = as.numeric(scale(precip_mm_30_day_sum ^ 2, center = TRUE, scale = TRUE)),
    temp_c_30_z2 = as.numeric(scale(temp_c_30_day_mean ^ 2, center = TRUE, scale = TRUE)),
    precip_mm_15_z2 = as.numeric(scale(precip_mm_15_day_sum ^ 2, center = TRUE, scale = TRUE)),
    temp_c_15_z2 = as.numeric(scale(temp_c_15_day_mean ^ 2, center = TRUE, scale = TRUE)),
  #   ##
  #   precip_mm_30_sp_z = (precip_mm_30_day_sum - precip_mm_30_day_sum_spmean) / precip_mm_30_day_sum_spsd,
  #   temp_c_30_sp_z = (temp_c_30_day_mean - temp_c_30_day_mean_spmean) / temp_c_30_day_mean_spsd,
  #   precip_mm_15_sp_z = (precip_mm_15_day_sum - precip_mm_15_day_sum_spmean) / precip_mm_15_day_sum_spsd,
  #   temp_c_15_sp_z = (temp_c_15_day_mean - temp_c_15_day_mean_spmean) / temp_c_15_day_mean_spsd,
  #   precip_mm_30_sp_z2 = (precip_mm_30_day_sum ^ 2 - precip_mm_30_day_sum2_spmean) / precip_mm_30_day_sum2_spsd,
  #   temp_c_30_sp_z2 = (temp_c_30_day_mean ^ 2 - temp_c_30_day_mean2_spmean) / temp_c_30_day_mean2_spsd,
  #   precip_mm_15_sp_z2 = (precip_mm_15_day_sum ^ 2 - precip_mm_15_day_sum2_spmean) / precip_mm_15_day_sum2_spsd,
  #   temp_c_15_sp_z2 = (temp_c_15_day_mean ^ 2 - temp_c_15_day_mean2_spmean) / temp_c_15_day_mean2_spsd
  ) %>%
  select(taxon_capture,
         everything())

# check z-scored means and sds
peace = brms_data %>%
  summarize(t_15_mean = mean(temp_c_15_z),
            t_15_sd = sd(temp_c_15_z),
            t_30_mean = mean(temp_c_30_z),
            t_30_sd = sd(temp_c_30_z),
            p_15_mean = mean(precip_mm_15_z),
            p_15_sd = sd(precip_mm_15_z),
            p_30_mean = mean(precip_mm_30_z),
            p_30_sd = sd(precip_mm_30_z),
            t2_15_mean = mean(temp_c_15_z2),
            t2_15_sd = sd(temp_c_15_z2),
            t2_30_mean = mean(temp_c_30_z2),
            t2_30_sd = sd(temp_c_30_z2),
            p2_15_mean = mean(precip_mm_15_z2),
            p2_15_sd = sd(precip_mm_15_z2),
            p2_30_mean = mean(precip_mm_30_z2),
            p2_30_sd = sd(precip_mm_30_z2))

# hope = brms_data %>%
#   group_by(taxon_capture) %>%
#   summarize(t_15_sp_mean = mean(temp_c_15_sp_z),
#             t_15_sp_sd = sd(temp_c_15_sp_z),
#             t_30_sp_mean = mean(temp_c_30_sp_z),
#             t_30_sp_sd = sd(temp_c_30_sp_z),
#             p_15_sp_mean = mean(precip_mm_15_sp_z),
#             p_15_sp_sd = sd(precip_mm_15_sp_z),
#             p_30_sp_mean = mean(precip_mm_30_sp_z),
#             p_30_sp_sd = sd(precip_mm_30_sp_z),
#             t2_15_sp_mean = mean(temp_c_15_sp_z2),
#             t2_15_sp_sd = sd(temp_c_15_sp_z2),
#             t2_30_sp_mean = mean(temp_c_30_sp_z2),
#             t2_30_sp_sd = sd(temp_c_30_sp_z2),
#             p2_15_sp_mean = mean(precip_mm_15_sp_z2),
#             p2_15_sp_sd = sd(precip_mm_15_sp_z2),
#             p2_30_sp_mean = mean(precip_mm_30_sp_z2),
#             p2_30_sp_sd = sd(precip_mm_30_sp_z2))

model_priors <- c(
  # b
  prior(normal(0, 5), class = "Intercept"),
  prior(normal(0, 3),   class = "b"),
  
  # hu
  prior(normal(0, 2), class = "Intercept", dpar = "hu"),
  prior(normal(0, 1), class = "b", dpar = "hu"),
  
  # Sigma
  prior(normal(0, 1), class = "Intercept", dpar = "sigma"),
  prior(normal(0, 0.3), class = "b", dpar = "sigma"),
  
  # Random effects
  prior(exponential(1), class = "sd")
)

model_priors_ga <- c(
  # b
  prior(normal(0, 5), class = "Intercept"),
  prior(normal(0, 3),   class = "link"),
  
  # hu
  prior(normal(0, 2), class = "Intercept", dpar = "link_hu"),
  prior(normal(0, 1), class = "b", dpar = "link_hu"),
  
  # Sigma
  prior(normal(0, 1), class = "Intercept", dpar = "link_shape"),
  prior(normal(0, 0.3), class = "b", dpar = "link_shape"),
  
  # Random effects
  prior(exponential(1), class = "sd")
)

model_priors_fixed <- c(
  # b
  prior(normal(0, 5), class = "Intercept"),
  prior(normal(0, 3),   class = "b"),
  
  # hu
  prior(normal(0, 2), class = "Intercept", dpar = "hu"),
  prior(normal(0, 1), class = "b", dpar = "hu"),
  
  # Sigma
  prior(normal(0, 1), class = "Intercept", dpar = "sigma"),
  prior(normal(0, 0.3), class = "b", dpar = "sigma")
)

set_cmdstan_path(Sys.getenv("cmd_stan_path"))
```

## Model fitting
```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show model fitting code"

null_zln <- brm(
  bd_load_co_density ~ 1,
  data = brms_data,
  family = hurdle_lognormal(),
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 124,
  cores = 4,
  backend = "cmdstanr"
)

null_zln <- add_criterion(null_zln, "loo")

saveRDS(null_zln, file = here("unified_models/unified_model_results_zln_null.rds"))
null_zln = readRDS(file = here("unified_models/unified_model_results_zln_null.rds"))
# 
# nullb_zln <- brm(
#   bd_load_co ~ 1,
#   data = brms_data,
#   family = hurdle_lognormal(),
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# nullb_zln <- add_criterion(nullb_zln, "loo")
# 
# saveRDS(nullb_zln, file = here("unified_models/unified_model_results_zln_nullb.rds"))
# nullb_zln = readRDS(file = here("unified_models/unified_model_results_zln_nullb.rds"))
# 
# m1_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture,
#     hu ~ 1 + life_stage_simple * taxon_capture,
#     sigma ~ 1 + taxon_capture
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors_fixed,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m1_zln <- add_criterion(m1_zln, "loo")
# 
# saveRDS(m1_zln, file = here("unified_models/unified_model_results_zln_life_tax.rds"))
# m1_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax.rds"))
# 

# m2_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     sigma ~ 1 + taxon_capture
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m2_zln <- add_criterion(m2_zln, "loo")
# 
# saveRDS(m2_zln, file = here("unified_models/unified_model_results_zln_life_tax_pop.rds"))
# m2_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax_pop.rds"))
# 
# m2b_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     sigma ~ 1 + taxon_capture + (1 | population)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m2b_zln <- add_criterion(m2b_zln, "loo")
# 
# saveRDS(m2b_zln, file = here("unified_models/unified_model_results_zln_life_tax_popb.rds"))
# m2b_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax_popb.rds"))

# m2b_zga <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     shape ~ 1 + taxon_capture + (1 | population)
#   ),
#   data = brms_data,
#   family = hurdle_gamma(),
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m2b_zga <- add_criterion(m2b_zga, "loo")
# 
# saveRDS(m2b_zga, file = here("unified_models/unified_model_results_zga_life_tax_popb.rds"))
# m2b_zga = readRDS(file = here("unified_models/unified_model_results_zga_life_tax_popb.rds"))
# 
# m3_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 + population) +  (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 + population) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture + (1 + population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m3_zln <- add_criterion(m3_zln, "loo")
# 
# saveRDS(m3_zln, file = here("unified_models/unified_model_results_zln_life_tax_poptls.rds"))
# m3_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax_poptls.rds"))
# 
# m3b_zln <- brm(
#   formula = bf(
#     bd_load_co ~ 1 + life_stage_simple * taxon_capture + (1 + population) +  (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 + population) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture + (1 + population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m3b_zln <- add_criterion(m3b_zln, "loo")
# 
# saveRDS(m3b_zln, file = here("unified_models/unified_model_results_zln_life_tax_poptls_nodens.rds"))
# m3b_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax_poptls_nodens.rds"))

# m4_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + temp_c_30_z + (1 + population) +  (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + temp_c_15_z + (1 + population) +  (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture + (1 + population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# m4_zln <- add_criterion(m4_zln, "loo")
# 
# saveRDS(m4_zln, file = "unified_models/unified_model_results_zln_life_tax_pop_env.rds")
# m4_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_pop_env.rds")
# 
# m5_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5_zln <- add_criterion(m5_zln, "loo")
# 
# saveRDS(m5_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls_env2.rds")
# m5_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls_env2.rds")
# 

# m5b_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + temp_c_30_z | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + temp_c_15_z | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5b_zln <- add_criterion(m5b_zln, "loo")
# 
# saveRDS(m5b_zln, file = "unified_models/unified_model_results_zln_life_tax_poptlsenv_env2.rds")
# m5b_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptlsenv_env2.rds")
# 
# m5c_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5c_zln <- add_criterion(m5c_zln, "loo")
# 
# saveRDS(m5c_zln, file = "unified_models/unified_model_results_zln_life_tax_poptlsenv2_env2.rds")
# m5c_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptlsenv2_env2.rds")

# m5d_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + temp_c_30_z | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + temp_c_15_z | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5d_zln <- add_criterion(m5d_zln, "loo")
# 
# saveRDS(m5d_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2env_env2.rds")
# m5d_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env_env2.rds")


# m5e_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + temp_c_30_z | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + temp_c_15_z | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + life_stage_simple * taxon_capture +  (1 | population) + (1 | population:taxon_capture:life_stage_simple)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5e_zln <- add_criterion(m5e_zln, "loo")
# 
# saveRDS(m5e_zln, file = "unified_models/unified_model_results_zln_life_tax_poptlsenv_env2_sigls.rds")
# m5e_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptlsenv_env2_sigls.rds")
# 
# m5f_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5f_zln <- add_criterion(m5f_zln, "loo")
# 
# saveRDS(m5f_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2.rds")
# m5f_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2.rds")

##
## I think I should drop the (1|population) terms and keep only the (1|population:taxon_capture:life_stage_simple) terms... maybe? Or does the fact that they are not neatly nested mean this not trully redundant?
##
##
m5g_zln <- brm(
  formula = bf(
    bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
    hu ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
    sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
  ),
  data = brms_data,
  family = hurdle_lognormal(),
  prior = model_priors,
  sample_prior = "yes",
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 124,
  cores = 4,
  backend = "cmdstanr"
)

m5g_zln <- add_criterion(m5g_zln, "loo")

saveRDS(m5g_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_30.rds")
m5g_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_30.rds")
# 
# m6_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_sp_z + precip_mm_15_sp_z2 + temp_c_15_sp_z + temp_c_15_sp_z2 + (1 | population) + (1 + precip_mm_15_sp_z + precip_mm_15_sp_z2 + temp_c_15_sp_z + temp_c_15_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m6_zln <- add_criterion(m6_zln, "loo")
# 
# saveRDS(m6_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_envsp2.rds")
# m6_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_envsp2.rds")
# 
# m6b_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_15_sp_z + precip_mm_15_sp_z2 + temp_c_15_sp_z + temp_c_15_sp_z2 | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m6b_zln <- add_criterion(m6b_zln, "loo")
# 
# saveRDS(m6b_zln, file = "unified_models/unified_model_results_zln_life_tax_popt2envsp2.rds")
# m6b_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_popt2envsp2.rds")
# 
# m6c_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m6c_zln <- add_criterion(m6c_zln, "loo")
# 
# saveRDS(m6c_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_envsp2_30.rds")
# m6c_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_envsp2_30.rds")
# 
# m6d_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m6d_zln <- add_criterion(m6d_zln, "loo")
# 
# saveRDS(m6d_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_30.rds")
# m6d_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_30.rds")
# 
# m7_zln <- brm(
#   formula = bf(
#     bd_load_co ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m7_zln <- add_criterion(m7_zln, "loo")
# 
# saveRDS(m7_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2_nodens.rds")
# m7_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2_nodens.rds")
# 


```

## Model comparison
```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show model comparison code"
# model comparison and selection

(loo_comp_out = loo_compare(null_zln,
                            # nullb_zln, # no density
                            # m1_zln,
                            # m2_zln,
                            # m2b_zln,
                            # m3_zln,
                            # m3b_zln,  # no density
                            # m4_zln,
                            # m5_zln,
                            # m5b_zln,
                            # m5c_zln,
                            # m5d_zln,
                            # m5e_zln,
                            # m5f_zln,
                            m5g_zln
                            # m6_zln,
                            # m6b_zln,
                            # m6c_zln,
                            # m6d_zln
                            ))
```


## Model checks

### Fitting and correlation checks
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show model checks code"
#| fig-height: 6
#| fig-width: 8

model_select = m5g_zln

summary(model_select)
plot((model_select), ask = FALSE)
plot(conditional_effects(model_select), ask = FALSE)
VarCorr(model_select)
```

### Posterios predictive checks
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show model checks code (fixed effects)"
#| fig-height: 6
#| fig-width: 8

color_scheme_set("viridis")
model_select = m5g_zln


pp_check(model_select,
         type = "dens_overlay",
         prefix = "ppc",
         ndraws = 100) +
  scale_x_log10()

# # break out by taxon
# pp_check(model_select,
#          type = "dens_overlay_grouped",
#          group = "taxon_capture",
#          ndraws = 100) +
#   scale_x_log10()


```

### Prior checks
```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show prior checks code"
#| fig-height: 6
#| fig-width: 8

mcmc_plot(m5g_zln, type = "dens", variable = c("prior_Intercept", "b_Intercept")) +
  xlim(-20, 20)

# mcmc_plot(m5f_zln, type = "dens", variable = c("prior_b", "b_taxon_capturecolostethus_panamansis", "b_taxon_capturelithobates_blairi", 
# "b_taxon_capturelithobates_sphenocephalus", "b_taxon_capturelithobates_warszewitschii", 
# "b_taxon_capturepseudacris_crucifer", "b_taxon_capturerana_catesbeiana", 
# "b_taxon_capturerana_clamitans", "b_taxon_capturerana_muscosa", 
# "b_taxon_capturerana_pipiens", "b_taxon_capturerana_sierrae", 
# "b_taxon_capturesilverstoneia_flotator")) +
#   xlim(-10, 10)

mcmc_plot(m5g_zln, type = "dens", variable = c("prior_Intercept_sigma", "b_sigma_Intercept")) +
  xlim(-2, 2)

# mcmc_plot(m5f_zln, type = "dens", variable = c("prior_b_sigma", "b_sigma_taxon_capturecolostethus_panamansis", "b_sigma_taxon_capturelithobates_blairi", 
# "b_sigma_taxon_capturelithobates_sphenocephalus", "b_sigma_taxon_capturelithobates_warszewitschii", 
# "b_sigma_taxon_capturepseudacris_crucifer", "b_sigma_taxon_capturerana_catesbeiana", 
# "b_sigma_taxon_capturerana_clamitans", "b_sigma_taxon_capturerana_muscosa", 
# "b_sigma_taxon_capturerana_pipiens", "b_sigma_taxon_capturerana_sierrae", 
# "b_sigma_taxon_capturesilverstoneia_flotator")) +
#   xlim(-2, 2)

mcmc_plot(m5g_zln, type = "dens", variable = c("prior_Intercept_hu", "b_hu_Intercept")) +
  xlim(-15, 15)

# mcmc_plot(m5f_zln, type = "dens", variable = c("prior_b_hu", "b_hu_life_stage_simplesubadult", "b_hu_taxon_capturecolostethus_panamansis", 
# "b_hu_taxon_capturelithobates_blairi", "b_hu_taxon_capturelithobates_sphenocephalus", 
# "b_hu_taxon_capturelithobates_warszewitschii", "b_hu_taxon_capturepseudacris_crucifer", 
# "b_hu_taxon_capturerana_catesbeiana", "b_hu_taxon_capturerana_clamitans", 
# "b_hu_taxon_capturerana_muscosa", "b_hu_taxon_capturerana_pipiens", 
# "b_hu_taxon_capturerana_sierrae", "b_hu_taxon_capturesilverstoneia_flotator")) +
#   xlim(-5, 5)

```

## Define analysis functions
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show processing & visualization functions"

get_species_name = function(taxon, newline = FALSE) {
  species_name = str_to_sentence(gsub("_", " ", gsub("usgs", "", taxon)))
  if (newline) {
    species_name = gsub(" ", "\n", species_name)
  }
  return(species_name)
}

wilson_ci = function(p_hat, n, conf.level = 0.95) {
  z <- qnorm(1 - (1 - conf.level) / 2)
  
  denominator <- 1 + (z^2 / n)
  centre_adjusted_probability <- p_hat + (z^2 / (2 * n))
  adjusted_standard_deviation <- sqrt((p_hat * (1 - p_hat) / n) + (z^2 / (4 * n^2)))
  
  lower_bound <- (centre_adjusted_probability - z * adjusted_standard_deviation) / denominator
  upper_bound <- (centre_adjusted_probability + z * adjusted_standard_deviation) / denominator
  
  return(c(lower_bound = lower_bound, upper_bound = upper_bound))
}

wilson_ci_lower = function(p_hat, n, conf.level = 0.95) {
  ws = wilson_ci(p_hat, n, conf.level)
  return(ws[1])
}

wilson_ci_upper = function(p_hat, n, conf.level = 0.95) {
  ws = wilson_ci(p_hat, n, conf.level)
  return(ws[2])
}


```

# Results

## Extract and organize draws
```{r}
model_select = m5g_zln

# emmeans parameter distributions
## prevalence
hu_specs = emmeans(model_select, 
                   ~ life_stage_simple | taxon_capture, 
                   dpar = "hu")

contrast(hu_specs, method = "revpairwise") 

hu_contrasts_specs = contrast(hu_specs, method = "revpairwise")

hu_group_draws = hu_specs %>%
  gather_emmeans_draws() %>%
  rename(hu = .value) %>%
  mutate(prevalence = 1 - plogis(hu))

hu_contrast_draws = hu_contrasts_specs %>%
  gather_emmeans_draws() %>%
  rename(hu_subadult = .value)

hu_contrast_agg = hu_contrast_draws %>%
    group_by(taxon_capture) %>%
  summarize(hdi_50 = hdi(hu_subadult, ci = 0.50),
         hdi_89 = hdi(hu_subadult, ci = 0.89),
         hdi_95 = hdi(hu_subadult, ci = 0.95),
         ratio_positive = sum(hu_subadult > 0) / n(),
         hu_median = median(hu_subadult),
         .groups = "drop") %>%
  mutate(
    hdi_zero_50 = case_when(
      hdi_50$CI_low > 0 | hdi_50$CI_high < 0 ~ TRUE,
      .default = FALSE
    ),
    hdi_zero_89 = case_when(
      hdi_89$CI_low > 0 | hdi_89$CI_high < 0 ~ TRUE,
      .default = FALSE
    ),
    hdi_zero_95 = case_when(
      hdi_95$CI_low > 0 | hdi_95$CI_high < 0 ~ TRUE,
      .default = FALSE
    ))

## b (mu)
b_specs = emmeans(model_select, 
                 ~ life_stage_simple | taxon_capture, 
                 dpar = "mu")

contrast(b_specs, method = "revpairwise") 

b_contrasts_specs = contrast(b_specs, method = "revpairwise")

b_group_draws = b_specs %>%
  gather_emmeans_draws() %>%
  rename(b = .value)

b_contrast_draws = b_contrasts_specs %>%
  gather_emmeans_draws() %>%
  rename(b_subadult = .value)

b_contrast_agg = b_contrast_draws %>%
    group_by(taxon_capture) %>%
  summarize(hdi_50 = hdi(b_subadult, ci = 0.50),
         hdi_89 = hdi(b_subadult, ci = 0.89),
         hdi_95 = hdi(b_subadult, ci = 0.95),
         ratio_positive = sum(b_subadult > 0) / n(),
         b_median = median(b_subadult),
         .groups = "drop") %>%
  mutate(
    hdi_zero_50 = case_when(
      hdi_50$CI_low > 0 | hdi_50$CI_high < 0 ~ TRUE,
      .default = FALSE
    ),
    hdi_zero_89 = case_when(
      hdi_89$CI_low > 0 | hdi_89$CI_high < 0 ~ TRUE,
      .default = FALSE
    ),
    hdi_zero_95 = case_when(
      hdi_95$CI_low > 0 | hdi_95$CI_high < 0 ~ TRUE,
      .default = FALSE
    ))

# combine
contrast_agg = hu_contrast_agg %>%
  left_join(b_contrast_agg, by = c("taxon_capture"), suffix = c("_hu", "_b"))

group_draws = b_group_draws %>%
  left_join(hu_group_draws, by = c(".chain", ".iteration", ".draw", 
                                   "life_stage_simple", "taxon_capture"))

group_draws_agg = group_draws %>%
  group_by(taxon_capture,
           life_stage_simple) %>%
  summarize(hu_median = median(hu),
            prevalence_median = median(prevalence),
            b_median = median(b),
            .groups = "drop")

# epred draws
newdata_basic = expand_grid(taxon_capture = sort(unique(brms_data$taxon_capture)),
                            life_stage_simple = c("subadult",
                                                  "adult")) %>%
  mutate(precip_mm_30_z = 0,
         precip_mm_30_z2 = 0,
         temp_c_30_z = 0,
         temp_c_30_z2 = 0)

# way too big, limit to 50 draws!
epred_original = epred_draws(model_select,
                             newdata = brms_data,
                             re_formula = NA,
                             ndraws = 200) %>%
  rename(bd_load_density_epred = .epred)

epred_fixed = epred_draws(model_select,
                          newdata = newdata_basic,
                          re_formula = NA) %>%
  rename(bd_load_density_epred = .epred)

# observation data
obs_data = brms_data %>%
  group_by(taxon_capture,
           life_stage_simple) %>%
   summarise(n = n(),
             n_pos = sum(bd_detected_co),
             bd_prevalence = sum(bd_detected_co) / n(),
             bd_prevalence_ci_lower = wilson_ci_lower(bd_prevalence, n(), 0.95),
             bd_prevalence_ci_upper = wilson_ci_upper(bd_prevalence, n(), 0.95),
             min_load_inf = min(bd_load_co_density[bd_detected_co > 0]),
             q_25_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.25),
             q_50_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.50),
             q_75_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.75),
             max_load_inf = max(bd_load_co_density[bd_detected_co > 0]),
             se = sqrt(bd_prevalence*(1-bd_prevalence)/n),
             .groups = "drop") %>%
  mutate(label_offset = recode_values(life_stage_simple,
                                          "adult" ~ 0,
                                          "subadult" ~ 1))


```
## Visualize outputs

### Prevalence
```{r}
#| fig-height: 8
#| fig-width: 12

# observed
ggplot(obs_data, aes(y = bd_prevalence, x = life_stage_simple, color = life_stage_simple)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = bd_prevalence_ci_lower, ymax = bd_prevalence_ci_upper),
                width = 0.2,
                position = position_dodge(width = 0.5)) +
  geom_text(aes(label = paste0("n=", n), y = 1 - (1 - label_offset) * 0.05),
          position = position_dodge(width = 0.5),
          size = 4) +
  ylim(0, 1) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# modeled
ggplot(group_draws, aes(y = prevalence, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")


```

### Load
```{r}
#| fig-height: 8
#| fig-width: 12

# observed
ggplot(brms_data %>%
         filter(bd_detected_co > 0), aes(y = bd_load_co_density, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  geom_text(data = obs_data, aes(label = paste0("n=", n_pos), y = 10e8 - (1 - label_offset) * 7 * 10e7),
            position = position_dodge(width = 0.5),
            size = 4) +
  scale_y_log10(limits = c(1, 1e10)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")

ggplot(epred_original, aes(y = bd_load_density_epred, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  # geom_boxplot() +
  scale_y_log10(limits = c(1, 1e10)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")

ggplot(epred_fixed, aes(y = bd_load_density_epred, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  # geom_boxplot() +
  scale_y_log10(limits = c(1, 1e10)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")



```

## Hypothesis testing
```{r}
#| eval: true
#| echo: false
#| warning: false
#| fig-height: 8
#| fig-width: 12


# visualize draws
ggplot(group_draws, aes(x = prevalence, y = b, color = life_stage_simple)) +
  geom_point(alpha = 0.1) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# single parameter distributions
ggplot(group_draws, aes(y = prevalence, x = life_stage_simple, color = life_stage_simple)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

ggplot(group_draws, aes(y = b, x = life_stage_simple, color = life_stage_simple)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# visualize contrasts
ggplot(hu_contrast_draws, aes(x = hu_subadult)) +
  stat_halfeye(fill = "#66c2a5",
               .width = c(0.5, 0.95),
               point_interval = mode_hdi) +
  facet_wrap("taxon_capture") +
  geom_vline(xintercept = 0, alpha = 1) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect on Bd Prevalence",
       x = "Bd Prevalence Subadult Effect Size (logit scale) [-]\nwith 50% & 95% High Density Interval",
       y = "Posterior Density [-]") +
  theme(text=element_text(size=14))

ggplot(b_contrast_draws, aes(x = b_subadult)) +
    stat_halfeye(fill = "#8da0cb",
               .width = c(0.5, 0.95),
               point_interval = mode_hdi) +
  facet_wrap("taxon_capture") +
  geom_vline(xintercept = 0, alpha = 1) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect on Bd Prevalence",
       x = "Bd Load Density Subadult Effect Size (logit scale) [-]\nwith 50% & 95% High Density Interval",
       y = "Posterior Density [-]") +
  theme(text=element_text(size=14))

# probability of positive effect
ggplot(contrast_agg, aes(x = 1-ratio_positive_hu, y = ratio_positive_b)) +
  geom_vline(xintercept = 0.5, color = "#aaa") +
  geom_hline(yintercept = 0.5, color = "#aaa") + 
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  xlim(0, 1) +
  ylim(0, 1) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect: Fraction Above Zero (FAZ)",
       x = "Bd Prevalence\nProbability of positive effect [-]",
       y = "Bd Load Density\nProbability of positive effect [-]") +
  scale_color_brewer(palette = "Set2")

ggplot(contrast_agg, aes(x = -hu_median, y = b_median)) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  theme_minimal() +
  labs(
       # title = "Subadult Effect: Fraction Above Zero (FAZ)",
       x = "Bd Prevalence\nMedian effect [-]",
       y = "Bd Load Density\nMedian effect [-]") +
  scale_color_brewer(palette = "Set2")

```

## Phylogenetic tree
```{r}
#| eval: true
#| echo: false
#| warning: false

# build tree from 
species_list = sort(get_species_name(bd_model_stats$taxon_capture))
tnrs_resolved = tnrs_match_names(species_list)
tnrs_info = taxonomy_taxon_info(ott_ids = ott_id(tnrs_resolved))
src = tax_sources(tnrs_info)
tr = tol_induced_subtree(ott_ids = ott_id(tnrs_resolved),
                         label_format = "name")

# source_studies <- map_dfr(tnrs_resolved$unique_name, function(species) {
#   result <- studies_find_trees(
#     property = "ot:ottTaxonName", 
#     value = species,
#     detailed = TRUE
#   )
#   
#   tibble(
#     species = species,
#     study_id = result$study_ids,
#     n_matched_trees = result$n_matched_trees,
#     study_year = result$study_year,
#     title = result$title
#   )
# })

# source_studies_wide = source_studies %>%
#   select(species, study_id) %>%
#   mutate(value = 1) %>%
#   pivot_wider(names_from = species,
#               values_from = value,
#               values_fill = 0) %>%
#   arrange(study_id)
# 
# source_studies_unique = source_studies %>%
#   select(-species) %>%
#   distinct() %>%
#   arrange(study_id)

tr_reordered = rotateConstr(tr, gsub(" ", "_", species_list))
tr_reordered$tip.label = get_species_name(tr_reordered$tip.label)

# plot tree
# ggtree(tr, layout = "rectangular") +
#   geom_tiplab(size = 3, align = TRUE) +
#   theme_tree2() +
#   scale_x_continuous(expand = expansion(mult = c(0, 0.3)),
#                      breaks = c(1, 3, 5, 7))

p_phylo_tree = ggtree(tr_reordered, layout = "rectangular") +
  geom_tiplab(size = 4, align = TRUE) +
  theme_tree2() +
  scale_x_continuous(expand = expansion(mult = c(0, 0.8)),
                     breaks = c(1, 3, 5, 7))

# ggsave(here("clh", "phologeny_tree.png"), plot = p_phylo_tree, width = 5, height = 6, units = "in")

p_phylo_tree


# correlate phylogeny with trends
cophen_dist = cophenetic(tr_reordered)
cophen_sorted = cophen_dist[sort(rownames(cophen_dist)), sort(colnames(cophen_dist))]
cophen_upper = cophen_sorted[upper.tri(cophen_sorted, diag = FALSE)]

hu_dist = dist(hu_contrast_agg$hu_median, upper = TRUE)
b_dist = dist(b_contrast_agg$b_median, upper = TRUE)

hu_rho = cor.test(hu_dist, cophen_upper, method = "spearman")
hu_rho
b_rho = cor.test(b_dist, cophen_upper, method = "spearman")
b_rho
```

## Collective results
```{r}
#| eval: true
#| echo: true
#| output: false
#| code-fold: true
#| code-summary: Show code

# m5g_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2.rds")

choice_model = m5g_zln

model_data = brms_data

draws_all = collect_draws(choice_model) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

density_mat_all = density_mat(log(0.5), 20, 100, draws_all) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

sim_load_all = simulate_bd_load(draws_all, 10, 1000) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

obs_stats_all = sample_stats(model_data) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

```

::: {.panel-tabset}
### Observed Prevalence
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# observed prevalence
ggplot(obs_stats_all, aes(y = bd_prevalence, x = species, color = life_stage_simple)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = bd_prevalence_ci_lower, ymax = bd_prevalence_ci_upper),
                width = 0.2,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  ggtitle("Observed prevalence") +
  xlab("Species") +
  ylab("Bd prevalence [-]\n(with 95% CI)") +
  labs(color = "Life stage") +
  geom_text(aes(label = paste0("n=", n),  y = label_y),
          position = position_dodge(width = 0.5)) +
  ylim(0, 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Modeled Prevalence 
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# model prevalence
ggplot(draws_all, aes(prevalence, x = species, color = life_stage_simple)) +
  geom_boxplot(coef = NULL) + 
  theme_minimal() +
  ggtitle("Modeled prevalence") +
  xlab("Species") +
  ylab("Bd prevalence [-]\n") +
  labs(color = "Life stage") +
  ylim(0, 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Observed Bd load
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# obsevation bd load
ggplot(model_data %>%
         mutate(species = get_species_name(taxon_capture, newline = TRUE)) %>%
         filter(bd_load_co > 0), aes(log_bd_load_co_density, x = species, color = life_stage_simple)) +
  geom_boxplot() + 
  theme_minimal() +
  ggtitle("Observed Bd-positive load") +
  xlab("Species") +
  ylab("log(Bd load density) [-]") +
  labs(color = "Life stage") +
  geom_text(
    data = obs_stats_all,
    aes(
      x = species,
      y = max(model_data$log_bd_load_co_density) * label_y * 1.05,
      label = paste0("n=", n_pos),
      group = life_stage_simple
    ),
    position = position_dodge(width = 0.75),
    vjust = 0
  ) +
  ylim(log(0.5), 19) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Modeled Bd load
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# modeled bd load
ggplot(sim_load_all %>%
         filter(log_bd_load_co_density > 0.5), aes(log_bd_load_co_density, x = species, color = life_stage_simple)) +
  geom_boxplot() + 
  theme_minimal() +
  ggtitle("Modeled Bd-positive load") +
  xlab("Species") +
  ylab("log(Bd load density) [-]") +
  labs(color = "Life stage") +
  ylim(log(0.5), 20) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Modeled Bd load PDF
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# PDF -- facet by species
ggplot(density_mat_all, aes(x = density,
                            y = pdf_median,
                            color = life_stage_simple, 
                            fill = life_stage_simple)) +
  geom_ribbon(aes(ymin = pdf_lower,
                  ymax = pdf_upper), 
              alpha = 0.15,
              color = NA) +
  geom_line() +
  theme_minimal() +
  ggtitle("Modeled PDF of Bd-positive load") +
  labs(color = "Life stage",
       fill = "Life stage") +
  xlab("log(Bd load) [-]") +
  ylab("Median probability density [-]\n(with 25th to 75th percentiles)") +
  scale_x_continuous(limits = c(log(0.5), 20), expand = c(0, 0)) +
  facet_grid(. ~ gsub(" ", "\n", species))

# # PDF -- facet by life_stage & season
# ggplot(density_mat_all, aes(x = density,
#                             y = pdf_median,
#                             color = taxon_capture, 
#                             fill = taxon_capture)) +
#   geom_ribbon(aes(ymin = pdf_lower,
#                   ymax = pdf_upper), 
#               alpha = 0.15,
#               color = NA) +
#   geom_line() +
#   theme_minimal() +
#   ggtitle("Modeled PDF of Bd-positive load") +
#   labs(color = "Species",
#        fill = "Species") +
#   xlab("log(Bd load) [-]") +
#   ylab("Median probability density [-]\n(with 25th to 75th percentiles)") +
#   scale_x_continuous(limits = c(log(0.3), 20), expand = c(0, 0)) +
#   facet_grid(life_stage_simple ~ season)
```

### Moded Bd load CCDF
```{r}
#| echo: false
#| warning: false
#| fig-height: 8
#| fig-width: 8
# CCDF -- facet by species
ggplot(density_mat_all, aes(x = density,
                            y = ccdf_median,
                            color = life_stage_simple, 
                            fill = life_stage_simple)) +
  geom_ribbon(aes(ymin = ccdf_lower,
                  ymax = ccdf_upper), 
              alpha = 0.15,
              color = NA) +
  geom_line() +
  theme_minimal() +
  ggtitle("Modeled CCDF of Bd-positive load") +
  labs(color = "Life stage",
       fill = "Life stage") +
  xlab("log(Bd load) [-]") +
  ylab("Median complementary cumulative density (1 - CDF) [-]\n(with 25th to 75th percentiles)") +
  scale_x_continuous(limits = c(log(0.5), 12), expand = c(0, 0)) +
  ylim(0, 0.75) +
  # facet_grid(. ~ gsub(" ", "\n", species))
  facet_wrap("species")

# CCDF -- facet by life_stage & season
# ggplot(density_mat_all, aes(x = density,
#                             y = ccdf_median,
#                             color = taxon_capture,
#                             fill = taxon_capture)) +
#   geom_ribbon(aes(ymin = ccdf_lower,
#                   ymax = ccdf_upper),
#               alpha = 0.15,
#               color = NA) +
#   geom_line() +
#   theme_minimal() +
#   ggtitle("Modeled CCDF of Bd-positive load") +
#   labs(color = "Species",
#        fill = "Species") +
#   xlab("log(Bd load) [-]") +
#   ylab("Median complementary cumulative density (1 - CDF) [-]\n(with 25th to 75th percentiles)") +
#   scale_x_continuous(limits = c(log(0.3), 20), expand = c(0, 0)) +
#   ylim(0, 1) +
#   facet_grid(. ~ life_stage_simple)
```

# Environmental conditions

## Distributions
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Environmental conditions distributions code"
#| fig-height: 6
#| fig-width: 8

# 15-day
# ggplot(brms_data, aes(x = temp_c_15_day_mean, y = precip_mm_15_day_sum, color = life_stage_simple)) +
#   geom_point(alpha = 0.1) +
#   stat_ellipse() +
#   facet_wrap(. ~ taxon_capture) +
#   xlim(-5, 40) +
#   ylim(0, 200) +
#   labs(x = "15-day temperature mean (degC)", 
#        y = "15-day precipitation total (mm)",
#        title = "Temperature v. precipitation") +
#   theme_minimal()

# 30-day by species
ggplot(brms_data, aes(x = temp_c_30_day_mean, y = precip_mm_30_day_sum, color = life_stage_simple)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  facet_wrap(. ~ taxon_capture) +
  xlim(-5, 40) +
  ylim(0, 400) +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()

# 30 day collective
ggplot(brms_data, aes(x = temp_c_30_day_mean, y = precip_mm_30_day_sum, color = taxon_capture)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()

# 30 day collective, scaled to climate by species
ggplot(brms_data, aes(x = temp_c_30_sp_z, y = precip_mm_30_sp_z, color = taxon_capture)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()


# ggplot(brms_data, aes(x = temp_c_30_sp_z2, y = precip_mm_30_sp_z2, color = life_stage_simple)) +
#   geom_point(alpha = 0.1) +
#   stat_ellipse() +
#   facet_wrap(. ~ taxon_capture) +
#   labs(x = "30-day temperature mean (degC)", 
#        y = "30-day precipitation total (mm)",
#        title = "Temperature v. precipitation") +
#   theme_minimal()
```

## Tempurature effects
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show temperature effects code"
#| fig-height: 6
#| fig-width: 8

select_model = m5f_zln
# select_model = m6_zln

climate_draws = spread_draws(select_model, `b_temp_c.*|b_precip_mm.*|b_hu_temp_c.*|b_hu_precip_mm.*|b_Intercept|b_hu_Intercept`, regex = TRUE)

# Temperature dependence
temp_range = seq(-5, 45, length.out = 100)

temp_grid = expand_grid(temp_c_15_day_mean = temp_range, .draw = 1:nrow(climate_draws)) %>%
  mutate(temp_c_30_day_mean = temp_c_15_day_mean) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    temp_c_15_z = (temp_c_15_day_mean - brms_scales$temp_c_15_day_mean_mean) / brms_scales$temp_c_15_day_mean_sd,
    temp_c_15_z2 = (temp_c_15_day_mean^2 - brms_scales$temp_c_15_day_mean2_mean) / brms_scales$temp_c_15_day_mean2_sd,
    temp_c_30_z = (temp_c_30_day_mean - brms_scales$temp_c_30_day_mean_mean) / brms_scales$temp_c_30_day_mean_sd,
    temp_c_30_z2 = (temp_c_30_day_mean^2 - brms_scales$temp_c_30_day_mean2_mean) / brms_scales$temp_c_30_day_mean2_sd,
    b_pred = b_temp_c_30_z * temp_c_30_z + b_temp_c_30_z2 * temp_c_30_z2 + b_Intercept,
    hu_pred = b_hu_temp_c_15_z * temp_c_15_z + b_hu_temp_c_15_z2 * temp_c_15_z2 + b_hu_Intercept
  ) %>%
  group_by(temp_c_15_day_mean) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  ) %>%
  rename(temp_c = temp_c_15_day_mean)

# prevalence
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day mean temperature (degC)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. temperature") +
  theme_minimal()

# load
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day mean temperature (degC)", 
       y = "log(+ Bd load)",
       title = "Bd load v. temperature") +
  theme_minimal()
```

## Precipitation effects
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show precipitation effects code"
#| fig-height: 6
#| fig-width: 8
# precipitation depencence

daily_precip_range = seq(0, 20, length.out = 100)

precip_grid = expand_grid(daily_precip = daily_precip_range, .draw = 1:nrow(climate_draws)) %>%
  mutate(precip_mm_15_day_sum = daily_precip * 15,
         precip_mm_30_day_sum = daily_precip * 30) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    precip_mm_15_z = (precip_mm_15_day_sum - brms_scales$precip_mm_15_day_sum_mean) / brms_scales$precip_mm_15_day_sum_sd,
    precip_mm_15_z2 = (precip_mm_15_day_sum^2 - brms_scales$precip_mm_15_day_sum2_mean) / brms_scales$precip_mm_15_day_sum2_sd,
    precip_mm_30_z = (precip_mm_30_day_sum - brms_scales$precip_mm_30_day_sum_mean) / brms_scales$precip_mm_30_day_sum_sd,
    precip_mm_30_z2 = (precip_mm_30_day_sum^2 - brms_scales$precip_mm_30_day_sum2_mean) / brms_scales$precip_mm_30_day_sum2_sd,
    b_pred = b_precip_mm_30_z * precip_mm_30_z + b_precip_mm_30_z2 * precip_mm_30_z2 + b_Intercept,
    hu_pred = b_hu_precip_mm_15_z * precip_mm_15_z + b_hu_precip_mm_15_z2 * precip_mm_15_z2 + b_hu_Intercept
  ) %>%
  group_by(daily_precip,
           precip_mm_15_day_sum,
           precip_mm_30_day_sum) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  )

# prevalence
ggplot(precip_grid, aes(x = precip_mm_15_day_sum)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day total precipitation (mm)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. precipitation") +
  theme_minimal()

# load
ggplot(precip_grid, aes(x = precip_mm_30_day_sum)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day total precipitation (mm)", 
       y = "log(+ Bd load)",
       title = "Bd load v. precipitation") +
  theme_minimal()

```

## Tempurature effects (scaled by species climate)
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show temperature effects code"
#| fig-height: 6
#| fig-width: 8


select_model = m6_zln

# fixed effects
climate_draws = spread_draws(select_model, `b_temp_c.*|b_precip_mm.*|b_hu_temp_c.*|b_hu_precip_mm.*|b_Intercept|b_hu_Intercept`, regex = TRUE)

# Temperature dependence
temp_range = seq(-3, 3, length.out = 100)

# taxa = brms_data %>%
#   select(taxon_capture) %>%
#   distinct() %>%
#   pull(taxon_capture) %>%
#   sort()

## Stumped by how to back-calculate the temp_z2 values from the temp_z values... I think this has to be done on a species-by-species basis...

temp_grid = expand_grid(temp_z_15_mean = temp_range,
                        .draw = 1:nrow(climate_draws)) %>%
  mutate(temp_z_30_mean = temp_z_15_mean) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    temp_c_15_z = (temp_c_15_day_mean - brms_scales$temp_c_15_day_mean_mean) / brms_scales$temp_c_15_day_mean_sd,
    temp_c_15_z2 = (temp_c_15_day_mean^2 - brms_scales$temp_c_15_day_mean2_mean) / brms_scales$temp_c_15_day_mean2_sd,
    temp_c_30_z = (temp_c_30_day_mean - brms_scales$temp_c_30_day_mean_mean) / brms_scales$temp_c_30_day_mean_sd,
    temp_c_30_z2 = (temp_c_30_day_mean^2 - brms_scales$temp_c_30_day_mean2_mean) / brms_scales$temp_c_30_day_mean2_sd,
    b_pred = b_temp_c_30_sp_z * temp_c_30_z + b_temp_c_30_sp_z2 * temp_c_30_z2 + b_Intercept,
    hu_pred = b_hu_temp_c_15_sp_z * temp_c_15_z + b_hu_temp_c_15_sp_z2 * temp_c_15_z2 + b_hu_Intercept
  ) %>%
  group_by(temp_c_15_day_mean) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  ) %>%
  rename(temp_c = temp_c_15_day_mean)

# prevalence
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day mean temperature (degC)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. temperature") +
  theme_minimal()

# load
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day mean temperature (degC)", 
       y = "log(+ Bd load)",
       title = "Bd load v. temperature") +
  theme_minimal()
```

## Precipitation effects (scaled by species climate)
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show precipitation effects code"
#| fig-height: 6
#| fig-width: 8
# precipitation depencence

daily_precip_range = seq(0, 20, length.out = 100)

precip_grid = expand_grid(daily_precip = daily_precip_range, .draw = 1:nrow(climate_draws)) %>%
  mutate(precip_mm_15_day_sum = daily_precip * 15,
         precip_mm_30_day_sum = daily_precip * 30) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    precip_mm_15_z = (precip_mm_15_day_sum - brms_scales$precip_mm_15_day_sum_mean) / brms_scales$precip_mm_15_day_sum_sd,
    precip_mm_15_z2 = (precip_mm_15_day_sum^2 - brms_scales$precip_mm_15_day_sum2_mean) / brms_scales$precip_mm_15_day_sum2_sd,
    precip_mm_30_z = (precip_mm_30_day_sum - brms_scales$precip_mm_30_day_sum_mean) / brms_scales$precip_mm_30_day_sum_sd,
    precip_mm_30_z2 = (precip_mm_30_day_sum^2 - brms_scales$precip_mm_30_day_sum2_mean) / brms_scales$precip_mm_30_day_sum2_sd,
    b_pred = b_precip_mm_30_z * precip_mm_30_z + b_precip_mm_30_z2 * precip_mm_30_z2 + b_Intercept,
    hu_pred = b_hu_precip_mm_15_z * precip_mm_15_z + b_hu_precip_mm_15_z2 * precip_mm_15_z2 + b_hu_Intercept
  ) %>%
  group_by(daily_precip,
           precip_mm_15_day_sum,
           precip_mm_30_day_sum) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  )

# prevalence
ggplot(precip_grid, aes(x = precip_mm_15_day_sum)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day total precipitation (mm)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. precipitation") +
  theme_minimal()

# load
ggplot(precip_grid, aes(x = precip_mm_30_day_sum)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day total precipitation (mm)", 
       y = "log(+ Bd load)",
       title = "Bd load v. precipitation") +
  theme_minimal()

```