---
title: "Size vs. Bd modeling: Panamá preliminary results"
author: Cob Staines
date: 2025-08-20
format:
  html:
    toc: true
df-print: kable
bibliography: references.bib
csl: ecology.csl
---

## Setup
```{r}
#| eval: true
#| output: false
#| code-fold: true
#| code-summary: "Show setup code"

# setup
librarian::shelf(tidyverse, dbplyr, here, janitor, RPostgres, DBI, RIBBiTR-BII/ribbitrrr, brms, loo, bayesplot, tidybayes, ggplot2)

## Connect to DB
dbcon <- hopToDB("ribbitr")

# load populations
data_pop = read.csv(here("data", "sn_site_pop_2025-05-15.csv")) %>%
  rename(population = pop_id) %>%
  mutate(population = ifelse(!is.na(population), paste0(as.character(population), "_pop"), NA_character_))

# table pointers
db_bd = tbl(dbcon, Id("survey_data", "bd_qpcr_results"))
db_sample = tbl(dbcon, Id("survey_data", "sample"))
db_capture = tbl(dbcon, Id("survey_data", "capture"))
db_survey = tbl(dbcon, Id("survey_data", "survey"))
db_visit = tbl(dbcon, Id("survey_data", "visit"))
db_site = tbl(dbcon, Id("survey_data", "site"))
db_region = tbl(dbcon, Id("survey_data", "region"))
db_country = tbl(dbcon, Id("survey_data", "country"))
db_hist = tbl(dbcon, Id("survey_data", "site_history"))

bd_sample = db_sample %>%
  right_join(db_capture, by = "capture_id") %>%
  left_join(db_survey, by = "survey_id") %>%
  left_join(db_visit, by = "visit_id") %>%
  left_join(db_site, by = "site_id") %>%
  left_join(db_region, by = "region_id") %>%
  left_join(db_country, by = "country_id") %>%
  filter(sample_type == "bd")
```

## Methods

### Data selection & classification

0. Beginning with all Rana muscosa & Rana Sierrae data, grouped by population
1. Filter out post-translocation data
2. Population selection criteria:
  a. minimum 15 swabbed subadult and adult
  b. minimum 3 survey years
3. Classify by:
  a. Bd Naive: low (or no detected) Bd prevalence & load, high population
  b. Epizootic: high Bd prevalence & load, followed by low population
  c. Post-Epizootic: mid Bd prevalence & load
    i. Persistant (low VES & Capture)
    ii. Recovering (high VES & capture)
4. Review:
  a. Yosemite all post-epizootic at start of swabbing, except for Galiston (Epizootic ~2020)
  b. Compare with SEKI outbreak map fro agreement
  c. R. muscosa -- is Mulki the only post-epizootic site? If so, skip this analysis for R muscosa
  
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show data collection code"

# identify populations
bd_population = bd_sample %>%
  filter(country == "usa",
         region == "california",
         life_stage != "tadpole",
         !is.na(taxon_capture %in% c("rana_muscosa", "rana_sierrae"))) %>%
  left_join(data_pop, by = "site", copy = TRUE) %>%
  mutate(population = ifelse(is.na(population), site, population),
         tax_pop = paste0(taxon_capture, "_", population),
         year = year(date),
         life_stage_simple = life_stage,
         life_stage_simple = case_when(
           is.na(svl_mm) ~ life_stage,
           (svl_mm >= 40) ~ "adult",
           (svl_mm < 40) ~ "subadult",
           .default = life_stage))

# identify translocation events
site_history = db_hist %>%
  left_join(db_site, by = "site_id") %>%
  left_join(db_region, by = "region_id") %>%
  filter(region == "california",
         treatment_type %in% c("translocation_release", "reintroduction_release")) %>%
  collect()

# selection criteria
## - 3 years of sampling
## - at least 15 each adult and subadult per population

bd_population_valid = bd_population %>%
  group_by(population) %>%
  mutate(year_first_bd = min(year[bd_detected]),
            .groups = "drop") %>%
  summarise(year_first = min(year),
            year_first_bd = min(year_first_bd),
            year_last = max(year),
            year_count = n_distinct(year),
            year_count_since_bd = n_distinct(year[year >= year_first_bd]),
            total_n_since_bd = sum(as.integer(year >= year_first_bd)),
            adult_n_since_bd = sum(as.integer(year >= year_first_bd & life_stage_simple == "adult")),
            subadult_n_since_bd = sum(as.integer(year >= year_first_bd & life_stage_simple == "subadult")),
            .groups = "drop") %>%
    filter(year_count_since_bd >= 10,
         adult_n_since_bd > 100,
         subadult_n_since_bd > 100) %>%
  collect() %>%
  arrange(desc(year_count_since_bd))
  
  
```
  


