---
title: "Size vs. Bd modeling: Panamá preliminary results"
author: Cob Staines
date: 2025-09-29
format:
  html:
    toc: true
df-print: kable
bibliography: references.bib
csl: ecology.csl
---

# setup

```{r}
#| eval: true
#| output: false
#| code-fold: true
#| code-summary: "Show setup code"

# setup
librarian::shelf(tidyverse, dbplyr, here, janitor, RPostgres, DBI, RIBBiTR-BII/ribbitrrr, brms, tidybayes, ggplot2, ggrepel, bayesplot, runner, cmdstanr, bayestestR, ropensci/rotl, ggtree, ape, patchwork, viridis)

```

## Data selection

```{r}
#| eval: false
#| warning: false
#| code-fold: true
#| code-summary: "Show data collection code"

## Connect to DB
dbcon = hopToDB("ribbitr", hopReg = FALSE)

# table pointers
db_bd = tbl(dbcon, Id("survey_data", "bd_qpcr_results"))
db_sample = tbl(dbcon, Id("survey_data", "sample"))
db_capture = tbl(dbcon, Id("survey_data", "capture"))
db_survey = tbl(dbcon, Id("survey_data", "survey"))
db_visit = tbl(dbcon, Id("survey_data", "visit"))
db_site = tbl(dbcon, Id("survey_data", "site"))
db_region = tbl(dbcon, Id("survey_data", "region"))
db_country = tbl(dbcon, Id("survey_data", "country"))

# preliminary data
bd_sample = db_sample %>%
  right_join(db_capture, by = "capture_id") %>%
  left_join(db_survey, by = "survey_id") %>%
  left_join(db_visit, by = "visit_id") %>%
  left_join(db_site, by = "site_id") %>%
  left_join(db_region, by = "region_id") %>%
  left_join(db_country, by = "country_id") %>%
  filter(sample_type == "bd",
         taxon_capture %in% c("rana_muscosa",
                              "rana_sierrae")) %>%
  inner_join(db_bd %>%
              filter(!is.na(bd_detected)) %>%
              group_by(sample_id) %>%
              slice_sample(n = 1) %>%
              ungroup() %>%
              select(sample_id,
                     bd_detected,
                     bd_its1_copies_per_swab), by = "sample_id") %>%
  rename(bd_load = bd_its1_copies_per_swab) %>%
  mutate(bd_detected_co = bd_load >= 6,
         bd_load_co = if_else(bd_load < 6,
                                  0,
                                  bd_load),
         bd_load_co_density = bd_load_co / svl_mm)

# load populations
site_year_pop = read.csv(here("data", "sn_resilience_trajectory_classification_pop_lookup_2026-02-28.csv")) %>%
  mutate(site = as.character(site))

# load resilience trajectory classifications
pop_year_class = read.csv(here("data", "sn_resilience_trajectory_classification_2026-02-28.csv"))

# load environmental data
env_agg = read_csv(here("data", "unified_env_agg.csv"))

# define life stage simple
bd_life_stage = bd_sample %>%
  filter(life_stage %in% c("adult",
                           "subadult",
                           NA)) %>%
  mutate(life_stage_simple = case_when(
    is.na(svl_mm) ~ life_stage,
    (svl_mm >= 40) ~ "adult",
    (svl_mm < 40) ~ "subadult",
    .default = NA),
    year = year(date)) %>%
  filter(!is.na(life_stage_simple))

bd_model = bd_life_stage %>%
  filter(site %in% site_year_pop$site) %>%
  select(sample_id,
         bd_detected,
         bd_load,
         bd_detected_co,
         bd_load_co,
         bd_load_co_density,
         sample_name,
         capture_id,
         taxon_capture,
         life_stage,
         life_stage_simple,
         svl_mm,
         body_mass_g,
         sex,
         date,
         year,
         site,
         site_id,
         region,
         country) %>%
  collect() %>%
  left_join(site_year_pop, by = c("site", "year")) %>%
  inner_join(pop_year_class %>%
              select(population,
                     year,
                     res_class), by = c("population", "year")) %>%
  left_join(env_agg %>%
              select(site,
                     date,
                     precip_mm_30_day_sum,
                     temp_c_30_day_mean), by = c("site", "date")) %>%
  filter(!is.na(bd_load_co_density))

```

# Descriptive stats
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show descriptive stats code"

bd_model_stats = bd_model %>%
  group_by(taxon_capture,
           res_class,
           life_stage_simple) %>%
  summarize(n = n(),
            .groups= "drop")

```

# Modeling

## Model setup
```{r}
#| code-fold: true
#| code-summary: "Show model setup code"

# unified, scaled model
brms_scales = bd_model %>%
  group_by() %>%
  summarize(precip_mm_30_day_sum_mean = mean(precip_mm_30_day_sum),
            precip_mm_30_day_sum_sd = sd(precip_mm_30_day_sum),
            temp_c_30_day_mean_mean = mean(temp_c_30_day_mean),
            temp_c_30_day_mean_sd = sd(temp_c_30_day_mean),
            precip_mm_30_day_sum2_mean = mean(precip_mm_30_day_sum ^ 2),
            precip_mm_30_day_sum2_sd = sd(precip_mm_30_day_sum ^ 2),
            temp_c_30_day_mean2_mean = mean(temp_c_30_day_mean ^ 2),
            temp_c_30_day_mean2_sd = sd(temp_c_30_day_mean ^ 2))

brms_rescale = function(df, scales) {
  df_out = df %>%
    mutate( precip_mm_30_z = (precip_mm_30_day_sum - scales$precip_mm_30_day_sum_mean)/scales$precip_mm_30_day_sum_sd,
            temp_c_30_z = (temp_c_30_day_mean - scales$temp_c_30_day_mean_mean)/scales$temp_c_30_day_mean_sd,
            precip_mm_30_z2 = (precip_mm_30_day_sum^2 - scales$precip_mm_30_day_sum2_mean)/scales$precip_mm_30_day_sum2_sd,
            temp_c_30_z2 = (temp_c_30_day_mean^2 - scales$temp_c_30_day_mean2_mean)/scales$temp_c_30_day_mean2_sd
    )
}

brms_data = bd_model %>%
  mutate(
    precip_mm_30_z = as.numeric(scale(precip_mm_30_day_sum, center = TRUE, scale = TRUE)),
    temp_c_30_z = as.numeric(scale(temp_c_30_day_mean, center = TRUE, scale = TRUE)),
    precip_mm_30_z2 = as.numeric(scale(precip_mm_30_day_sum ^ 2, center = TRUE, scale = TRUE)),
    temp_c_30_z2 = as.numeric(scale(temp_c_30_day_mean ^ 2, center = TRUE, scale = TRUE))
  )


model_priors <- c(
  # b
  prior(normal(0, 5), class = "Intercept"),
  prior(normal(0, 2),   class = "b"),
  
  # hu
  prior(normal(0, 2), class = "Intercept", dpar = "hu"),
  prior(normal(0, 1), class = "b", dpar = "hu"),
  
  # Sigma
  prior(normal(0, 1), class = "Intercept", dpar = "sigma"),
  prior(normal(0, 0.3), class = "b", dpar = "sigma"),
  
  # Random effects
  prior(exponential(1), class = "sd")
)

set_cmdstan_path(Sys.getenv("cmd_stan_path"))
```

## Model fitting
```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show model fitting code"

# null model
null_zln <- brm(
  bd_load_co_density ~ 1,
  data = brms_data,
  family = hurdle_lognormal(),
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 124,
  cores = 4,
  backend = "cmdstanr"
)

null_zln = add_criterion(null_zln, "loo")

saveRDS(null_zln, file = here("snrt_models/snrt_model_results_zln_null.rds"))
null_zln = readRDS(file = here("snrt_models/snrt_model_results_zln_null.rds"))


# fixed env
m6_zln <- brm(
  formula = bf(
    bd_load_co_density ~ 1 + life_stage_simple * taxon_capture * res_class + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture:life_stage_simple) + (1 | population:res_class:life_stage_simple),
    hu ~ 1 + life_stage_simple * taxon_capture * res_class + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture:life_stage_simple) + (1 | population:res_class:life_stage_simple),
    sigma ~ 1 + life_stage_simple * taxon_capture * res_class + (1 | population:res_class:life_stage_simple)
  ),
  data = brms_data,
  family = hurdle_lognormal(),
  prior = model_priors,
  sample_prior = "yes",
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 124,
  cores = 4,
  backend = "cmdstanr"
)

m6_zln = add_criterion(m6_zln, "loo")

saveRDS(m6_zln, file = "snrt_models/snrt_model_results_zln_life_tax_poptls2envsp2_envsp2.rds")
m6_zln = readRDS(file = "snrt_models/snrt_model_results_zln_life_tax_poptls2envsp2_envsp2.rds")

```


## Model comparison

```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show model comparison code"
# model comparison and selection

(loo_comp_out = loo_compare(null_zln,
                            m6_zln))
```

## Model report
```{r}
model_select = m6_zln

summary(model_select)
plot((model_select), ask = FALSE)
plot(conditional_effects(model_select), ask = FALSE)

```

## pp_check
```{r}
model_select = m6_zln

# all
pp_check(model_select) +
  scale_x_log10()

# break out by taxon_capture, life_stage_simple, res_class
brms_data$taxon_life_res = paste(brms_data$taxon_capture, brms_data$life_stage, brms_data$res_class, sep ="_")

brms_data$taxon_life_res = interaction(brms_data$taxon_capture, brms_data$life_stage, brms_data$res_class, drop = TRUE)

pp_check(model_select,
         type = "dens_overlay_grouped",
         group = "taxon_capture",
         ndraws = 1000) +
  scale_x_log10()

pp_check(model_select,
         type = "dens_overlay_grouped",
         group = "life_stage_simple",
         ndraws = 1000) +
  scale_x_log10()

pp_check(model_select,
         type = "dens_overlay_grouped",
         group = "res_class",
         ndraws = 1000) +
  scale_x_log10()


```


### Prior checks
```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show prior checks code"
#| fig-height: 6
#| fig-width: 8

model_select = m6_zln

mcmc_plot(model_select, type = "dens", variable = c("prior_b", "b_Intercept", "b_taxon_capturerana_sierrae")) +
  xlim(-10, 10)

mcmc_plot(model_select, type = "dens", variable = c("prior_Intercept_sigma", "b_sigma_Intercept", "b_sigma_taxon_capturerana_sierrae")) +
  xlim(-2, 2)

mcmc_plot(model_select, type = "dens", variable = c("prior_Intercept_hu", "b_hu_Intercept", "b_hu_taxon_capturerana_sierrae")) +
  xlim(-5, 5)


```

## Processing and visualization
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show processing & visualization functions"

get_species_name = function(taxon, newline = FALSE) {
  species_name = str_to_sentence(gsub("_", " ", taxon))
  if (newline) {
    species_name = gsub(" ", "\n", species_name)
  }
  return(species_name)
}

```

# Results
Results are broken out first by species, then combined for comparison between species

## Hypothesis testing
```{r}
#| eval: true
#| echo: false
#| warning: false
#| fig-height: 8
#| fig-width: 12

# model_select = m3_zln
# model_select = m5f_zln
# model_select = m5f_zln
model_select = m6_zln

# prevalence
hu_taxon_subadult = spread_draws(model_select, `b_hu_life_stage_simplesubadult.*`, regex = TRUE) %>%
  mutate(`b_hu_life_stage_simplesubadult:taxon_captureanaxyrus_americanus` = 0) %>%
  pivot_longer(starts_with("b_hu_life_stage_simplesubadult:taxon_capture"),
               names_to = "parameter",
               values_to = "hu_adj") %>%
  mutate(taxon_capture = gsub("b_hu_life_stage_simplesubadult:taxon_capture", "", parameter),
         hu_subadult = b_hu_life_stage_simplesubadult + hu_adj,
         species = get_species_name(taxon_capture)) %>%
  group_by(species) %>%
  mutate(sp_sd = sd(hu_subadult)) %>%
  ungroup() %>%
  mutate(hu_subadult_sc = hu_subadult / sp_sd)

log_th = hu_scale

hu_taxon_subadult_rope = hu_taxon_subadult %>%
  mutate(in_rope = abs(hu_subadult) <= log_th) %>%
  group_by(species) %>%
  summarize(n = n(),
            pd = pd(hu_subadult)$pd,
            hdi_50 = hdi(hu_subadult, ci = 0.50),
            hdi_89 = hdi(hu_subadult, ci = 0.89),
            hdi_95 = hdi(hu_subadult, ci = 0.95),
            ratio_rope = sum(in_rope) / n(),
            ratio_positive = sum(hu_subadult > 0) / n(),
            ratio_negative = sum(hu_subadult < 0) / n(),
            ratio_net = ratio_positive - ratio_negative,
            hu_median = median(hu_subadult),
            hu_mean = mean(hu_subadult),
            hu_sd = sd(hu_subadult),
            hu_var = var(hu_subadult),
            .groups = "drop") %>%
  mutate(
    hdi_zero_50 = case_when(
      hdi_50$CI_low > 0 | hdi_50$CI_high < 0 ~ TRUE,
      .default = FALSE
    ),
    hdi_zero_89 = case_when(
      hdi_89$CI_low > 0 | hdi_89$CI_high < 0 ~ TRUE,
      .default = FALSE
    ),
    hdi_zero_95 = case_when(
      hdi_95$CI_low > 0 | hdi_95$CI_high < 0 ~ TRUE,
      .default = FALSE
    ),
    hdi_rope_50 = case_when(
      hdi_50$CI_low > log_th | hdi_50$CI_high < -log_th ~ TRUE,
      .default = FALSE
    ),
    hdi_rope_89 = case_when(
      hdi_89$CI_low > log_th | hdi_89$CI_high < -log_th ~ TRUE,
      .default = FALSE
    ),
    hdi_rope_95 = case_when(
      hdi_95$CI_low > log_th | hdi_95$CI_high < -log_th ~ TRUE,
      .default = FALSE
    )
  )

hu_effect_weak = hu_taxon_subadult_rope %>%
  filter(hdi_zero_50)

# rope prevalence (hu)
p_hu_dist = ggplot(hu_taxon_subadult, aes(x = hu_subadult)) +
  geom_density(fill = "#66c2a5", alpha = 1) +
  facet_wrap("species") +
  # geom_vline(xintercept=0) +
  annotate("rect", 
           xmin = -log_th, xmax = log_th, 
           ymin = -Inf, ymax = Inf,
           fill = "#fc8d62", alpha = 0.5) +
  geom_vline(xintercept = 0, alpha = 0.5) +
  geom_text(data = hu_taxon_subadult_rope,
            aes(x = 2,
                y = 1,
                label = paste0("ROPE: ", as.character(round(ratio_rope, digits = 3)))),
            size = 3.5) +
  xlim(-2.7, 3.5) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect on Bd Prevalence",
       x = "Bd Prevalence - Subadult Effect Size (logit scale) [-]",
       y = "Posterior Density [-] [-]") +
  theme(text=element_text(size=14))

ggsave(here("clh", "prevalence_hu_distributions.png"), plot = p_hu_dist, width = 12, height = 6, units = "in")

p_hu_dist

# zero prevalence (hu)
p_hu_dist_hdi = ggplot(hu_taxon_subadult, aes(x = hu_subadult)) +
  geom_density(fill = "#66c2a5", alpha = 1) +
  facet_wrap("species") +
  # geom_vline(xintercept=0) +
  geom_rect(data = hu_taxon_subadult_rope, inherit.aes = FALSE,
            aes(xmin = hdi_95$CI_low, xmax = hdi_95$CI_high,
                ymin = -0.15, ymax = 0),
            fill = "#bbb", alpha = 1) +
  geom_rect(data = hu_taxon_subadult_rope, inherit.aes = FALSE,
            aes(xmin = hdi_50$CI_low, xmax = hdi_50$CI_high,
                ymin = -0.15, ymax = 0),
            fill = "#666", alpha = 1) + 
  geom_vline(xintercept = 0, alpha = 1) +
  xlim(-2.7, 3.5) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect on Bd Prevalence",
       x = "Bd Prevalence Subadult Effect Size (logit scale) [-]\nwith 50% & 95% High Density Interval",
       y = "Posterior Density [-]") +
  theme(text=element_text(size=14))

ggsave(here("clh", "prevalence_hu_distributions_hdi.png"), plot = p_hu_dist_hdi, width = 12, height = 6, units = "in")

p_hu_dist_hdi

# posterior median sigma
# sigma_draws <- posterior_epred(model_select, dpar = "sigma")
# sigma_mean_per_draw <- apply(sigma_draws, 1, mean)
# median_b_sigma <- median(sigma_mean_per_draw)


# load
b_taxon_subadult = spread_draws(model_select, `b_life_stage_simplesubadult.*`, regex = TRUE) %>%
  mutate(`b_life_stage_simplesubadult:taxon_captureanaxyrus_americanus` = 0) %>%
  pivot_longer(starts_with("b_life_stage_simplesubadult:taxon_capture"),
               names_to = "parameter",
               values_to = "b_adj") %>%
  mutate(taxon_capture = gsub("b_life_stage_simplesubadult:taxon_capture", "", parameter),
         b_subadult = b_life_stage_simplesubadult + b_adj,
         species = get_species_name(taxon_capture))



b_taxon_subadult_rope = b_taxon_subadult %>%
  mutate(in_rope = abs(b_subadult_sc) <= 0.1) %>%
  group_by(species) %>%
  summarize(n = n(),
            pd = pd(b_subadult)$pd,
            hdi_50 = hdi(b_subadult, ci = 0.50),
            hdi_89 = hdi(b_subadult, ci = 0.89),
            hdi_95 = hdi(b_subadult, ci = 0.95),
            # ratio_rope = sum(in_rope) / n(),
            # ratio_positive = sum(b_subadult_sc > 0) / n(),
            # ratio_negative = sum(b_subadult_sc < 0) / n(),
            # ratio_net = ratio_positive - ratio_negative,
            b_median = median(b_subadult),
            b_mean = mean(b_subadult),
            b_sd = sd(b_subadult),
            b_var = var(b_subadult),
            # b_sc_median = median(b_subadult_sc),
            # b_sc_mean = mean(b_subadult_sc),
            .groups = "drop") %>%
  mutate(
   hdi_zero_50 = case_when(
      hdi_50$CI_low > 0 | hdi_50$CI_high < 0 ~ TRUE,
      .default = FALSE
    ),
    hdi_zero_89 = case_when(
      hdi_89$CI_low > 0 | hdi_89$CI_high < 0 ~ TRUE,
      .default = FALSE
    ),
    hdi_zero_95 = case_when(
      hdi_95$CI_low > 0 | hdi_95$CI_high < 0 ~ TRUE,
      .default = FALSE
    ),
    hdi_rope_50 = case_when(
      hdi_50$CI_low > log_th | hdi_50$CI_high < -log_th ~ TRUE,
      .default = FALSE
    ),
    hdi_rope_89 = case_when(
      hdi_89$CI_low > log_th | hdi_89$CI_high < -log_th ~ TRUE,
      .default = FALSE
    ),
    hdi_rope_95 = case_when(
      hdi_95$CI_low > log_th | hdi_95$CI_high < -log_th ~ TRUE,
      .default = FALSE
    )
   )

b_effect_weak = b_taxon_subadult_rope %>%
  filter(hdi_zero_50)

# rope load (b)
p_b_dist = ggplot(b_taxon_subadult, aes(x = b_subadult)) +
  # geom_vline(xintercept = 0, alpha = 1) +
  geom_density(fill = "#8da0cb", alpha = 1) +
  geom_rect(data = b_taxon_subadult_rope, 
            aes(xmin = -0.1 * median_b_sigma, xmax = 0.1 * median_b_sigma, ymin = -Inf, ymax = Inf),
            fill = "#fc8d62", alpha = 0.5, inherit.aes = FALSE) +
  geom_vline(xintercept = 0, alpha = 0.5) +
  geom_text(data = b_taxon_subadult_rope,
            aes(x = 3,
                y = 0.72,
                label = paste0("ROPE: ", as.character(round(ratio_rope, digits = 3)))),
            size = 3.5) +
  facet_wrap("species") +
  theme_minimal() +
  xlim(-4, 6) + 
  labs(
       # title = "Subadult Effect: Bd Load",
       x = "Bd Load Density Subadult Effect Size (log scale) [-]",
       y = "Posterior Density [-]") +
  theme(text=element_text(size=14))

ggsave(here("clh", "prevalence_b_distributions.png"), plot = p_b_dist, width = 12, height = 6, units = "in")

p_b_dist

# zero load (b)
p_b_dist_hdi = ggplot(b_taxon_subadult, aes(x = b_subadult)) +
  # geom_vline(xintercept = 0, alpha = 1) +
  geom_density(fill = "#8da0cb", alpha = 1) +
  geom_rect(data = b_taxon_subadult_rope, inherit.aes = FALSE,
            aes(xmin = hdi_95$CI_low, xmax = hdi_95$CI_high,
                ymin = -0.08, ymax = 0),
            fill = "#bbb", alpha = 1) +
  geom_rect(data = b_taxon_subadult_rope, inherit.aes = FALSE,
            aes(xmin = hdi_50$CI_low, xmax = hdi_50$CI_high,
                ymin = -0.08, ymax = 0),
            fill = "#666", alpha = 1) + 
  geom_vline(xintercept = 0, alpha = 1) +
  facet_wrap("species") +
  theme_minimal() +
  xlim(-4, 6) + 
  labs(
       # title = "Subadult Effect: Bd Load",
       x = "Bd Load Density Subadult Effect Size (log scale) [-]\nwith 50% & 95% High Density Interval",
       y = "Posterior Density [-]") +
  theme(text=element_text(size=14))

ggsave(here("clh", "prevalence_b_distributions_hdi.png"), plot = p_b_dist_hdi, width = 12, height = 6, units = "in")

p_b_dist_hdi
  
  

taxon_subadult = b_taxon_subadult %>%
  left_join(hu_taxon_subadult, by = c(".draw", ".iteration", ".chain", "taxon_capture", "species"))

# Fraction in ROPE
taxon_subadult_rope = hu_taxon_subadult_rope %>%
  full_join(b_taxon_subadult_rope, by = "species", suffix = c("_hu", "_b")) %>%
  left_join(bd_model_env_select_stats %>%
              mutate(species = get_species_name(taxon_capture)), by = "species") %>%
  mutate(region = recode_values(country,
                                "panama" ~ "Central America",
                                "usa" ~ "North America",
                                default = NA))

ggplot(taxon_subadult_rope, aes(x = ratio_rope_hu, y = ratio_rope_b, color = species)) +
  geom_vline(xintercept = 0.025, color = "blue") +
  geom_hline(yintercept = 0.025, color = "blue") + 
  geom_point() +
  geom_text_repel(aes(label = species)) +
  theme_minimal() +
  labs(title = "Subadult Effect: ROPE",
       x = "Bd Prevalence ROPE fraction (logit scale)",
       y = "Subadult Bd Load ROPE fraction (log scale)")

# HU median sd
ggplot(taxon_subadult_rope, aes(y = hu_median, x = ratio_rope_hu, color = species)) +
  geom_point() +
  geom_text_repel(aes(label = species)) +
  theme_minimal()

# B median sd
ggplot(taxon_subadult_rope, aes(y = b_median, x = ratio_rope_b, color = species)) +
  geom_point() +
  geom_text_repel(aes(label = species)) +
  theme_minimal()


# probability of direction
ggplot(taxon_subadult_rope, aes(x = pd_hu, y = pd_b, color = species)) +
  geom_point() +
  geom_text_repel(aes(label = species)) +
  theme_minimal() +
  labs(title = "Subadult Effect: Probability of Direction (Pd)",
       x = "Bd Prevalence Pd (logit scale)",
       y = "Subadult Bd Load Pd (log scale)")

# probability of positive effect
n_groups <- 19
full_pal <- plasma(2 * n_groups)  # Generate double for subsetting
upper_half <- full_pal[1:n_groups]  # Upper ~half

p_ppos = ggplot(taxon_subadult_rope, aes(x = ratio_positive_hu, y = ratio_positive_b)) +
  geom_vline(xintercept = 0.5, color = "#aaa") +
  geom_hline(yintercept = 0.5, color = "#aaa") + 
  geom_point(size = 4, color = "#660099") +
  geom_text_repel(aes(label = species), size = 4, color = "#660099", seed = 123) +
  xlim(0, 1) +
  ylim(0, 1) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect: Fraction Above Zero (FAZ)",
       x = "Bd Prevalence\nProbability of positive effect [-]",
       y = "Bd Load Density\nProbability of positive effect [-]") +
  theme(legend.position="none") +
  theme(text=element_text(size=16))

ggsave(here("clh", "probability_positive_effect.png"), plot = p_ppos, width = 12, height = 6, units = "in")
p_ppos

# # colored
# ggplot(taxon_subadult_rope, aes(x = ratio_positive_hu, y = ratio_positive_b, color = country)) +
#   geom_vline(xintercept = 0.5, color = "#aaa") +
#   geom_hline(yintercept = 0.5, color = "#aaa") + 
#   geom_point(size = 4) +
#   geom_text_repel(aes(label = species), size = 4, seed = 123) +
#   xlim(0, 1) +
#   ylim(0, 1) +
#   theme_minimal() +
#   labs(
#        # title = "Subadult Effect: Fraction Above Zero (FAZ)",
#        x = "Bd Prevalence\nProbability of positive effect [-]",
#        y = "Bd Load Density\nProbability of positive effect [-]") +
#   theme(legend.position="none") +
#   theme(text=element_text(size=16))

# no text
p_ppos_no_text = ggplot(taxon_subadult_rope, aes(x = ratio_positive_hu, y = ratio_positive_b)) +
  geom_vline(xintercept = 0.5, color = "#aaa") +
  geom_hline(yintercept = 0.5, color = "#aaa") + 
  geom_point(size = 4, color = "#660099") +
  xlim(0, 1) +
  ylim(0, 1) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect: Fraction Above Zero (FAZ)",
       x = "Bd Prevalence\nProbability of positive effect [-]",
       y = "Bd Load Density\nProbability of positive effect [-]") +
  theme(legend.position="none") +
  theme(text=element_text(size=16))

ggsave(here("clh", "probability_positive_effect_no_text.png"), plot = p_ppos_no_text, width = 12, height = 6, units = "in")
p_ppos_no_text

p_ppos_no_text_classic = ggplot(taxon_subadult_rope, aes(x = ratio_positive_hu, y = ratio_positive_b)) +
  geom_point(size = 4, color = "#660099") +
  xlim(0, 1) +
  ylim(0, 1) +
  theme_classic() +
  labs(
       x = "Bd Prevalence\nProbability of positive effect [-]",
       y = "Bd Load Density\nProbability of positive effect [-]") +
  theme(legend.position="none") +
  theme(text=element_text(size=16)) +
  scale_color_manual(values = upper_half)

ggsave(here("clh", "probability_positive_effect_no_text_classic.png"), plot = p_ppos_no_text_classic, width = 12, height = 6, units = "in")

p_ppos_classic = ggplot(taxon_subadult_rope, aes(x = ratio_positive_hu, y = ratio_positive_b)) +
  geom_point(size = 4, color = "#660099") +
  geom_text_repel(aes(label = species), size = 4, color = "#660099", seed = 123) +
  xlim(0, 1) +
  ylim(0, 1) +
  theme_classic() +
  labs(
       x = "Bd Prevalence\nProbability of positive effect [-]",
       y = "Bd Load Density\nProbability of positive effect [-]") +
  theme(legend.position="none") +
  theme(text=element_text(size=16)) +
  scale_color_manual(values = upper_half)

ggsave(here("clh", "probability_positive_effect_classic.png"), plot = p_ppos_classic, width = 12, height = 6, units = "in")

# p_ppos_grid = ggplot(taxon_subadult_rope, aes(x = ratio_positive_hu, y = ratio_positive_b, color = species)) +
#   geom_vline(xintercept = 0.5, color = "#aaa") +
#   geom_hline(yintercept = 0.5, color = "#aaa") + 
#   geom_vline(xintercept = 0.10, color = "#222", alpha = 0.5, size = 4) +
#   geom_hline(yintercept = 0.10, color = "#222", alpha = 0.5, size = 4) + 
#   geom_vline(xintercept = 0.90, color = "#222", alpha = 0.5, size = 4) +
#   geom_hline(yintercept = 0.90, color = "#222", alpha = 0.5, size = 4) + 
#   geom_point(size = 4) +
#   geom_text_repel(aes(label = species), size = 4, seed = 123) +
#   xlim(0, 1) +
#   ylim(0, 1) +
#   theme_minimal() +
#   labs(
#        # title = "Subadult Effect: Fraction Above Zero (FAZ)",
#        x = "Bd Prevalence:\nProbability of positive effect [-]",
#        y = "Bd Load:\nProbability of positive effect [-]") +
#   theme(legend.position="none") +
#   theme(text=element_text(size=16)) +
#   scale_color_manual(values = upper_half)
# ggsave(here("clh", "probability_positive_effect_grid.png"), plot = p_ppos_grid, width = 12, height = 6, units = "in")
# p_ppos_grid

ggplot(taxon_subadult_rope, aes(x = ratio_positive_hu, y = ratio_positive_b, color = species)) +
  geom_vline(xintercept = 0.5, color = "#aaa") +
  geom_hline(yintercept = 0.5, color = "#aaa") +
  geom_vline(xintercept = 0.025, color = "blue", alpha = 0.5) +
  geom_hline(yintercept = 0.025, color = "blue", alpha = 0.5) + 
  geom_vline(xintercept = 0.975, color = "blue", alpha = 0.5) +
  geom_hline(yintercept = 0.975, color = "blue", alpha = 0.5) + 
  # geom_rect(data = data.frame(xmin = 0.025, xmax = 0.975, ymin = 0.025, ymax = 0.975),
  #         aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
  #         fill = "blue", alpha = 0.1, inherit.aes = FALSE) +
  geom_point() +
  geom_text_repel(aes(label = species)) +
  xlim(0, 1) +
  ylim(0, 1) +
  theme_minimal() +
  labs(title = "Subadult Effect: Fraction Above Zero (FAZ)",
       x = "Bd Prevalence FAZ (logit scale)",
       y = "Bd Load FAZ (log scale)") +
  theme(legend.position="none")


# net direction
ggplot(taxon_subadult_rope, aes(x = ratio_net_hu, y = ratio_net_b, color = species)) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point() +
  geom_text_repel(aes(label = species)) +
  xlim(-1, 1) +
  ylim(-1, 1) +
  theme_minimal() +
  labs(title = "Subadult Effect: Net Direction (Nd)",
       x = "Bd Prevalence Nd (logit scale)",
       y = "Bd Load Nd (log scale)") +
  theme(legend.position="none")

# median
p_b_hu_median = ggplot(taxon_subadult_rope, aes(x = hu_median, y = b_median)) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point(size = 4, color = "#660099") +
  geom_text_repel(aes(label = species), size = 4, color = "#660099", seed = 135) +
  # xlim(-1.6, 1.6) +
  # ylim(-3.5, 3.5) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect",
       x = "Bd Prevalence\nMedian Subadult Effect Size (logit scale) [-]",
       y = "Bd Load Density\nMedian Subadult Effect Size (log scale) [-]") +
  theme(legend.position="none")

ggsave(here("clh", "median_effect.png"), plot = p_b_hu_median, width = 12, height = 6, units = "in")

p_b_hu_median

# median no text
p_b_hu_median_no_text = ggplot(taxon_subadult_rope, aes(x = hu_median, y = b_median)) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point(size = 4, color = "#660099") +
  xlim(-1.6, 1.6) +
  ylim(-3.5, 3.5) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect",
       x = "Bd Prevalence\nMedian Subadult Effect Size (logit scale) [-]",
       y = "Bd Load Density\nMedian Subadult Effect Size (log scale) [-]") +
  theme(legend.position="none") +
  theme(text=element_text(size=16))

ggsave(here("clh", "median_effect_no_text.png"), plot = p_b_hu_median_no_text, width = 12, height = 6, units = "in")

p_b_hu_median_no_text

# median colored
p_b_hu_median_country = ggplot(taxon_subadult_rope, aes(x = hu_median, y = b_median, color = region)) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point(size = 4) +
  geom_text_repel(aes(label = species), size = 4, seed = 135) +
  # xlim(-1.6, 1.6) +
  # ylim(-3.5, 3.5) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect",
       x = "Bd Prevalence\nMedian Subadult Effect Size (logit scale) [-]",
       y = "Bd Load Density\nMedian Subadult Effect Size (log scale) [-]",
       color = "Region") +
  theme(legend.position = c(0.9, 0.9),
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black")) +
  theme(text=element_text(size=16)) +
  scale_color_brewer(palette = "Dark2", type = "qual")

ggsave(here("clh", "median_effect_country.png"), plot = p_b_hu_median_country, width = 12, height = 6, units = "in")

p_b_hu_median_country

# median with ellipse
p_b_hu_median_ellipse = ggplot(taxon_subadult_rope, aes(x = hu_median, y = b_median, color = species)) +
  stat_ellipse(data = taxon_subadult, aes(x = hu_subadult, y = b_subadult), level = 0.95, alpha = 0.3) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point(size = 3) +
  geom_text_repel(aes(label = species), size = 4, seed = 135) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect",
       x = "Bd Prevalence\nMedian Subadult Effect Size w/ 95% Ellipse [-]",
       y = "Bd Load Density\nMedian Subadult Effect Size  w/ 95% Ellipse [-]") +
  theme(legend.position="none") +
  theme(text=element_text(size=16)) +
  scale_color_manual(values = rep("#660099", 19))

ggsave(here("clh", "median_effect_ellipse.png"), plot = p_b_hu_median_ellipse, width = 12, height = 6, units = "in")

p_b_hu_median_ellipse

# median with ellipse no text
p_b_hu_median_ellipse_no_text = ggplot(taxon_subadult_rope, aes(x = hu_median, y = b_median, color = species)) +
  stat_ellipse(data = taxon_subadult, aes(x = hu_subadult, y = b_subadult), level = 0.95, alpha = 0.3) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point(size = 3) +
  # geom_text_repel(aes(label = species), size = 4, seed = 135) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect",
       x = "Bd Prevalence\nMedian Subadult Effect Size w/ 95% Ellipse [-]",
       y = "Bd Load Density\nMedian Subadult Effect Size  w/ 95% Ellipse [-]") +
  theme(legend.position="none") +
  theme(text=element_text(size=16)) +
  scale_color_manual(values = rep("#660099", 19))

ggsave(here("clh", "median_effect_ellipse_no_text.png"), plot = p_b_hu_median_ellipse_no_text, width = 12, height = 6, units = "in")

p_b_hu_median_ellipse_no_text




# violins
hu_order = hu_taxon_subadult_rope %>%
  arrange(hu_median) %>%
  pull(species)

hu_taxon_subadult$taxon_capture = factor(hu_taxon_subadult$species, levels = hu_order)

hu_violin = ggplot(hu_taxon_subadult, aes(y = taxon_capture, x = hu_subadult)) +
  geom_violin(draw_quantiles = 0.5, fill = "grey80") +
  theme_minimal()


b_order = b_taxon_subadult_rope %>%
  arrange(b_median) %>%
  pull(species)

b_taxon_subadult$taxon_capture = factor(b_taxon_subadult$species, levels = b_order)

b_violin = ggplot(b_taxon_subadult, aes(x = taxon_capture, y = b_subadult)) +
  geom_violin(draw_quantiles = 0.5, fill = "grey80") +
  theme_minimal() +
  scale_x_discrete(position = "top") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0, hjust=0))

# HDI

hu_taxon_subadult_rope$taxon_capture = factor(hu_taxon_subadult_rope$species, levels = rev(hu_order))

hu_box = ggplot(hu_taxon_subadult_rope, aes(x = taxon_capture)) +
  geom_boxplot(
    aes(ymin = hdi_89$CI_low, ymax = hdi_89$CI_high,
        lower = hdi_89$CI_low, upper = hdi_89$CI_high, middle = hu_median),
    stat = "identity", fill = "black", color = "white") +
  theme_minimal() +
  annotate("rect", 
           ymin = -log_th, ymax = log_th, 
           xmin = -Inf, xmax = Inf,
           fill = "red", alpha = 0.5) +
  geom_hline(yintercept = 0, color = "red") +
  coord_flip()

b_taxon_subadult_rope$taxon_capture = factor(b_taxon_subadult_rope$species, levels = b_order)

b_box = ggplot(b_taxon_subadult_rope, aes(x = taxon_capture)) +
  geom_boxplot(
    aes(ymin = hdi_89$CI_low, ymax = hdi_89$CI_high,
        lower = hdi_89$CI_low, upper = hdi_89$CI_high, middle = b_median),
    stat = "identity", fill = "black", color = "white") +
  theme_minimal() +
  scale_x_discrete(position = "top") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0, hjust=0)) +
    annotate("rect", 
           ymin = -0.1 * median_b_sigma, ymax = 0.1 * median_b_sigma, 
           xmin = -Inf, xmax = Inf,
           fill = "red", alpha = 0.5) +
  geom_hline(yintercept = 0, color = "red")

# cophen heatmap

```

## absolute indicator values
```{r}

get_variables(model_select)

hu_taxon_adult = spread_draws(model_select, `b_hu_Intercept|b_hu_taxon_capture.*`, regex = TRUE) %>%
  mutate(`b_hu_taxon_captureanaxyrus_americanus` = 0) %>%
  pivot_longer(starts_with("b_hu_taxon_capture"),
               names_to = "parameter",
               values_to = "hu_adj") %>%
  mutate(taxon_capture = gsub("b_hu_taxon_capture", "", parameter),
         hu_adult = b_hu_Intercept + hu_adj,
         species = get_species_name(taxon_capture))

b_taxon_adult = spread_draws(model_select, `b_Intercept|b_taxon_capture.*`, regex = TRUE) %>%
  mutate(`b_taxon_captureanaxyrus_americanus` = 0) %>%
  pivot_longer(starts_with("b_taxon_capture"),
               names_to = "parameter",
               values_to = "b_adj") %>%
  mutate(taxon_capture = gsub("b_taxon_capture", "", parameter),
         b_adult = b_Intercept + b_adj,
         species = get_species_name(taxon_capture))

sigma_taxon = spread_draws(model_select, `b_sigma_Intercept|b_sigma_taxon_capture.*`, regex = TRUE) %>%
  mutate(`b_sigma_taxon_captureanaxyrus_americanus` = 0) %>%
  pivot_longer(starts_with("b_sigma_taxon_capture"),
               names_to = "parameter",
               values_to = "sigma_adj") %>%
  mutate(taxon_capture = gsub("b_sigma_taxon_capture", "", parameter),
         sigma = b_sigma_Intercept + sigma_adj,
         species = get_species_name(taxon_capture))

param_wide = hu_taxon_adult %>%
  left_join(hu_taxon_subadult %>%
              select(.draw,
                     taxon_capture,
                     hu_subadult) %>%
              rename(hu_subadult_marginal = hu_subadult), by = c(".draw", "taxon_capture")) %>%
  left_join(b_taxon_adult %>%
              select(.draw,
                     taxon_capture,
                     b_adult), by = c(".draw", "taxon_capture")) %>%
  left_join(b_taxon_subadult %>%
              select(.draw,
                     taxon_capture,
                     b_subadult) %>%
              rename(b_subadult_marginal = b_subadult), by = c(".draw", "taxon_capture")) %>%
    left_join(sigma_taxon %>%
              select(.draw,
                     taxon_capture,
                     sigma), by = c(".draw", "taxon_capture")) %>%
  mutate(hu_subadult = hu_adult + hu_subadult_marginal,
         b_subadult = b_adult + b_subadult_marginal)

param_long_hu = param_wide %>%
  pivot_longer(c("hu_adult",
                 "hu_subadult"),
               names_to = "life_stage",
               values_to = "hu")

param_long_b = param_wide %>%
  pivot_longer(c("b_adult",
                 "b_subadult"),
               names_to = "life_stage",
               values_to = "b")

param_wide_sum = param_wide %>%
  group_by(taxon_capture, species) %>%
  summarize(hu_adult_median = median(hu_adult),
            hu_subadult_median = median(hu_subadult),
            b_adult_median = median(b_adult),
            b_subadult_median = median(b_subadult),
            .groups = "drop")

ggplot(param_wide_sum, aes(x = hu_adult_median, y = hu_subadult_median, color = species)) +
  # stat_ellipse(data = taxon_subadult, aes(x = hu_subadult, y = b_subadult), level = 0.95, alpha = 0.3) +
  geom_point(size = 3) +
  geom_text_repel(aes(label = species), size = 4, seed = 135) +
  theme_minimal() +
  theme(legend.position="none") +
  theme(text=element_text(size=16)) +
  scale_color_manual(values = rep("#660099", 19)) +
  geom_abline(slope=1, intercept=0)

ggplot(param_wide_sum, aes(x = b_adult_median, y = b_subadult_median, color = species)) +
  # stat_ellipse(data = taxon_subadult, aes(x = hu_subadult, y = b_subadult), level = 0.95, alpha = 0.3) +
  geom_point(size = 3) +
  geom_text_repel(aes(label = species), size = 4, seed = 135) +
  theme_minimal() +
  theme(legend.position="none") +
  theme(text=element_text(size=16)) +
  scale_color_manual(values = rep("#660099", 19)) +
  geom_abline(slope=1, intercept=0)

ggplot(param_long_hu, aes(x = species, y = hu, color = life_stage)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

ggplot(param_long_b, aes(x = species, y = b, color = life_stage)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))


```

## Phylogenetic tree
```{r}
#| eval: true
#| echo: false
#| warning: false

# build tree from 
species_list = sort(get_species_name(bd_model_env_select_stats$taxon_capture))
tnrs_resolved = tnrs_match_names(species_list)
tnrs_info = taxonomy_taxon_info(ott_ids = ott_id(tnrs_resolved))
src = tax_sources(tnrs_info)
tr = tol_induced_subtree(ott_ids = ott_id(tnrs_resolved),
                         label_format = "name")

# source_studies <- map_dfr(tnrs_resolved$unique_name, function(species) {
#   result <- studies_find_trees(
#     property = "ot:ottTaxonName", 
#     value = species,
#     detailed = TRUE
#   )
#   
#   tibble(
#     species = species,
#     study_id = result$study_ids,
#     n_matched_trees = result$n_matched_trees,
#     study_year = result$study_year,
#     title = result$title
#   )
# })

# source_studies_wide = source_studies %>%
#   select(species, study_id) %>%
#   mutate(value = 1) %>%
#   pivot_wider(names_from = species,
#               values_from = value,
#               values_fill = 0) %>%
#   arrange(study_id)
# 
# source_studies_unique = source_studies %>%
#   select(-species) %>%
#   distinct() %>%
#   arrange(study_id)

tr_reordered = rotateConstr(tr, gsub(" ", "_", species_list))
tr_reordered$tip.label = get_species_name(tr_reordered$tip.label)

# plot tree
# ggtree(tr, layout = "rectangular") +
#   geom_tiplab(size = 3, align = TRUE) +
#   theme_tree2() +
#   scale_x_continuous(expand = expansion(mult = c(0, 0.3)),
#                      breaks = c(1, 3, 5, 7))

p_phylo_tree = ggtree(tr_reordered, layout = "rectangular") +
  geom_tiplab(size = 4, align = TRUE) +
  theme_tree2() +
  scale_x_continuous(expand = expansion(mult = c(0, 0.8)),
                     breaks = c(1, 3, 5, 7))

ggsave(here("clh", "phologeny_tree.png"), plot = p_phylo_tree, width = 5, height = 6, units = "in")

p_phylo_tree


# correlate phylogeny with trends
cophen_dist = cophenetic(tr_reordered)
cophen_sorted = cophen_dist[sort(rownames(cophen_dist)), sort(colnames(cophen_dist))]
cophen_upper = cophen_sorted[upper.tri(cophen_sorted, diag = FALSE)]

hu_dist = dist(hu_taxon_subadult_rope$hu_median)
b_dist = dist(b_taxon_subadult_rope$b_median)

hu_rho = cor.test(hu_dist, cophen_upper, method = "spearman")
hu_rho
b_rho = cor.test(b_dist, cophen_upper, method = "spearman")
b_rho
```

## Collective results
```{r}
#| eval: true
#| echo: true
#| output: false
#| code-fold: true
#| code-summary: Show code

# m5g_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2.rds")

choice_model = m5g_zln

model_data = brms_data

draws_all = collect_draws(choice_model) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

density_mat_all = density_mat(log(0.5), 20, 100, draws_all) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

sim_load_all = simulate_bd_load(draws_all, 10, 1000) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

obs_stats_all = sample_stats(model_data) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

```

::: {.panel-tabset}
### Observed Prevalence
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# observed prevalence
ggplot(obs_stats_all, aes(y = bd_prevalence, x = species, color = life_stage_simple)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = bd_prevalence_ci_lower, ymax = bd_prevalence_ci_upper),
                width = 0.2,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  ggtitle("Observed prevalence") +
  xlab("Species") +
  ylab("Bd prevalence [-]\n(with 95% CI)") +
  labs(color = "Life stage") +
  geom_text(aes(label = paste0("n=", n),  y = label_y),
          position = position_dodge(width = 0.5)) +
  ylim(0, 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Modeled Prevalence 
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# model prevalence
ggplot(draws_all, aes(prevalence, x = species, color = life_stage_simple)) +
  geom_boxplot(coef = NULL) + 
  theme_minimal() +
  ggtitle("Modeled prevalence") +
  xlab("Species") +
  ylab("Bd prevalence [-]\n") +
  labs(color = "Life stage") +
  ylim(0, 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Observed Bd load
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# obsevation bd load
ggplot(model_data %>%
         mutate(species = get_species_name(taxon_capture, newline = TRUE)) %>%
         filter(bd_load_co > 0), aes(log_bd_load_co_density, x = species, color = life_stage_simple)) +
  geom_boxplot() + 
  theme_minimal() +
  ggtitle("Observed Bd-positive load") +
  xlab("Species") +
  ylab("log(Bd load density) [-]") +
  labs(color = "Life stage") +
  geom_text(
    data = obs_stats_all,
    aes(
      x = species,
      y = max(model_data$log_bd_load_co_density) * label_y * 1.05,
      label = paste0("n=", n_pos),
      group = life_stage_simple
    ),
    position = position_dodge(width = 0.75),
    vjust = 0
  ) +
  ylim(log(0.5), 19) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Modeled Bd load
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# modeled bd load
ggplot(sim_load_all %>%
         filter(log_bd_load_co_density > 0.5), aes(log_bd_load_co_density, x = species, color = life_stage_simple)) +
  geom_boxplot() + 
  theme_minimal() +
  ggtitle("Modeled Bd-positive load") +
  xlab("Species") +
  ylab("log(Bd load density) [-]") +
  labs(color = "Life stage") +
  ylim(log(0.5), 20) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Modeled Bd load PDF
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# PDF -- facet by species
ggplot(density_mat_all, aes(x = density,
                            y = pdf_median,
                            color = life_stage_simple, 
                            fill = life_stage_simple)) +
  geom_ribbon(aes(ymin = pdf_lower,
                  ymax = pdf_upper), 
              alpha = 0.15,
              color = NA) +
  geom_line() +
  theme_minimal() +
  ggtitle("Modeled PDF of Bd-positive load") +
  labs(color = "Life stage",
       fill = "Life stage") +
  xlab("log(Bd load) [-]") +
  ylab("Median probability density [-]\n(with 25th to 75th percentiles)") +
  scale_x_continuous(limits = c(log(0.5), 20), expand = c(0, 0)) +
  facet_grid(. ~ gsub(" ", "\n", species))

# # PDF -- facet by life_stage & season
# ggplot(density_mat_all, aes(x = density,
#                             y = pdf_median,
#                             color = taxon_capture, 
#                             fill = taxon_capture)) +
#   geom_ribbon(aes(ymin = pdf_lower,
#                   ymax = pdf_upper), 
#               alpha = 0.15,
#               color = NA) +
#   geom_line() +
#   theme_minimal() +
#   ggtitle("Modeled PDF of Bd-positive load") +
#   labs(color = "Species",
#        fill = "Species") +
#   xlab("log(Bd load) [-]") +
#   ylab("Median probability density [-]\n(with 25th to 75th percentiles)") +
#   scale_x_continuous(limits = c(log(0.3), 20), expand = c(0, 0)) +
#   facet_grid(life_stage_simple ~ season)
```

### Moded Bd load CCDF
```{r}
#| echo: false
#| warning: false
#| fig-height: 8
#| fig-width: 8
# CCDF -- facet by species
ggplot(density_mat_all, aes(x = density,
                            y = ccdf_median,
                            color = life_stage_simple, 
                            fill = life_stage_simple)) +
  geom_ribbon(aes(ymin = ccdf_lower,
                  ymax = ccdf_upper), 
              alpha = 0.15,
              color = NA) +
  geom_line() +
  theme_minimal() +
  ggtitle("Modeled CCDF of Bd-positive load") +
  labs(color = "Life stage",
       fill = "Life stage") +
  xlab("log(Bd load) [-]") +
  ylab("Median complementary cumulative density (1 - CDF) [-]\n(with 25th to 75th percentiles)") +
  scale_x_continuous(limits = c(log(0.5), 12), expand = c(0, 0)) +
  ylim(0, 0.75) +
  # facet_grid(. ~ gsub(" ", "\n", species))
  facet_wrap("species")

# CCDF -- facet by life_stage & season
# ggplot(density_mat_all, aes(x = density,
#                             y = ccdf_median,
#                             color = taxon_capture,
#                             fill = taxon_capture)) +
#   geom_ribbon(aes(ymin = ccdf_lower,
#                   ymax = ccdf_upper),
#               alpha = 0.15,
#               color = NA) +
#   geom_line() +
#   theme_minimal() +
#   ggtitle("Modeled CCDF of Bd-positive load") +
#   labs(color = "Species",
#        fill = "Species") +
#   xlab("log(Bd load) [-]") +
#   ylab("Median complementary cumulative density (1 - CDF) [-]\n(with 25th to 75th percentiles)") +
#   scale_x_continuous(limits = c(log(0.3), 20), expand = c(0, 0)) +
#   ylim(0, 1) +
#   facet_grid(. ~ life_stage_simple)
```

# Environmental conditions

## Distributions
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Environmental conditions distributions code"
#| fig-height: 6
#| fig-width: 8

# 15-day
# ggplot(brms_data, aes(x = temp_c_15_day_mean, y = precip_mm_15_day_sum, color = life_stage_simple)) +
#   geom_point(alpha = 0.1) +
#   stat_ellipse() +
#   facet_wrap(. ~ taxon_capture) +
#   xlim(-5, 40) +
#   ylim(0, 200) +
#   labs(x = "15-day temperature mean (degC)", 
#        y = "15-day precipitation total (mm)",
#        title = "Temperature v. precipitation") +
#   theme_minimal()

# 30-day by species
ggplot(brms_data, aes(x = temp_c_30_day_mean, y = precip_mm_30_day_sum, color = life_stage_simple)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  facet_wrap(. ~ taxon_capture) +
  xlim(-5, 40) +
  ylim(0, 400) +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()

# 30 day collective
ggplot(brms_data, aes(x = temp_c_30_day_mean, y = precip_mm_30_day_sum, color = taxon_capture)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()

# 30 day collective, scaled to climate by species
ggplot(brms_data, aes(x = temp_c_30_sp_z, y = precip_mm_30_sp_z, color = taxon_capture)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()


# ggplot(brms_data, aes(x = temp_c_30_sp_z2, y = precip_mm_30_sp_z2, color = life_stage_simple)) +
#   geom_point(alpha = 0.1) +
#   stat_ellipse() +
#   facet_wrap(. ~ taxon_capture) +
#   labs(x = "30-day temperature mean (degC)", 
#        y = "30-day precipitation total (mm)",
#        title = "Temperature v. precipitation") +
#   theme_minimal()
```

## Tempurature effects
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show temperature effects code"
#| fig-height: 6
#| fig-width: 8

select_model = m5f_zln
# select_model = m6_zln

climate_draws = spread_draws(select_model, `b_temp_c.*|b_precip_mm.*|b_hu_temp_c.*|b_hu_precip_mm.*|b_Intercept|b_hu_Intercept`, regex = TRUE)

# Temperature dependence
temp_range = seq(-5, 45, length.out = 100)

temp_grid = expand_grid(temp_c_15_day_mean = temp_range, .draw = 1:nrow(climate_draws)) %>%
  mutate(temp_c_30_day_mean = temp_c_15_day_mean) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    temp_c_15_z = (temp_c_15_day_mean - brms_scales$temp_c_15_day_mean_mean) / brms_scales$temp_c_15_day_mean_sd,
    temp_c_15_z2 = (temp_c_15_day_mean^2 - brms_scales$temp_c_15_day_mean2_mean) / brms_scales$temp_c_15_day_mean2_sd,
    temp_c_30_z = (temp_c_30_day_mean - brms_scales$temp_c_30_day_mean_mean) / brms_scales$temp_c_30_day_mean_sd,
    temp_c_30_z2 = (temp_c_30_day_mean^2 - brms_scales$temp_c_30_day_mean2_mean) / brms_scales$temp_c_30_day_mean2_sd,
    b_pred = b_temp_c_30_z * temp_c_30_z + b_temp_c_30_z2 * temp_c_30_z2 + b_Intercept,
    hu_pred = b_hu_temp_c_15_z * temp_c_15_z + b_hu_temp_c_15_z2 * temp_c_15_z2 + b_hu_Intercept
  ) %>%
  group_by(temp_c_15_day_mean) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  ) %>%
  rename(temp_c = temp_c_15_day_mean)

# prevalence
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day mean temperature (degC)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. temperature") +
  theme_minimal()

# load
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day mean temperature (degC)", 
       y = "log(+ Bd load)",
       title = "Bd load v. temperature") +
  theme_minimal()
```

## Precipitation effects
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show precipitation effects code"
#| fig-height: 6
#| fig-width: 8
# precipitation depencence

daily_precip_range = seq(0, 20, length.out = 100)

precip_grid = expand_grid(daily_precip = daily_precip_range, .draw = 1:nrow(climate_draws)) %>%
  mutate(precip_mm_15_day_sum = daily_precip * 15,
         precip_mm_30_day_sum = daily_precip * 30) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    precip_mm_15_z = (precip_mm_15_day_sum - brms_scales$precip_mm_15_day_sum_mean) / brms_scales$precip_mm_15_day_sum_sd,
    precip_mm_15_z2 = (precip_mm_15_day_sum^2 - brms_scales$precip_mm_15_day_sum2_mean) / brms_scales$precip_mm_15_day_sum2_sd,
    precip_mm_30_z = (precip_mm_30_day_sum - brms_scales$precip_mm_30_day_sum_mean) / brms_scales$precip_mm_30_day_sum_sd,
    precip_mm_30_z2 = (precip_mm_30_day_sum^2 - brms_scales$precip_mm_30_day_sum2_mean) / brms_scales$precip_mm_30_day_sum2_sd,
    b_pred = b_precip_mm_30_z * precip_mm_30_z + b_precip_mm_30_z2 * precip_mm_30_z2 + b_Intercept,
    hu_pred = b_hu_precip_mm_15_z * precip_mm_15_z + b_hu_precip_mm_15_z2 * precip_mm_15_z2 + b_hu_Intercept
  ) %>%
  group_by(daily_precip,
           precip_mm_15_day_sum,
           precip_mm_30_day_sum) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  )

# prevalence
ggplot(precip_grid, aes(x = precip_mm_15_day_sum)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day total precipitation (mm)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. precipitation") +
  theme_minimal()

# load
ggplot(precip_grid, aes(x = precip_mm_30_day_sum)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day total precipitation (mm)", 
       y = "log(+ Bd load)",
       title = "Bd load v. precipitation") +
  theme_minimal()

```

## Tempurature effects (scaled by species climate)
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show temperature effects code"
#| fig-height: 6
#| fig-width: 8


select_model = m6_zln

# fixed effects
climate_draws = spread_draws(select_model, `b_temp_c.*|b_precip_mm.*|b_hu_temp_c.*|b_hu_precip_mm.*|b_Intercept|b_hu_Intercept`, regex = TRUE)

# Temperature dependence
temp_range = seq(-3, 3, length.out = 100)

# taxa = brms_data %>%
#   select(taxon_capture) %>%
#   distinct() %>%
#   pull(taxon_capture) %>%
#   sort()

## Stumped by how to back-calculate the temp_z2 values from the temp_z values... I think this has to be done on a species-by-species basis...

temp_grid = expand_grid(temp_z_15_mean = temp_range,
                        .draw = 1:nrow(climate_draws)) %>%
  mutate(temp_z_30_mean = temp_z_15_mean) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    temp_c_15_z = (temp_c_15_day_mean - brms_scales$temp_c_15_day_mean_mean) / brms_scales$temp_c_15_day_mean_sd,
    temp_c_15_z2 = (temp_c_15_day_mean^2 - brms_scales$temp_c_15_day_mean2_mean) / brms_scales$temp_c_15_day_mean2_sd,
    temp_c_30_z = (temp_c_30_day_mean - brms_scales$temp_c_30_day_mean_mean) / brms_scales$temp_c_30_day_mean_sd,
    temp_c_30_z2 = (temp_c_30_day_mean^2 - brms_scales$temp_c_30_day_mean2_mean) / brms_scales$temp_c_30_day_mean2_sd,
    b_pred = b_temp_c_30_sp_z * temp_c_30_z + b_temp_c_30_sp_z2 * temp_c_30_z2 + b_Intercept,
    hu_pred = b_hu_temp_c_15_sp_z * temp_c_15_z + b_hu_temp_c_15_sp_z2 * temp_c_15_z2 + b_hu_Intercept
  ) %>%
  group_by(temp_c_15_day_mean) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  ) %>%
  rename(temp_c = temp_c_15_day_mean)

# prevalence
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day mean temperature (degC)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. temperature") +
  theme_minimal()

# load
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day mean temperature (degC)", 
       y = "log(+ Bd load)",
       title = "Bd load v. temperature") +
  theme_minimal()
```

## Precipitation effects (scaled by species climate)
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show precipitation effects code"
#| fig-height: 6
#| fig-width: 8
# precipitation depencence

daily_precip_range = seq(0, 20, length.out = 100)

precip_grid = expand_grid(daily_precip = daily_precip_range, .draw = 1:nrow(climate_draws)) %>%
  mutate(precip_mm_15_day_sum = daily_precip * 15,
         precip_mm_30_day_sum = daily_precip * 30) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    precip_mm_15_z = (precip_mm_15_day_sum - brms_scales$precip_mm_15_day_sum_mean) / brms_scales$precip_mm_15_day_sum_sd,
    precip_mm_15_z2 = (precip_mm_15_day_sum^2 - brms_scales$precip_mm_15_day_sum2_mean) / brms_scales$precip_mm_15_day_sum2_sd,
    precip_mm_30_z = (precip_mm_30_day_sum - brms_scales$precip_mm_30_day_sum_mean) / brms_scales$precip_mm_30_day_sum_sd,
    precip_mm_30_z2 = (precip_mm_30_day_sum^2 - brms_scales$precip_mm_30_day_sum2_mean) / brms_scales$precip_mm_30_day_sum2_sd,
    b_pred = b_precip_mm_30_z * precip_mm_30_z + b_precip_mm_30_z2 * precip_mm_30_z2 + b_Intercept,
    hu_pred = b_hu_precip_mm_15_z * precip_mm_15_z + b_hu_precip_mm_15_z2 * precip_mm_15_z2 + b_hu_Intercept
  ) %>%
  group_by(daily_precip,
           precip_mm_15_day_sum,
           precip_mm_30_day_sum) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  )

# prevalence
ggplot(precip_grid, aes(x = precip_mm_15_day_sum)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day total precipitation (mm)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. precipitation") +
  theme_minimal()

# load
ggplot(precip_grid, aes(x = precip_mm_30_day_sum)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day total precipitation (mm)", 
       y = "log(+ Bd load)",
       title = "Bd load v. precipitation") +
  theme_minimal()

```