---
title: "Size vs. Bd modeling: Panamá preliminary results"
author: Cob Staines
date: 2025-08-20
format:
  html:
    toc: true
df-print: kable
bibliography: references.bib
csl: ecology.csl
---

## Setup
```{r}
#| eval: true
#| output: false
#| code-fold: true
#| code-summary: "Show setup code"

# setup
librarian::shelf(tidyverse, dbplyr, here, janitor, RPostgres, DBI, RIBBiTR-BII/ribbitrrr, brms, loo, bayesplot, tidybayes, ggplot2)

## Connect to DB
dbcon <- hopToDB("ribbitr")

# load populations
data_pop = read.csv(here("data", "sn_site_pop_2025-05-15.csv")) %>%
  rename(population = pop_id) %>%
  mutate(population = ifelse(!is.na(population), paste0(as.character(population), "_pop"), NA_character_),
         population = if_else(is.na(population), site, population))

# table pointers
db_bd = tbl(dbcon, Id("survey_data", "bd_qpcr_results"))
db_sample = tbl(dbcon, Id("survey_data", "sample"))
db_capture = tbl(dbcon, Id("survey_data", "capture"))
db_ves = tbl(dbcon, Id("survey_data", "ves"))
db_survey = tbl(dbcon, Id("survey_data", "survey"))
db_visit = tbl(dbcon, Id("survey_data", "visit"))
db_site = tbl(dbcon, Id("survey_data", "site"))
db_region = tbl(dbcon, Id("survey_data", "region"))
db_country = tbl(dbcon, Id("survey_data", "country"))
db_hist = tbl(dbcon, Id("survey_data", "site_history"))

bd_sample = db_sample %>%
  right_join(db_capture, by = "capture_id") %>%
  left_join(db_survey, by = "survey_id") %>%
  left_join(db_visit, by = "visit_id") %>%
  left_join(db_site, by = "site_id") %>%
  left_join(db_region, by = "region_id") %>%
  left_join(db_country, by = "country_id") %>%
  filter(sample_type == "bd")
```

# define functions
```{r}
# fuzzy grouping for identifying survey periods in time
fuzzy_grouping = function(x, delta){
  # Make groups of x if differences are less than or equal to delta
  # 
  # Parameters
  # ----------
  # x : array of values to be grouped
  # delta : difference in value that defines a group
  
  
  if(length(x) > 1){
    
    run_ind = diff(x) <= delta
    grouping = array(NA, dim=length(run_ind + 1))
    grouping[1] = 1
    for(i in 1:length(run_ind)){
      
      if(run_ind[i] == FALSE | is.na(run_ind[i])){
        grouping[i + 1] = grouping[i] + 1
      }
      else{
        grouping[i + 1] = grouping[i]
      }
      
    }
  } else{
    grouping = 1
  }
  
  return(grouping)

}

# binomial confidence interval
wilson_ci = function(x, n, conf.level = 0.95) {
  p_hat <- x / n
  z <- qnorm(1 - (1 - conf.level) / 2)
  
  denominator <- 1 + (z^2 / n)
  centre_adjusted_probability <- p_hat + (z^2 / (2 * n))
  adjusted_standard_deviation <- sqrt((p_hat * (1 - p_hat) / n) + (z^2 / (4 * n^2)))
  
  lower_bound <- (centre_adjusted_probability - z * adjusted_standard_deviation) / denominator
  upper_bound <- (centre_adjusted_probability + z * adjusted_standard_deviation) / denominator
  
  return(c(lower_bound = lower_bound, upper_bound = upper_bound))
}

wilson_ci_lower = function(x, n, conf.level = 0.95) {
  ws = wilson_ci(x, n, conf.level)
  return(ws[1])
}

wilson_ci_upper = function(x, n, conf.level = 0.95) {
  ws = wilson_ci(x, n, conf.level)
  return(ws[2])
}

# normal confidence interval
normal_ci = function(x, conf.level = 0.95) {
  n <- length(x)
  mean_x <- mean(x)
  sd_x <- sd(x)
  
  alpha <- 1 - conf.level
  t_value <- qt(1 - alpha/2, df = n - 1)
  margin_error <- t_value * sd_x / sqrt(n)
  
  lower <- mean_x - margin_error
  upper <- mean_x + margin_error
  
  return(c(lower_bound = lower, upper_bound = upper))
}

normal_ci_lower = function(x, conf.level = 0.95) {
  no = normal_ci(x, conf.level)
  return(no[1])
}

normal_ci_upper = function(x, conf.level = 0.95) {
  no = normal_ci(x, conf.level)
  return(no[2])
}


```

## Methods

### Data selection & classification

0. Beginning with all Rana muscosa & Rana Sierrae data, grouped by population
1. Filter out post-translocation data
2. Population selection criteria:
  a. minimum 15 swabbed subadult and adult
  b. minimum 3 survey years
3. Classify by:
  a. Bd Naive: low (or no detected) Bd prevalence & load, high population
  b. Epizootic: high Bd prevalence & load, followed by low population
  c. Post-Epizootic: mid Bd prevalence & load
    i. Persistant (low population)
    ii. Recovering (high population)
4. Review:
  a. Yosemite all post-epizootic at start of swabbing, except for Galiston (Epizootic ~2020)
  b. Compare with SEKI outbreak map fro agreement
  c. R. muscosa -- is Mulki the only post-epizootic site? If so, skip this analysis for R muscosa
  
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show data collection code"

# identify populations
bd_population = bd_sample %>%
  filter(country == "usa",
         region == "california",
         life_stage != "tadpole",
         !is.na(taxon_capture %in% c("rana_muscosa", "rana_sierrae"))) %>%
  left_join(data_pop, by = "site", copy = TRUE) %>%
  mutate(tax_pop = paste0(taxon_capture, "_", population),
         year = year(date),
         life_stage_simple = life_stage,
         life_stage_simple = case_when(
           is.na(svl_mm) ~ life_stage,
           (svl_mm >= 40) ~ "adult",
           (svl_mm < 40) ~ "subadult",
           .default = life_stage))

# selection criteria for populations
## - 3 years of sampling
## - at least 15 each adult and subadult per population

bd_population_valid = bd_population %>%
  collect() %>%
  group_by(population) %>%
  summarise(n_years = n_distinct(year),
            n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop") %>%
  filter(n_years >= 3,
         n_adult >= 15,
         n_subadult >= 15)

bd_site_valid = data_pop %>%
  filter(population %in% bd_population_valid$population)

bd_capture_valid = bd_population %>%
  filter(population %in% bd_population_valid$population)

# all years of data for valid pops
data_year_valid = bd_capture_valid %>%
  distinct(population, year) %>%
  collect()
  

# all population - year combinations
pop_year = expand_grid(bd_population_valid$population,
                       unique(data_year_valid$year)) %>%
  rename(population = 1,
         year = 2)

# identify translocation events
release_events = db_hist %>%
  left_join(db_site, by = "site_id") %>%
  left_join(db_region, by = "region_id") %>%
  filter(region == "california",
         treatment_type %in% c("translocation_release", "reintroduction_release")) %>%
  collect() %>%
  left_join(data_pop, by = "site") %>%
  mutate(year = year(treatment_date),
         release_event = TRUE) %>%
  select(population,
         year,
         release_event) %>%
  distinct()

# pass on to pop year
pop_year_tx = pop_year %>%
  left_join(release_events, by = c("population", "year"))

```

# Calculate adult abundance indicators
```{r}
# sites of interest
site_oi = db_site %>%
  filter(site %in% bd_site_valid$site)

# visits of interest
visit_oi = db_visit %>%
  left_join(site_oi, by = "site_id") %>%
  mutate(year = year(date)) %>%
  filter(visit_status == "suitable",
         year %in% year_valid$year) %>%
  select(visit_id,
         site,
         site_id,
         year,
         date,
         comments_visit) %>%
  collect() %>%
  left_join(data_pop, by = "site") %>%
  arrange(population,
          date) %>%
  group_by(population, year) %>%
  mutate(survey_period = fuzzy_grouping(as.integer(date), 7)) %>% # consider surveys within a rolling 7 days to be within the same survey period
  ungroup()

# join with survey table
survey_oi = db_survey %>%
  filter(visit_id %in% visit_oi$visit_id,
         detection_type %in% c("capture", "visual")) %>%
  select(detection_type,
         detection_subtype,
         survey_quality,
         description,
         comments_survey,
         number_observers,
         survey_id,
         visit_id) %>%
  collect() %>%
  mutate(detection_type = ifelse(detection_type == "capture",
                                 detection_subtype,
                                 detection_type)) %>%
  select(-detection_subtype) %>%
  right_join(visit_oi, by = "visit_id") %>%
  arrange(site,
          date,
          survey_period)

# # Survey periods of interest
# 
# Selection criteria when more than 1 survey period available for a given & year:
# 
# 0) greatest number of sites within survey period (for populations with multiple sites)
# 1) greatest number of survey types (capture, ves, swab) within survey period
# 2) greatest mean survey_quality within survey period: good (3) > fair (2) > poor (1) > NA (0)
# 3) greatest number of surveys within survey period
# 4) randomly select one
# 
# * Note that this prioritizes using concurrent surveys between different detection types, but it is possible for some detection types for a given site/year to have different survey periods when concurrent surveys are not available.
# 
# ** where more than 1 VES survey was conducted in a survey period (rare) the 1st survey is used.

# set random seed for consistent selection
set.seed(482)

# calculate selection criteria
survey_oi_priority = survey_oi %>%
  mutate(survey_quality_score = case_match(survey_quality,
                                           "good" ~ 3,
                                           "fair" ~ 2,
                                           "poor" ~ 1,
                                           NA ~ 0,
                                           .default = NA)) %>%
  group_by(site,
           year) %>%
  mutate(period_count = n_distinct(survey_period)) %>%
  ungroup() %>%
  group_by(site,
           year,
           survey_period) %>%
   mutate(detection_type_count = n_distinct(detection_type),
         mean_survey_quality_score = mean(survey_quality_score),
         survey_count = n(),
         random_val = runif(1)) %>%
  ungroup()

# select survey periods for cmr
period_select_cmr = survey_oi_priority %>%
  filter(detection_type == "cmr") %>%
  distinct(population,
           site,
           year,
           survey_period,
           detection_type_count,
           mean_survey_quality_score,
           survey_count,
           random_val,
           detection_type) %>%
  arrange(site,
          year,
          survey_period,
          detection_type_count,
          mean_survey_quality_score,
          survey_count,
          random_val) %>%
  group_by(site,
           year) %>%
  slice_head(n = 1) %>%
  ungroup()

# select survey periods for swab
period_select_swab = survey_oi_priority %>%
  filter(detection_type == "swab") %>%
  distinct(population,
           site,
           year,
           survey_period,
           detection_type_count,
           mean_survey_quality_score,
           survey_count,
           random_val,
           detection_type) %>%
  arrange(site,
          year,
          survey_period,
          detection_type_count,
          mean_survey_quality_score,
          survey_count,
          random_val) %>%
  group_by(site,
           year) %>%
  slice_head(n = 1) %>%
  ungroup()

# select survey periods for visual
period_select_visual = survey_oi_priority %>%
  filter(detection_type == "visual") %>%
  distinct(population,
           site,
           year,
           survey_period,
           detection_type_count,
           mean_survey_quality_score,
           survey_count,
           random_val,
           detection_type) %>%
  arrange(site,
          year,
          survey_period,
          detection_type_count,
          mean_survey_quality_score,
          survey_count,
          random_val) %>%
  group_by(site,
           year) %>%
  slice_head(n = 1) %>%
  ungroup()

# bring selected periods together across detection types
period_select = bind_rows(period_select_cmr,
                          period_select_swab,
                          period_select_visual) %>%
  select(population,
         site,
         year,
         survey_period,
         detection_type) %>%
  rename(survey_period_select = survey_period)

# check for duplicates
dupes_period_select = get_dupes(period_select, site, year, detection_type)

# wider table for visual inspection of available surveys
period_select_wide = period_select %>%
  mutate(detection_type = paste0(detection_type, "_survey_period")) %>%
  pivot_wider(id_cols = c(population, site, year),
              names_from = detection_type,
              values_from = survey_period_select) %>%
  arrange(site,
          year)

# visually check selections where multiple periods present for a site/year, to see if any should be flagged
check_period_select = survey_oi_priority %>%
  left_join(period_select, by = c("population", "site", "year", "detection_type")) %>%
  filter(period_count > 1) %>%
  select(population,
         site,
         year,
         survey_period,
         survey_period_select,
         detection_type,
         description,
         comments_survey,
         comments_visit,
         detection_type_count,
         mean_survey_quality_score,
         survey_count,
         random_val,
         everything()) %>%
  arrange(site,
          year,
          survey_period,
          detection_type)

# surveys selected using criteria, across all detection types
survey_select = survey_oi_priority %>%
  right_join(period_select, by = c("population", "site", "year", "survey_period" = "survey_period_select", "detection_type")) %>%
  arrange(site,
          date,
          detection_type) %>%
  group_by(site,
           year,
           detection_type) %>%
  mutate(id = row_number()) %>%
  ungroup() %>%
  filter((detection_type == "visual" & id == 1) | detection_type %in% c("cmr", "swab"))

# adult frog abundance

# ## visual
# visual detection per survey period (1 survey per survey period)
# 
# ## cmr
# cmr detection within survey period (across all surveys in survey period)

# visual data pull
data_visual = survey_select %>%
  filter(detection_type == "visual") %>%
  select(population,
         site,
         date,
         survey_period,
         year,
         survey_id,
         visit_id,
         detection_type,
         comments_survey,
         comments_visit) %>%
  left_join(db_ves %>%
               filter(survey_id %in% survey_select$survey_id,
                      life_stage == "adult",
                      visual_animal_state != "dead") %>%
               select(count_ves,
                      ves_id,
                      survey_id) %>%
               collect(), by = "survey_id") %>%
  mutate(count_ves = ifelse(is.na(count_ves), 0, count_ves))

# visual abundance calc
abundance_visual = data_visual %>%
  group_by(population,
           year,
           date,
           ) %>%
  rename(date_ves = date) %>%
  summarise(adult_count_visual = sum(count_ves),
            .groups = "drop")

# cmr data pull
data_cmr = survey_select %>%
  filter(detection_type == "cmr") %>%
  select(population,
         site,
         date,
         survey_period,
         year,
         survey_id,
         visit_id,
         detection_type,
         comments_survey,
         comments_visit) %>%
  inner_join(db_capture %>%
               filter(survey_id %in% survey_select$survey_id,
                     life_stage == "adult",
                     capture_animal_state != "dead") %>%
               select(cmr_id,
                      capture_id,
                      survey_id,
                      life_stage,
                      capture_animal_state) %>%
               collect(), by = "survey_id")

# cmr abundance calc
abundance_cmr = data_cmr %>%
  group_by(population,
           year) %>%
  summarise(date_cmr_first = first(date),
            cmr_days = n_distinct(date),
            adult_count_cmr = n_distinct(cmr_id),
            .groups = "drop")

# join visual and cmr abundance data
abundance_stats = abundance_visual %>%
  full_join(abundance_cmr, by = c("population", "year"))

# join with pop_year_tx
pop_year_ab = pop_year_tx %>%
  left_join(abundance_stats %>%
              select(population,
                     year,
                     adult_count_cmr,
                     adult_count_visual), by = c("population", "year"))

```


# Calculate adult Bd stats
```{r}
bd_stats = bd_capture_valid %>%
  filter(life_stage_simple == "adult") %>%
  inner_join(db_bd %>%
              filter(!is.na(detected)) %>%
              group_by(sample_id) %>%
              summarise(bd_swab_count = n(),
                        bd_detected = any(detected),
                        bd_mean_its1_copies_per_swab = mean(bd_its1_copies_per_swab),
                        .groups = "drop"),
             by = "sample_id") %>%
  collect() %>%
  group_by(population,
           year) %>%
  summarise(adult_swab_n = n(),
            adult_bd_prevalence = sum(bd_detected) / n(),
            adult_bd_prevalence_ci_lower = wilson_ci_lower(sum(bd_detected), n(), conf.level = 0.95),
            adult_bd_prevalence_ci_upper = wilson_ci_upper(sum(bd_detected), n(), conf.level = 0.95),
            adult_positive_bd_load_median = median(bd_mean_its1_copies_per_swab[bd_detected]),
            adult_positive_bd_load_ci_lower = exp(normal_ci_lower(log(bd_mean_its1_copies_per_swab[bd_detected]), conf.level = 0.95)),
            adult_positive_bd_load_ci_upper = exp(normal_ci_upper(log(bd_mean_its1_copies_per_swab[bd_detected]), conf.level = 0.95)),
            .groups = "drop")

# join with pop_year_ab
pop_year_bd = pop_year_ab %>%
  left_join(bd_stats, by = c("population", "year")) %>%
  arrange(population, year)

```

