---
title: "Variation in Bd Intensity Through Size of Adult Amphibians"
format: html
editor: source
---

Guiding Question: How does Bd intensity change with size of adult amphibians across RIBBiTR survey locations?

-   Include toads?

-   Need to determine appropriate species per location with significant swab results. How many data points needed?


## load packages

```{r}

if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(tidyverse, here, DBI, RPostgres)

```

## connect to database

```{r}
tryCatch({
    drv <- dbDriver("Postgres")
    print("Connecting to Database…")
    connection <- dbConnect(drv,
                 dbname = Sys.getenv("aws_dbname"),
                 host = Sys.getenv("aws_host"),
                 port = Sys.getenv("aws_port"),
                 user = Sys.getenv("aws_user"),
                 password = Sys.getenv("aws_password"),
                 timezone=NULL)
    print("Database Connected!")
    },
    error=function(cond) {
            print("Unable to connect to Database.")
    })

#search path
dbExecute(connection, "set search_path to survey_data")
```

## query data of interest

```{r}



q <- "select l.location, r.region, s.site, v.date, s2.detection_type, c.species_capture, 
      c.svl_mm, c.life_stage, qbr.average_copy_number, qbr.detected 
      from location l 
      join region r on l.location_id = r.location_id 
      join site s on r.region_id = s.region_id 
      join visit v on s.site_id = v.site_id 
      join survey s2 on v.visit_id = s2.visit_id 
      join capture c on s2.survey_id = c.survey_id 
      join qpcr_bd_results qbr on c.bd_swab_id = qbr.bd_swab_id
      where c.svl_mm is not null
      and c.life_stage = 'adult'
      and qbr.detected = '1'";

raw_data <- dbGetQuery(connection, q) 

```

## Find good species list
```{r}

infec_size <- raw_data %>% 
  group_by(species_capture) %>% 
  summarise(n = n()) %>% 
  filter(n > 10) %>% 
  filter(!species_capture %in% c("anaxyrus_americanus", "notophthalmus_viridescens", "anaxyrus_fowleri"))   #remove toad and newt

species_list <- infec_size$species_capture

```


## Wrangle data

```{r}

data <- raw_data %>% 
  filter(species_capture %in% species_list) %>% 
  mutate(average_copy_number = log10(average_copy_number + 1))

## Explore
# ggplot(data = data, aes(x = svl_mm, y = average_copy_number)) +
#   geom_point() +
#   geom_smooth() +
#   facet_wrap(~species_capture)

```


## lm() Species Loop
```{r}

plots <- list()

for (i in 1:length(species_list)) {
  
  
  # data for each model
  dat <- data %>% 
    select(species_capture, average_copy_number, svl_mm) %>% 
    filter(species_capture == species_list[i]) 
  
  # run lm average_copy_number ~ svl_mm per species
  mod <- lm(average_copy_number ~ svl_mm, data=dat)
  
  # create model outputs as objects in environment
  assign(species_list[i], mod)
  
  # predict
  svl_mm = seq(min(dat$svl_mm), max(dat$svl_mm), len=50)
  #average_copy_number = seq(min(dat$average_copy_number), max(dat$average_copy_number), len = 50)
  newdata <- data.frame(svl_mm = svl_mm)
  mod_pred = predict(mod, newdata = newdata, se.fit = T)
  newdata$pred <- mod_pred$fit
  newdata$upper_se <- mod_pred$fit + (2 * mod_pred$se.fit)
  newdata$lower_se <- mod_pred$fit - (2 * mod_pred$se.fit)
  
  
  # make GAM plot
  p <- ggplot(newdata) + 
   geom_line(aes(x=svl_mm, y=pred)) +
   geom_point(data=dat, aes(x=svl_mm, y=average_copy_number), size=2, alpha = 0.2) +
   geom_ribbon(aes(ymin = lower_se, ymax= upper_se, x = svl_mm), alpha = 0.2, fill = "blue")+
   xlab("Snout Vent Length (mm)") + ylab("Log10 Infection Intensity") + theme_classic() +
   coord_cartesian(ylim = c(0,7)) + ggtitle(paste(species_list[i], ", slope:", summary(mod)$coefficients[2,1]))
    
 # save GAM plot in list
  plots[[i]] <- p  
  
  # residuals vs fitted
  plot(mod, 1, main = species_list[i])
  
  # Error checking during loop
  print(paste("done with", species_list[i]))
  
}


```

## Examine Model Plots
```{r}

plots[1:length(plots)]

```

## Examine residuals vs fitted
```{r}

res_vs_fit[1:length(res_vs_fit)]

```

