---
title: "01_survey_data"
output: html_document
---

# Introduction

## Motivation

Patterns of greater Bd load in recently metamorphosed individuals have been documented in some cases[@adams_chytridiomycosis-induced_2020] [@humphries_immune_2022] [@humphries_chytridiomycosis_2024] [@humphries_chytridiomycosis_2025] (and the opposite pattern in others[@bradley_host_2019]), with possible implications for:

  - resilience trajectory assessment (Q1)
  - population dynamics (Q2)
  - assessing Bd resistance/tolerance mechanisms (Q3)
  
... with the goal of complementing ongoing RIBBiTR research efforts

## Working hypotheses:

  a. Recently-metamorphosed individuals are more likely to be infected with Bd than older individuals across species
  b. Recently-metamorphosed individuals are more likely to carry higher Bd loads than older individuals across species

# Methods
## Setup

```{r}
#| eval: true
#| output: false
#| code-fold: true
#| code-summary: "Show setup code"

# setup
librarian::shelf(tidyverse, dbplyr, here, janitor, RPostgres, DBI, RIBBiTR-BII/ribbitrrr, terra, rio)

```
## Data selection
### Classifying by life stage
These analyses were carried out for post-metamorphic individuals only. Individuals were classified as "adult" or "subadult" with a threshold snout-vent-length (SVL [mm]), identified independently for each species from the literature (see data collection code for specific thresholds used and their sources).
  
### Classification by population
"Populations" were defined at the region level within Panamá. In the Sierra Nevada, sites were grouped where evidence supported connectivity at he scale of individuals. Elsewhere in North America, populations were defined at the site level.

### Species selection criteria:
Species were included if supported by least 14 swabbed adults and subadults (each).

### Bd density and Zero-cutoff
I scaled all Bd quantities by the Snout-Vent-Length (mm) to arrive at a Bd density per linear distance of skin swabbed (within species), to correct for linear effects of larger swabbing area in larger frogs. I also set any Bd ITS copy numbers < 10 to zero, to represent uncertainty in qPCR results and support consistency in zero-inflation modeling.


### RABO
```{r}
# Amphibian Disease Portal
adp_raw = read_csv(here("data", "amphibian_disease_data_processed.csv"))

adp_filter = adp_raw %>%
  filter(!is.na(length),
         !is.na(decimalLatitude),
         diseaseTested == "Bd",
         testMethod %in% c("quantitative PCR", "realtime PCR"),
         basisOfRecord == "LivingSpecimen")

adp_group = adp_filter %>%
  group_by(projectId, diagnosticLab, family, order, genus, specificEpithet, length >= 38) %>%
  count()

peace = adp_filter %>%
  select(testMethod) %>%
  distinct()

adp_select = adp_filter %>%
  filter(genus == "Rana",
         specificEpithet == "boylii",
         diagnosticLab == "Briggs Lab, University of California Santa Barbara")

rabo_clean = adp_select %>%
  mutate(svl_mm = length,
         life_stage_simple = case_when(
           svl_mm >= 38 ~ "adult",
           svl_mm < 38 ~ "subadult"),
         date = as.Date(verbatimEventDate),
         bd_its_copies = quantityDetected * as.numeric(dilutionFactor) * 60,
         taxon_capture = "rana_boylii",
         site = tolower(gsub(",", "", gsub(" ", "_", locality))),
         population = "alameda",
         region = "california",
         country = "usa",
         project = "RABO")

rabo_sites = rabo_clean %>%
  group_by(site) %>%
  summarize(site_latitude = mean(decimalLatitude, na.rm = TRUE),
            site_longitude = mean(decimalLongitude, na.rm = TRUE))

rabo_valid = rabo_clean %>%
  select(taxon_capture,
         date,
         site,
         population,
         dilutionFactor,
         quantityDetected,
         bd_its_copies,
         life_stage_simple,
         sex,
         svl_mm,
         region,
         country,
         project) %>%
  left_join(rabo_sites, by = "site")

```

### RACA
```{r}

# csv data
##  RACA data
raca_clean = read.csv(here("data", "JPSL_RACA_Swabbed_Individuals_2026-03-03.csv")) %>%
  clean_names() %>%
  filter(life_stage %in% c("A", "S", "Y")) %>%
  dplyr::rename(site = site_name,
         body_mass_g = weight_g,
         site_utme = utme_site,
         site_utmn = utmn_site,
         bd_its_copies = bd_swab_results_its) %>%
  mutate(life_stage_simple = case_when(
    svl_mm >= 50 ~ "adult",
    svl_mm < 50 ~ "subadult"),
    date = as.Date(date),
    site = tolower(gsub(" ", "_", site)),
    site_utm_zone = 11,  # verify
    population = site,
    region = "W",
    country = "usa",
    project = "RACA",
    taxon_capture = "rana_cascadae") %>%
  filter(!is.na(bd_its_copies))


raca_sites = raca_clean %>%
  select(site,
         site_utme,
         site_utmn) %>%
  distinct()

# convert to lat lon
raca_points = cbind(raca_sites$site_utme,
                    raca_sites$site_utmn)

v_utm = vect(raca_points, crs="+proj=utm +zone=10 +datum=WGS84  +units=m")
v_ll = project(v_utm, "+proj=longlat +datum=WGS84")
v_lonlat = geom(v_ll)[, c("x", "y")]

raca_sites_ll = raca_sites %>%
  bind_cols(v_lonlat) %>%
  dplyr::rename(site_longitude = x,
         site_latitude = y) %>%
  select(site,
         site_latitude,
         site_longitude)

raca_population_valid = raca_clean %>%
  group_by(population) %>%
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop")

raca_valid = raca_clean %>%
  left_join(raca_sites_ll, by = "site") %>%
  filter(population %in% raca_population_valid$population) %>%
  select(taxon_capture,
         date,
         site,
         population,
         bd_its_copies,
         life_stage_simple,
         life_stage,
         sex,
         svl_mm,
         site_latitude,
         site_longitude,
         region,
         country,
         project)

```

### RALU
```{r}
ralu_raw = import_list(here("data", "Data Russell et al 2010 DAO.xlsx"))

ralu_sites = ralu_raw[[2]] %>%
  clean_names() %>%
  mutate(site = tolower(pond_name))

ralu_points = cbind(ralu_sites$x_nad83_zone11n,
                    ralu_sites$y_nad83_zone11n)

w_utm = vect(ralu_points, crs="+proj=utm +zone=11 +datum=WGS84  +units=m")
w_ll = project(w_utm, "+proj=longlat +datum=WGS84")
w_lonlat = geom(w_ll)[, c("x", "y")]

ralu_sites_ll = ralu_sites %>%
  bind_cols(w_lonlat) %>%
  dplyr::rename(site_longitude = x,
         site_latitude = y) %>%
  select(site,
         site_latitude,
         site_longitude)


ralu_clean = ralu_raw[[4]] %>%
  clean_names() %>%
  dplyr::rename(life_stage = age_class) %>%
  mutate(taxon_capture = "rana_luteiventris",
         date = as.Date(bd_sample_date),
         site = tolower(pond_name),
         population = site,
         region = "NW",
         country = "usa",
         life_stage_simple = recode_values(life_stage,
                                           "ADULT" ~ "adult",
                                           "JUVENILE" ~ "subadult",
                                           "METAMORPH" ~ "metamorph"),
         bd_its_copies = actual_bd_load * 60,
         sex = tolower(sex),
         project = "RALU",
         svl_mm = 25  # using smallest age glass SVL, as researchers used standardized swab lengths.
         ) %>%
  filter(!is.na(life_stage_simple),
         !is.na(bd_its_copies),
         life_stage_simple %in% c("adult", "subadult"))

ralu_population_valid = ralu_clean %>%
  group_by(population) %>%
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop")

ralu_valid = ralu_clean %>%
  left_join(ralu_sites_ll, by = "site") %>%
  select(taxon_capture,
         date,
         site,
         population,
         bd_its_copies,
         life_stage_simple,
         life_stage,
         sex,
         svl_mm,
         site_latitude,
         site_longitude,
         region,
         country,
         project)
```

### USGS
```{r}

## laod USGS data
usgs_raw = read.csv(here("data", "USGSBsalBdDataFull.csv"))

usgs_clean = usgs_raw %>%
  clean_names() %>%
  dplyr::rename(site_latitude = latitude,
         site_longitude = longitude,
         svl_mm = body_length,
         life_stage = age,
         region = reg,
         bd_its_copies = copy_num) %>%
  mutate(taxon_capture = tolower(gsub(" ", "_", scientific_name)),
         life_stage_simple = tolower(life_stage),
         life_stage_simple = recode_values(life_stage_simple,
                                        "juvenile" ~ "subadult",
                                        default = life_stage_simple),
         date = mdy(collection_date),
         site = paste0("usgs_", site_id),
         population = site,
         project = "usgs",
         country = "usa") %>%
  filter(!is.na(taxon_capture),
         !is.na(bd_its_copies),
         !is.na(svl_mm),
         life_stage_simple %in% c("subadult", "adult"))

usgs_taxon_population_valid = usgs_clean %>%
  group_by(taxon_capture, population) %>%
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop")

usgs_taxon_valid = usgs_clean %>%
  inner_join(usgs_taxon_population_valid %>%
               select(taxon_capture,
                      population), by = c("taxon_capture", "population")) %>%
  group_by(taxon_capture) %>%
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop") %>%
  filter(n_adult >= 14,
         n_subadult >= 14,
         taxon_capture != "lithobates_chiricahuensis")

usgs_valid = usgs_clean %>%
  inner_join(usgs_taxon_population_valid %>%
               select(taxon_capture,
                      population), by = c("taxon_capture", "population")) %>%
  filter(taxon_capture %in% usgs_taxon_valid$taxon_capture) %>%
  select(taxon_capture,
         svl_mm,
         sex,
         life_stage_simple,
         life_stage,
         date,
         bd_its_copies,
         population,
         site,
         site_latitude,
         site_longitude,
         region,
         country,
         project)

usgs_spp = usgs_valid %>%
  group_by(taxon_capture,
           life_stage_simple) %>%
  count()

```

### RIBBiTR, etc.
```{r}

## SN populations
sn_populations = read.csv(here("data", "sn_site_pop_2_2025-10-31.csv")) %>%
  dplyr::rename(population = pop_id) %>%
  mutate(population_sn = if_else(!is.na(population), paste0(as.character(population), "_pop"), site),
         site = as.character(site)) %>%
  select(site,
         population_sn)


population_valid_pa = c("altos_de_campana",
                        "caribbean",
                        "el_cope",
                        "el_valle",
                        "santa_fe")

# RIBBiTR data

## Connect to DB
dbcon = hopToDB("ribbitr", hopReg = FALSE)

# table pointers
db_bd = tbl(dbcon, Id("survey_data", "bd_qpcr_results"))
db_sample = tbl(dbcon, Id("survey_data", "sample"))
db_capture = tbl(dbcon, Id("survey_data", "capture"))
db_survey = tbl(dbcon, Id("survey_data", "survey"))
db_visit = tbl(dbcon, Id("survey_data", "visit"))
db_project = tbl(dbcon, Id("survey_data", "project"))
db_site = tbl(dbcon, Id("survey_data", "site"))
db_region = tbl(dbcon, Id("survey_data", "region"))
db_country = tbl(dbcon, Id("survey_data", "country"))


# preliminary data
bd_sample = db_sample %>%
  right_join(db_capture, by = "capture_id") %>%
  left_join(db_survey, by = "survey_id") %>%
  left_join(db_visit, by = "visit_id") %>%
  left_join(db_site, by = "site_id") %>%
  left_join(db_region, by = "region_id") %>%
  left_join(db_country, by = "country_id") %>%
  left_join(db_project, by = "project_id") %>%
  filter(sample_type == "bd",
         !is.na(taxon_capture)) %>%
  inner_join(db_bd %>%
              filter(!is.na(bd_detected)) %>%
               # limit to one qPCR result per sample
              group_by(sample_id) %>%
              slice_sample(n = 1) %>%
              ungroup() %>%
              select(sample_id,
                     bd_detected,
                     bd_its1_copies_per_swab), by = "sample_id")

peace = bd_sample %>%
  group_by(taxon_capture) %>%
  count() %>%
  collect()

# define populations
bd_population = bd_sample %>%
  left_join(sn_populations, by = "site", copy = TRUE) %>%
  mutate(population = case_when(
    country %in% c("brazil", "panama") ~ region,
    region == "california" ~ population_sn,
    .default = site
  )) %>%
  filter(country != "panama" | (country == "panama" & population %in% population_valid_pa),
         !is.na(site_latitude))

# define life stage simple
bd_life_stage = bd_population %>%
  filter(life_stage %in% c("adult",
                           "subadult",
                           "juvenile",
                           NA)) %>%
  mutate(life_stage_simple = case_when(
    # us
    # https://animaldiversity.org/accounts/Acris_blanchardi/
    ((taxon_capture == "acris_blanchardi") & (svl_mm >= 16)) ~ "adult",
    ((taxon_capture == "acris_blanchardi") & (svl_mm < 16)) ~ "subadult",
    # Wright and Wright 1949
    ((taxon_capture == "anaxyrus_americanus") & (svl_mm >= 54)) ~ "adult",
    ((taxon_capture == "anaxyrus_americanus") & (svl_mm < 54)) ~ "subadult",
    # Lannoo 2005 (Amphibian Declines: The Conservation Status of United States Species)
    ((taxon_capture == "anaxyrus_canorus") & (svl_mm >= 48)) ~ "adult",
    ((taxon_capture == "anaxyrus_canorus") & (svl_mm < 48)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "anaxyrus_debilis") & (svl_mm >= 37)) ~ "adult",
    ((taxon_capture == "anaxyrus_debilis") & (svl_mm < 37)) ~ "subadult",
    # Conant and Collins 1991
    ((taxon_capture == "anaxyrus_fowleri") & (svl_mm >= 51)) ~ "adult",
    ((taxon_capture == "anaxyrus_fowleri") & (svl_mm < 51)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "hyla_chrysoscelis") & (svl_mm >= 32)) ~ "adult",
    ((taxon_capture == "hyla_chrysoscelis") & (svl_mm < 32)) ~ "subadult",
    # Conant & Collins 1998
    ((taxon_capture == "hyla_cinerea") & (svl_mm >= 32)) ~ "adult",
    ((taxon_capture == "hyla_cinerea") & (svl_mm < 32)) ~ "subadult",
    # Johnson 1987
    ((taxon_capture == "hyla_versicolor") & (svl_mm >= 32)) ~ "adult",
    ((taxon_capture == "hyla_versicolor") & (svl_mm < 32)) ~ "subadult",
    # Minton, 2001 / Lanoo 2005
    ((taxon_capture == "lithobates_blairi") & (svl_mm >= 51)) ~ "adult",
    ((taxon_capture == "lithobates_blairi") & (svl_mm < 51)) ~ "subadult",
    # Platz and Mecham 1963
    ((taxon_capture == "lithobates_chiricahuensis") & (svl_mm >= 57)) ~ "adult",
    ((taxon_capture == "lithobates_chiricahuensis") & (svl_mm < 57)) ~ "subadult",
    # Minton 2001
    ((taxon_capture == "lithobates_palustris") & (svl_mm >= 43)) ~ "adult",
    ((taxon_capture == "lithobates_palustris") & (svl_mm < 43)) ~ "subadult",
    # Conant and Collins, 1991
    ((taxon_capture == "lithobates_sphenocephalus") & (svl_mm >= 51)) ~ "adult",
    ((taxon_capture == "lithobates_sphenocephalus") & (svl_mm < 51)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "pseudacris_crucifer") & (svl_mm >= 20)) ~ "adult",
    ((taxon_capture == "pseudacris_crucifer") & (svl_mm < 20)) ~ "subadult",
    # Hulse, C. and McCoy C. J. and Ellen Censky ,1998.
    ((taxon_capture == "pseudacris_feriarum") & (svl_mm >= 19)) ~ "adult",
    ((taxon_capture == "pseudacris_feriarum") & (svl_mm < 19)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "pseudacris_regilla") & (svl_mm >= 25)) ~ "adult",
    ((taxon_capture == "pseudacris_regilla") & (svl_mm < 25)) ~ "subadult",
    # Howard 1981
    ((taxon_capture == "rana_catesbeiana") & (svl_mm >= 95)) ~ "adult",
    ((taxon_capture == "rana_catesbeiana") & (svl_mm < 95)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "rana_clamitans") & (svl_mm >= 60)) ~ "adult",
    ((taxon_capture == "rana_clamitans") & (svl_mm < 60)) ~ "subadult",
    # Knapp
    ((taxon_capture == "rana_muscosa") & is.na(svl_mm)) ~ life_stage,
    ((taxon_capture == "rana_muscosa") & (svl_mm >= 40)) ~ "adult",
    ((taxon_capture == "rana_muscosa") & (svl_mm < 40)) ~ "subadult",
    # https://www.nrc.gov/docs/ML1428/ML14282A848.pdf
    ((taxon_capture == "rana_pipiens") & (svl_mm >= 51)) ~ "adult",
    ((taxon_capture == "rana_pipiens") & (svl_mm < 51)) ~ "subadult",
    # Knapp
    ((taxon_capture == "rana_sierrae") & is.na(svl_mm)) ~ life_stage,
    ((taxon_capture == "rana_sierrae") & (svl_mm >= 40)) ~ "adult",
    ((taxon_capture == "rana_sierrae") & (svl_mm < 40)) ~ "subadult",
    # Martof 1963
    ((taxon_capture == "rana_sylvatica") & (svl_mm >= 37)) ~ "adult",
    ((taxon_capture == "rana_sylvatica") & (svl_mm < 37)) ~ "subadult",
    # Conant and Collins, 1991
    ((taxon_capture == "scaphiopus_holbrookii") & (svl_mm >= 44)) ~ "adult",
    ((taxon_capture == "scaphiopus_holbrookii") & (svl_mm < 44)) ~ "subadult",
    # Stebbins, 1985
    ((taxon_capture == "spea_multiplicata") & (svl_mm >= 37)) ~ "adult",
    ((taxon_capture == "spea_multiplicata") & (svl_mm < 37)) ~ "subadult",
    
    # brasil
    # Heyer, W. R., and R. B. Cocroft. 1986
    ((taxon_capture == "hylodes_phyllodes") & (svl_mm >= 27.5)) ~ "adult",
    ((taxon_capture == "hylodes_phyllodes") & (svl_mm < 27.5)) ~ "subadult",
    # Kwet, A., and M. Solé. 2005 (21)
    ((taxon_capture == "ischnocnema_henselii") & (svl_mm >= 23.0)) ~ "adult",
    ((taxon_capture == "ischnocnema_henselii") & (svl_mm < 23.0)) ~ "subadult",
    # Kwet, A., and M. Solé. 2005
    ((taxon_capture == "brachycephalus_pitanga") & (svl_mm >= 10.8)) ~ "adult",
    ((taxon_capture == "brachycephalus_pitanga") & (svl_mm <10.8)) ~ "subadult",
    
    ((taxon_capture == "boana_faber") & (svl_mm >= 70)) ~ "adult",
    ((taxon_capture == "boana_faber") & (svl_mm < 70)) ~ "subadult",
    # country == "brazil" & life_stage == "adult" ~ "adult",
    # country == "brazil" & life_stage %in% c("subadult", "juvenile") ~ "subadult",
    
    # panama
    # Savage 1972 & Crump 1988
    ((taxon_capture == "atelopus_varius") & (svl_mm >= 25)) ~ "adult",
    ((taxon_capture == "atelopus_varius") & (svl_mm < 25)) ~ "subadult",
    # Lynch and Myers 1983
    ((taxon_capture == "craugastor_crassidigitus") & (svl_mm >= 20.2)) ~ "adult",
    ((taxon_capture == "craugastor_crassidigitus") & (svl_mm < 20.2)) ~ "subadult",
    # Lynch and Myers 1983
    ((taxon_capture == "craugastor_fitzingeri") & (svl_mm >= 23.5)) ~ "adult",
    ((taxon_capture == "craugastor_fitzingeri") & (svl_mm < 23.5)) ~ "subadult",
    # Amphibiaweb
    ((taxon_capture == "colostethus_panamansis") & (svl_mm >= 18.8)) ~ "adult",
    ((taxon_capture == "colostethus_panamansis") & (svl_mm < 18.8)) ~ "subadult",
    # Leenders 2001
    ((taxon_capture == "dendrobates_auratus") & (svl_mm >= 25)) ~ "adult",
    ((taxon_capture == "dendrobates_auratus") & (svl_mm < 25)) ~ "subadult",
    # Amphibiaweb
    ((taxon_capture == "espadarana_prosoblepon") & (svl_mm >= 21)) ~ "adult",
    ((taxon_capture == "espadarana_prosoblepon") & (svl_mm < 21)) ~ "subadult",
    # Basante 2025
    ((taxon_capture == "lithobates_warszewitschii") & (svl_mm >= 35)) ~ "adult",
    ((taxon_capture == "lithobates_warszewitschii") & (svl_mm < 35)) ~ "subadult",
    # Wang 2008
    ((taxon_capture == "pristimantis_ridens") & (svl_mm >= 16)) ~ "adult",
    ((taxon_capture == "pristimantis_ridens") & (svl_mm < 16)) ~ "subadult",
    # Savage 2002
    ((taxon_capture == "rhaebo_haematiticus") & (svl_mm >= 42)) ~ "adult",
    ((taxon_capture == "rhaebo_haematiticus") & (svl_mm < 42)) ~ "subadult",
    # Easteal 1963
    ((taxon_capture == "rhinella_marina") & (svl_mm >= 85)) ~ "adult",
    ((taxon_capture == "rhinella_marina") & (svl_mm < 85)) ~ "subadult",
    # Savage 2002
    ((taxon_capture == "sachatamia_albomaculata") & (svl_mm >= 20.5)) ~ "adult",
    ((taxon_capture == "sachatamia_albomaculata") & (svl_mm < 20.5)) ~ "subadult",
    # Ibunez & Smith 1995
    ((taxon_capture == "silverstoneia_flotator") & (svl_mm >= 15.1)) ~ "adult",
    ((taxon_capture == "silverstoneia_flotator") & (svl_mm < 15.1)) ~ "subadult",
    # Grant & Myers 2013
    ((taxon_capture == "silverstoneia_nubicola") & (svl_mm >= 15.4)) ~ "adult",
    ((taxon_capture == "silverstoneia_nubicola") & (svl_mm < 15.4)) ~ "subadult",
    # Duellman, W. E. 2001
    ((taxon_capture == "smilisca_sila") & (svl_mm >= 31.6)) ~ "adult",
    ((taxon_capture == "smilisca_sila") & (svl_mm < 31.6)) ~ "subadult",
    .default = NA_character_
  )) %>%
  filter(!is.na(life_stage_simple),
         !is.na(svl_mm))

# identify valid species & populations
## drop pops without both adults and juveniles
## drop species without 14 or more adults and juveniles each
pop_taxa_valid = bd_life_stage %>%
  group_by(taxon_capture, population) %>%
  summarize(n_adult = sum(as.integer(life_stage_simple == "adult")),
            n_subadult = sum(as.integer(life_stage_simple == "subadult")),
            .groups = "drop")
# %>%
#   filter(n_adult > 0,
#          n_subadult > 0)

taxa_valid = bd_life_stage %>%
  inner_join(pop_taxa_valid %>%
               select(taxon_capture,
                      population), by = c("taxon_capture", "population")) %>%
  group_by(taxon_capture) %>%
  summarize(n_adult = sum(as.integer(life_stage_simple == "adult"), na.rm = TRUE),
            n_subadult = sum(as.integer(life_stage_simple == "subadult"), na.rm = TRUE),
            .groups = "drop") %>%
  filter(n_adult >= 14,
         n_subadult >= 14)

# identify valid samples
bd_sample_valid = bd_life_stage %>%
  mutate(year = year(date),
         project = project_id) %>%
  inner_join(pop_taxa_valid, by = c("taxon_capture", "population")) %>%
  inner_join(taxa_valid, by = "taxon_capture") %>%
  # limit to 1 bd sample per capture
  group_by(capture_id) %>%
  slice_sample(n = 1) %>%
  ungroup() %>%
  select(sample_id,
         bd_detected,
         bd_its1_copies_per_swab,
         sample_name,
         capture_id,
         taxon_capture,
         life_stage,
         life_stage_simple,
         svl_mm,
         body_mass_g,
         sex,
         date,
         site,
         site_latitude,
         site_longitude,
         site_elevation_m,
         population,
         region,
         country,
         project) %>%
  collect()

# final data selection and cleaning for model
ribbitr_valid = bd_sample_valid %>%
  dplyr::rename(bd_its_copies = bd_its1_copies_per_swab) %>%
  select(bd_its_copies,
         taxon_capture,
         life_stage,
         life_stage_simple,
         svl_mm,
         body_mass_g,
         sex,
         date,
         site,
         site_latitude,
         site_longitude,
         site_elevation_m,
         population,
         region,
         country,
         project)

```

### HYPHY
```{r}
### Hylodes phyllodes
hyphy_raw = import_list(here("data", "CDMR_H.phyllodes_metadata_Oct2_LKS.xlsx"))[[1]]

hyphy_clean = hyphy_raw %>%
  clean_names() %>%
  filter(timepoint == 0) %>%
  # filter(timepoint == 0 | treatment == "Saline control") %>%
  mutate(
    date = as.Date(capture_date),
    svl_mm = as.numeric(svl),
    life_stage_simple = case_when(
    svl_mm < 27.5 ~ "subadult",
    svl_mm >= 27.5 ~ "adult",
    .default = NA),
    taxon_capture = "hylodes_phyllodes",
    site = gsub("aw", "s", tolower(transect)),
    population = region,
    country = "brazil",
    bd_its_copies = bd_load,
    project = "HYPHY"
  ) %>%
  filter(!is.na(life_stage_simple),
         !is.na(bd_its_copies),
         life_stage_simple %in% c("adult", "subadult"))

hyphy_population_valid = hyphy_clean %>%
  group_by(population) %>%
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop")

hyphy_sites = db_site %>%
  filter(site %in% hyphy_clean$site) %>%
  select(site,
         site_latitude,
         site_longitude,
         site_elevation_m) %>%
  collect()

hyphy_valid = hyphy_clean %>%
  left_join(hyphy_sites, by = "site") %>%
  select(taxon_capture,
         date,
         site,
         site_latitude,
         site_longitude,
         population,
         bd_its_copies,
         life_stage_simple,
         svl_mm,
         region,
         country,
         project)
```

### BOFA
```{r}
### Boana faber
bofa_raw = read_tsv(here("data", "Boana_faber_capture_data_for_Cob.csv"))

bofa_clean = bofa_raw %>%
  clean_names() %>%
  mutate(
    date = as.Date(capture_date, format = "%d/%m/%Y"),
    svl_mm = as.numeric(svl),
    life_stage_simple = case_when(
      svl_mm < 63 ~ "subadult",
      svl_mm >= 63 ~ "adult",
      .default = NA),
    bd_its_copies = coalesce(bd_loads, as.numeric(bd_q1), as.numeric(bd_q2)),
    bd_its_copies = if_else(is.na(bd_its_copies), 0, bd_its_copies),
    taxon_capture = "boana_faber",
    site = tolower(gsub(" ", "_", capture_site)),
    site = if_else(grepl("o_paulo$", site), "joao_paulo", site),
    population = case_when(
      site == "lago_anta" ~ "santa_virginia",
      .default = site),
    region = case_when(
      site == "lago_anta" ~ "santa_virginia",
      .default = "sao_luis_paraitinga"),
    country = "brazil",
    project = "BOFA"
  ) %>%
  filter(!is.na(life_stage_simple),
         !is.na(bd_its_copies),
         life_stage_simple %in% c("adult", "subadult"),
         site != "roberto")

bofa_sites = bofa_clean %>%
  group_by(site) %>%
  summarize(site_latitude = mean(lat, na.rm = TRUE),
            site_longitude = mean(long, na.rm = TRUE)) %>%
  # manually correct lat/lon for lago_anta, to align with other datasets
  mutate(site_latitude = if_else(site == "lago_anta", -23.38511, site_latitude),
         site_longitude = if_else(site == "lago_anta", -45.18707, site_longitude))

bofa_valid = bofa_clean %>%
  left_join(bofa_sites, by = "site") %>%
  select(taxon_capture,
         date,
         site,
         site_latitude,
         site_longitude,
         population,
         bd_its_copies,
         life_stage_simple,
         svl_mm,
         region,
         country,
         project)

```

### Combine datasets
```{r}

column_check = compare_df_cols(raca_valid,
                               ralu_valid,
                               usgs_valid,
                               ribbitr_valid,
                               hyphy_valid)

bd_model = bind_rows(raca_valid,
                     ralu_valid,
                     usgs_valid,
                     ribbitr_valid,
                     hyphy_valid) %>%
  mutate(year = year(date),
         bd_load_co = if_else(bd_its_copies < 10,
                              0,
                              bd_its_copies),
         bd_detected_co = bd_load_co > 0,
         bd_load_co_density = bd_load_co / svl_mm)

# checks
sum(is.na(bd_model$bd_load_co_density))
sum(is.na(bd_model$taxon_capture))
sum(is.na(bd_model$population))

(bd_model_stats = bd_model %>%
  group_by(taxon_capture, country) %>%
  summarise(regions = paste(sort(unique(region)), collapse = ", "),
            population_n = n_distinct(population),
            site_n = n_distinct(site),
            year_first = year(min(date)),
            year_last = year(max(date)),
            year_count = n_distinct(year(date)),
            adult_n = sum(life_stage_simple == "adult"),
            subadult_n = sum(life_stage_simple == "subadult"),
            .groups = "drop") %>%
  arrange(country,
          taxon_capture))
```
## Environmental data
To support comparing Bd prevalence and loads between sites and species across a range of environmental conditions, I incorporated local (~5 km) daily maximum temperature (CHIRTS) and daily precipitation (CHIRPS) time series. Following Sonn et al. 2019 [sonn2019effects], I created 15-day and 30-day means (for tempurature) and sums (for precipitation). Preliminary model fitting showed stronger relationships between Bd prevalence and 15-day lag products, while Bd load showed stronger relationships with 30-day lag products, similar to Sonn et al. 2019.


### Join with env. data
```{r}
#| eval: false
#| warning: false
#| code-fold: true
#| code-summary: "Show environmental data mergin code"

env_agg = read_csv(here("data", "unified_env_agg.csv"))


bd_model_env = bd_model %>%
  left_join(env_agg, by = c("site", "date"))

brms_data = bd_model_env %>%
  mutate(
    precip_mm_10_z = as.numeric(scale(precip_mm_10_day_sum, center = TRUE, scale = TRUE)),
    precip_mm_15_z = as.numeric(scale(precip_mm_15_day_sum, center = TRUE, scale = TRUE)),
    precip_mm_30_z = as.numeric(scale(precip_mm_30_day_sum, center = TRUE, scale = TRUE)),
    temp_c_10_z = as.numeric(scale(temp_c_10_day_mean, center = TRUE, scale = TRUE)),
    temp_c_15_z = as.numeric(scale(temp_c_15_day_mean, center = TRUE, scale = TRUE)),
    temp_c_30_z = as.numeric(scale(temp_c_30_day_mean, center = TRUE, scale = TRUE))) %>%
  select(taxon_capture,
         everything())

```
# save data
```{r}
save(brms_data, file = here("size_bd_unified", "intermediates", paste0("bd_model_env_unified_", today(), ".RData")))
write_csv(bd_model_stats, here("size_bd_unified", "intermediates", "size_bd_modeling_unified_stats.csv"))

```
