---
title: "Size vs. Bd modeling: Panamá preliminary results"
author: Cob Staines
date: 2025-09-29
format:
  html:
    toc: true
df-print: kable
bibliography: references.bib
csl: ecology.csl
---
# Introduction

## Motivation

Patterns of greater Bd load in recently metamorphosed individuals have been documented in some cases[@adams_chytridiomycosis-induced_2020] [@humphries_immune_2022] [@humphries_chytridiomycosis_2024] [@humphries_chytridiomycosis_2025] (and the opposite pattern in others[@bradley_host_2019]), with possible implications for:

  - resilience trajectory assessment (Q1)
  - population dynamics (Q2)
  - assessing Bd resistance/tolerance mechanisms (Q3)
  
... with the goal of complementing ongoing RIBBiTR research efforts

## Working hypotheses:

  a. Recently-metamorphosed individuals are more likely to be infected with Bd than older individuals across species
  b. Recently-metamorphosed individuals are more likely to carry higher Bd loads than older individuals across species

# Methods
## Setup

```{r}
#| eval: true
#| output: false
#| code-fold: true
#| code-summary: "Show setup code"

# setup
librarian::shelf(tidyverse, dbplyr, here, janitor, RPostgres, DBI, RIBBiTR-BII/ribbitrrr, brms, tidybayes, bayesplot, cmdstanr, bayestestR, rio)

```
## Data selection
### Classifying by life stage
These analyses were carried out for post-metamorphic individuals only. Individuals were classified as "adult" or "subadult" with a threshold snout-vent-length (SVL [mm]), identified independently for each species from the literature (see data collection code for specific thresholds used and their sources).
  
### Classification by population
"Populations" were defined at the region level within Panamá. In the Sierra Nevada, sites were grouped where evidence supported connectivity at he scale of individuals. Elsewhere in North America, populations were defined at the site level.

### Species selection criteria:
Species were included if supported by least 14 swabbed adults and subadults (each).

### Bd density and Zero-cutoff
I scaled all Bd quantities by the Snout-Vent-Length (mm) to arrive at a Bd density per linear distance of skin swabbed (within species), to correct for linear effects of larger swabbing area in larger frogs. I also set any Bd ITS copy numbers < 10 to zero, to represent uncertainty in qPCR results and support consistency in zero-inflation modeling.

### RACA
```{r}

# csv data
##  RACA data
raca_clean = read.csv(here("data", "JPSL_RACA_Swabbed_Individuals_2026-03-03.csv")) %>%
  clean_names() %>%
  filter(life_stage %in% c("A", "S", "Y")) %>%
  dplyr::rename(site = site_name,
         body_mass_g = weight_g,
         site_utme = utme_site,
         site_utmn = utmn_site,
         bd_its_copies = bd_swab_results_its) %>%
  mutate(life_stage_simple = case_when(
    svl_mm >= 50 ~ "adult",
    svl_mm < 50 ~ "subadult"),
    date = as.Date(date),
    site = tolower(gsub(" ", "_", site)),
    site_utm_zone = 11,  # verify
    population = site,
    region = "W",
    country = "usa",
    project = "RACA",
    taxon_capture = "rana_cascadae") %>%
  filter(!is.na(bd_its_copies))

raca_population_valid = raca_clean %>%
  group_by(population) %>%
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop")

raca_valid = raca_clean %>%
  filter(population %in% raca_population_valid$population) %>%
  select(taxon_capture,
         date,
         site,
         population,
         bd_its_copies,
         life_stage_simple,
         life_stage,
         sex,
         svl_mm,
         site_latitude,
         site_longitude,
         region,
         country,
         project)

```

### RALU
```{r}
ralu_raw = import_list(here("data", "Data Russell et al 2010 DAO.xlsx"))

ralu_clean = ralu_raw[[4]] %>%
  clean_names() %>%
  dplyr::rename(life_stage = age_class) %>%
  mutate(taxon_capture = "rana_luteiventris",
         date = as.Date(bd_sample_date),
         site = tolower(pond_name),
         population = site,
         region = "NW",
         country = "usa",
         life_stage_simple = recode_values(life_stage,
                                           "ADULT" ~ "adult",
                                           "JUVENILE" ~ "subadult",
                                           "METAMORPH" ~ "metamorph"),
         bd_its_copies = actual_bd_load * 60,
         sex = tolower(sex),
         project = "RALU",
         svl_mm = 25  # using smallest age glass SVL, as researchers used standardized swab lengths.
         ) %>%
  filter(!is.na(life_stage_simple),
         !is.na(bd_its_copies),
         life_stage_simple %in% c("adult", "subadult"))

ralu_population_valid = ralu_clean %>%
  group_by(population) %>%
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop")

ralu_valid = ralu_clean %>%
  select(taxon_capture,
         date,
         site,
         population,
         bd_its_copies,
         life_stage_simple,
         life_stage,
         sex,
         svl_mm,
         site_latitude,
         site_longitude,
         region,
         country,
         project)
```

### USGS
```{r}

## laod USGS data
usgs_raw = read.csv(here("data", "USGSBsalBdDataFull.csv"))

usgs_clean = usgs_raw %>%
  clean_names() %>%
  dplyr::rename(site_latitude = latitude,
         site_longitude = longitude,
         svl_mm = body_length,
         life_stage = age,
         region = reg,
         bd_its_copies = copy_num) %>%
  mutate(taxon_capture = tolower(gsub(" ", "_", scientific_name)),
         life_stage_simple = tolower(life_stage),
         life_stage_simple = recode_values(life_stage_simple,
                                        "juvenile" ~ "subadult",
                                        default = life_stage_simple),
         date = mdy(collection_date),
         site = paste0("usgs_", site_id),
         population = site,
         project = "usgs",
         country = "usa") %>%
  filter(!is.na(taxon_capture),
         !is.na(bd_its_copies),
         !is.na(svl_mm),
         life_stage_simple %in% c("subadult", "adult"))

usgs_taxon_population_valid = usgs_clean %>%
  group_by(taxon_capture, population) %>%
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop")

usgs_taxon_valid = usgs_clean %>%
  inner_join(usgs_taxon_population_valid %>%
               select(taxon_capture,
                      population), by = c("taxon_capture", "population")) %>%
  group_by(taxon_capture) %>%
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop") %>%
  filter(n_adult >= 14,
         n_subadult >= 14,
         taxon_capture != "lithobates_chiricahuensis")

usgs_valid = usgs_clean %>%
  inner_join(usgs_taxon_population_valid %>%
               select(taxon_capture,
                      population), by = c("taxon_capture", "population")) %>%
  filter(taxon_capture %in% usgs_taxon_valid$taxon_capture) %>%
  select(taxon_capture,
         svl_mm,
         sex,
         life_stage_simple,
         life_stage,
         date,
         bd_its_copies,
         population,
         site,
         site_latitude,
         site_longitude,
         region,
         country,
         project)

usgs_spp = usgs_valid %>%
  group_by(taxon_capture,
           life_stage_simple) %>%
  count()

```

### RIBBiTR, etc.
```{r}

## SN populations
sn_populations = read.csv(here("data", "sn_site_pop_2_2025-10-31.csv")) %>%
  dplyr::rename(population = pop_id) %>%
  mutate(population_sn = if_else(!is.na(population), paste0(as.character(population), "_pop"), site),
         site = as.character(site)) %>%
  select(site,
         population_sn)


population_valid_pa = c("altos_de_campana",
                        "caribbean",
                        "el_cope",
                        "el_valle",
                        "santa_fe")

# RIBBiTR data

## Connect to DB
dbcon = hopToDB("ribbitr", hopReg = FALSE)

# table pointers
db_bd = tbl(dbcon, Id("survey_data", "bd_qpcr_results"))
db_sample = tbl(dbcon, Id("survey_data", "sample"))
db_capture = tbl(dbcon, Id("survey_data", "capture"))
db_survey = tbl(dbcon, Id("survey_data", "survey"))
db_visit = tbl(dbcon, Id("survey_data", "visit"))
db_project = tbl(dbcon, Id("survey_data", "project"))
db_site = tbl(dbcon, Id("survey_data", "site"))
db_region = tbl(dbcon, Id("survey_data", "region"))
db_country = tbl(dbcon, Id("survey_data", "country"))


# preliminary data
bd_sample = db_sample %>%
  right_join(db_capture, by = "capture_id") %>%
  left_join(db_survey, by = "survey_id") %>%
  left_join(db_visit, by = "visit_id") %>%
  left_join(db_site, by = "site_id") %>%
  left_join(db_region, by = "region_id") %>%
  left_join(db_country, by = "country_id") %>%
  left_join(db_project, by = "project_id") %>%
  filter(sample_type == "bd",
         !is.na(taxon_capture)) %>%
  inner_join(db_bd %>%
              filter(!is.na(bd_detected)) %>%
               # limit to one qPCR result per sample
              group_by(sample_id) %>%
              slice_sample(n = 1) %>%
              ungroup() %>%
              select(sample_id,
                     bd_detected,
                     bd_its1_copies_per_swab), by = "sample_id")

# define populations
bd_population = bd_sample %>%
  left_join(sn_populations, by = "site", copy = TRUE) %>%
  mutate(population = case_when(
    country %in% c("brazil", "panama") ~ region,
    region == "california" ~ population_sn,
    .default = site
  )) %>%
  filter(country != "panama" | (country == "panama" & population %in% population_valid_pa),
         !is.na(site_latitude))

# define life stage simple
bd_life_stage = bd_population %>%
  filter(life_stage %in% c("adult",
                           "subadult",
                           "juvenile",
                           NA)) %>%
  mutate(life_stage_simple = case_when(
    # us
    # https://animaldiversity.org/accounts/Acris_blanchardi/
    ((taxon_capture == "acris_blanchardi") & (svl_mm >= 16)) ~ "adult",
    ((taxon_capture == "acris_blanchardi") & (svl_mm < 16)) ~ "subadult",
    # Wright and Wright 1949
    ((taxon_capture == "anaxyrus_americanus") & (svl_mm >= 54)) ~ "adult",
    ((taxon_capture == "anaxyrus_americanus") & (svl_mm < 54)) ~ "subadult",
    # Lannoo 2005 (Amphibian Declines: The Conservation Status of United States Species)
    ((taxon_capture == "anaxyrus_canorus") & (svl_mm >= 48)) ~ "adult",
    ((taxon_capture == "anaxyrus_canorus") & (svl_mm < 48)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "anaxyrus_debilis") & (svl_mm >= 37)) ~ "adult",
    ((taxon_capture == "anaxyrus_debilis") & (svl_mm < 37)) ~ "subadult",
    # Conant and Collins 1991
    ((taxon_capture == "anaxyrus_fowleri") & (svl_mm >= 51)) ~ "adult",
    ((taxon_capture == "anaxyrus_fowleri") & (svl_mm < 51)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "hyla_chrysoscelis") & (svl_mm >= 32)) ~ "adult",
    ((taxon_capture == "hyla_chrysoscelis") & (svl_mm < 32)) ~ "subadult",
    # Conant & Collins 1998
    ((taxon_capture == "hyla_cinerea") & (svl_mm >= 32)) ~ "adult",
    ((taxon_capture == "hyla_cinerea") & (svl_mm < 32)) ~ "subadult",
    # Johnson 1987
    ((taxon_capture == "hyla_versicolor") & (svl_mm >= 32)) ~ "adult",
    ((taxon_capture == "hyla_versicolor") & (svl_mm < 32)) ~ "subadult",
    # Minton, 2001 / Lanoo 2005
    ((taxon_capture == "lithobates_blairi") & (svl_mm >= 51)) ~ "adult",
    ((taxon_capture == "lithobates_blairi") & (svl_mm < 51)) ~ "subadult",
    # Platz and Mecham 1963
    ((taxon_capture == "lithobates_chiricahuensis") & (svl_mm >= 57)) ~ "adult",
    ((taxon_capture == "lithobates_chiricahuensis") & (svl_mm < 57)) ~ "subadult",
    # Minton 2001
    ((taxon_capture == "lithobates_palustris") & (svl_mm >= 43)) ~ "adult",
    ((taxon_capture == "lithobates_palustris") & (svl_mm < 43)) ~ "subadult",
    # Conant and Collins, 1991
    ((taxon_capture == "lithobates_sphenocephalus") & (svl_mm >= 51)) ~ "adult",
    ((taxon_capture == "lithobates_sphenocephalus") & (svl_mm < 51)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "pseudacris_crucifer") & (svl_mm >= 20)) ~ "adult",
    ((taxon_capture == "pseudacris_crucifer") & (svl_mm < 20)) ~ "subadult",
    # Hulse, C. and McCoy C. J. and Ellen Censky ,1998.
    ((taxon_capture == "pseudacris_feriarum") & (svl_mm >= 19)) ~ "adult",
    ((taxon_capture == "pseudacris_feriarum") & (svl_mm < 19)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "pseudacris_regilla") & (svl_mm >= 25)) ~ "adult",
    ((taxon_capture == "pseudacris_regilla") & (svl_mm < 25)) ~ "subadult",
    # Howard 1981
    ((taxon_capture == "rana_catesbeiana") & (svl_mm >= 95)) ~ "adult",
    ((taxon_capture == "rana_catesbeiana") & (svl_mm < 95)) ~ "subadult",
    # Lannoo 2005
    ((taxon_capture == "rana_clamitans") & (svl_mm >= 60)) ~ "adult",
    ((taxon_capture == "rana_clamitans") & (svl_mm < 60)) ~ "subadult",
    # Knapp
    ((taxon_capture == "rana_muscosa") & is.na(svl_mm)) ~ life_stage,
    ((taxon_capture == "rana_muscosa") & (svl_mm >= 40)) ~ "adult",
    ((taxon_capture == "rana_muscosa") & (svl_mm < 40)) ~ "subadult",
    # https://www.nrc.gov/docs/ML1428/ML14282A848.pdf
    ((taxon_capture == "rana_pipiens") & (svl_mm >= 51)) ~ "adult",
    ((taxon_capture == "rana_pipiens") & (svl_mm < 51)) ~ "subadult",
    # Knapp
    ((taxon_capture == "rana_sierrae") & is.na(svl_mm)) ~ life_stage,
    ((taxon_capture == "rana_sierrae") & (svl_mm >= 40)) ~ "adult",
    ((taxon_capture == "rana_sierrae") & (svl_mm < 40)) ~ "subadult",
    # Martof 1963
    ((taxon_capture == "rana_sylvatica") & (svl_mm >= 37)) ~ "adult",
    ((taxon_capture == "rana_sylvatica") & (svl_mm < 37)) ~ "subadult",
    # Conant and Collins, 1991
    ((taxon_capture == "scaphiopus_holbrookii") & (svl_mm >= 44)) ~ "adult",
    ((taxon_capture == "scaphiopus_holbrookii") & (svl_mm < 44)) ~ "subadult",
    # Stebbins, 1985
    ((taxon_capture == "spea_multiplicata") & (svl_mm >= 37)) ~ "adult",
    ((taxon_capture == "spea_multiplicata") & (svl_mm < 37)) ~ "subadult",
    
    # brasil
    # Heyer, W. R., and R. B. Cocroft. 1986
    ((taxon_capture == "hylodes_phyllodes") & (svl_mm >= 27.5)) ~ "adult",
    ((taxon_capture == "hylodes_phyllodes") & (svl_mm < 27.5)) ~ "subadult",
    # Kwet, A., and M. Solé. 2005
    ((taxon_capture == "ischnocnema_henselii") & (svl_mm >= 23.0)) ~ "adult",
    ((taxon_capture == "ischnocnema_henselii") & (svl_mm < 23.0)) ~ "subadult",
    # Kwet, A., and M. Solé. 2005
    ((taxon_capture == "brachycephalus_pitanga") & (svl_mm >= 10.8)) ~ "adult",
    ((taxon_capture == "brachycephalus_pitanga") & (svl_mm <10.8)) ~ "subadult",
    ((taxon_capture == "boana_faber") & (svl_mm >= 70)) ~ "adult",
    ((taxon_capture == "boana_faber") & (svl_mm < 70)) ~ "subadult",
   
    # country == "brazil" & life_stage == "adult" ~ "adult",
    # country == "brazil" & life_stage %in% c("subadult", "juvenile") ~ "subadult",
    
    # panama
    # Savage 1972 & Crump 1988
    ((taxon_capture == "atelopus_varius") & (svl_mm >= 25)) ~ "adult",
    ((taxon_capture == "atelopus_varius") & (svl_mm < 25)) ~ "subadult",
    # Lynch and Myers 1983
    ((taxon_capture == "craugastor_crassidigitus") & (svl_mm >= 20.2)) ~ "adult",
    ((taxon_capture == "craugastor_crassidigitus") & (svl_mm < 20.2)) ~ "subadult",
    # Lynch and Myers 1983
    ((taxon_capture == "craugastor_fitzingeri") & (svl_mm >= 23.5)) ~ "adult",
    ((taxon_capture == "craugastor_fitzingeri") & (svl_mm < 23.5)) ~ "subadult",
    # Amphibiaweb
    ((taxon_capture == "colostethus_panamansis") & (svl_mm >= 18.8)) ~ "adult",
    ((taxon_capture == "colostethus_panamansis") & (svl_mm < 18.8)) ~ "subadult",
    # Leenders 2001
    ((taxon_capture == "dendrobates_auratus") & (svl_mm >= 25)) ~ "adult",
    ((taxon_capture == "dendrobates_auratus") & (svl_mm < 25)) ~ "subadult",
    # Amphibiaweb
    ((taxon_capture == "espadarana_prosoblepon") & (svl_mm >= 21)) ~ "adult",
    ((taxon_capture == "espadarana_prosoblepon") & (svl_mm < 21)) ~ "subadult",
    # Basante 2025
    ((taxon_capture == "lithobates_warszewitschii") & (svl_mm >= 35)) ~ "adult",
    ((taxon_capture == "lithobates_warszewitschii") & (svl_mm < 35)) ~ "subadult",
    # Wang 2008
    ((taxon_capture == "pristimantis_ridens") & (svl_mm >= 16)) ~ "adult",
    ((taxon_capture == "pristimantis_ridens") & (svl_mm < 16)) ~ "subadult",
    # Savage 2002
    ((taxon_capture == "rhaebo_haematiticus") & (svl_mm >= 42)) ~ "adult",
    ((taxon_capture == "rhaebo_haematiticus") & (svl_mm < 42)) ~ "subadult",
    # Easteal 1963
    ((taxon_capture == "rhinella_marina") & (svl_mm >= 85)) ~ "adult",
    ((taxon_capture == "rhinella_marina") & (svl_mm < 85)) ~ "subadult",
    # Savage 2002
    ((taxon_capture == "sachatamia_albomaculata") & (svl_mm >= 20.5)) ~ "adult",
    ((taxon_capture == "sachatamia_albomaculata") & (svl_mm < 20.5)) ~ "subadult",
    # Ibunez & Smith 1995
    ((taxon_capture == "silverstoneia_flotator") & (svl_mm >= 15.1)) ~ "adult",
    ((taxon_capture == "silverstoneia_flotator") & (svl_mm < 15.1)) ~ "subadult",
    # Grant & Myers 2013
    ((taxon_capture == "silverstoneia_nubicola") & (svl_mm >= 15.4)) ~ "adult",
    ((taxon_capture == "silverstoneia_nubicola") & (svl_mm < 15.4)) ~ "subadult",
    # Duellman, W. E. 2001
    ((taxon_capture == "smilisca_sila") & (svl_mm >= 31.6)) ~ "adult",
    ((taxon_capture == "smilisca_sila") & (svl_mm < 31.6)) ~ "subadult",
    .default = NA_character_
  )) %>%
  filter(!is.na(life_stage_simple),
         !is.na(svl_mm))

# identify valid species & populations
## drop pops without both adults and juveniles
## drop species without 14 or more adults and juveniles each
pop_taxa_valid = bd_life_stage %>%
  group_by(taxon_capture, population) %>%
  summarize(n_adult = sum(as.integer(life_stage_simple == "adult")),
            n_subadult = sum(as.integer(life_stage_simple == "subadult")),
            .groups = "drop")
# %>%
#   filter(n_adult > 0,
#          n_subadult > 0)

taxa_valid = bd_life_stage %>%
  inner_join(pop_taxa_valid %>%
               select(taxon_capture,
                      population), by = c("taxon_capture", "population")) %>%
  group_by(taxon_capture) %>%
  summarize(n_adult = sum(as.integer(life_stage_simple == "adult"), na.rm = TRUE),
            n_subadult = sum(as.integer(life_stage_simple == "subadult"), na.rm = TRUE),
            .groups = "drop") %>%
  filter(n_adult >= 14,
         n_subadult >= 14)

# identify valid samples
bd_sample_valid = bd_life_stage %>%
  mutate(year = year(date),
         project = project_id) %>%
  inner_join(pop_taxa_valid, by = c("taxon_capture", "population")) %>%
  inner_join(taxa_valid, by = "taxon_capture") %>%
  # limit to 1 bd sample per capture
  group_by(capture_id) %>%
  slice_sample(n = 1) %>%
  ungroup() %>%
  select(sample_id,
         bd_detected,
         bd_its1_copies_per_swab,
         sample_name,
         capture_id,
         taxon_capture,
         life_stage,
         life_stage_simple,
         svl_mm,
         body_mass_g,
         sex,
         date,
         year,
         site,
         site_latitude,
         site_longitude,
         population,
         region,
         country,
         project) %>%
  collect()

# final data selection and cleaning for model
ribbitr_valid = bd_sample_valid %>%
  dplyr::rename(bd_its_copies = bd_its1_copies_per_swab) %>%
  select(bd_its_copies,
         taxon_capture,
         life_stage,
         life_stage_simple,
         svl_mm,
         body_mass_g,
         sex,
         date,
         year,
         site,
         site_latitude,
         site_longitude,
         population,
         region,
         country,
         project)

```

### HYPHY
```{r}
### Hylodes phyllodes
hyphy_raw = import_list(here("data", "CDMR_H.phyllodes_metadata_Oct2_LKS.xlsx"))

hyphy_clean = hyphy_raw[[1]] %>%
  clean_names() %>%
  filter(timepoint == 0) %>%
  # filter(timepoint == 0 | treatment == "Saline control") %>%
  dplyr::rename(svl_mm = svl,
                date = capture_date) %>%
  mutate(life_stage_simple = case_when(
    svl_mm < 27.5 ~ "subadult",
    svl_mm >= 27.5 ~ "adult",
    .default = NA),
    taxon_capture = "hylodes_phyllodes",
    site = gsub("aw", "s", tolower(transect)),
    population = region,
    country = "brazil",
    bd_its_copies = bd_load,
    project = "HYPHY"
  ) %>%
  filter(!is.na(life_stage_simple),
         !is.na(bd_its_copies),
         life_stage_simple %in% c("adult", "subadult"))

hyphy_population_valid = hyphy_clean %>%
  group_by(population) %>%
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop")

peace = hyphy_clean %>%
  select(site) %>%
  distinct()

hope = hyphy_clean %>%
  summarize(date_min = min(date),
            date_max = max(date))

hyphy_valid = hyphy_clean %>%
  select(taxon_capture,
         date,
         site,
         population,
         bd_its_copies,
         life_stage_simple,
         svl_mm,
         region,
         country,
         project)
```

### Combine datasets
```{r}

column_check = compare_df_cols(raca_valid,
                               ralu_valid,
                               usgs_valid,
                               ribbitr_valid)

bd_model = bind_rows(raca_valid,
                     ralu_valid,
                     usgs_valid,
                     ribbitr_valid) %>%
  mutate(bd_load_co = if_else(bd_its_copies < 10,
                              0,
                              bd_its_copies),
         bd_detected_co = bd_load_co > 0,
         bd_load_co_density = bd_load_co / svl_mm)

(bd_model_stats = bd_model %>%
  group_by(taxon_capture, country) %>%
  summarise(regions = paste(sort(unique(region)), collapse = ", "),
            population_n = n_distinct(population),
            site_n = n_distinct(site),
            year_first = year(min(date)),
            year_last = year(max(date)),
            year_count = n_distinct(year(date)),
            adult_n = sum(life_stage_simple == "adult"),
            subadult_n = sum(life_stage_simple == "subadult"),
            .groups = "drop") %>%
  arrange(country,
          taxon_capture))

# write_csv(bd_model_stats, here("data", "size_bd_modeling_unified_stats.csv"))

```

## Environmental data
To support comparing Bd prevalence and loads between sites and species across a range of environmental conditions, I incorporated local (~5 km) daily maximum temperature (CHIRTS) and daily precipitation (CHIRPS) time series. Following Sonn et al. 2019 [sonn2019effects], I created 15-day and 30-day means (for tempurature) and sums (for precipitation). Preliminary model fitting showed stronger relationships between Bd prevalence and 15-day lag products, while Bd load showed stronger relationships with 30-day lag products, similar to Sonn et al. 2019.


### Join with env. data
```{r}
#| eval: false
#| warning: false
#| code-fold: true
#| code-summary: "Show environmental data mergin code"

env_agg = read_csv(here("data", "unified_env_agg.csv"))


bd_model_env = bd_model %>%
  left_join(env_agg, by = c("site", "date"))

brms_data = bd_model_env %>%
  mutate(
    precip_mm_30_z = as.numeric(scale(precip_mm_30_day_sum, center = TRUE, scale = TRUE)),
    temp_c_30_z = as.numeric(scale(temp_c_30_day_mean, center = TRUE, scale = TRUE)),
    precip_mm_15_z = as.numeric(scale(precip_mm_15_day_sum, center = TRUE, scale = TRUE)),
    temp_c_15_z = as.numeric(scale(temp_c_15_day_mean, center = TRUE, scale = TRUE)),
    precip_mm_30_z2 = as.numeric(scale(precip_mm_30_day_sum ^ 2, center = TRUE, scale = TRUE)),
    temp_c_30_z2 = as.numeric(scale(temp_c_30_day_mean ^ 2, center = TRUE, scale = TRUE)),
    precip_mm_15_z2 = as.numeric(scale(precip_mm_15_day_sum ^ 2, center = TRUE, scale = TRUE)),
    temp_c_15_z2 = as.numeric(scale(temp_c_15_day_mean ^ 2, center = TRUE, scale = TRUE))) %>%
  select(taxon_capture,
         everything())

save(brms_data, file = here("data", paste0("bd_model_env_unified_", today(), ".RData")))

```

## Load saved data
```{r}
#| eval: true
#| warning: false
#| echo: false
#| code-fold: true
#| code-summary: "Show data loading code"

# load bd_model_env
load(here("data", "bd_model_env_unified_2026-03-22.RData")) 

# checks
sum(is.na(brms_data$bd_load_co_density))
sum(is.na(brms_data$precip_mm_30_day_sum))
sum(is.na(brms_data$temperature_max_c))

bd_model_env_stats = brms_data %>%
  group_by(taxon_capture, country) %>%
  summarise(regions = paste(sort(unique(region)), collapse = ", "),
            population_n = n_distinct(population),
            site_n = n_distinct(site),
            year_first = year(min(date)),
            year_last = year(max(date)),
            year_count = n_distinct(year(date)),
            adult_n = sum(life_stage_simple == "adult"),
            adult_n_pos = sum(life_stage_simple == "adult" & bd_detected_co),
            subadult_n = sum(life_stage_simple == "subadult"),
            subadult_n_pos = sum(life_stage_simple == "subadult" & bd_detected_co),
            .groups = "drop") %>%
  arrange(country,
          taxon_capture)
```

# Modeling
I modeled Bd load using a Baysian zero-inflated (or "hurdle") log-normal model [@schrock2025fungi] with fixed effects of species, life stage, and recent precipitation and temperature, and considering population, species, life stage and recent precipitation and temperature as a random effects. The "hurdle" infection probability (hu) and the variability in Bd load (sigma) are allowed to vary with life stage.

## Model setup
```{r}
#| code-fold: true
#| code-summary: "Show model setup code"

# unified, scaled model
brms_scales = brms_data %>%
  group_by() %>%
  summarize(precip_mm_15_day_sum_mean = mean(precip_mm_15_day_sum),
            precip_mm_15_day_sum_sd = sd(precip_mm_15_day_sum),
            precip_mm_30_day_sum_mean = mean(precip_mm_30_day_sum),
            precip_mm_30_day_sum_sd = sd(precip_mm_30_day_sum),
            temp_c_15_day_mean_mean = mean(temp_c_15_day_mean),
            temp_c_15_day_mean_sd = sd(temp_c_15_day_mean),
            temp_c_30_day_mean_mean = mean(temp_c_30_day_mean),
            temp_c_30_day_mean_sd = sd(temp_c_30_day_mean),
            precip_mm_15_day_sum2_mean = mean(precip_mm_15_day_sum ^ 2),
            precip_mm_15_day_sum2_sd = sd(precip_mm_15_day_sum ^ 2),
            precip_mm_30_day_sum2_mean = mean(precip_mm_30_day_sum ^ 2),
            precip_mm_30_day_sum2_sd = sd(precip_mm_30_day_sum ^ 2),
            temp_c_15_day_mean2_mean = mean(temp_c_15_day_mean ^ 2),
            temp_c_15_day_mean2_sd = sd(temp_c_15_day_mean ^ 2),
            temp_c_30_day_mean2_mean = mean(temp_c_30_day_mean ^ 2),
            temp_c_30_day_mean2_sd = sd(temp_c_30_day_mean ^ 2))

brms_rescale = function(df, scales) {
  df_out = df %>%
    mutate(precip_mm_15_z = (precip_mm_15_day_sum - scales$precip_mm_15_day_sum_mean)/scales$precip_mm_15_day_sum_sd,
           precip_mm_30_z = (precip_mm_30_day_sum - scales$precip_mm_30_day_sum_mean)/scales$precip_mm_30_day_sum_sd,
           temp_c_15_z = (temp_c_15_day_mean - scales$temp_c_15_day_mean_mean)/scales$temp_c_15_day_mean_sd,
           temp_c_30_z = (temp_c_30_day_mean - scales$temp_c_30_day_mean_mean)/scales$temp_c_30_day_mean_sd,
           precip_mm_15_z2 = (precip_mm_15_day_sum^2 - scales$precip_mm_15_day_sum2_mean)/scales$precip_mm_15_day_sum2_sd,
           precip_mm_30_z2 = (precip_mm_30_day_sum^2 - scales$precip_mm_30_day_sum2_mean)/scales$precip_mm_30_day_sum2_sd,
           temp_c_15_z2 = (temp_c_15_day_mean^2 - scales$temp_c_15_day_mean2_mean)/scales$temp_c_15_day_mean2_sd,
           temp_c_30_z2 = (temp_c_30_day_mean^2 - scales$temp_c_30_day_mean2_mean)/scales$temp_c_30_day_mean2_sd
           )
}

model_priors <- c(
  # b
  prior(normal(0, 5), class = "Intercept"),
  prior(normal(0, 3),   class = "b"),
  
  # hu
  prior(normal(0, 2), class = "Intercept", dpar = "hu"),
  prior(normal(0, 1), class = "b", dpar = "hu"),
  
  # Sigma
  prior(normal(0, 1), class = "Intercept", dpar = "sigma"),
  prior(normal(0, 0.3), class = "b", dpar = "sigma"),
  
  # Random effects
  prior(exponential(1), class = "sd")
)

model_priors_ga <- c(
  # b
  prior(normal(0, 5), class = "Intercept"),
  prior(normal(0, 3),   class = "link"),
  
  # hu
  prior(normal(0, 2), class = "Intercept", dpar = "link_hu"),
  prior(normal(0, 1), class = "b", dpar = "link_hu"),
  
  # Sigma
  prior(normal(0, 1), class = "Intercept", dpar = "link_shape"),
  prior(normal(0, 0.3), class = "b", dpar = "link_shape"),
  
  # Random effects
  prior(exponential(1), class = "sd")
)

model_priors_fixed <- c(
  # b
  prior(normal(0, 5), class = "Intercept"),
  prior(normal(0, 3),   class = "b"),
  
  # hu
  prior(normal(0, 2), class = "Intercept", dpar = "hu"),
  prior(normal(0, 1), class = "b", dpar = "hu"),
  
  # Sigma
  prior(normal(0, 1), class = "Intercept", dpar = "sigma"),
  prior(normal(0, 0.3), class = "b", dpar = "sigma")
)

set_cmdstan_path(Sys.getenv("cmd_stan_path"))
```

## Model fitting
```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show model fitting code"

null_zln <- brm(
  bd_load_co_density ~ 1,
  data = brms_data,
  family = hurdle_lognormal(),
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 124,
  cores = 4,
  backend = "cmdstanr"
)

null_zln <- add_criterion(null_zln, "loo")

saveRDS(null_zln, file = here("unified_models/unified_model_results_zln_null.rds"))
null_zln = readRDS(file = here("unified_models/unified_model_results_zln_null.rds"))
# 
# nullb_zln <- brm(
#   bd_load_co ~ 1,
#   data = brms_data,
#   family = hurdle_lognormal(),
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# nullb_zln <- add_criterion(nullb_zln, "loo")
# 
# saveRDS(nullb_zln, file = here("unified_models/unified_model_results_zln_nullb.rds"))
# nullb_zln = readRDS(file = here("unified_models/unified_model_results_zln_nullb.rds"))
# 
# m1_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture,
#     hu ~ 1 + life_stage_simple * taxon_capture,
#     sigma ~ 1 + taxon_capture
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors_fixed,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m1_zln <- add_criterion(m1_zln, "loo")
# 
# saveRDS(m1_zln, file = here("unified_models/unified_model_results_zln_life_tax.rds"))
# m1_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax.rds"))
# 

# m2_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     sigma ~ 1 + taxon_capture
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m2_zln <- add_criterion(m2_zln, "loo")
# 
# saveRDS(m2_zln, file = here("unified_models/unified_model_results_zln_life_tax_pop.rds"))
# m2_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax_pop.rds"))
# 
# m2b_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     sigma ~ 1 + taxon_capture + (1 | population)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m2b_zln <- add_criterion(m2b_zln, "loo")
# 
# saveRDS(m2b_zln, file = here("unified_models/unified_model_results_zln_life_tax_popb.rds"))
# m2b_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax_popb.rds"))

# m2b_zga <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     shape ~ 1 + taxon_capture + (1 | population)
#   ),
#   data = brms_data,
#   family = hurdle_gamma(),
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m2b_zga <- add_criterion(m2b_zga, "loo")
# 
# saveRDS(m2b_zga, file = here("unified_models/unified_model_results_zga_life_tax_popb.rds"))
# m2b_zga = readRDS(file = here("unified_models/unified_model_results_zga_life_tax_popb.rds"))
# 
# m3_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 + population) +  (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 + population) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture + (1 + population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m3_zln <- add_criterion(m3_zln, "loo")
# 
# saveRDS(m3_zln, file = here("unified_models/unified_model_results_zln_life_tax_poptls.rds"))
# m3_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax_poptls.rds"))
# 
# m3b_zln <- brm(
#   formula = bf(
#     bd_load_co ~ 1 + life_stage_simple * taxon_capture + (1 + population) +  (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 + population) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture + (1 + population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m3b_zln <- add_criterion(m3b_zln, "loo")
# 
# saveRDS(m3b_zln, file = here("unified_models/unified_model_results_zln_life_tax_poptls_nodens.rds"))
# m3b_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax_poptls_nodens.rds"))

# m4_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + temp_c_30_z + (1 + population) +  (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + temp_c_15_z + (1 + population) +  (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture + (1 + population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# m4_zln <- add_criterion(m4_zln, "loo")
# 
# saveRDS(m4_zln, file = "unified_models/unified_model_results_zln_life_tax_pop_env.rds")
# m4_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_pop_env.rds")
# 
# m5_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5_zln <- add_criterion(m5_zln, "loo")
# 
# saveRDS(m5_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls_env2.rds")
# m5_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls_env2.rds")
# 

# m5b_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + temp_c_30_z | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + temp_c_15_z | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5b_zln <- add_criterion(m5b_zln, "loo")
# 
# saveRDS(m5b_zln, file = "unified_models/unified_model_results_zln_life_tax_poptlsenv_env2.rds")
# m5b_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptlsenv_env2.rds")
# 
# m5c_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5c_zln <- add_criterion(m5c_zln, "loo")
# 
# saveRDS(m5c_zln, file = "unified_models/unified_model_results_zln_life_tax_poptlsenv2_env2.rds")
# m5c_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptlsenv2_env2.rds")

# m5d_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + temp_c_30_z | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + temp_c_15_z | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5d_zln <- add_criterion(m5d_zln, "loo")
# 
# saveRDS(m5d_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2env_env2.rds")
# m5d_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env_env2.rds")


# m5e_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + temp_c_30_z | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + temp_c_15_z | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + life_stage_simple * taxon_capture +  (1 | population) + (1 | population:taxon_capture:life_stage_simple)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5e_zln <- add_criterion(m5e_zln, "loo")
# 
# saveRDS(m5e_zln, file = "unified_models/unified_model_results_zln_life_tax_poptlsenv_env2_sigls.rds")
# m5e_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptlsenv_env2_sigls.rds")
# 
# m5f_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5f_zln <- add_criterion(m5f_zln, "loo")
# 
# saveRDS(m5f_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2.rds")
# m5f_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2.rds")

##
## I think I should drop the (1|population) terms and keep only the (1|population:taxon_capture:life_stage_simple) terms... maybe? Or does the fact that they are not neatly nested mean this not trully redundant?
##
##
m5g_zln <- brm(
  formula = bf(
    bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
    hu ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
    sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
  ),
  data = brms_data,
  family = hurdle_lognormal(),
  prior = model_priors,
  sample_prior = "yes",
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 124,
  cores = 4,
  backend = "cmdstanr"
)

m5g_zln <- add_criterion(m5g_zln, "loo")

saveRDS(m5g_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_30.rds")
m5g_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_30.rds")
# 
# m6_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_sp_z + precip_mm_15_sp_z2 + temp_c_15_sp_z + temp_c_15_sp_z2 + (1 | population) + (1 + precip_mm_15_sp_z + precip_mm_15_sp_z2 + temp_c_15_sp_z + temp_c_15_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m6_zln <- add_criterion(m6_zln, "loo")
# 
# saveRDS(m6_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_envsp2.rds")
# m6_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_envsp2.rds")
# 
# m6b_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_15_sp_z + precip_mm_15_sp_z2 + temp_c_15_sp_z + temp_c_15_sp_z2 | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m6b_zln <- add_criterion(m6b_zln, "loo")
# 
# saveRDS(m6b_zln, file = "unified_models/unified_model_results_zln_life_tax_popt2envsp2.rds")
# m6b_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_popt2envsp2.rds")
# 
# m6c_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m6c_zln <- add_criterion(m6c_zln, "loo")
# 
# saveRDS(m6c_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_envsp2_30.rds")
# m6c_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_envsp2_30.rds")
# 
# m6d_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m6d_zln <- add_criterion(m6d_zln, "loo")
# 
# saveRDS(m6d_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_30.rds")
# m6d_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_30.rds")
# 
# m7_zln <- brm(
#   formula = bf(
#     bd_load_co ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m7_zln <- add_criterion(m7_zln, "loo")
# 
# saveRDS(m7_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2_nodens.rds")
# m7_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2_nodens.rds")
# 


```

## Model comparison
```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show model comparison code"
# model comparison and selection

(loo_comp_out = loo_compare(null_zln,
                            # nullb_zln, # no density
                            # m1_zln,
                            # m2_zln,
                            # m2b_zln,
                            # m3_zln,
                            # m3b_zln,  # no density
                            # m4_zln,
                            # m5_zln,
                            # m5b_zln,
                            # m5c_zln,
                            # m5d_zln,
                            # m5e_zln,
                            # m5f_zln,
                            m5g_zln
                            # m6_zln,
                            # m6b_zln,
                            # m6c_zln,
                            # m6d_zln
                            ))
```


## Model checks

### Fitting and correlation checks
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show model checks code"
#| fig-height: 6
#| fig-width: 8

model_select = m5g_zln

summary(model_select)
plot((model_select), ask = FALSE)
plot(conditional_effects(model_select), ask = FALSE)
VarCorr(model_select)
```

### Posterios predictive checks
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show model checks code (fixed effects)"
#| fig-height: 6
#| fig-width: 8

color_scheme_set("viridis")
model_select = m5g_zln


pp_check(model_select,
         type = "dens_overlay",
         prefix = "ppc",
         ndraws = 100) +
  scale_x_log10()

# # break out by taxon
# pp_check(model_select,
#          type = "dens_overlay_grouped",
#          group = "taxon_capture",
#          ndraws = 100) +
#   scale_x_log10()


```

### Prior checks
```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show prior checks code"
#| fig-height: 6
#| fig-width: 8

mcmc_plot(m5g_zln, type = "dens", variable = c("prior_Intercept", "b_Intercept")) +
  xlim(-20, 20)

# mcmc_plot(m5f_zln, type = "dens", variable = c("prior_b", "b_taxon_capturecolostethus_panamansis", "b_taxon_capturelithobates_blairi", 
# "b_taxon_capturelithobates_sphenocephalus", "b_taxon_capturelithobates_warszewitschii", 
# "b_taxon_capturepseudacris_crucifer", "b_taxon_capturerana_catesbeiana", 
# "b_taxon_capturerana_clamitans", "b_taxon_capturerana_muscosa", 
# "b_taxon_capturerana_pipiens", "b_taxon_capturerana_sierrae", 
# "b_taxon_capturesilverstoneia_flotator")) +
#   xlim(-10, 10)

mcmc_plot(m5g_zln, type = "dens", variable = c("prior_Intercept_sigma", "b_sigma_Intercept")) +
  xlim(-2, 2)

# mcmc_plot(m5f_zln, type = "dens", variable = c("prior_b_sigma", "b_sigma_taxon_capturecolostethus_panamansis", "b_sigma_taxon_capturelithobates_blairi", 
# "b_sigma_taxon_capturelithobates_sphenocephalus", "b_sigma_taxon_capturelithobates_warszewitschii", 
# "b_sigma_taxon_capturepseudacris_crucifer", "b_sigma_taxon_capturerana_catesbeiana", 
# "b_sigma_taxon_capturerana_clamitans", "b_sigma_taxon_capturerana_muscosa", 
# "b_sigma_taxon_capturerana_pipiens", "b_sigma_taxon_capturerana_sierrae", 
# "b_sigma_taxon_capturesilverstoneia_flotator")) +
#   xlim(-2, 2)

mcmc_plot(m5g_zln, type = "dens", variable = c("prior_Intercept_hu", "b_hu_Intercept")) +
  xlim(-15, 15)

# mcmc_plot(m5f_zln, type = "dens", variable = c("prior_b_hu", "b_hu_life_stage_simplesubadult", "b_hu_taxon_capturecolostethus_panamansis", 
# "b_hu_taxon_capturelithobates_blairi", "b_hu_taxon_capturelithobates_sphenocephalus", 
# "b_hu_taxon_capturelithobates_warszewitschii", "b_hu_taxon_capturepseudacris_crucifer", 
# "b_hu_taxon_capturerana_catesbeiana", "b_hu_taxon_capturerana_clamitans", 
# "b_hu_taxon_capturerana_muscosa", "b_hu_taxon_capturerana_pipiens", 
# "b_hu_taxon_capturerana_sierrae", "b_hu_taxon_capturesilverstoneia_flotator")) +
#   xlim(-5, 5)

```
