---
title: "Size vs. Bd modeling: Panamá preliminary results"
author: Cob Staines
date: 2025-09-29
format:
  html:
    toc: true
df-print: kable
bibliography: references.bib
csl: ecology.csl
---

# Introduction

## Motivation

Patterns of greater Bd load in recently metamorphosed individuals have been documented in some cases[@adams_chytridiomycosis-induced_2020] [@humphries_immune_2022] [@humphries_chytridiomycosis_2024] [@humphries_chytridiomycosis_2025] (and the opposite pattern in others[@bradley_host_2019]), with possible implications for:

  - resilience trajectory assessment (Q1)
  - population dynamics (Q2)
  - assessing Bd resistance/tolerance mechanisms (Q3)
  
... with the goal of complementing ongoing RIBBiTR research efforts

## Working hypotheses:

  a. Recently-metamorphosed individuals are more likely to be infected with Bd than older individuals across species
  b. Recently-metamorphosed individuals are more likely to carry higher Bd loads than older individuals across species

# Methods
## Setup

```{r}
#| eval: true
#| output: false
#| code-fold: true
#| code-summary: "Show setup code"

# setup
librarian::shelf(tidyverse, here, janitor, brms, tidybayes, bayesplot, cmdstanr, bayestestR)

```


## Load saved data
```{r}
#| eval: true
#| warning: false
#| echo: false
#| code-fold: true
#| code-summary: "Show data loading code"

# load bd_model_env
load(here("size_bd_unified", "intermediates", "bd_model_env_unified_2026-04-22.RData")) 

# checks
sum(is.na(brms_data$bd_load_co_density))
sum(is.na(brms_data$precip_mm_30_day_sum))
sum(is.na(brms_data$temperature_max_c))
sum(is.na(brms_data$past_year_weekly_precipitation_mm_sd))
sum(is.na(brms_data$past_year_weekly_tempurature_max_c_sd))

bd_model_env_stats = brms_data %>%
  group_by(taxon_capture, country) %>%
  summarise(regions = paste(sort(unique(region)), collapse = ", "),
            population_n = n_distinct(population),
            site_n = n_distinct(site),
            year_first = year(min(date)),
            year_last = year(max(date)),
            year_count = n_distinct(year(date)),
            adult_n = sum(life_stage_simple == "adult"),
            adult_n_pos = sum(life_stage_simple == "adult" & bd_detected_co),
            subadult_n = sum(life_stage_simple == "subadult"),
            subadult_n_pos = sum(life_stage_simple == "subadult" & bd_detected_co),
            .groups = "drop") %>%
  arrange(country,
          taxon_capture)
```

# Modeling
I modeled Bd load using a Baysian zero-inflated (or "hurdle") log-normal model [@schrock2025fungi] with fixed effects of species, life stage, and recent precipitation and temperature, and considering population, species, life stage and recent precipitation and temperature as a random effects. The "hurdle" infection probability (hu) and the variability in Bd load (sigma) are allowed to vary with life stage.

## Model setup
```{r}
#| code-fold: true
#| code-summary: "Show model setup code"

set_cmdstan_path(Sys.getenv("cmd_stan_path"))

model_priors <- c(
  # b
  prior(normal(5, 4), class = "Intercept"),
  prior(normal(0, 3),   class = "b"),
  
  # hu
  prior(normal(0, 2), class = "Intercept", dpar = "hu"),
  prior(normal(0, 2), class = "b", dpar = "hu"),
  
  # Sigma
  prior(normal(0, 1), class = "Intercept", dpar = "sigma"),
  prior(normal(0, 0.3), class = "b", dpar = "sigma"),
  
  # Random effects
  prior(exponential(1), class = "sd")
)

# model_priors_ga <- c(
#   # b
#   prior(normal(0, 5), class = "Intercept"),
#   prior(normal(0, 3),   class = "link"),
#   
#   # hu
#   prior(normal(0, 2), class = "Intercept", dpar = "link_hu"),
#   prior(normal(0, 1), class = "b", dpar = "link_hu"),
#   
#   # Sigma
#   prior(normal(0, 1), class = "Intercept", dpar = "link_shape"),
#   prior(normal(0, 0.3), class = "b", dpar = "link_shape"),
#   
#   # Random effects
#   prior(exponential(1), class = "sd")
# )
# 
# model_priors_fixed <- c(
#   # b
#   prior(normal(0, 5), class = "Intercept"),
#   prior(normal(0, 3),   class = "b"),
#   
#   # hu
#   prior(normal(0, 2), class = "Intercept", dpar = "hu"),
#   prior(normal(0, 1), class = "b", dpar = "hu"),
#   
#   # Sigma
#   prior(normal(0, 1), class = "Intercept", dpar = "sigma"),
#   prior(normal(0, 0.3), class = "b", dpar = "sigma")
# )

```

## Model fitting
```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show model fitting code"

null_zln <- brm(
  bd_load_co_density ~ 1,
  data = brms_data,
  family = hurdle_lognormal(),
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 124,
  cores = 4,
  backend = "cmdstanr"
)

null_zln <- add_criterion(null_zln, "loo")

saveRDS(null_zln, file = here("size_bd_unified/unified_models/unified_model_results_zln_null.rds"))
null_zln = readRDS(file = here("size_bd_unified/unified_models/unified_model_results_zln_null.rds"))
# 
# nullb_zln <- brm(
#   bd_load_co ~ 1,
#   data = brms_data,
#   family = hurdle_lognormal(),
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# nullb_zln <- add_criterion(nullb_zln, "loo")
# 
# saveRDS(nullb_zln, file = here("unified_models/unified_model_results_zln_nullb.rds"))
# nullb_zln = readRDS(file = here("unified_models/unified_model_results_zln_nullb.rds"))
# 
# m1_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture,
#     hu ~ 1 + life_stage_simple * taxon_capture,
#     sigma ~ 1 + taxon_capture
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors_fixed,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m1_zln <- add_criterion(m1_zln, "loo")
# 
# saveRDS(m1_zln, file = here("unified_models/unified_model_results_zln_life_tax.rds"))
# m1_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax.rds"))
# 

# m2_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     sigma ~ 1 + taxon_capture
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m2_zln <- add_criterion(m2_zln, "loo")
# 
# saveRDS(m2_zln, file = here("unified_models/unified_model_results_zln_life_tax_pop.rds"))
# m2_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax_pop.rds"))
# 
# m2b_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     sigma ~ 1 + taxon_capture + (1 | population)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m2b_zln <- add_criterion(m2b_zln, "loo")
# 
# saveRDS(m2b_zln, file = here("unified_models/unified_model_results_zln_life_tax_popb.rds"))
# m2b_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax_popb.rds"))

# m2b_zga <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population),
#     shape ~ 1 + taxon_capture + (1 | population)
#   ),
#   data = brms_data,
#   family = hurdle_gamma(),
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m2b_zga <- add_criterion(m2b_zga, "loo")
# 
# saveRDS(m2b_zga, file = here("unified_models/unified_model_results_zga_life_tax_popb.rds"))
# m2b_zga = readRDS(file = here("unified_models/unified_model_results_zga_life_tax_popb.rds"))
# 
# m3_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 + population) +  (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 + population) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture + (1 + population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m3_zln <- add_criterion(m3_zln, "loo")
# 
# saveRDS(m3_zln, file = here("unified_models/unified_model_results_zln_life_tax_poptls.rds"))
# m3_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax_poptls.rds"))
# 
# m3b_zln <- brm(
#   formula = bf(
#     bd_load_co ~ 1 + life_stage_simple * taxon_capture + (1 + population) +  (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 + population) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture + (1 + population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m3b_zln <- add_criterion(m3b_zln, "loo")
# 
# saveRDS(m3b_zln, file = here("unified_models/unified_model_results_zln_life_tax_poptls_nodens.rds"))
# m3b_zln = readRDS(file = here("unified_models/unified_model_results_zln_life_tax_poptls_nodens.rds"))

# m4_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + temp_c_30_z + (1 + population) +  (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + temp_c_15_z + (1 + population) +  (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture + (1 + population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# m4_zln <- add_criterion(m4_zln, "loo")
# 
# saveRDS(m4_zln, file = "unified_models/unified_model_results_zln_life_tax_pop_env.rds")
# m4_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_pop_env.rds")
# 
# m5_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5_zln <- add_criterion(m5_zln, "loo")
# 
# saveRDS(m5_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls_env2.rds")
# m5_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls_env2.rds")
# 

# m5b_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + temp_c_30_z | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + temp_c_15_z | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5b_zln <- add_criterion(m5b_zln, "loo")
# 
# saveRDS(m5b_zln, file = "unified_models/unified_model_results_zln_life_tax_poptlsenv_env2.rds")
# m5b_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptlsenv_env2.rds")
# 
# m5c_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5c_zln <- add_criterion(m5c_zln, "loo")
# 
# saveRDS(m5c_zln, file = "unified_models/unified_model_results_zln_life_tax_poptlsenv2_env2.rds")
# m5c_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptlsenv2_env2.rds")

# m5d_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + temp_c_30_z | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + temp_c_15_z | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5d_zln <- add_criterion(m5d_zln, "loo")
# 
# saveRDS(m5d_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2env_env2.rds")
# m5d_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env_env2.rds")


# m5e_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + temp_c_30_z | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + temp_c_15_z | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + life_stage_simple * taxon_capture +  (1 | population) + (1 | population:taxon_capture:life_stage_simple)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5e_zln <- add_criterion(m5e_zln, "loo")
# 
# saveRDS(m5e_zln, file = "unified_models/unified_model_results_zln_life_tax_poptlsenv_env2_sigls.rds")
# m5e_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptlsenv_env2_sigls.rds")
# 
# m5f_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5f_zln <- add_criterion(m5f_zln, "loo")
# 
# saveRDS(m5f_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2.rds")
# m5f_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2.rds")

##
## I think I should drop the (1|population) terms and keep only the (1|population:taxon_capture:life_stage_simple) terms... maybe? Or does the fact that they are not neatly nested mean this not trully redundant?
##
##
# m5g_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m5g_zln <- add_criterion(m5g_zln, "loo")
# 
# saveRDS(m5g_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_30.rds")
# m5g_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_30.rds")
# 
# m6_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_sp_z + precip_mm_15_sp_z2 + temp_c_15_sp_z + temp_c_15_sp_z2 + (1 | population) + (1 + precip_mm_15_sp_z + precip_mm_15_sp_z2 + temp_c_15_sp_z + temp_c_15_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m6_zln <- add_criterion(m6_zln, "loo")
# 
# saveRDS(m6_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_envsp2.rds")
# m6_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_envsp2.rds")
# 
# m6b_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_15_sp_z + precip_mm_15_sp_z2 + temp_c_15_sp_z + temp_c_15_sp_z2 | taxon_capture) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m6b_zln <- add_criterion(m6b_zln, "loo")
# 
# saveRDS(m6b_zln, file = "unified_models/unified_model_results_zln_life_tax_popt2envsp2.rds")
# m6b_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_popt2envsp2.rds")
# 
# m6c_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m6c_zln <- add_criterion(m6c_zln, "loo")
# 
# saveRDS(m6c_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_envsp2_30.rds")
# m6c_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_envsp2_30.rds")
# 
# m6d_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + (1 | population) + (1 + precip_mm_30_sp_z + precip_mm_30_sp_z2 + temp_c_30_sp_z + temp_c_30_sp_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m6d_zln <- add_criterion(m6d_zln, "loo")
# 
# saveRDS(m6d_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_30.rds")
# m6d_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2envsp2_30.rds")
# 
# m7_zln <- brm(
#   formula = bf(
#     bd_load_co ~ 1 + life_stage_simple * taxon_capture + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 + (1 | population) + (1 + precip_mm_30_z + precip_mm_30_z2 + temp_c_30_z + temp_c_30_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 + (1 | population) + (1 + precip_mm_15_z + precip_mm_15_z2 + temp_c_15_z + temp_c_15_z2 | taxon_capture:life_stage_simple) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.95, max_treedepth = 12),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 124,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m7_zln <- add_criterion(m7_zln, "loo")
# 
# saveRDS(m7_zln, file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2_nodens.rds")
# m7_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2_nodens.rds")
# 

# m8_zln <- brm(
#   formula = bf(
#     bd_load_co_density ~ 1 + life_stage_simple * taxon_capture + (precip_mm_10_z + I(precip_mm_10_z^2) + temp_c_30_z + I(temp_c_30_z^2)) * taxon_capture + (1 | population) + (1 | population:taxon_capture:life_stage_simple),
#     hu ~ 1 + life_stage_simple * taxon_capture + (precip_mm_10_z + I(precip_mm_10_z^2) + temp_c_30_z + I(temp_c_30_z^2)) * taxon_capture + (1 | population) + (1 | population:taxon_capture:life_stage_simple),
#     sigma ~ 1 + taxon_capture +  (1 | population) + (1 | population:taxon_capture)
#   ),
#   data = brms_data,
#   family = hurdle_lognormal(),
#   prior = model_priors,
#   sample_prior = "yes",
#   control = list(adapt_delta = 0.98),
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   seed = 125,
#   cores = 4,
#   backend = "cmdstanr"
# )
# 
# m8_zln <- add_criterion(m8_zln, "loo")
# 
# saveRDS(m8_zln, file = "unified_models/unified_model_results_zln_life_tax_env2_rpoptls_p10t30.rds")
# m8_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_env2_rpoptls_p10t30.rds")

# quadratic fixed temp and precip, linear random
m9a_zln <- brm(
  formula = bf(
    bd_load_co_density ~ 1 + 
      life_stage_simple * taxon_capture +
      (precip_mm_10_z + I(precip_mm_10_z^2) + temp_c_30_z + I(temp_c_30_z^2)) +
      (0 + precip_mm_10_z + temp_c_30_z | taxon_capture) +
      (1 | population) + 
      (1 | population:taxon_capture:life_stage_simple),
    
    hu ~ 1 + 
      life_stage_simple * taxon_capture +
      (precip_mm_10_z + I(precip_mm_10_z^2) + temp_c_30_z + I(temp_c_30_z^2)) +
      (0 + precip_mm_10_z + temp_c_30_z | taxon_capture) +
      (1 | population) + 
      (1 | population:taxon_capture:life_stage_simple),
    
    sigma ~ 1 + taxon_capture + 
      (1 | population)
  ),
  data = brms_data,
  family = hurdle_lognormal(),
  prior = model_priors,
  sample_prior = "yes",
  control = list(adapt_delta = 0.98),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 125,
  cores = 4,
  backend = "cmdstanr"
)

m9a_zln <- add_criterion(m9a_zln, "loo")

saveRDS(m9a_zln, file = "unified_models/unified_model_results_zln_life_tax_env2_renvpoptls_p10t30.rds")
m9a_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_env2_renvpoptls_p10t30.rds")

# quadratic fixed and random temp and precip
m9b_zln <- brm(
  formula = bf(
    bd_load_co_density ~ 1 + 
      life_stage_simple * taxon_capture +
      (precip_mm_10_z + I(precip_mm_10_z^2) + temp_c_30_z + I(temp_c_30_z^2)) +
      (0 + precip_mm_10_z + I(precip_mm_10_z^2) + temp_c_30_z + I(temp_c_30_z^2) | taxon_capture) +
      (1 | population) + 
      (1 | population:taxon_capture:life_stage_simple),
    
    hu ~ 1 + 
      life_stage_simple * taxon_capture +
      (precip_mm_10_z + I(precip_mm_10_z^2) + temp_c_30_z + I(temp_c_30_z^2)) +
      (0 + precip_mm_10_z + I(precip_mm_10_z^2) + temp_c_30_z + I(temp_c_30_z^2) | taxon_capture) +
      (1 | population) + 
      (1 | population:taxon_capture:life_stage_simple),
    
    sigma ~ 1 + taxon_capture + 
      (1 | population)
  ),
  data = brms_data,
  family = hurdle_lognormal(),
  prior = model_priors,
  sample_prior = "yes",
  control = list(adapt_delta = 0.98),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 125,
  cores = 4,
  backend = "cmdstanr"
)

m9b_zln <- add_criterion(m9b_zln, "loo")

saveRDS(m9b_zln, file = "unified_models/unified_model_results_zln_life_tax_env2_renv2poptls_p10t30.rds")
m9b_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_env2_renv2poptls_p10t30.rds")

# 30-day precip lag
m9b_30_zln <- brm(
  formula = bf(
    bd_load_co_density ~ 1 + 
      life_stage_simple * taxon_capture +
      (precip_mm_30_z + I(precip_mm_30_z^2) + temp_c_30_z + I(temp_c_30_z^2)) +
      (0 + precip_mm_30_z + I(precip_mm_30_z^2) + temp_c_30_z + I(temp_c_30_z^2) | taxon_capture) +
      (1 | population) + 
      (1 | population:taxon_capture:life_stage_simple),
    
    hu ~ 1 + 
      life_stage_simple * taxon_capture +
      (precip_mm_30_z + I(precip_mm_30_z^2) + temp_c_30_z + I(temp_c_30_z^2)) +
      (0 + precip_mm_30_z + I(precip_mm_30_z^2) + temp_c_30_z + I(temp_c_30_z^2) | taxon_capture) +
      (1 | population) + 
      (1 | population:taxon_capture:life_stage_simple),
    
    sigma ~ 1 + taxon_capture + 
      (1 | population)
  ),
  data = brms_data,
  family = hurdle_lognormal(),
  prior = model_priors,
  sample_prior = "yes",
  control = list(adapt_delta = 0.98),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 125,
  cores = 4,
  backend = "cmdstanr"
)

m9b_30_zln <- add_criterion(m9b_30_zln, "loo")

saveRDS(m9b_30_zln, file = "unified_models/unified_model_results_zln_life_tax_env2_renv2poptls_p30t30.rds")
m9b_30_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_env2_renv2poptls_p30t30.rds")

# quadratic in temp, linear in precip
m9c_zln <- brm(
  formula = bf(
    bd_load_co_density ~ 1 + 
      life_stage_simple * taxon_capture +
      (precip_mm_10_z  + temp_c_30_z + I(temp_c_30_z^2)) +
      (0 + precip_mm_10_z + temp_c_30_z + I(temp_c_30_z^2) | taxon_capture) +
      (1 | population) + 
      (1 | population:taxon_capture:life_stage_simple),
    
    hu ~ 1 + 
      life_stage_simple * taxon_capture +
      (precip_mm_10_z + temp_c_30_z + I(temp_c_30_z^2)) +
      (0 + precip_mm_10_z + temp_c_30_z + I(temp_c_30_z^2) | taxon_capture) +
      (1 | population) + 
      (1 | population:taxon_capture:life_stage_simple),
    
    sigma ~ 1 + taxon_capture + 
      (1 | population)
  ),
  data = brms_data,
  family = hurdle_lognormal(),
  prior = model_priors,
  sample_prior = "yes",
  control = list(adapt_delta = 0.98),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 125,
  cores = 4,
  backend = "cmdstanr"
)

m9c_zln <- add_criterion(m9c_zln, "loo")

saveRDS(m9c_zln, file = "unified_models/unified_model_results_zln_life_tax_pt2_rpt2poptls_p10t30.rds")
m9c_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_pt2_rpt2poptls_p10t30.rds")

# quadratic temp, linear precip, year
m9d_zln <- brm(
  formula = bf(
    bd_load_co_density ~ 1 + 
      life_stage_simple * taxon_capture +
      (precip_mm_10_z  + temp_c_30_z + I(temp_c_30_z^2)) +
      (0 + precip_mm_10_z + temp_c_30_z + I(temp_c_30_z^2) | taxon_capture) +
      (1 | population:year) + 
      (1 | population:taxon_capture:life_stage_simple),
    
    hu ~ 1 + 
      life_stage_simple * taxon_capture +
      (precip_mm_10_z + temp_c_30_z + I(temp_c_30_z^2)) +
      (0 + precip_mm_10_z + temp_c_30_z + I(temp_c_30_z^2) | taxon_capture) +
      (1 | population:year) + 
      (1 | population:taxon_capture:life_stage_simple),
    
    sigma ~ 1 + taxon_capture + 
      (1 | population:year)
  ),
  data = brms_data,
  family = hurdle_lognormal(),
  prior = model_priors,
  sample_prior = "yes",
  control = list(adapt_delta = 0.98),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 125,
  cores = 4,
  backend = "cmdstanr"
)

m9d_zln <- add_criterion(m9d_zln, "loo")

saveRDS(m9d_zln, file = "unified_models/unified_model_results_zln_life_tax_pt2_rpt2poptlsyr_p10t30.rds")
m9d_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_pt2_rpt2poptlsyr_p10t30.rds")

# quadratic temp, linear precip, year, nested taxon random
m9e_zln <- brm(
  formula = bf(
    bd_load_co_density ~ 1 + 
      life_stage_simple * taxon_capture +
      (precip_mm_10_z  + temp_c_30_z + I(temp_c_30_z^2)) +
      (0 + precip_mm_10_z + temp_c_30_z + I(temp_c_30_z^2) | taxon_capture) +
      (1 | population:year) + 
      (1 | taxon_capture/population:life_stage_simple),
    
    hu ~ 1 + 
      life_stage_simple * taxon_capture +
      (precip_mm_10_z + temp_c_30_z + I(temp_c_30_z^2)) +
      (0 + precip_mm_10_z + temp_c_30_z + I(temp_c_30_z^2) | taxon_capture) +
      (1 | population:year) + 
      (1 | taxon_capture/population:life_stage_simple),
    
    sigma ~ 1 + taxon_capture + 
      (1 | population:year)
  ),
  data = brms_data,
  family = hurdle_lognormal(),
  prior = model_priors,
  sample_prior = "yes",
  control = list(adapt_delta = 0.98),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 125,
  cores = 4,
  backend = "cmdstanr"
)

m9e_zln <- add_criterion(m9e_zln, "loo")

saveRDS(m9e_zln, file = "unified_models/unified_model_results_zln_life_tax_pt2_rpt2popt*lsyr_p10t30.rds")
m9e_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_pt2_rpt2popt*lsyr_p10t30.rds")

# quadratic temp, linear precip, year, climate var linear
m9f_zln <- brm(
  formula = bf(
    bd_load_co_density ~ 1 + 
      life_stage_simple * taxon_capture +
      past_year_weekly_precipitation_mm_sd + past_year_weekly_tempurature_max_c_sd +
      (precip_mm_10_z  + temp_c_30_z + I(temp_c_30_z^2)) +
      (0 + precip_mm_10_z + temp_c_30_z + I(temp_c_30_z^2) | taxon_capture) +
      (1 | population:year) + 
      (1 | population:taxon_capture:life_stage_simple),
    
    hu ~ 1 + 
      life_stage_simple * taxon_capture +
      past_year_weekly_precipitation_mm_sd + past_year_weekly_tempurature_max_c_sd +
      (precip_mm_10_z + temp_c_30_z + I(temp_c_30_z^2)) +
      (0 + precip_mm_10_z + temp_c_30_z + I(temp_c_30_z^2) | taxon_capture) +
      (1 | population:year) + 
      (1 | population:taxon_capture:life_stage_simple),
    
    sigma ~ 1 + taxon_capture + 
      past_year_weekly_precipitation_mm_sd + past_year_weekly_tempurature_max_c_sd +
      (1 | population:year)
  ),
  data = brms_data,
  family = hurdle_lognormal(),
  prior = model_priors,
  sample_prior = "yes",
  control = list(adapt_delta = 0.98),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 126,
  cores = 4,
  backend = "cmdstanr"
)

m9f_zln <- add_criterion(m9f_zln, "loo")

saveRDS(m9f_zln, file = "unified_models/unified_model_results_zln_life_tax_pt2_rpt2poptlsyrcl_p10t30.rds")
m9f_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_pt2_rpt2poptlsyrcl_p10t30.rds")

mxf_zln <- brm(
  formula = bf(
    bd_load_co_density ~ 1 + 
      past_year_weekly_precipitation_mm_sd + past_year_weekly_tempurature_max_c_sd +
      (1 | population:year),
    
    hu ~ 1 + 
      past_year_weekly_precipitation_mm_sd + past_year_weekly_tempurature_max_c_sd +
      (1 | population:year),
    
    sigma ~ 1 +
      past_year_weekly_precipitation_mm_sd + past_year_weekly_tempurature_max_c_sd +
      (1 | population:year)
  ),
  data = brms_data,
  family = hurdle_lognormal(),
  prior = model_priors,
  sample_prior = "yes",
  control = list(adapt_delta = 0.98),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 126,
  cores = 4,
  backend = "cmdstanr"
)

# mxf_zln <- add_criterion(mxf_zln, "loo")

saveRDS(mxf_zln, file = "unified_models/unified_model_results_zln_life_tax_cl.rds")
mxf_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_cl.rds")

```

## Model comparison
```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show model comparison code"
# model comparison and selection

(loo_comp_out = loo_compare(null_zln,
                            # nullb_zln, # no density
                            # m1_zln,
                            # m2_zln,
                            # m2b_zln,
                            # m3_zln,
                            # m3b_zln,  # no density
                            # m4_zln,
                            # m5_zln,
                            # m5b_zln,
                            # m5c_zln,
                            # m5d_zln,
                            # m5e_zln,
                            # m5f_zln,
                            # m5g_zln
                            # m6_zln,
                            # m6b_zln,
                            # m6c_zln,
                            # m6d_zln,
                            # m8_zln,
                            m9a_zln,
                            m9b_zln,
                            # m9b_30_zln
                            m9c_zln,
                            m9d_zln,
                            m9e_zln
                            ))
```


## Model checks

### Fitting and correlation checks
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show model checks code"
#| fig-height: 6
#| fig-width: 8

model_select = m9d_zln

summary(model_select)
plot((model_select), ask = FALSE)
plot(conditional_effects(model_select), ask = FALSE)
VarCorr(model_select)
bayes_R2(model_select)
```

### Posterios predictive checks
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show model checks code (fixed effects)"
#| fig-height: 10
#| fig-width: 10

color_scheme_set("viridis")
# model_select = m5g_zln


pp_check(model_select,
         type = "dens_overlay",
         prefix = "ppc",
         ndraws = 100) +
  scale_x_log10()

# # break out by taxon
# pp_check(model_select,
#          type = "dens_overlay_grouped",
#          group = "taxon_capture",
#          ndraws = 100) +
#   scale_x_log10()

pp_check(model_select, group = "taxon_capture", type = "dens_overlay_grouped") +
  scale_x_log10()

```

### Prior checks
```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show prior checks code"
#| fig-height: 6
#| fig-width: 8

mcmc_plot(model_select, type = "dens", variable = c("prior_Intercept", "b_Intercept")) +
  xlim(-20, 20)

# mcmc_plot(m5f_zln, type = "dens", variable = c("prior_b", "b_taxon_capturecolostethus_panamansis", "b_taxon_capturelithobates_blairi", 
# "b_taxon_capturelithobates_sphenocephalus", "b_taxon_capturelithobates_warszewitschii", 
# "b_taxon_capturepseudacris_crucifer", "b_taxon_capturerana_catesbeiana", 
# "b_taxon_capturerana_clamitans", "b_taxon_capturerana_muscosa", 
# "b_taxon_capturerana_pipiens", "b_taxon_capturerana_sierrae", 
# "b_taxon_capturesilverstoneia_flotator")) +
#   xlim(-10, 10)

mcmc_plot(model_select, type = "dens", variable = c("prior_Intercept_sigma", "b_sigma_Intercept")) +
  xlim(-2, 2)

# mcmc_plot(m5f_zln, type = "dens", variable = c("prior_b_sigma", "b_sigma_taxon_capturecolostethus_panamansis", "b_sigma_taxon_capturelithobates_blairi", 
# "b_sigma_taxon_capturelithobates_sphenocephalus", "b_sigma_taxon_capturelithobates_warszewitschii", 
# "b_sigma_taxon_capturepseudacris_crucifer", "b_sigma_taxon_capturerana_catesbeiana", 
# "b_sigma_taxon_capturerana_clamitans", "b_sigma_taxon_capturerana_muscosa", 
# "b_sigma_taxon_capturerana_pipiens", "b_sigma_taxon_capturerana_sierrae", 
# "b_sigma_taxon_capturesilverstoneia_flotator")) +
#   xlim(-2, 2)

mcmc_plot(model_select, type = "dens", variable = c("prior_Intercept_hu", "b_hu_Intercept")) +
  xlim(-15, 15)

# mcmc_plot(m5f_zln, type = "dens", variable = c("prior_b_hu", "b_hu_life_stage_simplesubadult", "b_hu_taxon_capturecolostethus_panamansis", 
# "b_hu_taxon_capturelithobates_blairi", "b_hu_taxon_capturelithobates_sphenocephalus", 
# "b_hu_taxon_capturelithobates_warszewitschii", "b_hu_taxon_capturepseudacris_crucifer", 
# "b_hu_taxon_capturerana_catesbeiana", "b_hu_taxon_capturerana_clamitans", 
# "b_hu_taxon_capturerana_muscosa", "b_hu_taxon_capturerana_pipiens", 
# "b_hu_taxon_capturerana_sierrae", "b_hu_taxon_capturesilverstoneia_flotator")) +
#   xlim(-5, 5)

```
