---
title: "size_bd_unified_analysis"
format: html
---

# Load data & models
```{r}
librarian::shelf(tidyverse, here, janitor, brms, tidybayes, ggplot2, ggrepel, emmeans, rstatix, phylotools, rentrez, DECIPHER, phangorn, phytools, RColorBrewer)

# load bd_model_env
load(here("data", "bd_model_env_unified_2026-03-22.RData")) 

bd_model_env_stats = brms_data %>%
  group_by(taxon_capture, country) %>%
  summarise(regions = paste(sort(unique(region)), collapse = ", "),
            population_n = n_distinct(population),
            site_n = n_distinct(site),
            year_first = year(min(date)),
            year_last = year(max(date)),
            year_count = n_distinct(year(date)),
            adult_n = sum(life_stage_simple == "adult"),
            adult_n_pos = sum(life_stage_simple == "adult" & bd_detected_co),
            subadult_n = sum(life_stage_simple == "subadult"),
            subadult_n_pos = sum(life_stage_simple == "subadult" & bd_detected_co),
            .groups = "drop") %>%
  arrange(country,
          taxon_capture)

# load models
null_zln = readRDS(file = here("size_bd_unified/unified_models/unified_model_results_zln_null.rds"))
m5g_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_30.rds")

model_select = m5g_zln
```

## Define analysis functions
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show processing & visualization functions"

get_species_name = function(taxon, newline = FALSE) {
  species_name = str_to_sentence(gsub("_", " ", gsub("usgs", "", taxon)))
  if (newline) {
    species_name = gsub(" ", "\n", species_name)
  }
  return(species_name)
}

wilson_ci = function(p_hat, n, conf.level = 0.95) {
  z <- qnorm(1 - (1 - conf.level) / 2)
  
  denominator <- 1 + (z^2 / n)
  centre_adjusted_probability <- p_hat + (z^2 / (2 * n))
  adjusted_standard_deviation <- sqrt((p_hat * (1 - p_hat) / n) + (z^2 / (4 * n^2)))
  
  lower_bound <- (centre_adjusted_probability - z * adjusted_standard_deviation) / denominator
  upper_bound <- (centre_adjusted_probability + z * adjusted_standard_deviation) / denominator
  
  return(c(lower_bound = lower_bound, upper_bound = upper_bound))
}

wilson_ci_lower = function(p_hat, n, conf.level = 0.95) {
  ws = wilson_ci(p_hat, n, conf.level)
  return(ws[1])
}

wilson_ci_upper = function(p_hat, n, conf.level = 0.95) {
  ws = wilson_ci(p_hat, n, conf.level)
  return(ws[2])
}


```

# Results

## Extract draws
```{r}
#| eval: false
#| include: false
# 
# # b intercept
# b_specs = emmeans(model_select,
#                   ~ taxon_capture,
#                   dpar = "mu",
#                   re_formula = NA,
#                   weights = "equal",
#                   ndraws = 500)
# save(b_specs, file = here("size_bd_unified", "intermediates", paste0("bd_model_unified_bspecs", today(), ".RData")))
# 
# # b life stage
# gc()
# b_specs_ls = emmeans(model_select,
#                   ~ life_stage_simple | taxon_capture,
#                   dpar = "mu",
#                   re_formula = NA,
#                   ndraws = 500)
# save(b_specs_ls, file = here("size_bd_unified", "intermediates", paste0("bd_model_unified_bspecs_ls", today(), ".RData")))
# gc()
# 
# # hu intercept
# hu_specs = emmeans(model_select, 
#                    ~ taxon_capture, 
#                    dpar = "hu",
#                    re_formula = NA,
#                    weights = "equal",
#                    ndraws = 500)
# save(hu_specs, file = here("size_bd_unified", "intermediates", paste0("bd_model_unified_huspecs", today(), ".RData")))
# 
# # hu life stage
# hu_specs_ls = emmeans(model_select, 
#                    ~ life_stage_simple | taxon_capture, 
#                    dpar = "hu",
#                    re_formula = NA,
#                    ndraws = 500)
# save(hu_specs_ls, file = here("size_bd_unified", "intermediates", paste0("bd_model_unified_huspecs_ls", today(), ".RData")))
# 
# # sigma intercept
# sigma_specs = emmeans(model_select, 
#                    ~ taxon_capture, 
#                    dpar = "sigma",
#                    re_formula = NA,
#                    weights = "equal",
#                    ndraws = 500)
# save(sigma_specs, file = here("size_bd_unified", "intermediates", paste0("bd_model_unified_sigmaspecs", today(), ".RData")))

```


## Organize draws
```{r}
# load specs
load(here("size_bd_unified", "intermediates", "bd_model_unified_bspecs_2026-03-22.RData"))
load(here("size_bd_unified", "intermediates", "bd_model_unified_bspecs_ls_2026-03-22.RData"))
load(here("size_bd_unified", "intermediates", "bd_model_unified_huspecs_2026-03-22.RData"))
load(here("size_bd_unified", "intermediates", "bd_model_unified_huspecs_ls_2026-03-22.RData"))
load(here("size_bd_unified", "intermediates", "bd_model_unified_sigmaspecs_2026-03-22.RData"))

# hu
hu_contrasts_specs = contrast(hu_specs_ls, method = "revpairwise")

hu_draws = hu_specs %>%
  gather_emmeans_draws() %>%
  dplyr::rename(hu = .value) %>%
  mutate(prevalence = 1 - plogis(hu))

hu_group_draws = hu_specs_ls %>%
  gather_emmeans_draws() %>%
  dplyr::rename(hu = .value) %>%
  mutate(prevalence = 1 - plogis(hu))

hu_contrast_draws = hu_contrasts_specs %>%
  gather_emmeans_draws() %>%
  dplyr::rename(hu_subadult = .value)

hu_contrast_agg = hu_contrast_draws %>%
  group_by(taxon_capture, contrast) %>%  # check!
  summarize(hdi_50 = hdi(hu_subadult, ci = 0.50),
            hdi_89 = hdi(hu_subadult, ci = 0.89),
            hdi_95 = hdi(hu_subadult, ci = 0.95),
            ratio_positive = sum(hu_subadult > 0) / n(),
            hu_subadult_median = median(hu_subadult),
            .groups = "drop") %>%
  mutate(
    hdi_zero_50 = case_when(
      hdi_50[1] > 0 | hdi_50[2] < 0 ~ TRUE,
      .default = FALSE
    ),
    hdi_zero_89 = case_when(
      hdi_89[1] > 0 | hdi_89[2] < 0 ~ TRUE,
      .default = FALSE
    ),
    hdi_zero_95 = case_when(
      hdi_95[1] > 0 | hdi_95[2] < 0 ~ TRUE,
      .default = FALSE
    ))


# b
b_contrasts_specs = contrast(b_specs_ls, method = "revpairwise")

b_draws = b_specs %>%
  gather_emmeans_draws() %>%
  dplyr::rename(b = .value)

b_group_draws = b_specs_ls %>%
  gather_emmeans_draws() %>%
  dplyr::rename(b = .value)

b_contrast_draws = b_contrasts_specs %>%
  gather_emmeans_draws() %>%
  dplyr::rename(b_subadult = .value)

b_contrast_agg = b_contrast_draws %>%
  group_by(taxon_capture) %>%
  summarize(hdi_50 = hdi(b_subadult, ci = 0.50),
            hdi_89 = hdi(b_subadult, ci = 0.89),
            hdi_95 = hdi(b_subadult, ci = 0.95),
            ratio_positive = sum(b_subadult > 0) / n(),
            b_subadult_median = median(b_subadult),
            .groups = "drop") %>%
  mutate(
    hdi_zero_50 = case_when(
      hdi_50[1] > 0 | hdi_50[2] < 0 ~ TRUE,
      .default = FALSE
    ),
    hdi_zero_89 = case_when(
      hdi_89[1] > 0 | hdi_89[2] < 0 ~ TRUE,
      .default = FALSE
    ),
    hdi_zero_95 = case_when(
      hdi_95[1] > 0 | hdi_95[2] < 0 ~ TRUE,
      .default = FALSE
    ))

# sigma
sigma_draws = sigma_specs %>%
  gather_emmeans_draws() %>%
  dplyr::rename(sigma = .value)

# combine
contrast_agg = hu_contrast_agg %>%
  left_join(b_contrast_agg, by = c("taxon_capture"), suffix = c("_hu", "_b")) %>%
  mutate(net_effect_hu = ratio_positive_hu * 2 - 1,
         net_effect_b = ratio_positive_b * 2 - 1,
         species = get_species_name(taxon_capture))

intercept_draws = hu_draws %>%
  left_join(b_draws, by = c(".chain", ".iteration", ".draw", "taxon_capture")) %>%
  left_join(sigma_draws, by = c(".chain", ".iteration", ".draw", "taxon_capture"))

intercept_agg = intercept_draws %>%
  group_by(taxon_capture) %>%
  summarize(prevalence_median = median(prevalence),
            hu_median = median(hu),
            b_median = median(b),
            sigma_median = median(sigma),
            .groups = "drop")

combined_agg = intercept_agg %>%
  left_join(contrast_agg %>%
              select(taxon_capture,
                     hu_subadult_median,
                     b_subadult_median,
                     net_effect_hu,
                     net_effect_b), by = "taxon_capture") %>%
  mutate(species = get_species_name(taxon_capture))

group_draws = b_group_draws %>%
  left_join(hu_group_draws, by = c(".chain", ".iteration", ".draw", 
                                   "life_stage_simple", "taxon_capture"))

group_draws_agg = group_draws %>%
  group_by(taxon_capture,
           life_stage_simple) %>%
  summarize(hu_median = median(hu),
            prevalence_median = median(prevalence),
            b_median = median(b),
            .groups = "drop")

```

## Obsevations and predictions
```{r}

# epred draws
newdata_basic = expand_grid(taxon_capture = sort(unique(brms_data$taxon_capture)),
                            life_stage_simple = c("subadult",
                                                  "adult")) %>%
  mutate(precip_mm_30_z = 0,
         precip_mm_30_z2 = 0,
         temp_c_30_z = 0,
         temp_c_30_z2 = 0)

# way too big, limit to 50 draws!
epred_original = epred_draws(model_select,
                             newdata = brms_data,
                             re_formula = NA,
                             ndraws = 200) %>%
  dplyr::rename(bd_load_density_epred = .epred)

epred_fixed = epred_draws(model_select,
                          newdata = newdata_basic,
                          re_formula = NA) %>%
  dplyr::rename(bd_load_density_epred = .epred)

# observation data
obs_data = brms_data %>%
  group_by(taxon_capture,
           life_stage_simple) %>%
   summarise(n = n(),
             n_pos = sum(bd_detected_co),
             bd_prevalence = sum(bd_detected_co) / n(),
             bd_prevalence_ci_lower = wilson_ci_lower(bd_prevalence, n(), 0.95),
             bd_prevalence_ci_upper = wilson_ci_upper(bd_prevalence, n(), 0.95),
             min_load_inf = min(bd_load_co_density[bd_detected_co > 0]),
             q_25_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.25),
             q_50_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.50),
             q_75_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.75),
             max_load_inf = max(bd_load_co_density[bd_detected_co > 0]),
             se = sqrt(bd_prevalence*(1-bd_prevalence)/n),
             .groups = "drop") %>%
  mutate(label_offset = recode_values(life_stage_simple,
                                          "adult" ~ 0,
                                          "subadult" ~ 1))


```
## Visualize outputs

### Prevalence
```{r}
#| fig-height: 8
#| fig-width: 12

# observed
ggplot(obs_data, aes(y = bd_prevalence, x = life_stage_simple, color = life_stage_simple)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = bd_prevalence_ci_lower, ymax = bd_prevalence_ci_upper),
                width = 0.2,
                position = position_dodge(width = 0.5)) +
  geom_text(aes(label = paste0("n=", n), y = 1 - (1 - label_offset) * 0.05),
          position = position_dodge(width = 0.5),
          size = 4) +
  ylim(0, 1) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# modeled
ggplot(group_draws, aes(y = prevalence, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")


```

### Load
```{r}
#| fig-height: 8
#| fig-width: 12

# observed
ggplot(brms_data %>%
         filter(bd_detected_co > 0), aes(y = bd_load_co_density, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  geom_text(data = obs_data, aes(label = paste0("n=", n_pos), y = 10e8 - (1 - label_offset) * 7 * 10e7),
            position = position_dodge(width = 0.5),
            size = 4) +
  scale_y_log10(limits = c(1, 1e10)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")

ggplot(epred_original, aes(y = bd_load_density_epred, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  # geom_boxplot() +
  scale_y_log10(limits = c(1, 1e10)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")

ggplot(epred_fixed, aes(y = bd_load_density_epred, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  # geom_boxplot() +
  scale_y_log10(limits = c(1, 1e10)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")



```

## Hypothesis testing
```{r}
#| eval: true
#| echo: false
#| warning: false
#| fig-height: 8
#| fig-width: 12


# visualize draws
ggplot(group_draws, aes(x = prevalence, y = b, color = life_stage_simple)) +
  geom_point(alpha = 0.1) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# single parameter distributions
ggplot(group_draws, aes(y = prevalence, x = life_stage_simple, color = life_stage_simple)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

ggplot(group_draws, aes(y = b, x = life_stage_simple, color = life_stage_simple)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# visualize contrasts
ggplot(hu_contrast_draws, aes(x = hu_subadult)) +
  stat_halfeye(fill = "#66c2a5",
               .width = c(0.5, 0.95),
               point_interval = mode_hdi) +
  facet_wrap("taxon_capture") +
  geom_vline(xintercept = 0, alpha = 1) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect on Bd Prevalence",
       x = "Bd Prevalence Subadult Effect Size (logit scale) [-]\nwith 50% & 95% High Density Interval",
       y = "Posterior Density [-]") +
  theme(text=element_text(size=14))

ggplot(b_contrast_draws, aes(x = b_subadult)) +
    stat_halfeye(fill = "#8da0cb",
               .width = c(0.5, 0.95),
               point_interval = mode_hdi) +
  facet_wrap("taxon_capture") +
  geom_vline(xintercept = 0, alpha = 1) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect on Bd Prevalence",
       x = "Bd Load Density Subadult Effect Size (logit scale) [-]\nwith 50% & 95% High Density Interval",
       y = "Posterior Density [-]") +
  theme(text=element_text(size=14))

# probability of positive effect
ggplot(contrast_agg, aes(x = -net_effect_hu, y = net_effect_b)) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  xlim(-1, 1) +
  ylim(-1, 1) +
  theme_minimal() +
  labs(
       title = "Net Subadult Effect (Subadult - Adult)",
       x = "Bd Prevalence (log scale) [-]",
       y = "Bd Load Density (logit scale) [-]") +
  scale_color_brewer(palette = "Set2")

ggplot(combined_agg, aes(x = -hu_subadult_median, y = b_subadult_median)) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  theme_minimal() +
  labs(
       title = "Median Subadult Effect (Subadult - Adult)",
       x = "Bd Prevalence (log scale) [-]",
       y = "Bd Load Density (logit scale) [-]") +
  scale_color_brewer(palette = "Set2")

ggplot(combined_agg, aes(x = prevalence_median, y = sigma_median)) +
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  theme_minimal() +
  labs(
       title = "Prevalence & Aggregation",
       x = "Median Bd Prevalence [-]",
       y = "Median Bd Load Aggregation -- log(sigma) [-]"
       ) +
  scale_color_brewer(palette = "Set2")

ggplot(combined_agg, aes(x = hu_subadult_median, y = hu_median)) +
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  theme_minimal() +
  labs(
       title = "Prevalence",
       x = "Median hu (subadult - adult) [-]",
       y = "Median hu [-]"
       ) +
  scale_color_brewer(palette = "Set2")

```

# correlation analysis
```{r}
simple_cor = cor_mat(combined_agg %>%
                       select(hu_median,
                              sigma_median,
                              hu_subadult_median,
                              b_subadult_median), method = "spearman")

cor_get_pval(simple_cor)


```

## Phylogenetic tree
```{r}
species_list = brms_data %>%
  mutate(sn = get_species_name(taxon_capture)) %>%
  select(sn) %>%
  distinct() %>%
  arrange(sn) 

resolved = gna_verifier(species_list %>%
                          pull(sn),
                        data_source_ids = 4)

species_list_ncbi = species_list %>%
  left_join(resolved %>%
              select(submittedName,
                     matchedCanonicalSimple), by = c("sn" = "submittedName")) %>%
  mutate(genus = str_split_fixed(sn, " ", 2)[,1],
         species = str_split_fixed(sn, " ", 2)[,2],
         umm = sn == matchedCanonicalSimple)

# load sequences
seq_class =
  read.fasta(here("data","amphibia16s_classifier", "Amphibia16S_derep1_seqs_extracted",
                               list.files(here("data","amphibia16s_classifier", "Amphibia16S_derep1_seqs_extracted")), "data", "dna-sequences.fasta")) %>%
  dplyr::rename(taxon_id = seq.name) %>%
  arrange(taxon_id) %>%
  mutate(bp = str_length(seq.text),
         gene = "16S")

taxa_class = read_tsv(here("data", "amphibia16s_classifier", "Amphibia16S_derep1_taxa_extracted",
                               list.files(here("data", "amphibia16s_classifier", "Amphibia16S_derep1_taxa_extracted")), "data", "taxonomy.tsv")) %>%
  rename(taxon_id = `Feature ID`,
         taxon = Taxon) %>%
  separate(taxon, into = c("k", "p", "c", "o", "f", "g", "s"), sep = ";") %>%
  mutate(across(everything(), ~ str_remove(., "^[a-z]__"))) %>%
  rename(kingdom = k,
         phylum = p,
         class = c,
         order = o,
         family = f,
         genus = g,
         species = s) %>%
  mutate(domain = "Eukaryota") %>%
  arrange(taxon_id) %>%
  select(taxon_id,
         domain,
         everything())

taxa_class_summary = taxa_class %>%
  group_by(genus, species) %>%
  count()

compare_taxa_lists = species_list %>%
  left_join(taxa_class_summary, by = c("genus", "species"))

##


gene_marker <- "16S"
species_list = bd_model_env_stats %>%
  mutate(species = get_species_name(taxon_capture)) %>%
  pull(species)

fetch_sequence_reproducible <- function(species, marker, max_seqs = 10, 
                                        min_length = 100,
                                        target_min = 400,
                                        target_max = 700) {
  query <- paste0(species, "[Organism] AND (", marker, "[Title] OR ", 
                  marker, "[All Fields])")
  search <- entrez_search(db = "nucleotide", term = query, retmax = max_seqs)
  
  if (length(search$ids) == 0) {
    message("No sequence found for: ", species)
    return(NULL)
  }
  
  seqs <- entrez_fetch(db = "nucleotide", id = search$ids, rettype = "fasta")
  
  tmp <- tempfile(fileext = ".fasta")
  writeLines(seqs, tmp)
  
  dna <- tryCatch(
    read.dna(tmp, format = "fasta"),
    error = function(e) { message("Failed to parse FASTA for: ", species); return(NULL) }
  )
  
  if (is.null(dna)) return(NULL)
  if (is.matrix(dna)) dna <- as.list(dna)
  
  # Filter by minimum length
  lengths <- sapply(dna, length)
  dna <- dna[lengths >= min_length]
  
  if (length(dna) == 0) {
    message("No sequences above min_length for: ", species)
    return(NULL)
  }
  
  lengths <- sapply(dna, length)
  
  # Try preferred length range first
  in_range <- dna[lengths >= target_min & lengths <= target_max]
  
  if (length(in_range) > 0) {
    range_lengths <- sapply(in_range, length)
    selected <- in_range[which.max(range_lengths)]
    label <- "preferred range"
  } else {
    # Fall back to longest available
    selected <- dna[which.max(lengths)]
    label <- "fallback longest"
  }
  
  # Log the selected accession for reproducibility
  message(species, " [", label, "]: ", 
          sapply(selected, length), "bp — ", names(selected))
  
  return(selected)
}

# Run and capture selection log
raw_seqs <- lapply(species_list, function(sp) {
  Sys.sleep(0.4)
  fetch_sequence_reproducible(sp, marker = gene_marker, max_seqs = 50)
})
names(raw_seqs) <- species_list
raw_seqs <- Filter(Negate(is.null), raw_seqs)

# After fetching, decode and save immediately
decode_dnabin <- function(x) {
  tmp <- tempfile(fileext = ".fasta")
  write.dna(x, tmp, format = "fasta")
  lines <- readLines(tmp)
  seq_lines <- lines[!grepl("^>", lines)]
  seq_string <- paste(seq_lines, collapse = "")
  gsub("[[:space:]]", "", seq_string)
}

# Decode all sequences
combined_char <- sapply(names(raw_seqs), function(sp) {
  decode_dnabin(raw_seqs[[sp]])
})
names(combined_char) <- names(raw_seqs)

# Sanity check before alignment
substr(combined_char[1], 1, 50)  # should show ACGTs now

# Save as FASTA for reuse — no need to re-fetch next time
fasta_lines <- unlist(lapply(names(combined_char), function(sp) {
  c(paste0(">", sp), combined_char[[sp]])
}))
writeLines(fasta_lines, "selected_sequences.fasta")
message("Sequences saved to selected_sequences.fasta")

# To reload next time instead of re-fetching:
# dna_saved <- read.dna("selected_sequences.fasta", format = "fasta")
# combined_char <- sapply(as.character(dna_saved), paste, collapse = "")
# names(combined_char) <- names(raw_seqs)

# Align
aligned <- AlignSeqs(DNAStringSet(combined_char))
# 
# fetch_all_sequences <- function(species, marker, max_seqs = 50, min_length = 100) {
#   query <- paste0(species, "[Organism] AND (", marker, "[Title] OR ", marker, "[All Fields])")
#   search <- entrez_search(db = "nucleotide", term = query, retmax = max_seqs)
#   
#   if (length(search$ids) == 0) {
#     message("No sequences found for: ", species)
#     return(NULL)
#   }
#   
#   message(species, ": ", length(search$ids), " sequences found")
#   
#   seqs <- entrez_fetch(db = "nucleotide", id = search$ids, rettype = "fasta")
#   
#   tmp <- tempfile(fileext = ".fasta")
#   writeLines(seqs, tmp)
#   
#   dna <- tryCatch(
#     read.dna(tmp, format = "fasta"),
#     error = function(e) { message("Failed to parse FASTA for: ", species); return(NULL) }
#   )
#   
#   if (is.null(dna)) return(NULL)
#   
#   # Diagnose structure before converting
#   message(species, ": class=", class(dna), " dim=", 
#           paste(dim(dna), collapse="x"), " length=", length(dna))
#   
#   # Convert to list safely depending on structure
#   if (!is.null(dim(dna))) {
#     # It's a matrix — convert row by row preserving DNAbin class
#     dna_list <- lapply(seq_len(nrow(dna)), function(i) dna[i, , drop = FALSE])
#     names(dna_list) <- rownames(dna)
#   } else if (is.list(dna)) {
#     dna_list <- dna
#   } else {
#     # Single sequence
#     dna_list <- list(dna)
#     names(dna_list) <- "seq1"
#   }
#   
#   # Filter by minimum length
#   lengths <- sapply(dna_list, length)
#   dna_list <- dna_list[lengths >= min_length]
#   
#   if (length(dna_list) == 0) {
#     message("No sequences above min_length for: ", species)
#     return(NULL)
#   }
#   
#   message(species, ": ", length(dna_list), " sequences retained after length filter")
#   return(dna_list)
# }
# 
# # Fetch for all species
# raw_seqs_all <- lapply(species_list, function(sp) {
#   Sys.sleep(0.4)
#   fetch_all_sequences(sp, marker = gene_marker, max_seqs = 50)
# })
# names(raw_seqs_all) <- species_list
# raw_seqs_all <- Filter(Negate(is.null), raw_seqs_all)
# message(length(raw_seqs_all), " / ", length(species_list), " species retrieved")
# 
# ## checks for length
# # Check length distribution across all sequences in raw_seqs_all
# lengths_all <- lapply(names(raw_seqs_all), function(sp) {
#   seqs <- raw_seqs_all[[sp]]
#   lengths <- sapply(seqs, length)
#   data.frame(species = sp, length = lengths, title = names(seqs))
# })
# 
# lengths_df <- do.call(rbind, lengths_all)
# 
# # Summary
# summary(lengths_df$length)
# 
# # Distribution plot
# hist(lengths_df$length, breaks = 50, 
#      main = "Sequence length distribution across all species",
#      xlab = "Length (bp)")
# 
# # Per species summary
# lengths_df |> 
#   dplyr::group_by(species) |> 
#   dplyr::summarise(n = n(), min = min(length), median = median(length), max = max(length)) |>
#   print(n = Inf)
# 
# 
# select_by_length_range <- function(seqs, species_name,
#                                    preferred_min = 400, preferred_max = 700,
#                                    fallback_min  = 200, fallback_max  = 2000) {
#   lengths <- sapply(seqs, length)
#   
#   # First try preferred range
#   in_preferred <- lengths >= preferred_min & lengths <= preferred_max
#   
#   if (sum(in_preferred) > 0) {
#     candidates <- seqs[in_preferred]
#     label <- "preferred range"
#   } else {
#     # Fall back to broader range
#     in_fallback <- lengths >= fallback_min & lengths <= fallback_max
#     candidates <- seqs[in_fallback]
#     label <- "fallback range"
#     if (sum(in_fallback) == 0) {
#       candidates <- seqs  # last resort: use anything
#       label <- "no range filter"
#     }
#   }
#   
#   # Among candidates, take longest
#   cand_lengths <- sapply(candidates, length)
#   best_idx <- which.max(cand_lengths)
#   
#   message(species_name, " [", label, "]: ", 
#           cand_lengths[best_idx], "bp — ", names(candidates)[best_idx])
#   
#   tmp <- tempfile(fileext = ".fasta")
#   write.dna(candidates[best_idx], tmp, format = "fasta")
#   lines <- readLines(tmp)
#   seq_string <- paste(lines[!grepl("^>", lines)], collapse = "")
#   gsub("[[:space:]]", "", seq_string)
# }
# 
# selected_seqs <- mapply(select_by_length_range,
#                         raw_seqs_all,
#                         names(raw_seqs_all),
#                         SIMPLIFY = FALSE)

# # Now combined_char is already a named list of plain strings
# combined_char <- setNames(unlist(raw_seqs), names(raw_seqs))
# 
# # Remove any whitespace from sequences
# combined_char <- gsub("[[:space:]]", "", combined_char)
# 
# # Sanity check
# substr(combined_char[1], 1, 50)
# 
# # Now align
# aligned <- AlignSeqs(DNAStringSet(combined_char))

# --- 3. BUILD TREE ---
phydat <- as.phyDat(aligned)

dist_mat <- dist.ml(phydat, model = "JC69")

nj_tree <- NJ(dist_mat)

fit <- pml(nj_tree, phydat)
fit_opt <- optim.pml(fit,
                     model         = "GTR",
                     optInv        = TRUE,
                     optGamma      = TRUE,
                     rearrangement = "NNI")

tr_phylo <- fit_opt$tree

# --- 4. ROOT ---
tr_rooted <- midpoint(tr_phylo)

# --- 5. QUICK SANITY CHECKS ---
summary(tr_rooted$edge.length)
is.ultrametric(tr_rooted)  # expected FALSE for ML tree
plot(tr_rooted, cex = 0.6)

# Optional: compare topology to your original OTL tree
ggtree::ggtree(tr_rooted, layout = "rectangular") +
  ggtree::geom_tiplab(size = 3, align = FALSE) +
  ggtree::theme_tree2() +
  scale_x_continuous(expand = expansion(mult = c(0, 0.3)))

# Now has real branch lengths!
summary(tr_rooted$edge.length)
is.ultrametric(tr_rooted)

hu_sub_effect_named <- setNames(-combined_agg$net_effect_hu, combined_agg$species)
b_sub_effect_named  <- setNames(combined_agg$net_effect_b,  combined_agg$species)
hu_sub_med_named <- setNames(-combined_agg$hu_subadult_median, combined_agg$species)
b_sub_med_named  <- setNames(combined_agg$b_subadult_median,  combined_agg$species)
hu_med_named <- setNames(-combined_agg$hu_median, combined_agg$species)
b_med_named  <- setNames(combined_agg$b_median,  combined_agg$species)
prev_med_named <- setNames(combined_agg$prevalence_median, combined_agg$species)
sigma_med_named  <- setNames(combined_agg$sigma_median,  combined_agg$species)

blomberg_k_hu_sub_effect = phylosig(tr_rooted, hu_sub_effect_named, method = "K", test = TRUE, nsim = 1000)
blomberg_k_b_sub_effect = phylosig(tr_rooted, b_sub_effect_named, method = "K", test = TRUE, nsim = 1000)
blomberg_k_hu_sub_med = phylosig(tr_rooted, hu_sub_med_named, method = "K", test = TRUE, nsim = 1000)
blomberg_k_b_sub_med = phylosig(tr_rooted, b_sub_med_named, method = "K", test = TRUE, nsim = 1000)
blomberg_k_hu_med = phylosig(tr_rooted, hu_med_named, method = "K", test = TRUE, nsim = 1000)
blomberg_k_b_med = phylosig(tr_rooted, b_med_named, method = "K", test = TRUE, nsim = 1000)
blomberg_k_prev_med = phylosig(tr_rooted, prev_med_named, method = "K", test = TRUE, nsim = 1000)
blomberg_k_sigma_med = phylosig(tr_rooted, sigma_med_named, method = "K", test = TRUE, nsim = 1000)

tr_rooted$tip.label

# plots
hu_sub_effect_map = contMap(tr_rooted, hu_sub_effect_named,  title = "hu net effect", fsize = 0.6)
b_sub_effect_map = contMap(tr_rooted, b_sub_effect_named,   title = "b net_effect", fsize = 0.6)
hu_sub_med_map = contMap(tr_rooted, hu_sub_med_named,  title = "hu net median", fsize = 0.6)
b_sub_med_map = contMap(tr_rooted, b_sub_med_named,   title = "b net_median", fsize = 0.6)
hu_med_map = contMap(tr_rooted, hu_med_named,  title = "hu median", fsize = 0.6)
b_med_map = contMap(tr_rooted, b_med_named,   title = "b median", fsize = 0.6)
prev_med_map = contMap(tr_rooted, prev_med_named,  title = "prevalence median effect", fsize = 0.6)
sigma_med_map = contMap(tr_rooted, sigma_med_named,   title = "sigma median effect", fsize = 0.6)



hu_sub_effect_map <- setMap(hu_sub_effect_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))
b_sub_effect_map <- setMap(b_sub_effect_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))
hu_sub_med_map <- setMap(hu_sub_med_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))
b_sub_med_map <- setMap(b_sub_med_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))
hu_med_map <- setMap(hu_med_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))
b_med_map <- setMap(b_med_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))
prev_med_map <- setMap(prev_med_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))
sigma_med_map <- setMap(sigma_med_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))

plot(hu_sub_effect_map)
plot(b_sub_effect_map)
plot(hu_sub_med_map)
plot(b_sub_med_map)
plot(hu_med_map)
plot(b_med_map)
plot(prev_med_map)
plot(sigma_med_map)

```

## Collective results
```{r}
#| eval: true
#| echo: true
#| output: false
#| code-fold: true
#| code-summary: Show code

# m5g_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2.rds")

choice_model = m5g_zln

model_data = brms_data

draws_all = collect_draws(choice_model) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

density_mat_all = density_mat(log(0.5), 20, 100, draws_all) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

sim_load_all = simulate_bd_load(draws_all, 10, 1000) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

obs_stats_all = sample_stats(model_data) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

```

::: {.panel-tabset}
### Observed Prevalence
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# observed prevalence
ggplot(obs_stats_all, aes(y = bd_prevalence, x = species, color = life_stage_simple)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = bd_prevalence_ci_lower, ymax = bd_prevalence_ci_upper),
                width = 0.2,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  ggtitle("Observed prevalence") +
  xlab("Species") +
  ylab("Bd prevalence [-]\n(with 95% CI)") +
  labs(color = "Life stage") +
  geom_text(aes(label = paste0("n=", n),  y = label_y),
          position = position_dodge(width = 0.5)) +
  ylim(0, 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Modeled Prevalence 
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# model prevalence
ggplot(draws_all, aes(prevalence, x = species, color = life_stage_simple)) +
  geom_boxplot(coef = NULL) + 
  theme_minimal() +
  ggtitle("Modeled prevalence") +
  xlab("Species") +
  ylab("Bd prevalence [-]\n") +
  labs(color = "Life stage") +
  ylim(0, 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Observed Bd load
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# obsevation bd load
ggplot(model_data %>%
         mutate(species = get_species_name(taxon_capture, newline = TRUE)) %>%
         filter(bd_load_co > 0), aes(log_bd_load_co_density, x = species, color = life_stage_simple)) +
  geom_boxplot() + 
  theme_minimal() +
  ggtitle("Observed Bd-positive load") +
  xlab("Species") +
  ylab("log(Bd load density) [-]") +
  labs(color = "Life stage") +
  geom_text(
    data = obs_stats_all,
    aes(
      x = species,
      y = max(model_data$log_bd_load_co_density) * label_y * 1.05,
      label = paste0("n=", n_pos),
      group = life_stage_simple
    ),
    position = position_dodge(width = 0.75),
    vjust = 0
  ) +
  ylim(log(0.5), 19) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Modeled Bd load
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# modeled bd load
ggplot(sim_load_all %>%
         filter(log_bd_load_co_density > 0.5), aes(log_bd_load_co_density, x = species, color = life_stage_simple)) +
  geom_boxplot() + 
  theme_minimal() +
  ggtitle("Modeled Bd-positive load") +
  xlab("Species") +
  ylab("log(Bd load density) [-]") +
  labs(color = "Life stage") +
  ylim(log(0.5), 20) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Modeled Bd load PDF
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# PDF -- facet by species
ggplot(density_mat_all, aes(x = density,
                            y = pdf_median,
                            color = life_stage_simple, 
                            fill = life_stage_simple)) +
  geom_ribbon(aes(ymin = pdf_lower,
                  ymax = pdf_upper), 
              alpha = 0.15,
              color = NA) +
  geom_line() +
  theme_minimal() +
  ggtitle("Modeled PDF of Bd-positive load") +
  labs(color = "Life stage",
       fill = "Life stage") +
  xlab("log(Bd load) [-]") +
  ylab("Median probability density [-]\n(with 25th to 75th percentiles)") +
  scale_x_continuous(limits = c(log(0.5), 20), expand = c(0, 0)) +
  facet_grid(. ~ gsub(" ", "\n", species))

# # PDF -- facet by life_stage & season
# ggplot(density_mat_all, aes(x = density,
#                             y = pdf_median,
#                             color = taxon_capture, 
#                             fill = taxon_capture)) +
#   geom_ribbon(aes(ymin = pdf_lower,
#                   ymax = pdf_upper), 
#               alpha = 0.15,
#               color = NA) +
#   geom_line() +
#   theme_minimal() +
#   ggtitle("Modeled PDF of Bd-positive load") +
#   labs(color = "Species",
#        fill = "Species") +
#   xlab("log(Bd load) [-]") +
#   ylab("Median probability density [-]\n(with 25th to 75th percentiles)") +
#   scale_x_continuous(limits = c(log(0.3), 20), expand = c(0, 0)) +
#   facet_grid(life_stage_simple ~ season)
```

### Moded Bd load CCDF
```{r}
#| echo: false
#| warning: false
#| fig-height: 8
#| fig-width: 8
# CCDF -- facet by species
ggplot(density_mat_all, aes(x = density,
                            y = ccdf_median,
                            color = life_stage_simple, 
                            fill = life_stage_simple)) +
  geom_ribbon(aes(ymin = ccdf_lower,
                  ymax = ccdf_upper), 
              alpha = 0.15,
              color = NA) +
  geom_line() +
  theme_minimal() +
  ggtitle("Modeled CCDF of Bd-positive load") +
  labs(color = "Life stage",
       fill = "Life stage") +
  xlab("log(Bd load) [-]") +
  ylab("Median complementary cumulative density (1 - CDF) [-]\n(with 25th to 75th percentiles)") +
  scale_x_continuous(limits = c(log(0.5), 12), expand = c(0, 0)) +
  ylim(0, 0.75) +
  # facet_grid(. ~ gsub(" ", "\n", species))
  facet_wrap("species")

# CCDF -- facet by life_stage & season
# ggplot(density_mat_all, aes(x = density,
#                             y = ccdf_median,
#                             color = taxon_capture,
#                             fill = taxon_capture)) +
#   geom_ribbon(aes(ymin = ccdf_lower,
#                   ymax = ccdf_upper),
#               alpha = 0.15,
#               color = NA) +
#   geom_line() +
#   theme_minimal() +
#   ggtitle("Modeled CCDF of Bd-positive load") +
#   labs(color = "Species",
#        fill = "Species") +
#   xlab("log(Bd load) [-]") +
#   ylab("Median complementary cumulative density (1 - CDF) [-]\n(with 25th to 75th percentiles)") +
#   scale_x_continuous(limits = c(log(0.3), 20), expand = c(0, 0)) +
#   ylim(0, 1) +
#   facet_grid(. ~ life_stage_simple)
```

# Environmental conditions

## Distributions
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Environmental conditions distributions code"
#| fig-height: 6
#| fig-width: 8

# 15-day
# ggplot(brms_data, aes(x = temp_c_15_day_mean, y = precip_mm_15_day_sum, color = life_stage_simple)) +
#   geom_point(alpha = 0.1) +
#   stat_ellipse() +
#   facet_wrap(. ~ taxon_capture) +
#   xlim(-5, 40) +
#   ylim(0, 200) +
#   labs(x = "15-day temperature mean (degC)", 
#        y = "15-day precipitation total (mm)",
#        title = "Temperature v. precipitation") +
#   theme_minimal()

# 30-day by species
ggplot(brms_data, aes(x = temp_c_30_day_mean, y = precip_mm_30_day_sum, color = life_stage_simple)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  facet_wrap(. ~ taxon_capture) +
  xlim(-5, 40) +
  ylim(0, 400) +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()

# 30 day collective
ggplot(brms_data, aes(x = temp_c_30_day_mean, y = precip_mm_30_day_sum, color = taxon_capture)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()

# 30 day collective, scaled to climate by species
ggplot(brms_data, aes(x = temp_c_30_sp_z, y = precip_mm_30_sp_z, color = taxon_capture)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()


# ggplot(brms_data, aes(x = temp_c_30_sp_z2, y = precip_mm_30_sp_z2, color = life_stage_simple)) +
#   geom_point(alpha = 0.1) +
#   stat_ellipse() +
#   facet_wrap(. ~ taxon_capture) +
#   labs(x = "30-day temperature mean (degC)", 
#        y = "30-day precipitation total (mm)",
#        title = "Temperature v. precipitation") +
#   theme_minimal()
```

## Tempurature effects
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show temperature effects code"
#| fig-height: 6
#| fig-width: 8

select_model = m5f_zln
# select_model = m6_zln

climate_draws = spread_draws(select_model, `b_temp_c.*|b_precip_mm.*|b_hu_temp_c.*|b_hu_precip_mm.*|b_Intercept|b_hu_Intercept`, regex = TRUE)

# Temperature dependence
temp_range = seq(-5, 45, length.out = 100)

temp_grid = expand_grid(temp_c_15_day_mean = temp_range, .draw = 1:nrow(climate_draws)) %>%
  mutate(temp_c_30_day_mean = temp_c_15_day_mean) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    temp_c_15_z = (temp_c_15_day_mean - brms_scales$temp_c_15_day_mean_mean) / brms_scales$temp_c_15_day_mean_sd,
    temp_c_15_z2 = (temp_c_15_day_mean^2 - brms_scales$temp_c_15_day_mean2_mean) / brms_scales$temp_c_15_day_mean2_sd,
    temp_c_30_z = (temp_c_30_day_mean - brms_scales$temp_c_30_day_mean_mean) / brms_scales$temp_c_30_day_mean_sd,
    temp_c_30_z2 = (temp_c_30_day_mean^2 - brms_scales$temp_c_30_day_mean2_mean) / brms_scales$temp_c_30_day_mean2_sd,
    b_pred = b_temp_c_30_z * temp_c_30_z + b_temp_c_30_z2 * temp_c_30_z2 + b_Intercept,
    hu_pred = b_hu_temp_c_15_z * temp_c_15_z + b_hu_temp_c_15_z2 * temp_c_15_z2 + b_hu_Intercept
  ) %>%
  group_by(temp_c_15_day_mean) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  ) %>%
  dplyr::rename(temp_c = temp_c_15_day_mean)

# prevalence
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day mean temperature (degC)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. temperature") +
  theme_minimal()

# load
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day mean temperature (degC)", 
       y = "log(+ Bd load)",
       title = "Bd load v. temperature") +
  theme_minimal()
```

## Precipitation effects
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show precipitation effects code"
#| fig-height: 6
#| fig-width: 8
# precipitation depencence

daily_precip_range = seq(0, 20, length.out = 100)

precip_grid = expand_grid(daily_precip = daily_precip_range, .draw = 1:nrow(climate_draws)) %>%
  mutate(precip_mm_15_day_sum = daily_precip * 15,
         precip_mm_30_day_sum = daily_precip * 30) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    precip_mm_15_z = (precip_mm_15_day_sum - brms_scales$precip_mm_15_day_sum_mean) / brms_scales$precip_mm_15_day_sum_sd,
    precip_mm_15_z2 = (precip_mm_15_day_sum^2 - brms_scales$precip_mm_15_day_sum2_mean) / brms_scales$precip_mm_15_day_sum2_sd,
    precip_mm_30_z = (precip_mm_30_day_sum - brms_scales$precip_mm_30_day_sum_mean) / brms_scales$precip_mm_30_day_sum_sd,
    precip_mm_30_z2 = (precip_mm_30_day_sum^2 - brms_scales$precip_mm_30_day_sum2_mean) / brms_scales$precip_mm_30_day_sum2_sd,
    b_pred = b_precip_mm_30_z * precip_mm_30_z + b_precip_mm_30_z2 * precip_mm_30_z2 + b_Intercept,
    hu_pred = b_hu_precip_mm_15_z * precip_mm_15_z + b_hu_precip_mm_15_z2 * precip_mm_15_z2 + b_hu_Intercept
  ) %>%
  group_by(daily_precip,
           precip_mm_15_day_sum,
           precip_mm_30_day_sum) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  )

# prevalence
ggplot(precip_grid, aes(x = precip_mm_15_day_sum)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day total precipitation (mm)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. precipitation") +
  theme_minimal()

# load
ggplot(precip_grid, aes(x = precip_mm_30_day_sum)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day total precipitation (mm)", 
       y = "log(+ Bd load)",
       title = "Bd load v. precipitation") +
  theme_minimal()

```

## Tempurature effects (scaled by species climate)
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show temperature effects code"
#| fig-height: 6
#| fig-width: 8


select_model = m6_zln

# fixed effects
climate_draws = spread_draws(select_model, `b_temp_c.*|b_precip_mm.*|b_hu_temp_c.*|b_hu_precip_mm.*|b_Intercept|b_hu_Intercept`, regex = TRUE)

# Temperature dependence
temp_range = seq(-3, 3, length.out = 100)

# taxa = brms_data %>%
#   select(taxon_capture) %>%
#   distinct() %>%
#   pull(taxon_capture) %>%
#   sort()

## Stumped by how to back-calculate the temp_z2 values from the temp_z values... I think this has to be done on a species-by-species basis...

temp_grid = expand_grid(temp_z_15_mean = temp_range,
                        .draw = 1:nrow(climate_draws)) %>%
  mutate(temp_z_30_mean = temp_z_15_mean) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    temp_c_15_z = (temp_c_15_day_mean - brms_scales$temp_c_15_day_mean_mean) / brms_scales$temp_c_15_day_mean_sd,
    temp_c_15_z2 = (temp_c_15_day_mean^2 - brms_scales$temp_c_15_day_mean2_mean) / brms_scales$temp_c_15_day_mean2_sd,
    temp_c_30_z = (temp_c_30_day_mean - brms_scales$temp_c_30_day_mean_mean) / brms_scales$temp_c_30_day_mean_sd,
    temp_c_30_z2 = (temp_c_30_day_mean^2 - brms_scales$temp_c_30_day_mean2_mean) / brms_scales$temp_c_30_day_mean2_sd,
    b_pred = b_temp_c_30_sp_z * temp_c_30_z + b_temp_c_30_sp_z2 * temp_c_30_z2 + b_Intercept,
    hu_pred = b_hu_temp_c_15_sp_z * temp_c_15_z + b_hu_temp_c_15_sp_z2 * temp_c_15_z2 + b_hu_Intercept
  ) %>%
  group_by(temp_c_15_day_mean) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  ) %>%
  dplyr::rename(temp_c = temp_c_15_day_mean)

# prevalence
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day mean temperature (degC)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. temperature") +
  theme_minimal()

# load
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day mean temperature (degC)", 
       y = "log(+ Bd load)",
       title = "Bd load v. temperature") +
  theme_minimal()
```

## Precipitation effects (scaled by species climate)
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show precipitation effects code"
#| fig-height: 6
#| fig-width: 8
# precipitation depencence

daily_precip_range = seq(0, 20, length.out = 100)

precip_grid = expand_grid(daily_precip = daily_precip_range, .draw = 1:nrow(climate_draws)) %>%
  mutate(precip_mm_15_day_sum = daily_precip * 15,
         precip_mm_30_day_sum = daily_precip * 30) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    precip_mm_15_z = (precip_mm_15_day_sum - brms_scales$precip_mm_15_day_sum_mean) / brms_scales$precip_mm_15_day_sum_sd,
    precip_mm_15_z2 = (precip_mm_15_day_sum^2 - brms_scales$precip_mm_15_day_sum2_mean) / brms_scales$precip_mm_15_day_sum2_sd,
    precip_mm_30_z = (precip_mm_30_day_sum - brms_scales$precip_mm_30_day_sum_mean) / brms_scales$precip_mm_30_day_sum_sd,
    precip_mm_30_z2 = (precip_mm_30_day_sum^2 - brms_scales$precip_mm_30_day_sum2_mean) / brms_scales$precip_mm_30_day_sum2_sd,
    b_pred = b_precip_mm_30_z * precip_mm_30_z + b_precip_mm_30_z2 * precip_mm_30_z2 + b_Intercept,
    hu_pred = b_hu_precip_mm_15_z * precip_mm_15_z + b_hu_precip_mm_15_z2 * precip_mm_15_z2 + b_hu_Intercept
  ) %>%
  group_by(daily_precip,
           precip_mm_15_day_sum,
           precip_mm_30_day_sum) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  )

# prevalence
ggplot(precip_grid, aes(x = precip_mm_15_day_sum)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day total precipitation (mm)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. precipitation") +
  theme_minimal()

# load
ggplot(precip_grid, aes(x = precip_mm_30_day_sum)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day total precipitation (mm)", 
       y = "log(+ Bd load)",
       title = "Bd load v. precipitation") +
  theme_minimal()

```