---
title: "size_bd_unified_analysis"
format: html
---

# Load data & models
```{r}
librarian::shelf(tidyverse, here, janitor, brms, tidybayes, ggrepel, emmeans, rstatix, phylotools, rentrez, DECIPHER, phangorn, phytools, RColorBrewer)

# load bd_model_env
load(here("data", "bd_model_env_unified_2026-03-22.RData")) 

bd_model_env_stats = brms_data %>%
  group_by(taxon_capture, country) %>%
  summarise(regions = paste(sort(unique(region)), collapse = ", "),
            population_n = n_distinct(population),
            site_n = n_distinct(site),
            year_first = year(min(date)),
            year_last = year(max(date)),
            year_count = n_distinct(year(date)),
            adult_n = sum(life_stage_simple == "adult"),
            adult_n_pos = sum(life_stage_simple == "adult" & bd_detected_co),
            subadult_n = sum(life_stage_simple == "subadult"),
            subadult_n_pos = sum(life_stage_simple == "subadult" & bd_detected_co),
            .groups = "drop") %>%
  arrange(country,
          taxon_capture)

# load models
null_zln = readRDS(file = here("size_bd_unified/unified_models/unified_model_results_zln_null.rds"))
m5g_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_30.rds")

model_select = m5g_zln
```

## Define analysis functions
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show processing & visualization functions"

get_species_name = function(taxon, newline = FALSE) {
  species_name = str_to_sentence(gsub("_", " ", gsub("usgs", "", taxon)))
  if (newline) {
    species_name = gsub(" ", "\n", species_name)
  }
  return(species_name)
}

wilson_ci = function(p_hat, n, conf.level = 0.95) {
  z <- qnorm(1 - (1 - conf.level) / 2)
  
  denominator <- 1 + (z^2 / n)
  centre_adjusted_probability <- p_hat + (z^2 / (2 * n))
  adjusted_standard_deviation <- sqrt((p_hat * (1 - p_hat) / n) + (z^2 / (4 * n^2)))
  
  lower_bound <- (centre_adjusted_probability - z * adjusted_standard_deviation) / denominator
  upper_bound <- (centre_adjusted_probability + z * adjusted_standard_deviation) / denominator
  
  return(c(lower_bound = lower_bound, upper_bound = upper_bound))
}

wilson_ci_lower = function(p_hat, n, conf.level = 0.95) {
  ws = wilson_ci(p_hat, n, conf.level)
  return(ws[1])
}

wilson_ci_upper = function(p_hat, n, conf.level = 0.95) {
  ws = wilson_ci(p_hat, n, conf.level)
  return(ws[2])
}


```

# Results

## Extract draws
```{r}
#| eval: false
#| include: false
# 
# # b intercept
# b_specs = emmeans(model_select,
#                   ~ taxon_capture,
#                   dpar = "mu",
#                   re_formula = NA,
#                   weights = "equal",
#                   ndraws = 500)
# save(b_specs, file = here("size_bd_unified", "intermediates", paste0("bd_model_unified_bspecs", today(), ".RData")))
# 
# # b life stage
# gc()
# b_specs_ls = emmeans(model_select,
#                   ~ life_stage_simple | taxon_capture,
#                   dpar = "mu",
#                   re_formula = NA,
#                   ndraws = 500)
# save(b_specs_ls, file = here("size_bd_unified", "intermediates", paste0("bd_model_unified_bspecs_ls", today(), ".RData")))
# gc()
# 
# # hu intercept
# hu_specs = emmeans(model_select, 
#                    ~ taxon_capture, 
#                    dpar = "hu",
#                    re_formula = NA,
#                    weights = "equal",
#                    ndraws = 500)
# save(hu_specs, file = here("size_bd_unified", "intermediates", paste0("bd_model_unified_huspecs", today(), ".RData")))
# 
# # hu life stage
# hu_specs_ls = emmeans(model_select, 
#                    ~ life_stage_simple | taxon_capture, 
#                    dpar = "hu",
#                    re_formula = NA,
#                    ndraws = 500)
# save(hu_specs_ls, file = here("size_bd_unified", "intermediates", paste0("bd_model_unified_huspecs_ls", today(), ".RData")))
# 
# # sigma intercept
# sigma_specs = emmeans(model_select, 
#                    ~ taxon_capture, 
#                    dpar = "sigma",
#                    re_formula = NA,
#                    weights = "equal",
#                    ndraws = 500)
# save(sigma_specs, file = here("size_bd_unified", "intermediates", paste0("bd_model_unified_sigmaspecs", today(), ".RData")))

```


## Organize draws

### Linpred Draws
```{r}

nd_spp_lss = model_select$data |>
  distinct(taxon_capture, life_stage_simple)


# hu contrast
hu_spp_lss_draws = model_select |>
  linpred_draws(
    newdata    = nd_spp_lss,
    dpar       = "hu",
    re_formula = NA
  )

hu_spp_draws = hu_spp_lss_draws |>
  group_by(taxon_capture, .draw) |>
  summarise(hu = mean(hu), .groups = "drop")

hu_spp_lss_contrast_draws = hu_spp_lss_draws |>
  pivot_wider(
    id_cols     = c(taxon_capture, .draw),
    names_from  = life_stage_simple,
    values_from = hu
  ) |>
  mutate(hu_contrast = subadult - adult,
         contrast = "subadult - adult")

# mu 
mu_spp_lss_draws = model_select |>
  linpred_draws(
    newdata    = nd_spp_lss,
    dpar       = "mu",
    re_formula = NA
  )

mu_spp_lss_contrast_draws = mu_spp_lss_draws |>
  pivot_wider(
    id_cols     = c(taxon_capture, .draw),
    names_from  = life_stage_simple,
    values_from = mu
  ) |>
  mutate(mu_contrast = subadult - adult,
         contrast = "subadult - adult")

# sigma 
sigma_spp_lss_draws = model_select |>
  linpred_draws(
    newdata    = nd_spp_lss,
    dpar       = "sigma",
    re_formula = NA
  )

sigma_spp_draws = sigma_spp_lss_draws |>
  group_by(taxon_capture, .draw) |>
  summarise(sigma = mean(sigma), .groups = "drop")

f_draw_agg <- function(draws, group_vars, agg_var) {
  
  draws |>
    group_by(across(all_of(group_vars))) |>
    reframe(
      "{agg_var}_lower_50" := tidybayes::hdi(.data[[agg_var]], .width = 0.50)[, 1],
      "{agg_var}_upper_50" := tidybayes::hdi(.data[[agg_var]], .width = 0.50)[, 2],
      "{agg_var}_lower_95" := tidybayes::hdi(.data[[agg_var]], .width = 0.95)[, 1],
      "{agg_var}_upper_95" := tidybayes::hdi(.data[[agg_var]], .width = 0.95)[, 2],
      "{agg_var}_median"   := median(.data[[agg_var]]),
      .groups = "drop"
    )
}

hu_spp_agg = f_draw_agg(hu_spp_draws, c("taxon_capture"), "hu")
prevalence_spp_agg = f_draw_agg(hu_spp_draws %>%
                                    mutate(prevalence = 1 - plogis(hu)), c("taxon_capture"), "prevalence")
sigma_spp_agg = f_draw_agg(sigma_spp_draws, c("taxon_capture"), "sigma")
mu_spp_ls_contrast_agg = f_draw_agg(mu_spp_lss_contrast_draws, c("taxon_capture", "contrast"), "mu_contrast")
hu_spp_ls_contrast_agg = f_draw_agg(hu_spp_lss_contrast_draws, c("taxon_capture", "contrast"), "hu_contrast")

```

```{r}
# load specs
load(here("size_bd_unified", "intermediates", "bd_model_unified_bspecs_2026-03-22.RData"))
load(here("size_bd_unified", "intermediates", "bd_model_unified_bspecs_ls_2026-03-22.RData"))
load(here("size_bd_unified", "intermediates", "bd_model_unified_huspecs_2026-03-22.RData"))
load(here("size_bd_unified", "intermediates", "bd_model_unified_huspecs_ls_2026-03-22.RData"))
load(here("size_bd_unified", "intermediates", "bd_model_unified_sigmaspecs_2026-03-22.RData"))

# hu
hu_contrasts_specs = contrast(hu_specs_ls, method = "revpairwise")

hu_draws = hu_specs %>%
  gather_emmeans_draws() %>%
  dplyr::rename(hu = .value) %>%
  mutate(prevalence = 1 - plogis(hu))

hu_group_draws = hu_specs_ls %>%
  gather_emmeans_draws() %>%
  dplyr::rename(hu = .value) %>%
  mutate(prevalence = 1 - plogis(hu))

hu_group_agg = hu_group_draws %>%
  group_by(taxon_capture) %>%
  summarize(hdi_50 = tidybayes::hdi(hu, .width = 0.50),
            hdi_95 = tidybayes::hdi(hu, .width = 0.95),
            hu_median = median(hu),
            .groups = "drop")

hu_contrast_draws = hu_contrasts_specs %>%
  gather_emmeans_draws() %>%
  dplyr::rename(hu_subadult = .value)

hu_contrast_agg = hu_contrast_draws %>%
  group_by(taxon_capture, contrast) %>%  # check!
  summarize(hdi_50 = tidybayes::hdi(hu_subadult, .width = 0.50),
            hdi_95 = tidybayes::hdi(hu_subadult, .width = 0.95),
            ratio_positive = sum(hu_subadult > 0) / n(),
            hu_subadult_median = median(hu_subadult),
            .groups = "drop")


# b
b_contrasts_specs = contrast(b_specs_ls, method = "revpairwise")

b_draws = b_specs %>%
  gather_emmeans_draws() %>%
  dplyr::rename(b = .value)

b_group_draws = b_specs_ls %>%
  gather_emmeans_draws() %>%
  dplyr::rename(b = .value)

b_contrast_draws = b_contrasts_specs %>%
  gather_emmeans_draws() %>%
  dplyr::rename(b_subadult = .value)

b_contrast_agg = b_contrast_draws %>%
  group_by(taxon_capture) %>%
  summarize(hdi_50 = tidybayes::hdi(b_subadult, .width = 0.50),
            hdi_95 = tidybayes::hdi(b_subadult, .width = 0.95),
            ratio_positive = sum(b_subadult > 0) / n(),
            b_subadult_median = median(b_subadult),
            .groups = "drop")

# sigma
sigma_draws = sigma_specs %>%
  gather_emmeans_draws() %>%
  dplyr::rename(sigma = .value)


sigma_group_agg = sigma_draws %>%
  group_by(taxon_capture) %>%
  reframe(hdi_50 = tidybayes::hdi(sigma, .width = 0.50),
          hdi_95 = tidybayes::hdi(sigma, .width = 0.95),
          sigma_median = median(sigma),
          .groups = "drop")

# combine
contrast_agg = hu_contrast_agg %>%
  left_join(b_contrast_agg, by = c("taxon_capture"), suffix = c("_hu", "_b")) %>%
  mutate(net_effect_hu = ratio_positive_hu * 2 - 1,
         net_effect_b = ratio_positive_b * 2 - 1,
         species = get_species_name(taxon_capture))

intercept_draws = hu_draws %>%
  left_join(b_draws, by = c(".chain", ".iteration", ".draw", "taxon_capture")) %>%
  left_join(sigma_draws, by = c(".chain", ".iteration", ".draw", "taxon_capture"))

intercept_agg = intercept_draws %>%
  group_by(taxon_capture) %>%
  summarize(prevalence_median = median(prevalence),
            hu_median = median(hu),
            b_median = median(b),
            sigma_median = median(sigma),
            .groups = "drop")

combined_agg = intercept_agg %>%
  left_join(contrast_agg %>%
              select(taxon_capture,
                     hu_subadult_median,
                     b_subadult_median,
                     net_effect_hu,
                     net_effect_b), by = "taxon_capture") %>%
  mutate(species = get_species_name(taxon_capture))

group_draws = b_group_draws %>%
  left_join(hu_group_draws, by = c(".chain", ".iteration", ".draw", 
                                   "life_stage_simple", "taxon_capture"))

group_draws_agg = group_draws %>%
  group_by(taxon_capture,
           life_stage_simple) %>%
  summarize(hu_median = median(hu),
            hu_hdi = tidybayes::hdi(hu, ci = 0.89),
            prevalence_median = median(prevalence),
            b_median = median(b),
            b_hdi = tidybayes::hdi(b, ci = 0.89),
            .groups = "drop") %>%
  mutate(prevalence_hdi_1 = 1 - plogis(hu_hdi[, 2]),
         prevalence_hdi_2 = 1 - plogis(hu_hdi[, 1]))
         
```

## Obsevations and predictions
```{r}

# epred draws
newdata_basic = expand_grid(taxon_capture = sort(unique(brms_data$taxon_capture)),
                            life_stage_simple = c("subadult",
                                                  "adult")) %>%
  mutate(precip_mm_30_z = 0,
         precip_mm_30_z2 = 0,
         temp_c_30_z = 0,
         temp_c_30_z2 = 0)

# # way too big, limit to 50 draws!
# epred_original = epred_draws(model_select,
#                              newdata = brms_data,
#                              re_formula = NA,
#                              ndraws = 200) %>%
#   dplyr::rename(bd_load_density_epred = .epred)

epred_fixed = epred_draws(model_select,
                          newdata = newdata_basic,
                          re_formula = NA) %>%
  dplyr::rename(bd_load_density_epred = .epred)

# observation data
obs_data = brms_data %>%
  group_by(taxon_capture,
           life_stage_simple) %>%
   summarise(n = n(),
             n_pos = sum(bd_detected_co),
             bd_prevalence = sum(bd_detected_co) / n(),
             bd_prevalence_ci_lower = wilson_ci_lower(bd_prevalence, n(), 0.95),
             bd_prevalence_ci_upper = wilson_ci_upper(bd_prevalence, n(), 0.95),
             min_load_inf = min(bd_load_co_density[bd_detected_co > 0]),
             q_25_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.25),
             q_50_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.50),
             q_75_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.75),
             max_load_inf = max(bd_load_co_density[bd_detected_co > 0]),
             se = sqrt(bd_prevalence*(1-bd_prevalence)/n),
             .groups = "drop") %>%
  mutate(label_offset = recode_values(life_stage_simple,
                                          "adult" ~ 0,
                                          "subadult" ~ 1))


```
## Visualize outputs

### Prevalence
```{r}
#| fig-height: 8
#| fig-width: 12

# observed
ggplot(obs_data, aes(y = bd_prevalence, x = life_stage_simple, color = life_stage_simple)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = bd_prevalence_ci_lower, ymax = bd_prevalence_ci_upper),
                width = 0.2,
                position = position_dodge(width = 0.5)) +
  geom_text(aes(label = paste0("n=", n), y = 1 - (1 - label_offset) * 0.05),
          position = position_dodge(width = 0.5),
          size = 4) +
  ylim(0, 1) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# modeled
ggplot(group_draws, aes(y = prevalence, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")


```

### Load
```{r}
#| fig-height: 8
#| fig-width: 12

# observed
ggplot(brms_data %>%
         filter(bd_detected_co > 0), aes(y = bd_load_co_density, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  geom_text(data = obs_data, aes(label = paste0("n=", n_pos), y = 10e8 - (1 - label_offset) * 7 * 10e7),
            position = position_dodge(width = 0.5),
            size = 4) +
  scale_y_log10(limits = c(1, 1e10)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")

# ggplot(epred_original, aes(y = bd_load_density_epred, x = life_stage_simple, fill = life_stage_simple)) +
#   geom_violin() +
#   # geom_boxplot() +
#   scale_y_log10(limits = c(1, 1e10)) +
#   theme_minimal() +
#   facet_wrap("taxon_capture") +
#   scale_fill_brewer(palette = "Set2")

ggplot(epred_fixed, aes(y = bd_load_density_epred, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  # geom_boxplot() +
  scale_y_log10(limits = c(1, 1e10)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")



```

## Hypothesis testing
```{r}
#| eval: true
#| echo: false
#| warning: false
#| fig-height: 8
#| fig-width: 12


# visualize draws
ggplot(group_draws, aes(x = prevalence, y = b, color = life_stage_simple)) +
  geom_point(alpha = 0.1) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# single parameter distributions
ggplot(group_draws, aes(y = prevalence, x = life_stage_simple, color = life_stage_simple)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

ggplot(group_draws, aes(y = b, x = life_stage_simple, color = life_stage_simple)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# visualize contrasts
ggplot(hu_contrast_draws, aes(x = hu_subadult)) +
  stat_halfeye(fill = "#66c2a5",
               .width = c(0.5, 0.95),
               point_interval = mode_hdi) +
  facet_wrap("taxon_capture") +
  geom_vline(xintercept = 0, alpha = 1) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect on Bd Prevalence",
       x = "Bd Prevalence Subadult Effect Size (logit scale) [-]\nwith 50% & 95% High Density Interval",
       y = "Posterior Density [-]") +
  theme(text=element_text(size=14))

ggplot(b_contrast_draws, aes(x = b_subadult)) +
    stat_halfeye(fill = "#8da0cb",
               .width = c(0.5, 0.95),
               point_interval = mode_hdi) +
  facet_wrap("taxon_capture") +
  geom_vline(xintercept = 0, alpha = 1) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect on Bd Prevalence",
       x = "Bd Load Density Subadult Effect Size (logit scale) [-]\nwith 50% & 95% High Density Interval",
       y = "Posterior Density [-]") +
  theme(text=element_text(size=14))

# probability of positive effect
p_net_effect_scatter = ggplot(contrast_agg, aes(x = -net_effect_hu, y = net_effect_b)) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point() +
  geom_text_repel(aes(label = get_species_name(taxon_capture)),
                  color = "#333",
                  size = 3) +
  xlim(-1, 1) +
  ylim(-1, 1) +
  theme_minimal() +
  labs(
       title = "Net Subadult Effect (Subadult - Adult)",
       x = "Contrast of log odds of Bd detection\n(subadult - adult) [-]",
       y = "Contrast of log Bd load density\n(subadult - adult) [-]") +
  scale_color_brewer(palette = "Set2")

ggsave(here("size_bd_unified", "unified_figures", "03_net_effect_scatter.png"),
       p_net_effect_scatter,
       width = 8, height = 8)

p_net_effect_scatter

ggplot(combined_agg, aes(x = -hu_subadult_median, y = b_subadult_median)) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  theme_minimal() +
  labs(
       title = "Median Subadult Effect (Subadult - Adult)",
       x = "Bd Prevalence (log scale) [-]",
       y = "Bd Load Density (logit scale) [-]") +
  scale_color_brewer(palette = "Set2")

ggplot(combined_agg, aes(x = prevalence_median, y = sigma_median)) +
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  theme_minimal() +
  labs(
       title = "Prevalence & Aggregation",
       x = "Median Bd Prevalence [-]",
       y = "Median Bd Load Aggregation -- log(sigma) [-]"
       ) +
  scale_color_brewer(palette = "Set2")

ggplot(combined_agg, aes(x = hu_subadult_median, y = hu_median)) +
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  theme_minimal() +
  labs(
       title = "Prevalence",
       x = "Median hu (subadult - adult) [-]",
       y = "Median hu [-]"
       ) +
  scale_color_brewer(palette = "Set2")

ggplot(group_draws_agg, aes(y = prevalence_median, ymin = prevalence_hdi_1, ymax = prevalence_hdi_2, x = taxon_capture, color = life_stage_simple)) +
  geom_pointrange(position = position_dodge(width = 0.7)) +
  theme_minimal() +
  coord_flip()

```

## Cascade plots
```{r}
# hu_spp_agg = f_draw_agg(hu_spp_lss_draws, c("taxon_capture"), "hu")
# sigma_spp_agg = f_draw_agg(sigma_spp_draws, c("taxon_capture"), "sigma")
# mu_spp_ls_contrast_agg = f_draw_agg(mu_spp_lss_contrast_draws, c("taxon_capture", "contrast"), "mu_contrast")
# hu_spp_ls_contrast_agg = f_draw_agg(hu_spp_lss_contrast_draws, c("taxon_capture", "contrast"), "hu_contrast")

p_cascade_hu_contrast_linpred = ggplot(hu_spp_ls_contrast_agg, aes(x = -hu_contrast_median, xmin = -hu_contrast_lower_95, xmax = -hu_contrast_upper_95, reorder(get_species_name(taxon_capture), hu_contrast_median))) +
  geom_vline(xintercept = 0, color = "#555") +
  geom_rect(aes(xmin = -hu_contrast_lower_50, xmax = -hu_contrast_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  theme_minimal() +
  labs(title = "Bd Prevalence Contrast (Subadult - Adult)",
       x = "Contrast of log odds of Bd detection (Subadult - Adult) [-]",
       y = "Species")

# p_cascade_hu_contrast = ggplot(hu_contrast_agg, aes(x = -hu_subadult_median, xmin = -hdi_95[,1], xmax = -hdi_95[,2], reorder(get_species_name(taxon_capture), hu_subadult_median))) +
#   geom_vline(xintercept = 0, color = "#555") +
#   geom_rect(aes(xmin = -hdi_50[,1], xmax = -hdi_50[,2]), height = 0.2, color = "#000") +
#   geom_pointrange() +
#   theme_minimal() +
#   labs(title = "Bd Prevalence Contrast (Subadult - Adult)",
#        x = "Contrast of log odds of Bd detection (Subadult - Adult) [-]",
#        y = "Species")

p_cascade_mu_contrast_linpred = ggplot(mu_spp_ls_contrast_agg, aes(x = mu_contrast_median, xmin = mu_contrast_lower_95, xmax = mu_contrast_upper_95, reorder(get_species_name(taxon_capture), -mu_contrast_median))) +
  geom_vline(xintercept = 0, color = "#555") +
  geom_rect(aes(xmin = mu_contrast_lower_50, xmax = mu_contrast_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  theme_minimal() +
  labs(title = "Bd Load Contrast (Subadult - Adult)",
       x = "Contrast of log Bd load (Subadult - Adult) [-]",
       y = "Species")

# p_cascade_mu_contrast = ggplot(b_contrast_agg, aes(x = b_subadult_median, xmin = hdi_95[,1], xmax = hdi_95[,2], reorder(get_species_name(taxon_capture), -b_subadult_median))) +
#   geom_vline(xintercept = 0, color = "#555") +
#   geom_rect(aes(xmin = hdi_50[,1], xmax = hdi_50[,2]), height = 0.2, color = "#000") +
#   geom_pointrange() +
#   theme_minimal() +
#   labs(title = "Bd Load Contrast (Subadult - Adult)",
#        x = "Contrast of log Bd load (Subadult - Adult) [-]",
#        y = "Species")

p_cascade_prevalence_linpred = ggplot(prevalence_spp_agg, aes(x = prevalence_median, xmin = prevalence_lower_95, xmax = prevalence_upper_95, reorder(get_species_name(taxon_capture), -prevalence_median))) +
  geom_rect(aes(xmin = prevalence_lower_50, xmax = prevalence_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  xlim(0, 1) +
  theme_minimal() +
  labs(title = "Bd Prevalence",
       x = "Prevalence draws [-]",
       y = "Species")


# p_cascade_prevalence = ggplot(hu_group_agg, aes(x = 1 - plogis(hu_median), xmin = 1 - plogis(hdi_95[,1]), xmax = 1 - plogis(hdi_95[,2]), reorder(get_species_name(taxon_capture), hu_median))) +
#   geom_rect(aes(xmin = 1 - plogis(hdi_50[,1]), xmax = 1 - plogis(hdi_50[,2])), height = 0.2, color = "#000") +
#   geom_pointrange() +
#   xlim(0, 1) +
#   theme_minimal() +
#   labs(title = "Bd Prevalence",
#        x = "Prevalence draws [-]",
#        y = "Species")

p_cascade_sigma_linpred = ggplot(sigma_spp_agg, aes(x = sigma_median, xmin = sigma_lower_95, xmax = sigma_upper_95, reorder(get_species_name(taxon_capture), -sigma_median))) +
  geom_rect(aes(xmin = sigma_lower_50, xmax = sigma_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  theme_minimal() +
  labs(title = "Bd Load Dispersion",
       x = "Log load standard deviation (sigma) draws [-]",
       y = "Species")

# p_cascade_sigma = ggplot(sigma_group_agg, aes(x = sigma_median, xmin = hdi_95[,1], xmax = hdi_95[,2], reorder(get_species_name(taxon_capture), -sigma_median))) +
#   geom_rect(aes(xmin = hdi_50[,1], xmax = hdi_50[,2]), height = 0.2, color = "#000") +
#   geom_pointrange() +
#   theme_minimal() +
#   labs(title = "Bd Load Dispersion",
#        x = "Log load standard deviation (sigma) draws [-]",
#        y = "Species")

# save plots
ggsave(here("size_bd_unified", "unified_figures", "03_species_cascade_hu_contrast.png"),
       p_cascade_hu_contrast_linpred,
       width = 8, height = 6)

ggsave(here("size_bd_unified", "unified_figures", "03_species_cascade_mu_contrast.png"),
       p_cascade_mu_contrast_linpred,
       width = 8, height = 6)

ggsave(here("size_bd_unified", "unified_figures", "03_species_cascade_prevalence.png"),
       p_cascade_prevalence_linpred,
       width = 8, height = 6)

ggsave(here("size_bd_unified", "unified_figures", "03_species_cascade_sigma.png"),
       p_cascade_sigma_linpred,
       width = 8, height = 6)


p_cascade_hu_contrast_linpred
p_cascade_mu_contrast_linpred
p_cascade_prevalence_linpred
p_cascade_sigma_linpred

```

# correlation analysis
```{r}
simple_cor = cor_mat(combined_agg %>%
                       select(hu_median,
                              sigma_median,
                              hu_subadult_median,
                              b_subadult_median), method = "spearman")

cor_get_pval(simple_cor)


```

## Phylogenetic tree

### Jetz & Pyron 2018
https://datadryad.org/dataset/doi:10.5061/dryad.cc3n6j5#usage
```{r}
jp_raw = read.tree(file = here("data", "amph_shl_new_Consensus_7238.tre"))

# preview alignment
jk_species_list = bd_model_env_stats %>%
  mutate(sn = str_to_title(taxon_capture)) %>%
  select(sn) %>%
  mutate(genus = str_split_fixed(sn, "_", 2)[,1],
         species = str_split_fixed(sn, "_", 2)[,2])

jp_species_list = tibble(jp_sn = jp_raw$tip.label) %>%
  mutate(jp_genus = str_split_fixed(jp_sn, "_", 2)[,1],
         species = str_split_fixed(jp_sn, "_", 2)[,2],
         jp = TRUE)
  

peace = jk_species_list %>%
  left_join(jp_species_list, by = "species") %>%
  select(sn,
         jp_sn,
         genus,
         jp_genus,
         everything())

species_list_j = bd_model_env_stats %>%
  select(taxon_capture) %>%
  distinct() %>%
  mutate(sn = str_to_title(taxon_capture),
         jp_sn = recode_values(sn,
                               "Lithobates_warszewitschii" ~ "Rana_warszewitschii",
                               "Lithobates_blairi" ~ "Rana_blairi",
                               "Lithobates_chiricahuensis" ~ "Rana_chiricahuensis",
                               "Lithobates_palustris" ~ "Rana_palustris",
                               "Lithobates_sphenocephalus" ~ "Rana_sphenocephala",
                               default = sn))

# Prune to species list
jp_pruned = keep.tip(jp_raw, species_list_j$jp_sn)
combined_agg_j = combined_agg %>%
  left_join(species_list_j %>%
              select(taxon_capture,
                     jp_sn), by = "taxon_capture")

# Ccheck plot
plotTree(jp_pruned, fsize = 0.8)

is.binary(jp_pruned)

table(sapply(1:jp_pruned$Nnode, function(n) {
  length(Children(jp_pruned, n + Ntip(jp_pruned)))
}))

```

<!-- ### Create -->
<!-- ```{r} -->
<!-- gene_marker <- "16S" -->
<!-- species_list = bd_model_env_stats %>% -->
<!--   mutate(species = get_species_name(taxon_capture)) %>% -->
<!--   pull(species) -->

<!-- taxon_list = bd_model_env_stats %>% -->
<!--   pull(taxon_capture) -->

<!-- fetch_sequence_reproducible <- function(species, marker, max_seqs = 10,  -->
<!--                                         min_length = 100, -->
<!--                                         target_min = 400, -->
<!--                                         target_max = 700) { -->
<!--   query <- paste0(species, "[Organism] AND (", marker, "[Title] OR ",  -->
<!--                   marker, "[All Fields])") -->
<!--   search <- entrez_search(db = "nucleotide", term = query, retmax = max_seqs) -->

<!--   if (length(search$ids) == 0) { -->
<!--     message("No sequence found for: ", species) -->
<!--     return(NULL) -->
<!--   } -->

<!--   seqs <- entrez_fetch(db = "nucleotide", id = search$ids, rettype = "fasta") -->

<!--   tmp <- tempfile(fileext = ".fasta") -->
<!--   writeLines(seqs, tmp) -->

<!--   dna <- tryCatch( -->
<!--     read.dna(tmp, format = "fasta"), -->
<!--     error = function(e) { message("Failed to parse FASTA for: ", species); return(NULL) } -->
<!--   ) -->

<!--   if (is.null(dna)) return(NULL) -->
<!--   if (is.matrix(dna)) dna <- as.list(dna) -->

<!--   # Filter by minimum length -->
<!--   lengths <- sapply(dna, length) -->
<!--   dna <- dna[lengths >= min_length] -->

<!--   if (length(dna) == 0) { -->
<!--     message("No sequences above min_length for: ", species) -->
<!--     return(NULL) -->
<!--   } -->

<!--   lengths <- sapply(dna, length) -->

<!--   # Try preferred length range first -->
<!--   in_range <- dna[lengths >= target_min & lengths <= target_max] -->

<!--   if (length(in_range) > 0) { -->
<!--     range_lengths <- sapply(in_range, length) -->
<!--     selected <- in_range[which.max(range_lengths)] -->
<!--     label <- "preferred range" -->
<!--   } else { -->
<!--     # Fall back to longest available -->
<!--     selected <- dna[which.max(lengths)] -->
<!--     label <- "fallback longest" -->
<!--   } -->

<!--   # Log the selected accession for reproducibility -->
<!--   message(species, " [", label, "]: ",  -->
<!--           sapply(selected, length), "bp — ", names(selected)) -->

<!--   return(selected) -->
<!-- } -->

<!-- # Run and capture selection log -->
<!-- raw_seqs <- lapply(species_list, function(sp) { -->
<!--   Sys.sleep(0.4) -->
<!--   fetch_sequence_reproducible(sp, marker = gene_marker, max_seqs = 50) -->
<!-- }) -->
<!-- names(raw_seqs) <- species_list -->
<!-- raw_seqs <- Filter(Negate(is.null), raw_seqs) -->

<!-- # After fetching, decode and save immediately -->
<!-- decode_dnabin <- function(x) { -->
<!--   tmp <- tempfile(fileext = ".fasta") -->
<!--   write.dna(x, tmp, format = "fasta") -->
<!--   lines <- readLines(tmp) -->
<!--   seq_lines <- lines[!grepl("^>", lines)] -->
<!--   seq_string <- paste(seq_lines, collapse = "") -->
<!--   gsub("[[:space:]]", "", seq_string) -->
<!-- } -->

<!-- # Decode all sequences -->
<!-- combined_char <- sapply(names(raw_seqs), function(sp) { -->
<!--   decode_dnabin(raw_seqs[[sp]]) -->
<!-- }) -->
<!-- names(combined_char) <- names(raw_seqs) -->

<!-- # Sanity check before alignment -->
<!-- substr(combined_char[1], 1, 50)  # should show ACGTs now -->

<!-- # Save as FASTA for reuse — no need to re-fetch next time -->
<!-- fasta_lines <- unlist(lapply(names(combined_char), function(sp) { -->
<!--   c(paste0(">", sp), combined_char[[sp]]) -->
<!-- })) -->
<!-- writeLines(fasta_lines, here("size_bd_unified", "intermediates", "selected_sequences.fasta")) -->
<!-- message("Sequences saved to selected_sequences.fasta") -->

<!-- # To reload next time instead of re-fetching: -->
<!-- # dna_saved <- read.dna(here("size_bd_unified", "intermediates", "selected_sequences.fasta"), format = "fasta") -->
<!-- # combined_char <- sapply(as.character(dna_saved), paste, collapse = "") -->
<!-- # names(combined_char) <- taxon_list -->

<!-- # Align -->
<!-- aligned <- AlignSeqs(DNAStringSet(combined_char)) -->

<!-- # --- 3. BUILD TREE --- -->
<!-- phydat <- as.phyDat(aligned) -->

<!-- dist_mat <- dist.ml(phydat, model = "JC69") -->

<!-- nj_tree <- NJ(dist_mat) -->

<!-- fit <- pml(nj_tree, phydat) -->
<!-- fit_opt <- optim.pml(fit, -->
<!--                      model         = "GTR", -->
<!--                      optInv        = TRUE, -->
<!--                      optGamma      = TRUE, -->
<!--                      rearrangement = "NNI") -->

<!-- tr_phylo <- fit_opt$tree -->

<!-- # --- 4. ROOT --- -->
<!-- tr_rooted <- midpoint(tr_phylo) -->

<!-- # --- 5. QUICK SANITY CHECKS --- -->
<!-- summary(tr_rooted$edge.length) -->
<!-- is.ultrametric(tr_rooted)  # expected FALSE for ML tree -->
<!-- # plot(tr_rooted, cex = 0.6) -->

<!-- # Optional: compare topology to your original OTL tree -->
<!-- ggtree::ggtree(tr_rooted, layout = "rectangular") + -->
<!--   ggtree::geom_tiplab(size = 3, align = FALSE) + -->
<!--   ggtree::theme_tree2() + -->
<!--   scale_x_continuous(expand = expansion(mult = c(0, 0.3))) -->

<!-- # Now has real branch lengths! -->
<!-- summary(tr_rooted$edge.length) -->
<!-- is.ultrametric(tr_rooted) -->

<!-- phylo_covar = vcv.phylo(tr_rooted, corr = TRUE) -->

<!-- saveRDS(phylo_covar, here("size_bd_unified", "intermediates", "phylo_covar.rds")) -->

<!-- write.tree(tr_rooted, here("size_bd_unified", "intermediates", "phylo_tree.nwk")) -->
<!-- ``` -->

### Analyze
```{r}
tr_rooted = read.tree(here("size_bd_unified", "intermediates", "phylo_tree.nwk"))

# hu_sub_effect_named <- setNames(-combined_agg$net_effect_hu, combined_agg$species)
# b_sub_effect_named  <- setNames(combined_agg$net_effect_b,  combined_agg$species)
# hu_sub_med_named <- setNames(-combined_agg$hu_subadult_median, combined_agg$species)
# b_sub_med_named  <- setNames(combined_agg$b_subadult_median,  combined_agg$species)
# hu_med_named <- setNames(-combined_agg$hu_median, combined_agg$species)
# b_med_named  <- setNames(combined_agg$b_median,  combined_agg$species)
# prev_med_named <- setNames(combined_agg$prevalence_median, combined_agg$species)
# sigma_med_named  <- setNames(combined_agg$sigma_median,  combined_agg$species)


# blomberg_k_hu_sub_effect = phylosig(tr_rooted, hu_sub_effect_named, method = "K", test = TRUE, nsim = 1000)
# blomberg_k_b_sub_effect = phylosig(tr_rooted, b_sub_effect_named, method = "K", test = TRUE, nsim = 1000)
# blomberg_k_hu_sub_med = phylosig(tr_rooted, hu_sub_med_named, method = "K", test = TRUE, nsim = 1000)
# blomberg_k_b_sub_med = phylosig(tr_rooted, b_sub_med_named, method = "K", test = TRUE, nsim = 1000)
# blomberg_k_hu_med = phylosig(tr_rooted, hu_med_named, method = "K", test = TRUE, nsim = 1000)
# blomberg_k_b_med = phylosig(tr_rooted, b_med_named, method = "K", test = TRUE, nsim = 1000)
# blomberg_k_prev_med = phylosig(tr_rooted, prev_med_named, method = "K", test = TRUE, nsim = 1000)
# blomberg_k_sigma_med = phylosig(tr_rooted, sigma_med_named, method = "K", test = TRUE, nsim = 1000)

# tr_rooted$tip.label

# plots
# hu_sub_effect_map = contMap(jp_pruned, hu_sub_effect_named,  title = "hu net effect", fsize = 0.6)
# b_sub_effect_map = contMap(jp_pruned, b_sub_effect_named,   title = "b net_effect", fsize = 0.6)
# hu_sub_med_map = contMap(jp_pruned, hu_sub_med_named,  title = "hu net median", fsize = 0.6)
# b_sub_med_map = contMap(jp_pruned, b_sub_med_named,   title = "b net_median", fsize = 0.6)
# hu_med_map = contMap(jp_pruned, hu_med_named,  title = "hu median", fsize = 0.6)
# b_med_map = contMap(jp_pruned, b_med_named,   title = "b median", fsize = 0.6)
# prev_med_map = contMap(jp_pruned, prev_med_named,  title = "prevalence median effect", fsize = 0.6)
# sigma_med_map = contMap(jp_pruned, sigma_med_named,   title = "sigma median effect", fsize = 0.6)

# hu_sub_effect_map <- setMap(hu_sub_effect_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))
# b_sub_effect_map <- setMap(b_sub_effect_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))
# hu_sub_med_map <- setMap(hu_sub_med_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))
# b_sub_med_map <- setMap(b_sub_med_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))
# hu_med_map <- setMap(hu_med_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))
# b_med_map <- setMap(b_med_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))
# prev_med_map <- setMap(prev_med_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))
# sigma_med_map <- setMap(sigma_med_map, colors = colorRampPalette(brewer.pal(9, "PRGn"))(100))

# calculate blomberg K
blomberg_k = function(tree, var_values, species) {
  set_named = setNames(var_values, species)
  return(blomberg_k_sigma_med = phylosig(tree, set_named, method = "K", test = TRUE, nsim = 1000))
}


# define color maps
sym_div_cont_map = function(tree, var_values, species, center, p_title, style = "div") {
  set_named = setNames(var_values, species)
  
  sym_lim = max(abs(center - min(var_values)), abs(center - max(var_values)))
  
  map_1 = contMap(tree, set_named,
                   lims = c(center - sym_lim, center + sym_lim),
                   title = p_title, fsize = 0.4)
  
  if (style == "div") {
    colors = colorRampPalette(c("#762a83", "white", "#1b7837"))(101)
  } else if (style == "lin") {
    colors = colorRampPalette(c("white", "#1b7837"))(101)
  }
  map_2 = setMap(map_1, colors = colors)
  
  return(map_2)
}

hu_sub_med_k = blomberg_k(jp_pruned, combined_agg_j$hu_subadult_median, combined_agg_j$jp_sn)
b_sub_med_k = blomberg_k(jp_pruned, combined_agg_j$b_subadult_median, combined_agg_j$jp_sn)
# b_med_k = blomberg_k(jp_pruned, combined_agg_j$b_median, combined_agg_j$jp_sn)
hu_med_k = blomberg_k(jp_pruned, combined_agg_j$hu_median, combined_agg_j$jp_sn)
prev_med_k = blomberg_k(jp_pruned, combined_agg_j$prevalence_median, combined_agg_j$jp_sn)
sigma_med_k = blomberg_k(jp_pruned, combined_agg_j$sigma_median, combined_agg_j$jp_sn)

hu_sub_med_map = sym_div_cont_map(jp_pruned,
                                  combined_agg_j$hu_subadult_median,
                                  combined_agg_j$jp_sn,
                                  0,
                                  "Bd prevalence (hu): median subadult effect")

b_sub_med_map = sym_div_cont_map(jp_pruned,
                                 combined_agg_j$b_subadult_median,
                                 combined_agg_j$jp_sn,
                                 0,
                                 "Bd log load (mu): median subadult effect")

# b_med_map = sym_div_cont_map(jp_pruned,
#                               combined_agg_j$b_median,
#                               combined_agg_j$jp_sn,
#                               4.135,
#                               "Median Bd log load (mu)")
# 
# hu_med_map = sym_div_cont_map(jp_pruned,
#                               combined_agg_j$hu_median,
#                               combined_agg_j$jp_sn,
#                               0,
#                               "Median Bd prevalence (hu)")

prev_med_map = sym_div_cont_map(jp_pruned,
                                combined_agg_j$prevalence_median,
                                combined_agg_j$jp_sn,
                                0.5,
                                "Median Bd prevalence")

sigma_med_map = sym_div_cont_map(jp_pruned,
                                 combined_agg_j$sigma_median,
                                 combined_agg_j$jp_sn,
                                 0.8822,
                                 "Median Bd load dispersion (sigma)")


hu_sub_med_k
b_sub_med_k
# hu_med_k
prev_med_k
sigma_med_k

# plot(hu_sub_effect_map)
# plot(b_sub_effect_map)
plot(hu_sub_med_map)
plot(b_sub_med_map)
# plot(hu_med_map)
# plot(b_med_map)
plot(prev_med_map)
plot(sigma_med_map)

```

## Collective results
```{r}
#| eval: true
#| echo: true
#| output: false
#| code-fold: true
#| code-summary: Show code

# m5g_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2.rds")

choice_model = m5g_zln

model_data = brms_data

draws_all = collect_draws(choice_model) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

density_mat_all = density_mat(log(0.5), 20, 100, draws_all) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

sim_load_all = simulate_bd_load(draws_all, 10, 1000) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

obs_stats_all = sample_stats(model_data) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

```

::: {.panel-tabset}
### Observed Prevalence
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# observed prevalence
ggplot(obs_stats_all, aes(y = bd_prevalence, x = species, color = life_stage_simple)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = bd_prevalence_ci_lower, ymax = bd_prevalence_ci_upper),
                width = 0.2,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  ggtitle("Observed prevalence") +
  xlab("Species") +
  ylab("Bd prevalence [-]\n(with 95% CI)") +
  labs(color = "Life stage") +
  geom_text(aes(label = paste0("n=", n),  y = label_y),
          position = position_dodge(width = 0.5)) +
  ylim(0, 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Modeled Prevalence 
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# model prevalence
ggplot(draws_all, aes(prevalence, x = species, color = life_stage_simple)) +
  geom_boxplot(coef = NULL) + 
  theme_minimal() +
  ggtitle("Modeled prevalence") +
  xlab("Species") +
  ylab("Bd prevalence [-]\n") +
  labs(color = "Life stage") +
  ylim(0, 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Observed Bd load
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# obsevation bd load
ggplot(model_data %>%
         mutate(species = get_species_name(taxon_capture, newline = TRUE)) %>%
         filter(bd_load_co > 0), aes(log_bd_load_co_density, x = species, color = life_stage_simple)) +
  geom_boxplot() + 
  theme_minimal() +
  ggtitle("Observed Bd-positive load") +
  xlab("Species") +
  ylab("log(Bd load density) [-]") +
  labs(color = "Life stage") +
  geom_text(
    data = obs_stats_all,
    aes(
      x = species,
      y = max(model_data$log_bd_load_co_density) * label_y * 1.05,
      label = paste0("n=", n_pos),
      group = life_stage_simple
    ),
    position = position_dodge(width = 0.75),
    vjust = 0
  ) +
  ylim(log(0.5), 19) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Modeled Bd load
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# modeled bd load
ggplot(sim_load_all %>%
         filter(log_bd_load_co_density > 0.5), aes(log_bd_load_co_density, x = species, color = life_stage_simple)) +
  geom_boxplot() + 
  theme_minimal() +
  ggtitle("Modeled Bd-positive load") +
  xlab("Species") +
  ylab("log(Bd load density) [-]") +
  labs(color = "Life stage") +
  ylim(log(0.5), 20) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Modeled Bd load PDF
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# PDF -- facet by species
ggplot(density_mat_all, aes(x = density,
                            y = pdf_median,
                            color = life_stage_simple, 
                            fill = life_stage_simple)) +
  geom_ribbon(aes(ymin = pdf_lower,
                  ymax = pdf_upper), 
              alpha = 0.15,
              color = NA) +
  geom_line() +
  theme_minimal() +
  ggtitle("Modeled PDF of Bd-positive load") +
  labs(color = "Life stage",
       fill = "Life stage") +
  xlab("log(Bd load) [-]") +
  ylab("Median probability density [-]\n(with 25th to 75th percentiles)") +
  scale_x_continuous(limits = c(log(0.5), 20), expand = c(0, 0)) +
  facet_grid(. ~ gsub(" ", "\n", species))

# # PDF -- facet by life_stage & season
# ggplot(density_mat_all, aes(x = density,
#                             y = pdf_median,
#                             color = taxon_capture, 
#                             fill = taxon_capture)) +
#   geom_ribbon(aes(ymin = pdf_lower,
#                   ymax = pdf_upper), 
#               alpha = 0.15,
#               color = NA) +
#   geom_line() +
#   theme_minimal() +
#   ggtitle("Modeled PDF of Bd-positive load") +
#   labs(color = "Species",
#        fill = "Species") +
#   xlab("log(Bd load) [-]") +
#   ylab("Median probability density [-]\n(with 25th to 75th percentiles)") +
#   scale_x_continuous(limits = c(log(0.3), 20), expand = c(0, 0)) +
#   facet_grid(life_stage_simple ~ season)
```

### Moded Bd load CCDF
```{r}
#| echo: false
#| warning: false
#| fig-height: 8
#| fig-width: 8
# CCDF -- facet by species
ggplot(density_mat_all, aes(x = density,
                            y = ccdf_median,
                            color = life_stage_simple, 
                            fill = life_stage_simple)) +
  geom_ribbon(aes(ymin = ccdf_lower,
                  ymax = ccdf_upper), 
              alpha = 0.15,
              color = NA) +
  geom_line() +
  theme_minimal() +
  ggtitle("Modeled CCDF of Bd-positive load") +
  labs(color = "Life stage",
       fill = "Life stage") +
  xlab("log(Bd load) [-]") +
  ylab("Median complementary cumulative density (1 - CDF) [-]\n(with 25th to 75th percentiles)") +
  scale_x_continuous(limits = c(log(0.5), 12), expand = c(0, 0)) +
  ylim(0, 0.75) +
  # facet_grid(. ~ gsub(" ", "\n", species))
  facet_wrap("species")

# CCDF -- facet by life_stage & season
# ggplot(density_mat_all, aes(x = density,
#                             y = ccdf_median,
#                             color = taxon_capture,
#                             fill = taxon_capture)) +
#   geom_ribbon(aes(ymin = ccdf_lower,
#                   ymax = ccdf_upper),
#               alpha = 0.15,
#               color = NA) +
#   geom_line() +
#   theme_minimal() +
#   ggtitle("Modeled CCDF of Bd-positive load") +
#   labs(color = "Species",
#        fill = "Species") +
#   xlab("log(Bd load) [-]") +
#   ylab("Median complementary cumulative density (1 - CDF) [-]\n(with 25th to 75th percentiles)") +
#   scale_x_continuous(limits = c(log(0.3), 20), expand = c(0, 0)) +
#   ylim(0, 1) +
#   facet_grid(. ~ life_stage_simple)
```

# Environmental conditions

## Distributions
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Environmental conditions distributions code"
#| fig-height: 6
#| fig-width: 8

# 15-day
# ggplot(brms_data, aes(x = temp_c_15_day_mean, y = precip_mm_15_day_sum, color = life_stage_simple)) +
#   geom_point(alpha = 0.1) +
#   stat_ellipse() +
#   facet_wrap(. ~ taxon_capture) +
#   xlim(-5, 40) +
#   ylim(0, 200) +
#   labs(x = "15-day temperature mean (degC)", 
#        y = "15-day precipitation total (mm)",
#        title = "Temperature v. precipitation") +
#   theme_minimal()

# 30-day by species
ggplot(brms_data, aes(x = temp_c_30_day_mean, y = precip_mm_30_day_sum, color = life_stage_simple)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  facet_wrap(. ~ taxon_capture) +
  xlim(-5, 40) +
  ylim(0, 400) +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()

# 30 day collective
ggplot(brms_data, aes(x = temp_c_30_day_mean, y = precip_mm_30_day_sum, color = taxon_capture)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()

# 30 day collective, scaled to climate by species
ggplot(brms_data, aes(x = temp_c_30_sp_z, y = precip_mm_30_sp_z, color = taxon_capture)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()


# ggplot(brms_data, aes(x = temp_c_30_sp_z2, y = precip_mm_30_sp_z2, color = life_stage_simple)) +
#   geom_point(alpha = 0.1) +
#   stat_ellipse() +
#   facet_wrap(. ~ taxon_capture) +
#   labs(x = "30-day temperature mean (degC)", 
#        y = "30-day precipitation total (mm)",
#        title = "Temperature v. precipitation") +
#   theme_minimal()
```

## Tempurature effects
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show temperature effects code"
#| fig-height: 6
#| fig-width: 8

select_model = m5f_zln
# select_model = m6_zln

climate_draws = spread_draws(select_model, `b_temp_c.*|b_precip_mm.*|b_hu_temp_c.*|b_hu_precip_mm.*|b_Intercept|b_hu_Intercept`, regex = TRUE)

# Temperature dependence
temp_range = seq(-5, 45, length.out = 100)

temp_grid = expand_grid(temp_c_15_day_mean = temp_range, .draw = 1:nrow(climate_draws)) %>%
  mutate(temp_c_30_day_mean = temp_c_15_day_mean) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    temp_c_15_z = (temp_c_15_day_mean - brms_scales$temp_c_15_day_mean_mean) / brms_scales$temp_c_15_day_mean_sd,
    temp_c_15_z2 = (temp_c_15_day_mean^2 - brms_scales$temp_c_15_day_mean2_mean) / brms_scales$temp_c_15_day_mean2_sd,
    temp_c_30_z = (temp_c_30_day_mean - brms_scales$temp_c_30_day_mean_mean) / brms_scales$temp_c_30_day_mean_sd,
    temp_c_30_z2 = (temp_c_30_day_mean^2 - brms_scales$temp_c_30_day_mean2_mean) / brms_scales$temp_c_30_day_mean2_sd,
    b_pred = b_temp_c_30_z * temp_c_30_z + b_temp_c_30_z2 * temp_c_30_z2 + b_Intercept,
    hu_pred = b_hu_temp_c_15_z * temp_c_15_z + b_hu_temp_c_15_z2 * temp_c_15_z2 + b_hu_Intercept
  ) %>%
  group_by(temp_c_15_day_mean) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  ) %>%
  dplyr::rename(temp_c = temp_c_15_day_mean)

# prevalence
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day mean temperature (degC)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. temperature") +
  theme_minimal()

# load
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day mean temperature (degC)", 
       y = "log(+ Bd load)",
       title = "Bd load v. temperature") +
  theme_minimal()
```

## Precipitation effects
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show precipitation effects code"
#| fig-height: 6
#| fig-width: 8
# precipitation depencence

daily_precip_range = seq(0, 20, length.out = 100)

precip_grid = expand_grid(daily_precip = daily_precip_range, .draw = 1:nrow(climate_draws)) %>%
  mutate(precip_mm_15_day_sum = daily_precip * 15,
         precip_mm_30_day_sum = daily_precip * 30) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    precip_mm_15_z = (precip_mm_15_day_sum - brms_scales$precip_mm_15_day_sum_mean) / brms_scales$precip_mm_15_day_sum_sd,
    precip_mm_15_z2 = (precip_mm_15_day_sum^2 - brms_scales$precip_mm_15_day_sum2_mean) / brms_scales$precip_mm_15_day_sum2_sd,
    precip_mm_30_z = (precip_mm_30_day_sum - brms_scales$precip_mm_30_day_sum_mean) / brms_scales$precip_mm_30_day_sum_sd,
    precip_mm_30_z2 = (precip_mm_30_day_sum^2 - brms_scales$precip_mm_30_day_sum2_mean) / brms_scales$precip_mm_30_day_sum2_sd,
    b_pred = b_precip_mm_30_z * precip_mm_30_z + b_precip_mm_30_z2 * precip_mm_30_z2 + b_Intercept,
    hu_pred = b_hu_precip_mm_15_z * precip_mm_15_z + b_hu_precip_mm_15_z2 * precip_mm_15_z2 + b_hu_Intercept
  ) %>%
  group_by(daily_precip,
           precip_mm_15_day_sum,
           precip_mm_30_day_sum) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  )

# prevalence
ggplot(precip_grid, aes(x = precip_mm_15_day_sum)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day total precipitation (mm)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. precipitation") +
  theme_minimal()

# load
ggplot(precip_grid, aes(x = precip_mm_30_day_sum)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day total precipitation (mm)", 
       y = "log(+ Bd load)",
       title = "Bd load v. precipitation") +
  theme_minimal()

```

## Tempurature effects (scaled by species climate)
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show temperature effects code"
#| fig-height: 6
#| fig-width: 8


select_model = m6_zln

# fixed effects
climate_draws = spread_draws(select_model, `b_temp_c.*|b_precip_mm.*|b_hu_temp_c.*|b_hu_precip_mm.*|b_Intercept|b_hu_Intercept`, regex = TRUE)

# Temperature dependence
temp_range = seq(-3, 3, length.out = 100)

# taxa = brms_data %>%
#   select(taxon_capture) %>%
#   distinct() %>%
#   pull(taxon_capture) %>%
#   sort()

## Stumped by how to back-calculate the temp_z2 values from the temp_z values... I think this has to be done on a species-by-species basis...

temp_grid = expand_grid(temp_z_15_mean = temp_range,
                        .draw = 1:nrow(climate_draws)) %>%
  mutate(temp_z_30_mean = temp_z_15_mean) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    temp_c_15_z = (temp_c_15_day_mean - brms_scales$temp_c_15_day_mean_mean) / brms_scales$temp_c_15_day_mean_sd,
    temp_c_15_z2 = (temp_c_15_day_mean^2 - brms_scales$temp_c_15_day_mean2_mean) / brms_scales$temp_c_15_day_mean2_sd,
    temp_c_30_z = (temp_c_30_day_mean - brms_scales$temp_c_30_day_mean_mean) / brms_scales$temp_c_30_day_mean_sd,
    temp_c_30_z2 = (temp_c_30_day_mean^2 - brms_scales$temp_c_30_day_mean2_mean) / brms_scales$temp_c_30_day_mean2_sd,
    b_pred = b_temp_c_30_sp_z * temp_c_30_z + b_temp_c_30_sp_z2 * temp_c_30_z2 + b_Intercept,
    hu_pred = b_hu_temp_c_15_sp_z * temp_c_15_z + b_hu_temp_c_15_sp_z2 * temp_c_15_z2 + b_hu_Intercept
  ) %>%
  group_by(temp_c_15_day_mean) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  ) %>%
  dplyr::rename(temp_c = temp_c_15_day_mean)

# prevalence
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day mean temperature (degC)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. temperature") +
  theme_minimal()

# load
ggplot(temp_grid, aes(x = temp_c)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day mean temperature (degC)", 
       y = "log(+ Bd load)",
       title = "Bd load v. temperature") +
  theme_minimal()
```

## Precipitation effects (scaled by species climate)
```{r}
#| eval: true
#| code-fold: true
#| code-summary: "Show precipitation effects code"
#| fig-height: 6
#| fig-width: 8
# precipitation depencence

daily_precip_range = seq(0, 20, length.out = 100)

precip_grid = expand_grid(daily_precip = daily_precip_range, .draw = 1:nrow(climate_draws)) %>%
  mutate(precip_mm_15_day_sum = daily_precip * 15,
         precip_mm_30_day_sum = daily_precip * 30) %>%
  left_join(climate_draws, by = ".draw") %>%
  mutate(
    precip_mm_15_z = (precip_mm_15_day_sum - brms_scales$precip_mm_15_day_sum_mean) / brms_scales$precip_mm_15_day_sum_sd,
    precip_mm_15_z2 = (precip_mm_15_day_sum^2 - brms_scales$precip_mm_15_day_sum2_mean) / brms_scales$precip_mm_15_day_sum2_sd,
    precip_mm_30_z = (precip_mm_30_day_sum - brms_scales$precip_mm_30_day_sum_mean) / brms_scales$precip_mm_30_day_sum_sd,
    precip_mm_30_z2 = (precip_mm_30_day_sum^2 - brms_scales$precip_mm_30_day_sum2_mean) / brms_scales$precip_mm_30_day_sum2_sd,
    b_pred = b_precip_mm_30_z * precip_mm_30_z + b_precip_mm_30_z2 * precip_mm_30_z2 + b_Intercept,
    hu_pred = b_hu_precip_mm_15_z * precip_mm_15_z + b_hu_precip_mm_15_z2 * precip_mm_15_z2 + b_hu_Intercept
  ) %>%
  group_by(daily_precip,
           precip_mm_15_day_sum,
           precip_mm_30_day_sum) %>%
  summarise(
    b_mean = mean(b_pred),
    b_med = median(b_pred),
    b_lower_95 = quantile(b_pred, 0.025),
    b_upper_95 = quantile(b_pred, 0.975),
    b_lower_89 = quantile(b_pred, 0.055),
    b_upper_89 = quantile(b_pred, 0.945),
    hu_mean = mean(hu_pred),
    hu_med = median(hu_pred),
    hu_lower_95 = quantile(hu_pred, 0.025),
    hu_upper_95 = quantile(hu_pred, 0.975),
    hu_lower_89 = quantile(hu_pred, 0.055),
    hu_upper_89 = quantile(hu_pred, 0.945),
    p_mean = plogis(mean(hu_pred)),
    p_med = plogis(median(hu_pred)),
    p_lower_95 = plogis(quantile(hu_pred, 0.025)),
    p_upper_95 = plogis(quantile(hu_pred, 0.975)),
    p_lower_89 = plogis(quantile(hu_pred, 0.055)),
    p_upper_89 = plogis(quantile(hu_pred, 0.945)),
    .groups = "drop"
  )

# prevalence
ggplot(precip_grid, aes(x = precip_mm_15_day_sum)) +
  geom_ribbon(aes(ymin = p_lower_89, ymax = p_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = p_lower_95, ymax = p_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = p_mean), linewidth = 1) +
  labs(x = "15-day total precipitation (mm)", 
       y = "Probability of Bd detection",
       title = "Bd prevalence v. precipitation") +
  theme_minimal()

# load
ggplot(precip_grid, aes(x = precip_mm_30_day_sum)) +
  geom_ribbon(aes(ymin = b_lower_89, ymax = b_upper_89), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = b_lower_95, ymax = b_upper_95), 
              fill = "steelblue", alpha = 0.2) +
  geom_line(aes(y = b_mean), linewidth = 1) +
  labs(x = "30-day total precipitation (mm)", 
       y = "log(+ Bd load)",
       title = "Bd load v. precipitation") +
  theme_minimal()

```