---
title: "size_bd_unified_analysis"
format: html
---

# Load data & models
```{r}
librarian::shelf(tidyverse, here, janitor, brms, tidybayes, ggrepel, emmeans, rstatix, phylotools, rentrez, DECIPHER, phangorn, phytools, RColorBrewer)

# load bd_model_env
# load(here("size_bd_unified", "intermediates", "bd_model_env_unified_2026-04-03.RData")) 
load(here("size_bd_unified", "intermediates", "bd_model_env_unified_2026-04-03.RData")) 

bd_model_env_stats = brms_data %>%
  group_by(taxon_capture, country) %>%
  summarise(regions = paste(sort(unique(region)), collapse = ", "),
            population_n = n_distinct(population),
            site_n = n_distinct(site),
            year_first = year(min(date)),
            year_last = year(max(date)),
            year_count = n_distinct(year(date)),
            adult_n = sum(life_stage_simple == "adult"),
            adult_n_pos = sum(life_stage_simple == "adult" & bd_detected_co),
            subadult_n = sum(life_stage_simple == "subadult"),
            subadult_n_pos = sum(life_stage_simple == "subadult" & bd_detected_co),
            .groups = "drop") %>%
  arrange(country,
          taxon_capture)

# load models
null_zln = readRDS(file = here("size_bd_unified/unified_models/unified_model_results_zln_null.rds"))
m9a_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_env2_renvpoptls_30.rds")

model_select = m9a_zln
```

## Define analysis functions
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show processing & visualization functions"

get_species_name = function(taxon, newline = FALSE) {
  species_name = str_to_sentence(gsub("_", " ", gsub("usgs", "", taxon)))
  if (newline) {
    species_name = gsub(" ", "\n", species_name)
  }
  return(species_name)
}

wilson_ci = function(p_hat, n, conf.level = 0.95) {
  z <- qnorm(1 - (1 - conf.level) / 2)
  
  denominator <- 1 + (z^2 / n)
  centre_adjusted_probability <- p_hat + (z^2 / (2 * n))
  adjusted_standard_deviation <- sqrt((p_hat * (1 - p_hat) / n) + (z^2 / (4 * n^2)))
  
  lower_bound <- (centre_adjusted_probability - z * adjusted_standard_deviation) / denominator
  upper_bound <- (centre_adjusted_probability + z * adjusted_standard_deviation) / denominator
  
  return(c(lower_bound = lower_bound, upper_bound = upper_bound))
}

wilson_ci_lower = function(p_hat, n, conf.level = 0.95) {
  ws = wilson_ci(p_hat, n, conf.level)
  return(ws[1])
}

wilson_ci_upper = function(p_hat, n, conf.level = 0.95) {
  ws = wilson_ci(p_hat, n, conf.level)
  return(ws[2])
}


```

# Results

## Organize draws

### Linpred Draws
```{r}

# define linpred data
nd_env_subadult = model_select$data %>%
  group_by(taxon_capture, life_stage_simple) %>%
  summarize(precip_mm_10_z = median(precip_mm_10_z),
            temp_c_30_z = median(temp_c_30_z),
            .groups = "drop") %>%
  filter(life_stage_simple == "subadult")

nd_spp_lss = model_select$data |>
  distinct(taxon_capture, life_stage_simple) %>%
  left_join(nd_env_subadult %>%
              select(-life_stage_simple), by = "taxon_capture")


# hu contrast
hu_spp_lss_draws = model_select |>
  linpred_draws(
    newdata    = nd_spp_lss,
    dpar       = "hu",
    re_formula = NA
  ) %>%
  ungroup()

hu_spp_draws = hu_spp_lss_draws |>
  group_by(taxon_capture, .draw) |>
  summarise(hu = mean(hu), .groups = "drop")

hu_spp_lss_contrast_draws = hu_spp_lss_draws |>
  pivot_wider(
    id_cols     = c(taxon_capture, .draw),
    names_from  = life_stage_simple,
    values_from = hu
  ) |>
  mutate(hu_contrast = subadult - adult,
         contrast = "subadult - adult")

# mu 
mu_spp_lss_draws = model_select |>
  linpred_draws(
    newdata    = nd_spp_lss,
    dpar       = "mu",
    re_formula = NA
  ) %>% ungroup()

mu_spp_lss_contrast_draws = mu_spp_lss_draws |>
  pivot_wider(
    id_cols     = c(taxon_capture, .draw),
    names_from  = life_stage_simple,
    values_from = mu
  ) |>
  mutate(mu_contrast = subadult - adult,
         contrast = "subadult - adult")

# sigma 
sigma_spp_lss_draws = model_select |>
  linpred_draws(
    newdata    = nd_spp_lss,
    dpar       = "sigma",
    re_formula = NA
  ) %>% ungroup()

sigma_spp_draws = sigma_spp_lss_draws |>
  group_by(taxon_capture, .draw) |>
  summarise(sigma = mean(sigma), .groups = "drop")

f_draw_agg <- function(draws, group_vars, agg_var) {
  
  draws |>
    group_by(across(all_of(group_vars))) |>
    reframe(
      "{agg_var}_lower_50" := tidybayes::hdi(.data[[agg_var]], .width = 0.50)[, 1],
      "{agg_var}_upper_50" := tidybayes::hdi(.data[[agg_var]], .width = 0.50)[, 2],
      "{agg_var}_lower_95" := tidybayes::hdi(.data[[agg_var]], .width = 0.95)[, 1],
      "{agg_var}_upper_95" := tidybayes::hdi(.data[[agg_var]], .width = 0.95)[, 2],
      "{agg_var}_median"   := median(.data[[agg_var]]),
      "{agg_var}_net_effect" := (sum(.data[[agg_var]] > 0) / n()) * 2 - 1
    )
}

hu_spp_agg = f_draw_agg(hu_spp_draws, c("taxon_capture"), "hu")
prevalence_spp_agg = f_draw_agg(hu_spp_draws %>%
                                    mutate(prevalence = 1 - plogis(hu)), c("taxon_capture"), "prevalence")
sigma_spp_agg = f_draw_agg(sigma_spp_draws, c("taxon_capture"), "sigma")
mu_spp_lss_contrast_agg = f_draw_agg(mu_spp_lss_contrast_draws, c("taxon_capture", "contrast"), "mu_contrast")
hu_spp_lss_contrast_agg = f_draw_agg(hu_spp_lss_contrast_draws, c("taxon_capture", "contrast"), "hu_contrast")

# combine
spp_draws = hu_spp_draws %>%
  mutate(prevalence = 1 - plogis(hu)) %>%
  left_join(sigma_spp_draws, by = c(".draw", "taxon_capture"))


spp_lss_draws = hu_spp_lss_draws %>%
  select(taxon_capture,
         life_stage_simple,
         .draw,
         hu) %>%
  mutate(prevalence = 1 - plogis(hu)) %>%
  left_join(mu_spp_lss_draws %>%
              select(taxon_capture,
                     life_stage_simple,
                     .draw,
                     mu), by = c("taxon_capture", "life_stage_simple", ".draw")) %>%
  left_join(sigma_spp_lss_draws %>%
              select(taxon_capture,
                     life_stage_simple,
                     .draw,
                     sigma), by = c("taxon_capture", "life_stage_simple", ".draw"))


combined_agg = hu_spp_agg %>%
  left_join(prevalence_spp_agg, by = "taxon_capture") %>%
  left_join(sigma_spp_agg, by = "taxon_capture") %>%
  left_join(hu_spp_lss_contrast_agg, by = "taxon_capture") %>%
  left_join(mu_spp_lss_contrast_agg, by = c("contrast", "taxon_capture"))

combined_agg_single = combined_agg %>%
  select(taxon_capture,
         hu_median,
         hu_net_effect,
         prevalence_median,
         hu_contrast_median,
         hu_contrast_net_effect,
         mu_contrast_median,
         mu_contrast_net_effect,
         sigma_median) %>%
  distinct()

```

## Obsevations and predictions
```{r}

# # epred draws
# newdata_basic = expand_grid(taxon_capture = sort(unique(brms_data$taxon_capture)),
#                             life_stage_simple = c("subadult",
#                                                   "adult")) %>%
#   mutate(precip_mm_30_z = 0,
#          precip_mm_30_z2 = 0,
#          temp_c_30_z = 0,
#          temp_c_30_z2 = 0)

# # way too big, limit to 50 draws!
# epred_original = epred_draws(model_select,
#                              newdata = brms_data,
#                              re_formula = NA,
#                              ndraws = 200) %>%
#   dplyr::rename(bd_load_density_epred = .epred)

epred_fixed = epred_draws(model_select,
                          newdata = nd_spp_lss,
                          re_formula = NA) %>%
  dplyr::rename(bd_load_density_epred = .epred)

# observation data
obs_data = brms_data %>%
  group_by(taxon_capture,
           life_stage_simple) %>%
   summarise(n = n(),
             n_pos = sum(bd_detected_co),
             bd_prevalence = sum(bd_detected_co) / n(),
             bd_prevalence_ci_lower = wilson_ci_lower(bd_prevalence, n(), 0.95),
             bd_prevalence_ci_upper = wilson_ci_upper(bd_prevalence, n(), 0.95),
             min_load_inf = min(bd_load_co_density[bd_detected_co > 0]),
             q_25_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.25),
             q_50_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.50),
             q_75_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.75),
             max_load_inf = max(bd_load_co_density[bd_detected_co > 0]),
             se = sqrt(bd_prevalence*(1-bd_prevalence)/n),
             .groups = "drop") %>%
  mutate(label_offset = recode_values(life_stage_simple,
                                          "adult" ~ 0,
                                          "subadult" ~ 1))


```
## Visualize outputs

### Prevalence
```{r}
#| fig-height: 8
#| fig-width: 12

# observed
ggplot(obs_data, aes(y = bd_prevalence, x = life_stage_simple, color = life_stage_simple)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = bd_prevalence_ci_lower, ymax = bd_prevalence_ci_upper),
                width = 0.2,
                position = position_dodge(width = 0.5)) +
  geom_text(aes(label = paste0("n=", n), y = 1 - (1 - label_offset) * 0.05),
          position = position_dodge(width = 0.5),
          size = 4) +
  ylim(0, 1) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# modeled
ggplot(spp_lss_draws, aes(y = prevalence, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")


```

### Load
```{r}
#| fig-height: 8
#| fig-width: 12

# observed
ggplot(brms_data %>%
         filter(bd_detected_co > 0), aes(y = bd_load_co_density, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  geom_text(data = obs_data, aes(label = paste0("n=", n_pos), y = 10e8 - (1 - label_offset) * 7 * 10e7),
            position = position_dodge(width = 0.5),
            size = 4) +
  scale_y_log10(limits = c(1, 1e10)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")

# ggplot(epred_original, aes(y = bd_load_density_epred, x = life_stage_simple, fill = life_stage_simple)) +
#   geom_violin() +
#   # geom_boxplot() +
#   scale_y_log10(limits = c(1, 1e10)) +
#   theme_minimal() +
#   facet_wrap("taxon_capture") +
#   scale_fill_brewer(palette = "Set2")

ggplot(epred_fixed, aes(y = bd_load_density_epred, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  # geom_boxplot() +
  scale_y_log10(limits = c(1, 1e10)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")



```

### Hypothesis testing
```{r}
#| eval: true
#| echo: false
#| warning: false
#| fig-height: 8
#| fig-width: 12


# scatterplot draws
ggplot(spp_lss_draws, aes(x = prevalence, y = mu, color = life_stage_simple)) +
  geom_point(alpha = 0.1) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# single parameter distributions
ggplot(spp_lss_draws, aes(y = prevalence, x = life_stage_simple, color = life_stage_simple)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

ggplot(spp_lss_draws, aes(y = mu, x = life_stage_simple, color = life_stage_simple)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# visualize contrasts
ggplot(hu_spp_lss_contrast_draws, aes(x = hu_contrast)) +
  stat_halfeye(fill = "#66c2a5",
               .width = c(0.5, 0.95),
               point_interval = mode_hdi) +
  facet_wrap("taxon_capture") +
  geom_vline(xintercept = 0, alpha = 1) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect on Bd Prevalence",
       x = "Bd Prevalence Subadult Effect Size (logit scale) [-]\nwith 50% & 95% High Density Interval",
       y = "Posterior Density [-]") +
  theme(text=element_text(size=14))

ggplot(mu_spp_lss_contrast_draws, aes(x = mu_contrast)) +
    stat_halfeye(fill = "#8da0cb",
               .width = c(0.5, 0.95),
               point_interval = mode_hdi) +
  facet_wrap("taxon_capture") +
  geom_vline(xintercept = 0, alpha = 1) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect on Bd Prevalence",
       x = "Bd Load Density Subadult Effect Size (logit scale) [-]\nwith 50% & 95% High Density Interval",
       y = "Posterior Density [-]") +
  theme(text=element_text(size=14))

# net effect scatter
p_net_effect_scatter = ggplot(combined_agg_single, aes(x = -hu_contrast_net_effect, y = mu_contrast_net_effect)) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point() +
  geom_text_repel(aes(label = get_species_name(taxon_capture)),
                  color = "#333",
                  size = 3) +
  xlim(-1, 1) +
  ylim(-1, 1) +
  theme_minimal() +
  labs(
       title = "Net Subadult Effect (Subadult - Adult)",
       x = "Net contrast of log odds of Bd detection\n(subadult - adult) [-]",
       y = "Net contrast of log Bd load density\n(subadult - adult) [-]") +
  scale_color_brewer(palette = "Set2")

ggsave(here("size_bd_unified", "unified_figures", "03_net_effect_scatter.png"),
       p_net_effect_scatter,
       width = 8, height = 8)

p_net_effect_scatter

# median effect scatter of contrast
p_median_effect_scatter = ggplot(combined_agg_single, aes(x = -hu_contrast_median, y = mu_contrast_median)) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  theme_minimal() +
  labs(
       title = "Median Subadult Effect (Subadult - Adult)",
       x = "Median contrast of log odds of Bd detection\n(subadult - adult) [-]",
       y = "Median contrast of log Bd load density\n(subadult - adult) [-]") +
  scale_color_brewer(palette = "Set2")

ggsave(here("size_bd_unified", "unified_figures", "03_median_effect_scatter.png"),
       p_median_effect_scatter,
       width = 8, height = 8)

p_median_effect_scatter

# median effect scatter
ggplot(combined_agg_single, aes(x = prevalence_median, y = sigma_median)) +
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  theme_minimal() +
  labs(
       title = "Prevalence & Aggregation",
       x = "Median Bd Prevalence [-]",
       y = "Median Bd Load Aggregation -- log(sigma) [-]"
       ) +
  scale_color_brewer(palette = "Set2")

ggplot(combined_agg_single, aes(x = hu_contrast_median, y = hu_median)) +
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  theme_minimal() +
  labs(
       title = "Prevalence",
       x = "Median hu (subadult - adult) [-]",
       y = "Median hu [-]"
       ) +
  scale_color_brewer(palette = "Set2")

```

### Cascade plots
```{r}
# hu_spp_agg = f_draw_agg(hu_spp_lss_draws, c("taxon_capture"), "hu")
# sigma_spp_agg = f_draw_agg(sigma_spp_draws, c("taxon_capture"), "sigma")
# mu_spp_ls_contrast_agg = f_draw_agg(mu_spp_lss_contrast_draws, c("taxon_capture", "contrast"), "mu_contrast")
# hu_spp_ls_contrast_agg = f_draw_agg(hu_spp_lss_contrast_draws, c("taxon_capture", "contrast"), "hu_contrast")

p_cascade_hu_contrast_linpred = ggplot(hu_spp_lss_contrast_agg, aes(x = -hu_contrast_median, xmin = -hu_contrast_lower_95, xmax = -hu_contrast_upper_95, reorder(get_species_name(taxon_capture), hu_contrast_median))) +
  geom_vline(xintercept = 0, color = "#555") +
  geom_rect(aes(xmin = -hu_contrast_lower_50, xmax = -hu_contrast_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  theme_minimal() +
  labs(title = "Bd Prevalence Contrast (Subadult - Adult)",
       x = "Contrast of log odds of Bd detection (Subadult - Adult) [-]",
       y = "Species")

# p_cascade_hu_contrast = ggplot(hu_contrast_agg, aes(x = -hu_subadult_median, xmin = -hdi_95[,1], xmax = -hdi_95[,2], reorder(get_species_name(taxon_capture), hu_subadult_median))) +
#   geom_vline(xintercept = 0, color = "#555") +
#   geom_rect(aes(xmin = -hdi_50[,1], xmax = -hdi_50[,2]), height = 0.2, color = "#000") +
#   geom_pointrange() +
#   theme_minimal() +
#   labs(title = "Bd Prevalence Contrast (Subadult - Adult)",
#        x = "Contrast of log odds of Bd detection (Subadult - Adult) [-]",
#        y = "Species")

p_cascade_mu_contrast_linpred = ggplot(mu_spp_lss_contrast_agg, aes(x = mu_contrast_median, xmin = mu_contrast_lower_95, xmax = mu_contrast_upper_95, reorder(get_species_name(taxon_capture), -mu_contrast_median))) +
  geom_vline(xintercept = 0, color = "#555") +
  geom_rect(aes(xmin = mu_contrast_lower_50, xmax = mu_contrast_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  theme_minimal() +
  labs(title = "Bd Load Contrast (Subadult - Adult)",
       x = "Contrast of log Bd load (Subadult - Adult) [-]",
       y = "Species")

# p_cascade_mu_contrast = ggplot(b_contrast_agg, aes(x = b_subadult_median, xmin = hdi_95[,1], xmax = hdi_95[,2], reorder(get_species_name(taxon_capture), -b_subadult_median))) +
#   geom_vline(xintercept = 0, color = "#555") +
#   geom_rect(aes(xmin = hdi_50[,1], xmax = hdi_50[,2]), height = 0.2, color = "#000") +
#   geom_pointrange() +
#   theme_minimal() +
#   labs(title = "Bd Load Contrast (Subadult - Adult)",
#        x = "Contrast of log Bd load (Subadult - Adult) [-]",
#        y = "Species")

p_cascade_prevalence_linpred = ggplot(prevalence_spp_agg, aes(x = prevalence_median, xmin = prevalence_lower_95, xmax = prevalence_upper_95, reorder(get_species_name(taxon_capture), -prevalence_median))) +
  geom_rect(aes(xmin = prevalence_lower_50, xmax = prevalence_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  xlim(0, 1) +
  theme_minimal() +
  labs(title = "Bd Prevalence",
       x = "Prevalence draws [-]",
       y = "Species")


# p_cascade_prevalence = ggplot(hu_group_agg, aes(x = 1 - plogis(hu_median), xmin = 1 - plogis(hdi_95[,1]), xmax = 1 - plogis(hdi_95[,2]), reorder(get_species_name(taxon_capture), hu_median))) +
#   geom_rect(aes(xmin = 1 - plogis(hdi_50[,1]), xmax = 1 - plogis(hdi_50[,2])), height = 0.2, color = "#000") +
#   geom_pointrange() +
#   xlim(0, 1) +
#   theme_minimal() +
#   labs(title = "Bd Prevalence",
#        x = "Prevalence draws [-]",
#        y = "Species")

p_cascade_sigma_linpred = ggplot(sigma_spp_agg, aes(x = sigma_median, xmin = sigma_lower_95, xmax = sigma_upper_95, reorder(get_species_name(taxon_capture), -sigma_median))) +
  geom_rect(aes(xmin = sigma_lower_50, xmax = sigma_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  theme_minimal() +
  labs(title = "Bd Load Dispersion",
       x = "Log load standard deviation (sigma) draws [-]",
       y = "Species")

# p_cascade_sigma = ggplot(sigma_group_agg, aes(x = sigma_median, xmin = hdi_95[,1], xmax = hdi_95[,2], reorder(get_species_name(taxon_capture), -sigma_median))) +
#   geom_rect(aes(xmin = hdi_50[,1], xmax = hdi_50[,2]), height = 0.2, color = "#000") +
#   geom_pointrange() +
#   theme_minimal() +
#   labs(title = "Bd Load Dispersion",
#        x = "Log load standard deviation (sigma) draws [-]",
#        y = "Species")

# save plots
ggsave(here("size_bd_unified", "unified_figures", "03_species_cascade_hu_contrast.png"),
       p_cascade_hu_contrast_linpred,
       width = 8, height = 6)

ggsave(here("size_bd_unified", "unified_figures", "03_species_cascade_mu_contrast.png"),
       p_cascade_mu_contrast_linpred,
       width = 8, height = 6)

ggsave(here("size_bd_unified", "unified_figures", "03_species_cascade_prevalence.png"),
       p_cascade_prevalence_linpred,
       width = 8, height = 6)

ggsave(here("size_bd_unified", "unified_figures", "03_species_cascade_sigma.png"),
       p_cascade_sigma_linpred,
       width = 8, height = 6)


p_cascade_hu_contrast_linpred
p_cascade_mu_contrast_linpred
p_cascade_prevalence_linpred
p_cascade_sigma_linpred

```

# correlation analysis
```{r}
simple_cor = cor_mat(combined_agg_single %>%
                       select(hu_median,
                              sigma_median,
                              hu_contrast_median,
                              mu_contrast_median), method = "spearman")

cor_get_pval(simple_cor)


```

## Phylogenetic tree

### Jetz & Pyron 2018
https://datadryad.org/dataset/doi:10.5061/dryad.cc3n6j5#usage
```{r}
jp_raw = read.tree(file = here("data", "amph_shl_new_Consensus_7238.tre"))

# preview alignment
jk_species_list = bd_model_env_stats %>%
  mutate(sn = str_to_title(taxon_capture)) %>%
  select(sn) %>%
  mutate(genus = str_split_fixed(sn, "_", 2)[,1],
         species = str_split_fixed(sn, "_", 2)[,2])

jp_species_list = tibble(jp_sn = jp_raw$tip.label) %>%
  mutate(jp_genus = str_split_fixed(jp_sn, "_", 2)[,1],
         species = str_split_fixed(jp_sn, "_", 2)[,2],
         jp = TRUE)
  

peace = jk_species_list %>%
  left_join(jp_species_list, by = "species") %>%
  select(sn,
         jp_sn,
         genus,
         jp_genus,
         everything())

species_list_j = bd_model_env_stats %>%
  select(taxon_capture) %>%
  distinct() %>%
  mutate(sn = str_to_title(taxon_capture),
         jp_sn = recode_values(sn,
                               "Lithobates_warszewitschii" ~ "Rana_warszewitschii",
                               "Lithobates_blairi" ~ "Rana_blairi",
                               "Lithobates_chiricahuensis" ~ "Rana_chiricahuensis",
                               "Lithobates_palustris" ~ "Rana_palustris",
                               "Lithobates_sphenocephalus" ~ "Rana_sphenocephala",
                               default = sn))

# Prune to species list
jp_pruned = keep.tip(jp_raw, species_list_j$jp_sn)

combined_agg_j = combined_agg_single %>%
  left_join(species_list_j %>%
              select(taxon_capture,
                     jp_sn), by = "taxon_capture")

# Ccheck plot
plotTree(jp_pruned, fsize = 0.8)

is.binary(jp_pruned)

table(sapply(1:jp_pruned$Nnode, function(n) {
  length(Children(jp_pruned, n + Ntip(jp_pruned)))
}))

```

### Analyze
```{r}
tr_rooted = read.tree(here("size_bd_unified", "intermediates", "phylo_tree.nwk"))

# calculate blomberg K
blomberg_k = function(tree, var_values, species) {
  set_named = setNames(var_values, species)
  return(blomberg_k_sigma_med = phylosig(tree, set_named, method = "K", test = TRUE, nsim = 1000))
}

pagel_lambda = function(tree, var_values, species) {
  set_named = setNames(var_values, species)
  return(blomberg_k_sigma_med = phylosig(tree, set_named, method = "lambda", test = TRUE, nsim = 1000))
}

# blomberg k
hu_contrast_med_k = blomberg_k(jp_pruned, combined_agg_j$hu_contrast_median, combined_agg_j$jp_sn)
mu_contrast_med_k = blomberg_k(jp_pruned, combined_agg_j$mu_contrast_median, combined_agg_j$jp_sn)
prev_med_k = blomberg_k(jp_pruned, combined_agg_j$prevalence_median, combined_agg_j$jp_sn)
sigma_med_k = blomberg_k(jp_pruned, combined_agg_j$sigma_median, combined_agg_j$jp_sn)

hu_contrast_med_k
mu_contrast_med_k
prev_med_k
sigma_med_k

# pagel lambda
hu_contrast_med_lambda = pagel_lambda(jp_pruned, combined_agg_j$hu_contrast_median, combined_agg_j$jp_sn)
mu_contrast_med_lambda = pagel_lambda(jp_pruned, combined_agg_j$mu_contrast_median, combined_agg_j$jp_sn)
prev_med_lambda = pagel_lambda(jp_pruned, combined_agg_j$prevalence_median, combined_agg_j$jp_sn)
sigma_med_lambda = pagel_lambda(jp_pruned, combined_agg_j$sigma_median, combined_agg_j$jp_sn)

hu_contrast_med_lambda
mu_contrast_med_lambda
prev_med_lambda
sigma_med_lambda

# define color maps
sym_div_cont_map = function(tree, var_values, species, center, p_title, style = "div") {
  set_named = setNames(var_values, species)
  
  sym_lim = max(abs(center - min(var_values)), abs(center - max(var_values)))
  
  map_1 = contMap(tree, set_named,
                   lims = c(center - sym_lim, center + sym_lim),
                   title = p_title, fsize = 0.4)
  
  if (style == "div") {
    colors = colorRampPalette(c("#762a83", "white", "#1b7837"))(101)
  } else if (style == "lin") {
    colors = colorRampPalette(c("white", "#1b7837"))(101)
  }
  map_2 = setMap(map_1, colors = colors)
  
  return(map_2)
}

p_hu_contrast_med_map = sym_div_cont_map(jp_pruned,
                                  combined_agg_j$hu_contrast_median,
                                  combined_agg_j$jp_sn,
                                  0,
                                  "Bd prevalence (hu): median subadult effect")

p_mu_contrast_med_map = sym_div_cont_map(jp_pruned,
                                 combined_agg_j$mu_contrast_median,
                                 combined_agg_j$jp_sn,
                                 0,
                                 "Bd log load (mu): median subadult effect")

p_prev_med_map = sym_div_cont_map(jp_pruned,
                                combined_agg_j$prevalence_median,
                                combined_agg_j$jp_sn,
                                0.5,
                                "Median Bd prevalence")

p_sigma_med_map = sym_div_cont_map(jp_pruned,
                                 combined_agg_j$sigma_median,
                                 combined_agg_j$jp_sn,
                                 0.8822,
                                 "Median Bd load dispersion (sigma)")

png(file = here("size_bd_unified", "unified_figures", "03_tree_contmap_hu_contrast.png"),
    width = 6, height = 8, units = "in", res = 300)
plot(p_hu_contrast_med_map)
dev.off()

png(file = here("size_bd_unified", "unified_figures", "03_tree_contmap_mu_contrast.png"),
    width = 6, height = 8, units = "in", res = 300)
plot(p_mu_contrast_med_map)
dev.off()


png(file = here("size_bd_unified", "unified_figures", "03_tree_contmap_prevalence.png"),
    width = 6, height = 8, units = "in", res = 300)
plot(p_prev_med_map)
dev.off()


png(file = here("size_bd_unified", "unified_figures", "03_tree_contmap_sigma.png"),
    width = 6, height = 8, units = "in", res = 300)
plot(p_sigma_med_map)
dev.off()


plot(p_hu_contrast_med_map)
plot(p_mu_contrast_med_map)
plot(p_prev_med_map)
plot(p_sigma_med_map)

```

## Collective results
```{r}
#| eval: true
#| echo: true
#| output: false
#| code-fold: true
#| code-summary: Show code

# m5g_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2.rds")


model_data = brms_data

draws_all = collect_draws(model_select) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

density_mat_all = density_mat(log(0.5), 20, 100, draws_all) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

sim_load_all = simulate_bd_load(draws_all, 10, 1000) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

obs_stats_all = sample_stats(model_data) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

```


### Modeled Bd load PDF
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# PDF -- facet by species
ggplot(density_mat_all, aes(x = density,
                            y = pdf_median,
                            color = life_stage_simple, 
                            fill = life_stage_simple)) +
  geom_ribbon(aes(ymin = pdf_lower,
                  ymax = pdf_upper), 
              alpha = 0.15,
              color = NA) +
  geom_line() +
  theme_minimal() +
  ggtitle("Modeled PDF of Bd-positive load") +
  labs(color = "Life stage",
       fill = "Life stage") +
  xlab("log(Bd load) [-]") +
  ylab("Median probability density [-]\n(with 25th to 75th percentiles)") +
  scale_x_continuous(limits = c(log(0.5), 20), expand = c(0, 0)) +
  facet_grid(. ~ gsub(" ", "\n", species))

# # PDF -- facet by life_stage & season
# ggplot(density_mat_all, aes(x = density,
#                             y = pdf_median,
#                             color = taxon_capture, 
#                             fill = taxon_capture)) +
#   geom_ribbon(aes(ymin = pdf_lower,
#                   ymax = pdf_upper), 
#               alpha = 0.15,
#               color = NA) +
#   geom_line() +
#   theme_minimal() +
#   ggtitle("Modeled PDF of Bd-positive load") +
#   labs(color = "Species",
#        fill = "Species") +
#   xlab("log(Bd load) [-]") +
#   ylab("Median probability density [-]\n(with 25th to 75th percentiles)") +
#   scale_x_continuous(limits = c(log(0.3), 20), expand = c(0, 0)) +
#   facet_grid(life_stage_simple ~ season)
```

### Moded Bd load CCDF
```{r}
#| echo: false
#| warning: false
#| fig-height: 8
#| fig-width: 8
# CCDF -- facet by species
ggplot(density_mat_all, aes(x = density,
                            y = ccdf_median,
                            color = life_stage_simple, 
                            fill = life_stage_simple)) +
  geom_ribbon(aes(ymin = ccdf_lower,
                  ymax = ccdf_upper), 
              alpha = 0.15,
              color = NA) +
  geom_line() +
  theme_minimal() +
  ggtitle("Modeled CCDF of Bd-positive load") +
  labs(color = "Life stage",
       fill = "Life stage") +
  xlab("log(Bd load) [-]") +
  ylab("Median complementary cumulative density (1 - CDF) [-]\n(with 25th to 75th percentiles)") +
  scale_x_continuous(limits = c(log(0.5), 12), expand = c(0, 0)) +
  ylim(0, 0.75) +
  # facet_grid(. ~ gsub(" ", "\n", species))
  facet_wrap("species")

# CCDF -- facet by life_stage & season
# ggplot(density_mat_all, aes(x = density,
#                             y = ccdf_median,
#                             color = taxon_capture,
#                             fill = taxon_capture)) +
#   geom_ribbon(aes(ymin = ccdf_lower,
#                   ymax = ccdf_upper),
#               alpha = 0.15,
#               color = NA) +
#   geom_line() +
#   theme_minimal() +
#   ggtitle("Modeled CCDF of Bd-positive load") +
#   labs(color = "Species",
#        fill = "Species") +
#   xlab("log(Bd load) [-]") +
#   ylab("Median complementary cumulative density (1 - CDF) [-]\n(with 25th to 75th percentiles)") +
#   scale_x_continuous(limits = c(log(0.3), 20), expand = c(0, 0)) +
#   ylim(0, 1) +
#   facet_grid(. ~ life_stage_simple)
```

# Environmental conditions

## Distributions
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Environmental conditions distributions code"
#| fig-height: 6
#| fig-width: 8

# 15-day
# ggplot(brms_data, aes(x = temp_c_15_day_mean, y = precip_mm_15_day_sum, color = life_stage_simple)) +
#   geom_point(alpha = 0.1) +
#   stat_ellipse() +
#   facet_wrap(. ~ taxon_capture) +
#   xlim(-5, 40) +
#   ylim(0, 200) +
#   labs(x = "15-day temperature mean (degC)", 
#        y = "15-day precipitation total (mm)",
#        title = "Temperature v. precipitation") +
#   theme_minimal()

# 10 day precip, 30-day temp by species
ggplot(brms_data, aes(x = temp_c_30_day_mean, y = precip_mm_10_day_sum, color = life_stage_simple)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  facet_wrap(. ~ taxon_capture) +
  xlim(-5, 40) +
  ylim(0, 200) +
  labs(x = "10-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()

# 30 day precip, 30-day temp by species
ggplot(brms_data, aes(x = temp_c_30_day_mean, y = precip_mm_30_day_sum, color = life_stage_simple)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  facet_wrap(. ~ taxon_capture) +
  xlim(-5, 40) +
  ylim(0, 400) +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()

# 30 day collective
ggplot(brms_data, aes(x = temp_c_30_day_mean, y = precip_mm_30_day_sum, color = taxon_capture)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()

# 30 day collective, scaled to climate by species
ggplot(brms_data, aes(x = temp_c_30_sp_z, y = precip_mm_30_sp_z, color = taxon_capture)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()


# ggplot(brms_data, aes(x = temp_c_30_sp_z2, y = precip_mm_30_sp_z2, color = life_stage_simple)) +
#   geom_point(alpha = 0.1) +
#   stat_ellipse() +
#   facet_wrap(. ~ taxon_capture) +
#   labs(x = "30-day temperature mean (degC)", 
#        y = "30-day precipitation total (mm)",
#        title = "Temperature v. precipitation") +
#   theme_minimal()
```

## Climate

### Precip
```{r}

apply_scale <- function(x_new, sc_obj) {
  (x_new - attr(sc_obj, "scaled:center")) / attr(sc_obj, "scaled:scale")
}

remove_scale = function(x_sc, sc_obj) {
  x_sc * attr(sc_obj, "scaled:scale") + attr(sc_obj, "scaled:center")
}

precip_10_sc = scale(brms_data$precip_mm_10_day_sum)
temp_30_sc = scale(brms_data$temp_c_30_day_mean)

# define sequences over real-world ranges
precip_seq <- seq(min(brms_data$precip_mm_10_day_sum),
                  max(brms_data$precip_mm_10_day_sum),
                  length.out = 100)

# scale them using the original parameters
nd_precip <- model_select$data |>
  distinct(taxon_capture, life_stage_simple) |>
  cross_join(tibble(precip_mm_10_day_sum = precip_seq)) |>
  mutate(
    temp_c_30_day_mean   = remove_scale(0, temp_30_sc),
    precip_mm_10_z  = apply_scale(precip_mm_10_day_sum, precip_10_sc),
    temp_c_30_z     = apply_scale(temp_c_30_day_mean, temp_30_sc)
  )

# hu
hu_draws_precip <- model_select |>
  epred_draws(
    newdata    = nd_precip,
    dpar       = "hu",
    re_formula = ~ (1 | taxon_capture:life_stage_simple)
  )

hu_agg_precip = hu_draws_precip |>
  mutate(prevalence = 1 - plogis(hu)) |>
  group_by(taxon_capture, precip_mm_10_day_sum) |>
  median_qi(prevalence) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

# mu
mu_draws_precip <- model_select |>
  epred_draws(
    newdata    = nd_precip,
    dpar       = "mu",
    re_formula = ~ (1 | taxon_capture:life_stage_simple)
  )

mu_agg_precip = mu_draws_precip |>
  group_by(taxon_capture, precip_mm_10_day_sum) |>
  median_qi(mu) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

# visualize
p_precip_prev = ggplot(hu_agg_precip, aes(x = precip_mm_10_day_sum, y = prevalence, ymin = .lower, ymax = .upper)) +
  geom_line() +
  geom_ribbon(alpha = 0.3) +
  facet_wrap("species") +
  theme_minimal() +
  labs(title = "Bd Prevalence vs. Precipitation",
       x = "10-day lag precipitation [mm]",
       y = "Bd prevalence [-]")

p_precip_mu = ggplot(mu_agg_precip, aes(x = precip_mm_10_day_sum, y = mu, ymin = .lower, ymax = .upper)) +
  geom_line() +
  geom_ribbon(alpha = 0.3) +
  facet_wrap("species") +
  theme_minimal() +
  labs(title = "Bd Load vs. Precipitation",
       x = "10-day lag precipitation [mm]",
       y = "log(Bd load) [-]")

ggsave(here("size_bd_unified", "unified_figures", "03_precip_prevalence.png"),
       p_precip_prev,
       width = 8, height = 8)

ggsave(here("size_bd_unified", "unified_figures", "03_precip_mu.png"),
       p_precip_mu,
       width = 8, height = 8)

p_precip_prev
p_precip_mu

```

### Temp
```{r}
temp_seq   <- seq(min(brms_data$temp_c_30_day_mean),
                  max(brms_data$temp_c_30_day_mean),
                  length.out = 100)

nd_temp <- model_select$data |>
  distinct(taxon_capture, life_stage_simple) |>
  cross_join(tibble(temp_c_30_day_mean = temp_seq)) |>
  mutate(
    precip_mm_10_day_sum   = remove_scale(0, precip_10_sc),
    precip_mm_10_z  = apply_scale(precip_mm_10_day_sum, precip_10_sc),
    temp_c_30_z     = apply_scale(temp_c_30_day_mean, temp_30_sc),
  )


# hu
hu_draws_temp <- model_select |>
  epred_draws(
    newdata    = nd_temp,
    dpar       = "hu",
    re_formula = ~ (1 | taxon_capture:life_stage_simple)
  )

hu_agg_temp = hu_draws_temp |>
  mutate(prevalence = 1 - plogis(hu)) |>
  group_by(taxon_capture, temp_c_30_day_mean) |>
  median_qi(prevalence) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

# mu
mu_draws_temp <- model_select |>
  epred_draws(
    newdata    = nd_temp,
    dpar       = "mu",
    re_formula = ~ (1 | taxon_capture:life_stage_simple)
  )

mu_agg_temp = mu_draws_temp |>
  group_by(taxon_capture, temp_c_30_day_mean) |>
  median_qi(mu) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

# visualize
p_temp_prev = ggplot(hu_agg_temp, aes(x = temp_c_30_day_mean, y = prevalence, ymin = .lower, ymax = .upper)) +
  geom_line() +
  geom_ribbon(alpha = 0.3) +
  facet_wrap("species") +
  theme_minimal() +
  ylim(0, 1) +
  labs(title = "Bd Prevalence vs. Temperature",
       x = "30-day lag max daily temperature [C]",
       y = "Bd prevalence [-]")

p_temp_mu = ggplot(mu_agg_temp, aes(x = temp_c_30_day_mean, y = mu, ymin = .lower, ymax = .upper)) +
  geom_line() +
  geom_ribbon(alpha = 0.3) +
  facet_wrap("species") +
  theme_minimal() +
  labs(title = "Bd Load vs. Temperature",
       x = "30-day lag max daily temperature [C]",
       y = "log(Bd load) [-]")

ggsave(here("size_bd_unified", "unified_figures", "03_temp_prevalence.png"),
       p_temp_prev,
       width = 8, height = 8)

ggsave(here("size_bd_unified", "unified_figures", "03_temp_mu.png"),
       p_temp_mu,
       width = 8, height = 8)

p_temp_prev
p_temp_mu

```