---
title: "size_bd_unified_analysis"
format: html
---

# Load data & models
```{r}
librarian::shelf(tidyverse, here, janitor, brms, tidybayes, ggrepel, emmeans, rstatix, phylotools, rentrez, DECIPHER, phangorn, phytools, RColorBrewer, arrow, patchwork)

# load bd_model_env
load(here("size_bd_unified", "intermediates", "bd_model_env_unified_2026-04-22.RData"))

size_cutoff = read_csv(here("data", "Species notes - size cutoffs.csv"))

bd_model_env_stats = brms_data %>%
  group_by(taxon_capture, country) %>%
  summarise(regions = paste(sort(unique(region)), collapse = ", "),
            population_n = n_distinct(population),
            site_n = n_distinct(site),
            year_first = year(min(date)),
            year_last = year(max(date)),
            year_count = n_distinct(year(date)),
            adult_n = sum(life_stage_simple == "adult"),
            adult_n_pos = sum(life_stage_simple == "adult" & bd_detected_co),
            subadult_n = sum(life_stage_simple == "subadult"),
            subadult_n_pos = sum(life_stage_simple == "subadult" & bd_detected_co),
            .groups = "drop") %>%
  arrange(country,
          taxon_capture) %>%
  left_join(size_cutoff, by = "taxon_capture")

bd_model_env_stats_sub = bd_model_env_stats %>%
  select(taxon_capture,
         country,
         regions,
         adult_n,
         subadult_n)

# load models
# null_zln = readRDS(file = here("size_bd_unified/unified_models/unified_model_results_zln_null.rds"))
# m9a_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_env2_renvpoptls_p10t30.rds")
# m9b_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_env2_renv2poptls_p10t30.rds")
# m9c_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_pt2_rpt2poptls_p10t30.rds")
# m9d_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_pt2_rpt2poptlsyr_p10t30.rds")
# m9b_30_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_env2_renv2poptls_p30t30.rds")
# m9f_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_pt2_rpt2poptlsyrclt_p10t30.rds")
# m9h_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_pt2_rpt2poptlsyrkg2_p10t30.rds")
m9i_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_pt2_rpt2poptlsyrkg3_p10t30.rds")

model_select = m9i_zln
```

## Define analysis functions
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Show processing & visualization functions"

get_species_name = function(taxon, newline = FALSE) {
  species_name = str_to_sentence(gsub("_", " ", gsub("usgs", "", taxon)))
  if (newline) {
    species_name = gsub(" ", "\n", species_name)
  }
  return(species_name)
}

wilson_ci = function(p_hat, n, conf.level = 0.95) {
  z <- qnorm(1 - (1 - conf.level) / 2)
  
  denominator <- 1 + (z^2 / n)
  centre_adjusted_probability <- p_hat + (z^2 / (2 * n))
  adjusted_standard_deviation <- sqrt((p_hat * (1 - p_hat) / n) + (z^2 / (4 * n^2)))
  
  lower_bound <- (centre_adjusted_probability - z * adjusted_standard_deviation) / denominator
  upper_bound <- (centre_adjusted_probability + z * adjusted_standard_deviation) / denominator
  
  return(c(lower_bound = lower_bound, upper_bound = upper_bound))
}

wilson_ci_lower = function(p_hat, n, conf.level = 0.95) {
  ws = wilson_ci(p_hat, n, conf.level)
  return(ws[1])
}

wilson_ci_upper = function(p_hat, n, conf.level = 0.95) {
  ws = wilson_ci(p_hat, n, conf.level)
  return(ws[2])
}

mmode = function(x) {
  ux = unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

```

# Results

## Organize draws

### Linpred Draws
```{r}

# define linpred data
nd_env_subadult = model_select$data %>%
  group_by(taxon_capture, life_stage_simple) %>%
  summarize(precip_mm_10_z = median(precip_mm_10_z),
            temp_c_30_z = median(temp_c_30_z),
            kg_code_3 = mmode(kg_code_3),
            .groups = "drop") %>%
  filter(life_stage_simple == "subadult")

nd_spp_lss = model_select$data |>
  distinct(taxon_capture, life_stage_simple) %>%
  left_join(nd_env_subadult %>%
              select(-life_stage_simple), by = "taxon_capture")


# hu contrast
hu_spp_lss_draws = model_select |>
  linpred_draws(
    newdata    = nd_spp_lss,
    dpar       = "hu",
    re_formula = NA
  ) %>%
  ungroup()

hu_spp_draws = hu_spp_lss_draws |>
  group_by(taxon_capture, .draw) |>
  summarise(hu = mean(hu), .groups = "drop")

hu_spp_lss_contrast_draws = hu_spp_lss_draws |>
  pivot_wider(
    id_cols     = c(taxon_capture, .draw),
    names_from  = life_stage_simple,
    values_from = hu
  ) |>
  mutate(hu_contrast = subadult - adult,
         contrast = "subadult - adult")

# mu 
mu_spp_lss_draws = model_select |>
  linpred_draws(
    newdata    = nd_spp_lss,
    dpar       = "mu",
    re_formula = NA
  ) %>% ungroup()

mu_spp_lss_contrast_draws = mu_spp_lss_draws |>
  pivot_wider(
    id_cols     = c(taxon_capture, .draw),
    names_from  = life_stage_simple,
    values_from = mu
  ) |>
  mutate(mu_contrast = subadult - adult,
         contrast = "subadult - adult")

# sigma 
sigma_spp_lss_draws = model_select |>
  linpred_draws(
    newdata    = nd_spp_lss,
    dpar       = "sigma",
    re_formula = NA
  ) %>% ungroup()

sigma_spp_draws = sigma_spp_lss_draws |>
  group_by(taxon_capture, .draw) |>
  summarise(sigma = mean(sigma), .groups = "drop")

# f_draw_agg <- function(draws, group_vars, agg_var) {
#   
#   draws |>
#     group_by(across(all_of(group_vars))) |>
#     reframe(
#       "{agg_var}_lower_50" := tidybayes::hdi(.data[[agg_var]], .width = 0.50)[, 1],
#       "{agg_var}_upper_50" := tidybayes::hdi(.data[[agg_var]], .width = 0.50)[, 2],
#       "{agg_var}_lower_95" := tidybayes::hdi(.data[[agg_var]], .width = 0.95)[, 1],
#       "{agg_var}_upper_95" := tidybayes::hdi(.data[[agg_var]], .width = 0.95)[, 2],
#       "{agg_var}_median"   := median(.data[[agg_var]]),
#       "{agg_var}_net_effect" := (sum(.data[[agg_var]] > 0) / n()) * 2 - 1
#     )
# }
f_draw_agg <- function(draws, group_vars, agg_var) {
  
  draws |>
    group_by(across(all_of(group_vars))) |>
    reframe(
      "{agg_var}_lower_50" := tidybayes::median_qi(.data[[agg_var]], .width = 0.50)[, 2],
      "{agg_var}_upper_50" := tidybayes::median_qi(.data[[agg_var]], .width = 0.50)[, 3],
      "{agg_var}_lower_95" := tidybayes::median_qi(.data[[agg_var]], .width = 0.95)[, 2],
      "{agg_var}_upper_95" := tidybayes::median_qi(.data[[agg_var]], .width = 0.95)[, 3],
      "{agg_var}_median"   := median(.data[[agg_var]]),
      "{agg_var}_net_effect" := (sum(.data[[agg_var]] > 0) / n()) * 2 - 1
    )
}

hu_spp_agg = f_draw_agg(hu_spp_draws, c("taxon_capture"), "hu")
prevalence_spp_agg = f_draw_agg(hu_spp_draws %>%
                                    mutate(prevalence = 1 - plogis(hu)), c("taxon_capture"), "prevalence")
sigma_spp_agg = f_draw_agg(sigma_spp_draws, c("taxon_capture"), "sigma")
mu_spp_lss_contrast_agg = f_draw_agg(mu_spp_lss_contrast_draws, c("taxon_capture", "contrast"), "mu_contrast")
hu_spp_lss_contrast_agg = f_draw_agg(hu_spp_lss_contrast_draws, c("taxon_capture", "contrast"), "hu_contrast")

# combine
spp_draws = hu_spp_draws %>%
  mutate(prevalence = 1 - plogis(hu)) %>%
  left_join(sigma_spp_draws, by = c(".draw", "taxon_capture"))


spp_lss_draws = hu_spp_lss_draws %>%
  select(taxon_capture,
         life_stage_simple,
         .draw,
         hu) %>%
  mutate(prevalence = 1 - plogis(hu)) %>%
  left_join(mu_spp_lss_draws %>%
              select(taxon_capture,
                     life_stage_simple,
                     .draw,
                     mu), by = c("taxon_capture", "life_stage_simple", ".draw")) %>%
  left_join(sigma_spp_lss_draws %>%
              select(taxon_capture,
                     life_stage_simple,
                     .draw,
                     sigma), by = c("taxon_capture", "life_stage_simple", ".draw"))

write_parquet(spp_lss_draws, here("size_bd_unified", "intermediates", "unified_03_b_spp_lss_draws.parquet"))

combined_agg = hu_spp_agg %>%
  left_join(prevalence_spp_agg, by = "taxon_capture") %>%
  left_join(sigma_spp_agg, by = "taxon_capture") %>%
  left_join(hu_spp_lss_contrast_agg, by = "taxon_capture") %>%
  left_join(mu_spp_lss_contrast_agg, by = c("contrast", "taxon_capture"))

combined_agg_single = combined_agg %>%
  select(taxon_capture,
         hu_median,
         hu_net_effect,
         prevalence_median,
         hu_contrast_median,
         hu_contrast_net_effect,
         mu_contrast_median,
         mu_contrast_net_effect,
         sigma_median) %>%
  distinct()

```

## Obsevations and predictions
```{r}

# # epred draws
# newdata_basic = expand_grid(taxon_capture = sort(unique(brms_data$taxon_capture)),
#                             life_stage_simple = c("subadult",
#                                                   "adult")) %>%
#   mutate(precip_mm_30_z = 0,
#          precip_mm_30_z2 = 0,
#          temp_c_30_z = 0,
#          temp_c_30_z2 = 0)

# # way too big, limit to 50 draws!
# epred_original = epred_draws(model_select,
#                              newdata = brms_data,
#                              re_formula = NA,
#                              ndraws = 200) %>%
#   dplyr::rename(bd_load_density_epred = .epred)

epred_fixed = epred_draws(model_select,
                          newdata = nd_spp_lss,
                          re_formula = NA) %>%
  dplyr::rename(bd_load_density_epred = .epred)

# observation data
obs_data = brms_data %>%
  group_by(taxon_capture,
           life_stage_simple) %>%
   summarise(n = n(),
             n_pos = sum(bd_detected_co),
             bd_prevalence = sum(bd_detected_co) / n(),
             bd_prevalence_ci_lower = wilson_ci_lower(bd_prevalence, n(), 0.95),
             bd_prevalence_ci_upper = wilson_ci_upper(bd_prevalence, n(), 0.95),
             min_load_inf = min(bd_load_co_density[bd_detected_co > 0]),
             q_25_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.25),
             q_50_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.50),
             q_75_load_inf = quantile(bd_load_co_density[bd_detected_co > 0], probs = 0.75),
             max_load_inf = max(bd_load_co_density[bd_detected_co > 0]),
             se = sqrt(bd_prevalence*(1-bd_prevalence)/n),
             .groups = "drop") %>%
  mutate(label_offset = recode_values(life_stage_simple,
                                          "adult" ~ 0,
                                          "subadult" ~ 1))


```
## Visualize outputs

### Prevalence
```{r}
#| fig-height: 8
#| fig-width: 12

# observed
ggplot(obs_data, aes(y = bd_prevalence, x = life_stage_simple, color = life_stage_simple)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = bd_prevalence_ci_lower, ymax = bd_prevalence_ci_upper),
                width = 0.2,
                position = position_dodge(width = 0.5)) +
  geom_text(aes(label = paste0("n=", n), y = 1 - (1 - label_offset) * 0.05),
          position = position_dodge(width = 0.5),
          size = 4) +
  ylim(0, 1) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# modeled
ggplot(spp_lss_draws, aes(y = prevalence, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")


```

### Load
```{r}
#| fig-height: 8
#| fig-width: 12

# observed
ggplot(brms_data %>%
         filter(bd_detected_co > 0), aes(y = bd_load_co_density, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  geom_text(data = obs_data, aes(label = paste0("n=", n_pos), y = 10e8 - (1 - label_offset) * 7 * 10e7),
            position = position_dodge(width = 0.5),
            size = 4) +
  scale_y_log10(limits = c(1, 1e10)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")

# ggplot(epred_original, aes(y = bd_load_density_epred, x = life_stage_simple, fill = life_stage_simple)) +
#   geom_violin() +
#   # geom_boxplot() +
#   scale_y_log10(limits = c(1, 1e10)) +
#   theme_minimal() +
#   facet_wrap("taxon_capture") +
#   scale_fill_brewer(palette = "Set2")

ggplot(epred_fixed, aes(y = bd_load_density_epred, x = life_stage_simple, fill = life_stage_simple)) +
  geom_violin() +
  # geom_boxplot() +
  scale_y_log10(limits = c(1, 1e10)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_fill_brewer(palette = "Set2")



```

### Hypothesis testing
```{r}
#| eval: true
#| echo: false
#| warning: false
#| fig-height: 8
#| fig-width: 12


# scatterplot draws
ggplot(spp_lss_draws, aes(x = prevalence, y = mu, color = life_stage_simple)) +
  geom_point(alpha = 0.1) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# single parameter distributions
ggplot(spp_lss_draws, aes(y = prevalence, x = life_stage_simple, color = life_stage_simple)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

ggplot(spp_lss_draws, aes(y = mu, x = life_stage_simple, color = life_stage_simple)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  theme_minimal() +
  facet_wrap("taxon_capture") +
  scale_color_brewer(palette = "Set2")

# visualize contrasts
ggplot(hu_spp_lss_contrast_draws, aes(x = hu_contrast)) +
  stat_halfeye(fill = "#66c2a5",
               .width = c(0.5, 0.95),
               point_interval = mode_hdi) +
  facet_wrap("taxon_capture") +
  geom_vline(xintercept = 0, alpha = 1) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect on Bd Prevalence",
       x = "Bd Prevalence Subadult Effect Size (logit scale) [-]\nwith 50% & 95% High Density Interval",
       y = "Posterior Density [-]") +
  theme(text=element_text(size=14))

ggplot(mu_spp_lss_contrast_draws, aes(x = mu_contrast)) +
    stat_halfeye(fill = "#8da0cb",
               .width = c(0.5, 0.95),
               point_interval = mode_hdi) +
  facet_wrap("taxon_capture") +
  geom_vline(xintercept = 0, alpha = 1) +
  theme_minimal() +
  labs(
       # title = "Subadult Effect on Bd Prevalence",
       x = "Bd Load Density Subadult Effect Size (logit scale) [-]\nwith 50% & 95% High Density Interval",
       y = "Posterior Density [-]") +
  theme(text=element_text(size=14))

# net effect scatter
p_net_effect_scatter = ggplot(combined_agg_single, aes(x = -hu_contrast_net_effect, y = mu_contrast_net_effect)) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point() +
  geom_text_repel(aes(label = get_species_name(taxon_capture)),
                  color = "#333",
                  size = 3) +
  xlim(-1, 1) +
  ylim(-1, 1) +
  theme_minimal() +
  labs(
       title = "Net Subadult Effect (Subadult - Adult)",
       x = "Net contrast of log odds of Bd detection\n(subadult - adult) [-]",
       y = "Net contrast of log Bd load density\n(subadult - adult) [-]") +
  scale_color_brewer(palette = "Set2")

ggsave(here("size_bd_unified", "unified_figures", "03_net_effect_scatter.png"),
       p_net_effect_scatter,
       width = 8, height = 8)

# median effect scatter of contrast
p_median_effect_scatter = ggplot(combined_agg_single, aes(x = -hu_contrast_median, y = mu_contrast_median)) +
  geom_vline(xintercept = 0, color = "#aaa") +
  geom_hline(yintercept = 0, color = "#aaa") + 
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  theme_minimal() +
  labs(
       title = "Median Subadult Effect (Subadult - Adult)",
       x = "Median contrast of log odds of Bd detection\n(subadult - adult) [-]",
       y = "Median contrast of log Bd load density\n(subadult - adult) [-]") +
  scale_color_brewer(palette = "Set2")

ggsave(here("size_bd_unified", "unified_figures", "03_median_effect_scatter.png"),
       p_median_effect_scatter,
       width = 8, height = 8)


p_net_effect_scatter
p_median_effect_scatter

# median effect scatter
ggplot(combined_agg_single, aes(x = prevalence_median, y = sigma_median)) +
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  theme_minimal() +
  labs(
       title = "Prevalence & Aggregation",
       x = "Median Bd Prevalence [-]",
       y = "Median Bd Load Aggregation -- log(sigma) [-]"
       ) +
  scale_color_brewer(palette = "Set2")

ggplot(combined_agg_single, aes(x = hu_contrast_median, y = hu_median)) +
  geom_point() +
  geom_text_repel(aes(label = taxon_capture),
                  color = "#888") +
  theme_minimal() +
  labs(
       title = "Prevalence",
       x = "Median hu (subadult - adult) [-]",
       y = "Median hu [-]"
       ) +
  scale_color_brewer(palette = "Set2")

```

### Cascade plots
```{r}
# hu_spp_agg = f_draw_agg(hu_spp_lss_draws, c("taxon_capture"), "hu")
# sigma_spp_agg = f_draw_agg(sigma_spp_draws, c("taxon_capture"), "sigma")
# mu_spp_ls_contrast_agg = f_draw_agg(mu_spp_lss_contrast_draws, c("taxon_capture", "contrast"), "mu_contrast")
# hu_spp_ls_contrast_agg = f_draw_agg(hu_spp_lss_contrast_draws, c("taxon_capture", "contrast"), "hu_contrast")

p_cascade_hu_contrast_linpred = ggplot(hu_spp_lss_contrast_agg, aes(x = -hu_contrast_median, xmin = -hu_contrast_lower_95, xmax = -hu_contrast_upper_95, reorder(get_species_name(taxon_capture), hu_contrast_median))) +
  geom_vline(xintercept = 0, color = "#555") +
  geom_rect(aes(xmin = -hu_contrast_lower_50, xmax = -hu_contrast_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  theme_minimal() +
  labs(title = "Bd Prevalence Contrast (Subadult - Adult)",
       x = "Contrast of log odds of Bd detection (Subadult - Adult) [-]",
       y = "Species")

# p_cascade_hu_contrast = ggplot(hu_contrast_agg, aes(x = -hu_subadult_median, xmin = -hdi_95[,1], xmax = -hdi_95[,2], reorder(get_species_name(taxon_capture), hu_subadult_median))) +
#   geom_vline(xintercept = 0, color = "#555") +
#   geom_rect(aes(xmin = -hdi_50[,1], xmax = -hdi_50[,2]), height = 0.2, color = "#000") +
#   geom_pointrange() +
#   theme_minimal() +
#   labs(title = "Bd Prevalence Contrast (Subadult - Adult)",
#        x = "Contrast of log odds of Bd detection (Subadult - Adult) [-]",
#        y = "Species")

p_cascade_mu_contrast_linpred = ggplot(mu_spp_lss_contrast_agg, aes(x = mu_contrast_median, xmin = mu_contrast_lower_95, xmax = mu_contrast_upper_95, reorder(get_species_name(taxon_capture), -mu_contrast_median))) +
  geom_vline(xintercept = 0, color = "#555") +
  geom_rect(aes(xmin = mu_contrast_lower_50, xmax = mu_contrast_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  theme_minimal() +
  labs(title = "Bd Load Contrast (Subadult - Adult)",
       x = "Contrast of log Bd load (Subadult - Adult) [-]",
       y = "Species")

# p_cascade_mu_contrast = ggplot(b_contrast_agg, aes(x = b_subadult_median, xmin = hdi_95[,1], xmax = hdi_95[,2], reorder(get_species_name(taxon_capture), -b_subadult_median))) +
#   geom_vline(xintercept = 0, color = "#555") +
#   geom_rect(aes(xmin = hdi_50[,1], xmax = hdi_50[,2]), height = 0.2, color = "#000") +
#   geom_pointrange() +
#   theme_minimal() +
#   labs(title = "Bd Load Contrast (Subadult - Adult)",
#        x = "Contrast of log Bd load (Subadult - Adult) [-]",
#        y = "Species")

p_cascade_prevalence_linpred = ggplot(prevalence_spp_agg, aes(x = prevalence_median, xmin = prevalence_lower_95, xmax = prevalence_upper_95, reorder(get_species_name(taxon_capture), -prevalence_median))) +
  geom_rect(aes(xmin = prevalence_lower_50, xmax = prevalence_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  xlim(0, 1) +
  theme_minimal() +
  labs(title = "Bd Prevalence",
       x = "Prevalence draws [-]",
       y = "Species")


# p_cascade_prevalence = ggplot(hu_group_agg, aes(x = 1 - plogis(hu_median), xmin = 1 - plogis(hdi_95[,1]), xmax = 1 - plogis(hdi_95[,2]), reorder(get_species_name(taxon_capture), hu_median))) +
#   geom_rect(aes(xmin = 1 - plogis(hdi_50[,1]), xmax = 1 - plogis(hdi_50[,2])), height = 0.2, color = "#000") +
#   geom_pointrange() +
#   xlim(0, 1) +
#   theme_minimal() +
#   labs(title = "Bd Prevalence",
#        x = "Prevalence draws [-]",
#        y = "Species")

p_cascade_sigma_linpred = ggplot(sigma_spp_agg, aes(x = sigma_median, xmin = sigma_lower_95, xmax = sigma_upper_95, reorder(get_species_name(taxon_capture), -sigma_median))) +
  geom_rect(aes(xmin = sigma_lower_50, xmax = sigma_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  theme_minimal() +
  labs(title = "Bd Load Dispersion",
       x = "Log load standard deviation (sigma) draws [-]",
       y = "Species")

# p_cascade_sigma = ggplot(sigma_group_agg, aes(x = sigma_median, xmin = hdi_95[,1], xmax = hdi_95[,2], reorder(get_species_name(taxon_capture), -sigma_median))) +
#   geom_rect(aes(xmin = hdi_50[,1], xmax = hdi_50[,2]), height = 0.2, color = "#000") +
#   geom_pointrange() +
#   theme_minimal() +
#   labs(title = "Bd Load Dispersion",
#        x = "Log load standard deviation (sigma) draws [-]",
#        y = "Species")

# save plots
ggsave(here("size_bd_unified", "unified_figures", "03_species_cascade_hu_contrast.png"),
       p_cascade_hu_contrast_linpred,
       width = 8, height = 6)

ggsave(here("size_bd_unified", "unified_figures", "03_species_cascade_mu_contrast.png"),
       p_cascade_mu_contrast_linpred,
       width = 8, height = 6)

ggsave(here("size_bd_unified", "unified_figures", "03_species_cascade_prevalence.png"),
       p_cascade_prevalence_linpred,
       width = 8, height = 6)

ggsave(here("size_bd_unified", "unified_figures", "03_species_cascade_sigma.png"),
       p_cascade_sigma_linpred,
       width = 8, height = 6)


p_cascade_hu_contrast_linpred
p_cascade_mu_contrast_linpred
p_cascade_prevalence_linpred
p_cascade_sigma_linpred

```
# r2
```{r}
# get_subset_r2 = function(model_select, taxon) {
#   print(taxon)
#   s_data = model_select$data %>%
#     filter(taxon_capture == taxon)
#   
#   s_r2 = bayes_R2(model_select, new_data = s_data)
#   
#   gc()
#   
#   return(s_r2)
# }

get_subset_diagnostics = function(model_select, taxon) {
  print(taxon)
  
  taxon_idx = which(model_select$data$taxon_capture == taxon)
  observed = model_select$data$bd_load_co_density[taxon_idx]
  
  # Hurdle component - how well does it predict zero vs nonzero?
  fitted_hu = fitted(model_select, dpar = "hu")[taxon_idx, "Estimate"]
  observed_zero = as.integer(observed == 0)
  
  # Simple correlation between predicted P(zero) and actual zero/nonzero
  hu_cor = cor(fitted_hu, observed_zero)
  
  # Lognormal component - nonzero observations only
  nonzero_idx = observed > 0
  obs_log = log(observed[nonzero_idx])
  fitted_mu = fitted(model_select, dpar = "mu")[taxon_idx, "Estimate"]
  # mu is on log scale for lognormal
  
  ss_res = sum((obs_log - fitted_mu[nonzero_idx])^2)
  ss_tot = sum((obs_log - mean(obs_log))^2)
  r2_lognormal = 1 - (ss_res / ss_tot)
  
  gc()
  
  return(list(
    hu_correlation = hu_cor,
    r2_lognormal = r2_lognormal,
    n = length(observed),
    n_nonzero = sum(nonzero_idx),
    pct_zero = mean(observed == 0)
  ))
}

preds_1 = posterior_predict(model_select, newdata = model_select$data %>%
                                                    filter(taxon_capture == taxon_list[1]))

preds_2 = posterior_predict(model_select, newdata = model_select$data %>%
                                                    filter(taxon_capture == taxon_list[2]))

taxon_list = model_select$data %>%
  distinct(taxon_capture) %>%
  pull(taxon_capture) %>%
  sort()

r2_list = purrr::map(taxon_list, ~get_subset_r2(model_select, .x))

r2_diagnostics_list = purrr::map(taxon_list, ~get_subset_diagnostics(model_select, .x))

r2_estimate = map_dbl(r2_diagnostics_list, "r2_lognormal")

r2_df = tibble(taxon_capture = taxon_list,
               hu_correlation = map_dbl(r2_diagnostics_list, "hu_correlation"),
               r2_lognormal = map_dbl(r2_diagnostics_list, "r2_lognormal"))
```

# correlation analysis
```{r}
simple_cor = cor_mat(combined_agg_single %>%
                       select(hu_contrast_median,
                              mu_contrast_median,
                              hu_median,
                              sigma_median), method = "spearman")
simple_cor
cor_get_pval(simple_cor)


```

## Phylogenetic tree

### Jetz & Pyron 2018
https://datadryad.org/dataset/doi:10.5061/dryad.cc3n6j5#usage
```{r}
jp_all = read.nexus(file = here("data", "tree-pruner-a1f34500-26bc-45a2-a777-afe418ddad90", "output.nex"))

jp_sample_100 = jp_all[sample(1000, 100)]

consensus_100 = consensus(jp_sample_100)

jp_raw = consensus.edges(jp_sample_100, consensus.tree = consensus_100)

# plot(jp_raw)

# jp_raw = read.tree(file = here("data", "amph_shl_new_Consensus_7238.tre"))

# preview alignment
jk_species_list = bd_model_env_stats %>%
  mutate(sn = str_to_title(taxon_capture)) %>%
  select(sn) %>%
  mutate(genus = str_split_fixed(sn, "_", 2)[,1],
         species = str_split_fixed(sn, "_", 2)[,2])

jp_species_list = tibble(jp_sn = jp_raw$tip.label) %>%
  mutate(jp_genus = str_split_fixed(jp_sn, "_", 2)[,1],
         species = str_split_fixed(jp_sn, "_", 2)[,2],
         jp = TRUE)
  

peace = jk_species_list %>%
  left_join(jp_species_list, by = "species") %>%
  select(sn,
         jp_sn,
         genus,
         jp_genus,
         everything())

species_list_j = bd_model_env_stats %>%
  select(taxon_capture) %>%
  distinct() %>%
  mutate(sn = str_to_title(taxon_capture),
         jp_sn = recode_values(sn,
                               "Lithobates_warszewitschii" ~ "Rana_warszewitschii",
                               "Lithobates_blairi" ~ "Rana_blairi",
                               "Lithobates_chiricahuensis" ~ "Rana_chiricahuensis",
                               "Lithobates_palustris" ~ "Rana_palustris",
                               "Lithobates_sphenocephalus" ~ "Rana_sphenocephala",
                               default = sn))

# Prune to species list
jp_pruned = keep.tip(jp_raw, species_list_j$jp_sn)

combined_agg_j = combined_agg_single %>%
  left_join(species_list_j %>%
              select(taxon_capture,
                     jp_sn), by = "taxon_capture")

# Check plot
plotTree(jp_pruned, fsize = 0.8)

is.binary(jp_pruned)

table(sapply(1:jp_pruned$Nnode, function(n) {
  length(Children(jp_pruned, n + Ntip(jp_pruned)))
}))

new_tips = tibble(jp_sn = jp_pruned$tip.label) %>%
  left_join(species_list_j, by = "jp_sn") %>%
  pull(taxon_capture)

jp_pruned$tip.label = new_tips

jp_pruned$tip.label

jp_covar = ape::vcv.phylo(jp_pruned, corr = FALSE)
jp_corr = ape::vcv.phylo(jp_pruned, corr = TRUE)
rownames(jp_corr)

jp_corr_df = jp_corr %>%
  as_tibble() %>%
  mutate(mu_spp = rownames(jp_corr)) %>%
  pivot_longer(cols = c(-"mu_spp"),
               names_to = "hu_spp",
               values_to = "corr") %>%
  left_join(combined_agg_single %>%
              select(taxon_capture,
                     hu_contrast_median), by = c("hu_spp" = "taxon_capture")) %>%
  left_join(combined_agg_single %>%
              select(taxon_capture,
                     mu_contrast_median), by = c("mu_spp" = "taxon_capture")) %>%
  mutate(hu_spp = reorder(hu_spp, hu_contrast_median),
         mu_spp = reorder(mu_spp, mu_contrast_median))

jp_corr_hu_mean = jp_corr_df %>%
  group_by(hu_spp) %>%
  summarise(hu_corr_mean = mean(corr),
            .groups = "drop")

jp_corr_mu_mean = jp_corr_df %>%
  group_by(mu_spp) %>%
  summarise(mu_corr_mean = mean(corr),
            .groups = "drop")

p_phylo_heatmap = ggplot(jp_corr_df, aes(mu_spp, hu_spp)) +
  geom_tile(aes(fill = corr)) +
  # geom_text(aes(label = round(corr, 2))) +
  scale_fill_viridis_c() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1)) +
  scale_y_discrete(position = "right")
  
p_phylo_heatmap

# save data
write.tree(jp_pruned, file = here("size_bd_unified", "intermediates", "phylo_tree.nwk"))

```

### Analyze
```{r}
jp_pruned = read.tree(here("size_bd_unified", "intermediates", "phylo_tree.nwk"))

jp_pruned$tip.label

# calculate blomberg K
blomberg_k = function(tree, var_values, species) {
  set_named = setNames(var_values, species)
  return(blomberg_k_sigma_med = phylosig(tree, set_named, method = "K", test = TRUE, nsim = 1000))
}

pagel_lambda = function(tree, var_values, species) {
  set_named = setNames(var_values, species)
  return(blomberg_k_sigma_med = phylosig(tree, set_named, method = "lambda", test = TRUE, nsim = 1000))
}

# blomberg k
hu_contrast_med_k = blomberg_k(jp_pruned, combined_agg$hu_contrast_median, combined_agg$taxon_capture)
mu_contrast_med_k = blomberg_k(jp_pruned, combined_agg$mu_contrast_median, combined_agg$taxon_capture)
prev_med_k = blomberg_k(jp_pruned, combined_agg$prevalence_median, combined_agg$taxon_capture)
sigma_med_k = blomberg_k(jp_pruned, combined_agg$sigma_median, combined_agg$taxon_capture)

hu_contrast_med_k
mu_contrast_med_k
prev_med_k
sigma_med_k

# pagel lambda
hu_contrast_med_lambda = pagel_lambda(jp_pruned, combined_agg$hu_contrast_median, combined_agg$taxon_capture)
mu_contrast_med_lambda = pagel_lambda(jp_pruned, combined_agg$mu_contrast_median, combined_agg$taxon_capture)
prev_med_lambda = pagel_lambda(jp_pruned, combined_agg$prevalence_median, combined_agg$taxon_capture)
sigma_med_lambda = pagel_lambda(jp_pruned, combined_agg$sigma_median, combined_agg$taxon_capture)

hu_contrast_med_lambda
mu_contrast_med_lambda
prev_med_lambda
sigma_med_lambda

# define color maps
sym_div_cont_map = function(tree, var_values, species, center, p_title, style = "div") {
  set_named = setNames(var_values, species)
  
  sym_lim = max(abs(center - min(var_values)), abs(center - max(var_values)))
  
  map_1 = contMap(tree, set_named,
                   lims = c(center - sym_lim, center + sym_lim),
                   title = p_title, fsize = 0.4)
  
  if (style == "div") {
    colors = colorRampPalette(c("#762a83", "white", "#1b7837"))(101)
  } else if (style == "lin") {
    colors = colorRampPalette(c("white", "#1b7837"))(101)
  }
  map_2 = setMap(map_1, colors = colors)
  
  return(map_2)
}

p_hu_contrast_med_map = sym_div_cont_map(jp_pruned,
                                  combined_agg$hu_contrast_median,
                                  combined_agg$taxon_capture,
                                  0,
                                  "Bd prevalence (hu): median subadult effect")

p_mu_contrast_med_map = sym_div_cont_map(jp_pruned,
                                 combined_agg$mu_contrast_median,
                                 combined_agg$taxon_capture,
                                 0,
                                 "Bd log load (mu): median subadult effect")

p_prev_med_map = sym_div_cont_map(jp_pruned,
                                combined_agg$prevalence_median,
                                combined_agg$taxon_capture,
                                0.5,
                                "Median Bd prevalence")

p_sigma_med_map = sym_div_cont_map(jp_pruned,
                                 combined_agg$sigma_median,
                                 combined_agg$taxon_capture,
                                 0.8822,
                                 "Median Bd load dispersion (sigma)")

png(file = here("size_bd_unified", "unified_figures", "03_tree_contmap_hu_contrast.png"),
    width = 6, height = 8, units = "in", res = 300)
plot(p_hu_contrast_med_map)
dev.off()

png(file = here("size_bd_unified", "unified_figures", "03_tree_contmap_mu_contrast.png"),
    width = 6, height = 8, units = "in", res = 300)
plot(p_mu_contrast_med_map)
dev.off()


png(file = here("size_bd_unified", "unified_figures", "03_tree_contmap_prevalence.png"),
    width = 6, height = 8, units = "in", res = 300)
plot(p_prev_med_map)
dev.off()


png(file = here("size_bd_unified", "unified_figures", "03_tree_contmap_sigma.png"),
    width = 6, height = 8, units = "in", res = 300)
plot(p_sigma_med_map)
dev.off()


plot(p_hu_contrast_med_map)
plot(p_mu_contrast_med_map)
plot(p_prev_med_map)
plot(p_sigma_med_map)

```

## Collective results
```{r}
#| eval: true
#| echo: true
#| output: false
#| code-fold: true
#| code-summary: Show code

# m5g_zln = readRDS(file = "unified_models/unified_model_results_zln_life_tax_poptls2env2_env2.rds")


model_data = brms_data

draws_all = collect_draws(model_select) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

density_mat_all = density_mat(log(0.5), 20, 100, draws_all) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

sim_load_all = simulate_bd_load(draws_all, 10, 1000) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

obs_stats_all = sample_stats(model_data) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

```


### Modeled Bd load PDF
```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 6
# PDF -- facet by species
ggplot(density_mat_all, aes(x = density,
                            y = pdf_median,
                            color = life_stage_simple, 
                            fill = life_stage_simple)) +
  geom_ribbon(aes(ymin = pdf_lower,
                  ymax = pdf_upper), 
              alpha = 0.15,
              color = NA) +
  geom_line() +
  theme_minimal() +
  ggtitle("Modeled PDF of Bd-positive load") +
  labs(color = "Life stage",
       fill = "Life stage") +
  xlab("log(Bd load) [-]") +
  ylab("Median probability density [-]\n(with 25th to 75th percentiles)") +
  scale_x_continuous(limits = c(log(0.5), 20), expand = c(0, 0)) +
  facet_grid(. ~ gsub(" ", "\n", species))

# # PDF -- facet by life_stage & season
# ggplot(density_mat_all, aes(x = density,
#                             y = pdf_median,
#                             color = taxon_capture, 
#                             fill = taxon_capture)) +
#   geom_ribbon(aes(ymin = pdf_lower,
#                   ymax = pdf_upper), 
#               alpha = 0.15,
#               color = NA) +
#   geom_line() +
#   theme_minimal() +
#   ggtitle("Modeled PDF of Bd-positive load") +
#   labs(color = "Species",
#        fill = "Species") +
#   xlab("log(Bd load) [-]") +
#   ylab("Median probability density [-]\n(with 25th to 75th percentiles)") +
#   scale_x_continuous(limits = c(log(0.3), 20), expand = c(0, 0)) +
#   facet_grid(life_stage_simple ~ season)
```

### Moded Bd load CCDF
```{r}
#| echo: false
#| warning: false
#| fig-height: 8
#| fig-width: 8
# CCDF -- facet by species
ggplot(density_mat_all, aes(x = density,
                            y = ccdf_median,
                            color = life_stage_simple, 
                            fill = life_stage_simple)) +
  geom_ribbon(aes(ymin = ccdf_lower,
                  ymax = ccdf_upper), 
              alpha = 0.15,
              color = NA) +
  geom_line() +
  theme_minimal() +
  ggtitle("Modeled CCDF of Bd-positive load") +
  labs(color = "Life stage",
       fill = "Life stage") +
  xlab("log(Bd load) [-]") +
  ylab("Median complementary cumulative density (1 - CDF) [-]\n(with 25th to 75th percentiles)") +
  scale_x_continuous(limits = c(log(0.5), 12), expand = c(0, 0)) +
  ylim(0, 0.75) +
  # facet_grid(. ~ gsub(" ", "\n", species))
  facet_wrap("species")

# CCDF -- facet by life_stage & season
# ggplot(density_mat_all, aes(x = density,
#                             y = ccdf_median,
#                             color = taxon_capture,
#                             fill = taxon_capture)) +
#   geom_ribbon(aes(ymin = ccdf_lower,
#                   ymax = ccdf_upper),
#               alpha = 0.15,
#               color = NA) +
#   geom_line() +
#   theme_minimal() +
#   ggtitle("Modeled CCDF of Bd-positive load") +
#   labs(color = "Species",
#        fill = "Species") +
#   xlab("log(Bd load) [-]") +
#   ylab("Median complementary cumulative density (1 - CDF) [-]\n(with 25th to 75th percentiles)") +
#   scale_x_continuous(limits = c(log(0.3), 20), expand = c(0, 0)) +
#   ylim(0, 1) +
#   facet_grid(. ~ life_stage_simple)
```

# Environmental conditions

## Correlations

```{r}
simple_cor = cor_mat(brms_data %>%
                       select(precip_mm_10_day_sum,
                              precip_mm_30_day_sum,
                              temp_c_10_day_mean,
                              temp_c_30_day_mean), method = "pearson")
peace = simple_cor
cor_get_pval(simple_cor)


```

## Distributions
```{r}
#| eval: true
#| warning: false
#| code-fold: true
#| code-summary: "Environmental conditions distributions code"
#| fig-height: 6
#| fig-width: 8

f_draw_agg <- function(draws, group_vars, agg_var) {
  
  draws |>
    group_by(across(all_of(group_vars))) |>
    reframe(
      "{agg_var}_lower_50" := tidybayes::median_qi(.data[[agg_var]], .width = 0.50)[, 2],
      "{agg_var}_upper_50" := tidybayes::median_qi(.data[[agg_var]], .width = 0.50)[, 3],
      "{agg_var}_lower_95" := tidybayes::median_qi(.data[[agg_var]], .width = 0.95)[, 2],
      "{agg_var}_upper_95" := tidybayes::median_qi(.data[[agg_var]], .width = 0.95)[, 3],
      "{agg_var}_median"   := median(.data[[agg_var]]),
      "{agg_var}_net_effect" := (sum(.data[[agg_var]] > 0) / n()) * 2 - 1
    )
}

precip_spp_sample_summary = f_draw_agg(brms_data, c("taxon_capture", "life_stage_simple"), "precip_mm_10_day_sum") %>%
  mutate(offset = recode_values(life_stage_simple,
                                "adult" ~ 1,
                                "subadult" ~ 0))

temp_spp_sample_summary = f_draw_agg(brms_data, c("taxon_capture", "life_stage_simple"), "temp_c_30_day_mean") %>%
  mutate(offset = recode_values(life_stage_simple,
                                "adult" ~ 1,
                                "subadult" ~ 0))

# 10 day precip, 30-day temp by species
ggplot(brms_data, aes(x = temp_c_30_day_mean, y = precip_mm_10_day_sum, color = life_stage_simple)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  facet_wrap(. ~ taxon_capture) +
  xlim(-5, 40) +
  ylim(0, 200) +
  labs(x = "10-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()

# 30 day precip, 30-day temp by species
ggplot(brms_data, aes(x = temp_c_30_day_mean, y = precip_mm_30_day_sum, color = life_stage_simple)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  facet_wrap(. ~ taxon_capture) +
  xlim(-5, 40) +
  ylim(0, 400) +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()

# 30 day collective
ggplot(brms_data, aes(x = temp_c_30_day_mean, y = precip_mm_30_day_sum, color = taxon_capture)) +
  geom_point(alpha = 0.1) +
  stat_ellipse() +
  labs(x = "30-day temperature mean (degC)", 
       y = "30-day precipitation total (mm)",
       title = "Temperature v. precipitation") +
  theme_minimal()

```

## Climate

### Precip
```{r}

apply_scale <- function(x_new, sc_obj) {
  (x_new - attr(sc_obj, "scaled:center")) / attr(sc_obj, "scaled:scale")
}

remove_scale = function(x_sc, sc_obj) {
  x_sc * attr(sc_obj, "scaled:scale") + attr(sc_obj, "scaled:center")
}

precip_10_sc = scale(brms_data$precip_mm_10_day_sum)
precip_30_sc = scale(brms_data$precip_mm_30_day_sum)
temp_30_sc = scale(brms_data$temp_c_30_day_mean)

# define sequences over real-world ranges
precip_10_seq <- seq(min(brms_data$precip_mm_10_day_sum),
                  max(brms_data$precip_mm_10_day_sum),
                  length.out = 100)

precip_30_seq <- seq(min(brms_data$precip_mm_30_day_sum),
                  max(brms_data$precip_mm_30_day_sum),
                  length.out = 100)



# scale them using the original parameters
nd_precip <- model_select$data |>
  group_by(taxon_capture, life_stage_simple) %>%
  summarize(past_year_weekly_tempurature_max_c_sd = median(past_year_weekly_tempurature_max_c_sd),
            .groups = "drop") |>
  cross_join(tibble(precip_mm_10_day_sum = precip_10_seq)) |>
  mutate(
    temp_c_30_day_mean   = remove_scale(0, temp_30_sc),
    precip_mm_10_z  = apply_scale(precip_mm_10_day_sum, precip_10_sc),
    temp_c_30_z     = apply_scale(temp_c_30_day_mean, temp_30_sc)
  )

# hu
hu_draws_precip <- model_select |>
  linpred_draws(
    newdata    = nd_precip,
    dpar       = "hu",
    re_formula = ~ (1 | taxon_capture:life_stage_simple)
  )

hu_agg_precip = hu_draws_precip |>
  mutate(prevalence = 1 - plogis(hu)) |>
  group_by(taxon_capture, precip_mm_10_day_sum) |>
  median_qi(prevalence) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

# mu
mu_draws_precip <- model_select |>
  linpred_draws(
    newdata    = nd_precip,
    dpar       = "mu",
    re_formula = ~ (1 | taxon_capture:life_stage_simple)
  )

mu_agg_precip = mu_draws_precip |>
  group_by(taxon_capture, precip_mm_10_day_sum) |>
  median_qi(mu) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

agg_precip = hu_agg_precip %>%
  dplyr::rename(prevalence_lower = .lower,
                prevalence_upper = .upper,
                prevalence_width = .width) %>%
  select(taxon_capture,
         precip_mm_10_day_sum,
         prevalence,
         prevalence_lower,
         prevalence_upper,
         prevalence_width) %>%
  left_join(mu_agg_precip%>%
              dplyr::rename(mu_lower = .lower,
                            mu_upper = .upper,
                            mu_width = .width) %>%
              select(taxon_capture,
                     precip_mm_10_day_sum,
                     mu,
                     mu_lower,
                     mu_upper,
                     mu_width), by = c("taxon_capture", "precip_mm_10_day_sum"))

# visualize
p_precip_prev = ggplot(agg_precip, aes(x = precip_mm_10_day_sum, y = prevalence, ymin = prevalence_lower, ymax = prevalence_upper)) +
  geom_line() +
  geom_ribbon(alpha = 0.3) +
  facet_wrap("taxon_capture") +
  theme_minimal() +
  labs(title = "Bd Prevalence vs. Precipitation",
       x = "10-day lag precipitation [mm]",
       y = "Bd prevalence [-]")

# p_precip_prev_samp = ggplot(agg_precip) +
#   geom_line(aes(x = precip_mm_10_day_sum, y = prevalence)) +
#   geom_ribbon(aes(x = precip_mm_10_day_sum, ymin = prevalence_lower, ymax = prevalence_upper), alpha = 0.3) +
#   geom_rect(data = precip_spp_sample_summary, aes(xmin = precip_mm_10_day_sum_lower_50, xmax = precip_mm_10_day_sum_upper_50, y = -0.1 - 0.2 * offset, height = 0.05, color = life_stage_simple, )) +
#   geom_point(data = precip_spp_sample_summary, aes(x = precip_mm_10_day_sum_median, y = -0.1 - 0.2 * offset, color = life_stage_simple)) +
#   facet_wrap("taxon_capture") +
#   theme_minimal() +
#   labs(title = "Bd Prevalence vs. Precipitation",
#        x = "10-day lag precipitation [mm]",
#        y = "Bd prevalence [-]")

p_precip_mu = ggplot(agg_precip, aes(x = precip_mm_10_day_sum, y = mu, ymin = mu_lower, ymax = mu_upper)) +
  geom_line() +
  geom_ribbon(alpha = 0.3) +
  facet_wrap("taxon_capture") +
  theme_minimal() +
  labs(title = "Bd Load vs. Precipitation",
       x = "10-day lag precipitation [mm]",
       y = "log(Bd load) [-]")

ggsave(here("size_bd_unified", "unified_figures", "03_precip_prevalence.png"),
       p_precip_prev,
       width = 8, height = 8)

ggsave(here("size_bd_unified", "unified_figures", "03_precip_mu.png"),
       p_precip_mu,
       width = 8, height = 8)

p_precip_prev
p_precip_mu

```

### Temp
```{r}
temp_seq   <- seq(min(brms_data$temp_c_30_day_mean),
                  max(brms_data$temp_c_30_day_mean),
                  length.out = 100)

nd_temp <- model_select$data |>
  group_by(taxon_capture, life_stage_simple) %>%
  summarize(past_year_weekly_tempurature_max_c_sd = median(past_year_weekly_tempurature_max_c_sd),
            .groups = "drop") |>
  cross_join(tibble(temp_c_30_day_mean = temp_seq)) |>
  mutate(
    precip_mm_10_day_sum   = remove_scale(0, precip_10_sc),
    precip_mm_10_z  = apply_scale(precip_mm_10_day_sum, precip_10_sc),
    temp_c_30_z     = apply_scale(temp_c_30_day_mean, temp_30_sc),
  )


# hu
hu_draws_temp <- model_select |>
  linpred_draws(
    newdata    = nd_temp,
    dpar       = "hu",
    re_formula = ~ (1 | taxon_capture:life_stage_simple)
  )

hu_agg_temp = hu_draws_temp |>
  mutate(prevalence = 1 - plogis(hu)) |>
  group_by(taxon_capture, temp_c_30_day_mean) |>
  median_qi(prevalence) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

# mu
mu_draws_temp <- model_select |>
  linpred_draws(
    newdata    = nd_temp,
    dpar       = "mu",
    re_formula = ~ (1 | taxon_capture:life_stage_simple)
  )

mu_agg_temp = mu_draws_temp |>
  group_by(taxon_capture, temp_c_30_day_mean) |>
  median_qi(mu) %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

agg_temp = hu_agg_temp %>%
  dplyr::rename(prevalence_lower = .lower,
                prevalence_upper = .upper,
                prevalence_width = .width) %>%
  select(taxon_capture,
         temp_c_30_day_mean,
         prevalence,
         prevalence_lower,
         prevalence_upper,
         prevalence_width) %>%
  left_join(mu_agg_temp %>%
              dplyr::rename(mu_lower = .lower,
                            mu_upper = .upper,
                            mu_width = .width) %>%
              select(taxon_capture,
                     temp_c_30_day_mean,
                     mu,
                     mu_lower,
                     mu_upper,
                     mu_width), by = c("taxon_capture", "temp_c_30_day_mean"))

# visualize
p_temp_prev = ggplot(agg_temp, aes(x = temp_c_30_day_mean, y = prevalence, ymin = prevalence_lower, ymax = prevalence_upper)) +
  geom_line() +
  geom_ribbon(alpha = 0.3) +
  facet_wrap("taxon_capture") +
  theme_minimal() +
  ylim(0, 1) +
  labs(title = "Bd Prevalence vs. Temperature",
       x = "30-day lag max daily temperature [C]",
       y = "Bd prevalence [-]")

p_temp_mu = ggplot(agg_temp, aes(x = temp_c_30_day_mean, y = mu, ymin = mu_lower, ymax = mu_upper)) +
  geom_line() +
  geom_ribbon(alpha = 0.3) +
  facet_wrap("taxon_capture") +
  theme_minimal() +
  labs(title = "Bd Load vs. Temperature",
       x = "30-day lag max daily temperature [C]",
       y = "log(Bd load) [-]")

ggsave(here("size_bd_unified", "unified_figures", "03_temp_prevalence.png"),
       p_temp_prev,
       width = 8, height = 8)

ggsave(here("size_bd_unified", "unified_figures", "03_temp_mu.png"),
       p_temp_mu,
       width = 8, height = 8)

p_temp_prev
p_temp_mu

```

### KG zones
```{r}
nd_kg_env = model_select$data |>
  group_by(taxon_capture, life_stage_simple) |>
  summarise(precip_mm_10_z = median(precip_mm_10_z),
            temp_c_30_z = median(temp_c_30_z),
            .groups = "drop") |>
  filter(life_stage_simple == "subadult") %>%
  select(-life_stage_simple)

nd_kg = model_select$data |>
  distinct(taxon_capture, life_stage_simple, kg_code_3) |>
  left_join(nd_kg_env, by = "taxon_capture")

hu_draws_kg_spp_lss = model_select |>
  linpred_draws(
    newdata    = nd_kg,
    dpar       = "hu",
    re_formula = ~ (1 | kg_code_3)
  )

hu_agg_kg = hu_draws_kg_spp_lss |>
  mutate(prevalence = 1 - plogis(hu)) |>
  group_by(kg_code_3) |>
  median_qi(prevalence)


# mu
mu_draws_kg_spp_lss = model_select |>
  linpred_draws(
    newdata    = nd_kg,
    dpar       = "mu",
    re_formula = ~ (1 | kg_code_3)
  )

mu_agg_kg = mu_draws_kg_spp_lss |>
  group_by(kg_code_3) |>
  median_qi(mu)

agg_kg = hu_agg_kg %>%
  dplyr::rename(prevalence_lower = .lower,
                prevalence_upper = .upper,
                prevalence_width = .width) %>%
  select(kg_code_3,
         prevalence,
         prevalence_lower,
         prevalence_upper,
         prevalence_width) %>%
  left_join(mu_agg_kg %>%
              dplyr::rename(mu_lower = .lower,
                            mu_upper = .upper,
                            mu_width = .width) %>%
              select(kg_code_3,
                     mu,
                     mu_lower,
                     mu_upper,
                     mu_width), by = "kg_code_3") %>%
  mutate(kg_code_2 = substr(kg_code_3, start = 1, stop = 2),
         kg_code_1 = substr(kg_code_3, start = 1, stop = 1))

# visualize
ggplot(agg_kg, aes(x = kg_code_3, fill = kg_code_1)) +
  geom_boxplot(stat = "identity", aes(middle = prevalence, lower = prevalence_lower, upper = prevalence_upper, ymin = prevalence_lower, ymax = prevalence_upper))

ggplot(agg_kg, aes(x = kg_code_3, fill = kg_code_1)) +
  geom_boxplot(stat = "identity", aes(middle = mu, lower = mu_lower, upper = mu_upper, ymin = mu_lower, ymax = mu_upper))

p_y_temp_prev = ggplot(agg_y_temp, aes(x = past_year_weekly_tempurature_max_c_sd, y = prevalence, ymin = prevalence_lower, ymax = prevalence_upper)) +
  geom_line() +
  geom_ribbon(alpha = 0.3) +
  theme_minimal() +
  ylim(0, 1) +
  labs(title = "Bd Prevalence vs. Past year weekly temperature max SD",
       x = "Past year weekly tempurature max SD [C]",
       y = "Bd prevalence [-]")

p_y_temp_mu = ggplot(agg_y_temp, aes(x = past_year_weekly_tempurature_max_c_sd, y = mu, ymin = mu_lower, ymax = mu_upper)) +
  geom_line() +
  geom_ribbon(alpha = 0.3) +
  theme_minimal() +
  labs(title = "Bd Load vs. Past year weekly temperature max SD",
       x = "Past year weekly tempurature max SD [C]",
       y = "log(Bd load) [-]")

ggsave(here("size_bd_unified", "unified_figures", "03_temp_prevalence.png"),
       p_temp_prev,
       width = 8, height = 8)

ggsave(here("size_bd_unified", "unified_figures", "03_temp_mu.png"),
       p_temp_mu,
       width = 8, height = 8)

p_temp_prev
p_temp_mu
```

# Populations
```{r}
#| fig-width: 10
#| fig-height: 10

f_draw_agg <- function(draws, group_vars, agg_var) {
  
  draws |>
    group_by(across(all_of(group_vars))) |>
    reframe(
      "{agg_var}_lower_50" := tidybayes::median_qi(.data[[agg_var]], .width = 0.50)[, 2],
      "{agg_var}_upper_50" := tidybayes::median_qi(.data[[agg_var]], .width = 0.50)[, 3],
      "{agg_var}_lower_95" := tidybayes::median_qi(.data[[agg_var]], .width = 0.95)[, 2],
      "{agg_var}_upper_95" := tidybayes::median_qi(.data[[agg_var]], .width = 0.95)[, 3],
      "{agg_var}_median"   := median(.data[[agg_var]]),
      "{agg_var}_net_effect" := (sum(.data[[agg_var]] > 0) / n()) * 2 - 1
    )
}

# pop coords
pop_coords = brms_data %>%
  distinct(site, latitude, longitude, population) %>%
  group_by(population) %>%
  summarize(latitude = mean(latitude),
            longitude = mean(longitude),
            .groups = "drop")

# peace = as_tibble(get_variables(model_select))
# 
# "r_population:taxon_capture:life_stage_simple[10114_pop_anaxyrus_canorus_subadult,Intercept]"
# 
# "r_population:taxon_capture:life_stage_simple__hu[10477_pop_pseudacris_regilla_adult,Intercept]"

nd_pop_env = model_select$data |>
  group_by(taxon_capture, life_stage_simple) |>
  summarise(precip_mm_10_z = median(precip_mm_10_z),
            temp_c_30_z = median(temp_c_30_z),
            past_year_weekly_tempurature_max_c_sd = median(past_year_weekly_tempurature_max_c_sd),
            .groups = "drop") |>
  filter(life_stage_simple == "subadult") %>%
  select(-life_stage_simple)

nd_pop = model_select$data |>
  distinct(taxon_capture, life_stage_simple, population) |>
  left_join(nd_pop_env, by = "taxon_capture")

hu_draws_pop_spp_lss = model_select |>
  linpred_draws(
    newdata    = nd_pop,
    dpar       = "hu",
    re_formula = ~ (1 | population:taxon_capture:life_stage_simple)
  )

hu_pop_contrast_draws = hu_draws_pop_spp_lss |>
  pivot_wider(
    id_cols     = c(population, taxon_capture, .draw),
    names_from  = life_stage_simple,
    values_from = hu
  ) |>
  mutate(hu_contrast = subadult - adult,
         contrast = "subadult - adult")

hu_pop_contrast_agg = f_draw_agg(hu_pop_contrast_draws %>%
                                   filter(!is.na(hu_contrast)), c("taxon_capture", "population", "contrast"), "hu_contrast")
hu_pop_adult_agg = f_draw_agg(hu_pop_contrast_draws %>%
                                filter(!is.na(adult)) %>%
                                dplyr::rename(hu_adult = adult), c("taxon_capture", "population", "contrast"), "hu_adult")
hu_pop_subadult_agg = f_draw_agg(hu_pop_contrast_draws %>%
                                   filter(!is.na(subadult)) %>%
                                dplyr::rename(hu_subadult = subadult), c("taxon_capture", "population", "contrast"), "hu_subadult")

# mu
mu_draws_pop_spp_lss = model_select |>
  linpred_draws(
    newdata    = nd_pop,
    dpar       = "mu",
    re_formula = ~ (1 | population:taxon_capture:life_stage_simple)
  )

mu_pop_contrast_draws = mu_draws_pop_spp_lss |>
  pivot_wider(
    id_cols     = c(population, taxon_capture, .draw),
    names_from  = life_stage_simple,
    values_from = mu
  ) |>
  mutate(mu_contrast = subadult - adult,
         contrast = "subadult - adult")

mu_pop_contrast_agg = f_draw_agg(mu_pop_contrast_draws %>%
                                   filter(!is.na(mu_contrast)), c("taxon_capture", "population", "contrast"), "mu_contrast")
mu_pop_adult_agg = f_draw_agg(mu_pop_contrast_draws %>%
                                filter(!is.na(adult)) %>%
                                dplyr::rename(mu_adult = adult), c("taxon_capture", "population", "contrast"), "mu_adult")
mu_pop_subadult_agg = f_draw_agg(mu_pop_contrast_draws %>%
                                   filter(!is.na(subadult)) %>%
                                dplyr::rename(mu_subadult = subadult), c("taxon_capture", "population", "contrast"), "mu_subadult")

pop_agg_combined = hu_pop_contrast_agg %>%
  full_join(hu_pop_adult_agg, by = c("taxon_capture", "population", "contrast")) %>%
  full_join(hu_pop_subadult_agg, by = c("taxon_capture", "population", "contrast")) %>%
  left_join(mu_pop_contrast_agg, by = c("taxon_capture", "population", "contrast")) %>%
  left_join(mu_pop_adult_agg, by = c("taxon_capture", "population", "contrast")) %>%
  left_join(mu_pop_subadult_agg, by = c("taxon_capture", "population", "contrast")) %>%
  left_join(pop_coords, by = "population") %>%
  mutate(species = get_species_name(taxon_capture, newline = TRUE))

pop_agg_stats = pop_agg_combined %>%
  group_by(taxon_capture) %>%
  summarize(adult_n = sum(!is.na(hu_adult_median)),
            subadult_n = sum(!is.na(hu_subadult_median)),
            contrast_n = sum(!is.na(hu_contrast_median)))

pop_agg_stats_select = pop_agg_stats %>%
  filter(contrast_n >= 0)

ggplot(pop_agg_combined %>%
         filter(taxon_capture %in% pop_agg_stats_select$taxon_capture), aes(x = -hu_contrast_median, y = mu_contrast_median, color = taxon_capture)) +
  geom_point() +
  geom_text_repel(aes(label = population),
                  color = "#333",
                  size = 3)

ggplot(pop_agg_combined %>%
         filter(taxon_capture %in% pop_agg_stats_select$taxon_capture), aes(x = -hu_contrast_median, y = mu_contrast_median, color = taxon_capture)) +
  geom_point() +
  geom_text_repel(aes(label = population),
                  color = "#333",
                  size = 3) +
  facet_wrap("taxon_capture")

p_scatter_hu_life_stage_linpred_pop = ggplot(pop_agg_combined %>%
         filter(taxon_capture %in% pop_agg_stats_select$taxon_capture), aes(x = -hu_adult_median, y = -hu_subadult_median)) +
  geom_abline(intercept = 0,
              slope = 1,
              alpha = 0.5) + 
  geom_point() +
  geom_text_repel(aes(label = population),
                  color = "#333",
                  size = 3) +
  facet_wrap("species") +
  labs(title = "Adult v. Subadult Prevalence by Population",
       x = "Adult median log(Bd detection odds)",
       y = "Subadult median log(Bd detection odds)") +
  theme_minimal()

p_scatter_hu_life_stage_linpred_pop_sn = ggplot(pop_agg_combined %>%
         filter(taxon_capture %in% c("rana_muscosa", "rana_sierrae")) %>%
         arrange(desc(taxon_capture)), aes(x = -hu_adult_median, y = -hu_subadult_median, color = species)) +
  geom_abline(intercept = 0,
              slope = 1,
              alpha = 0.5) + 
  geom_point() +
  geom_text_repel(aes(label = population),
                  color = "#333",
                  size = 3) +
  labs(title = "Adult v. Subadult Prevalence by Population",
       x = "Adult median log(Bd detection odds)",
       y = "Subadult median log(Bd detection odds)") +
  theme_minimal()

p_scatter_mu_life_stage_linpred_pop = ggplot(pop_agg_combined %>%
         filter(taxon_capture %in% pop_agg_stats_select$taxon_capture), aes(x = mu_adult_median, y = mu_subadult_median)) +
  geom_abline(intercept = 0,
              slope = 1,
              alpha = 0.5) + 
  geom_point() +
  geom_text_repel(aes(label = population),
                  color = "#333",
                  size = 3) +
  facet_wrap("species") +
  labs(title = "Adult v. Subadult Load by Population",
       x = "Adult median log(Bd load)",
       y = "Subadult median log(Bd load)") +
  theme_minimal()

p_scatter_mu_life_stage_linpred_pop_sn = ggplot(pop_agg_combined %>%
         filter(taxon_capture %in% c("rana_muscosa", "rana_sierrae")) %>%
         arrange(desc(taxon_capture)), aes(x = mu_adult_median, y = mu_subadult_median, color = species)) +
  geom_abline(intercept = 0,
              slope = 1,
              alpha = 0.5) + 
  geom_point() +
  geom_text_repel(aes(label = population),
                  color = "#333",
                  size = 3) +
  labs(title = "Adult v. Subadult Load by Population",
       x = "Adult median log(Bd load)",
       y = "Subadult median log(Bd load)") +
  theme_minimal()


p_cascade_hu_contrast_linpred_pop_sn = ggplot(pop_agg_combined %>%
                                                filter(taxon_capture %in% c("rana_muscosa",
                                                                            "rana_sierrae"),
                                                       !is.na(hu_contrast_median)), aes(x = -hu_contrast_median, xmin = -hu_contrast_lower_95, xmax = -hu_contrast_upper_95, reorder(population, hu_contrast_median), color = taxon_capture)) +
  geom_vline(xintercept = 0, color = "#555") +
  geom_rect(aes(xmin = -hu_contrast_lower_50, xmax = -hu_contrast_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  theme_minimal() +
  labs(title = "Bd Prevalence Contrast (Subadult - Adult)",
       x = "Contrast of log odds of Bd detection (Subadult - Adult) [-]",
       y = "Species") +
  scale_color_brewer(palette = "Set2")

p_cascade_mu_contrast_linpred_pop_sn = ggplot(pop_agg_combined %>%
                                                filter(taxon_capture %in% c("rana_muscosa",
                                                                            "rana_sierrae"),
                                                       !is.na(mu_contrast_median)), aes(x = mu_contrast_median, xmin = mu_contrast_lower_95, xmax = mu_contrast_upper_95, reorder(population, -mu_contrast_median), color = taxon_capture)) +
  geom_vline(xintercept = 0, color = "#555") +
  geom_rect(aes(xmin = mu_contrast_lower_50, xmax = mu_contrast_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  theme_minimal() +
  labs(title = "Bd Load Contrast (Subadult - Adult)",
       x = "Contrast of log Bd load density (Subadult - Adult) [-]",
       y = "Species") +
  scale_color_brewer(palette = "Set2")

p_cascade_hu_contrast_linpred_pop_sn_lat = ggplot(pop_agg_combined %>%
                                                filter(taxon_capture %in% c("rana_muscosa",
                                                                            "rana_sierrae"),
                                                       !is.na(hu_contrast_median)), aes(x = -hu_contrast_median, xmin = -hu_contrast_lower_95, xmax = -hu_contrast_upper_95, reorder(population, latitude), color = taxon_capture)) +
  geom_vline(xintercept = 0, color = "#555") +
  geom_rect(aes(xmin = -hu_contrast_lower_50, xmax = -hu_contrast_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  theme_minimal() +
  labs(title = "Bd Prevalence Contrast (Subadult - Adult)",
       x = "Contrast of log odds of Bd detection (Subadult - Adult) [-]",
       y = "Species") +
  scale_color_brewer(palette = "Set2")

p_cascade_mu_contrast_linpred_pop_sn_lat = ggplot(pop_agg_combined %>%
                                                filter(taxon_capture %in% c("rana_muscosa",
                                                                            "rana_sierrae"),
                                                       !is.na(mu_contrast_median)), aes(x = mu_contrast_median, xmin = mu_contrast_lower_95, xmax = mu_contrast_upper_95, reorder(population, latitude), color = taxon_capture)) +
  geom_vline(xintercept = 0, color = "#555") +
  geom_rect(aes(xmin = mu_contrast_lower_50, xmax = mu_contrast_upper_50), height = 0.2, color = "#000") +
  geom_pointrange() +
  theme_minimal() +
  labs(title = "Bd Load Contrast (Subadult - Adult)",
       x = "Contrast of log Bd load density (Subadult - Adult) [-]",
       y = "Species") +
  scale_color_brewer(palette = "Set2")

ggsave(here("size_bd_unified", "unified_figures", "03_population_cascade_sn_prev_contrast.png"),
       p_cascade_hu_contrast_linpred_pop_sn,
       width = 8, height = 18)

ggsave(here("size_bd_unified", "unified_figures", "03_population_cascade_sn_load_contrast.png"),
       p_cascade_mu_contrast_linpred_pop_sn,
       width = 8, height = 18)

ggsave(here("size_bd_unified", "unified_figures", "03_population_cascade_sn_prev_contrast_lat.png"),
       p_cascade_hu_contrast_linpred_pop_sn_lat,
       width = 8, height = 18)

ggsave(here("size_bd_unified", "unified_figures", "03_population_cascade_sn_load_contrast_lat.png"),
       p_cascade_mu_contrast_linpred_pop_sn_lat,
       width = 8, height = 18)


pop_cor = cor_mat(pop_agg_combined %>%
                    filter(taxon_capture %in% c("rana_muscosa",
                                                "rana_sierrae")) %>%
                       select(hu_contrast_median,
                              mu_contrast_median,
                              latitude), method = "spearman")

cor_get_pval(pop_cor)

p_cascade_hu_contrast_linpred_pop_sn
p_cascade_mu_contrast_linpred_pop_sn
p_cascade_hu_contrast_linpred_pop_sn_lat
p_cascade_mu_contrast_linpred_pop_sn_lat

p_scatter_hu_life_stage_linpred_pop
p_scatter_hu_life_stage_linpred_pop_sn
p_scatter_mu_life_stage_linpred_pop
p_scatter_mu_life_stage_linpred_pop_sn

ggsave(here("size_bd_unified", "unified_figures", "03_population_prev_adult_subadult.png"),
       p_scatter_hu_life_stage_linpred_pop,
       width = 10, height = 10)

ggsave(here("size_bd_unified", "unified_figures", "03_population_prev_adult_subadult_sn.png"),
       p_scatter_hu_life_stage_linpred_pop_sn,
       width = 10, height = 10)

ggsave(here("size_bd_unified", "unified_figures", "03_population_load_adult_subadult.png"),
       p_scatter_mu_life_stage_linpred_pop,
       width = 10, height = 10)

ggsave(here("size_bd_unified", "unified_figures", "03_population_load_adult_subadult_sn.png"),
       p_scatter_mu_life_stage_linpred_pop_sn,
       width = 10, height = 10)


```

## population year
```{r}

f_draw_agg <- function(draws, group_vars, agg_var) {
  
  draws |>
    group_by(across(all_of(group_vars))) |>
    reframe(
      "{agg_var}_lower_50" := tidybayes::median_qi(.data[[agg_var]], .width = 0.50)[, 2],
      "{agg_var}_upper_50" := tidybayes::median_qi(.data[[agg_var]], .width = 0.50)[, 3],
      "{agg_var}_lower_95" := tidybayes::median_qi(.data[[agg_var]], .width = 0.95)[, 2],
      "{agg_var}_upper_95" := tidybayes::median_qi(.data[[agg_var]], .width = 0.95)[, 3],
      "{agg_var}_median"   := median(.data[[agg_var]]),
      "{agg_var}_net_effect" := (sum(.data[[agg_var]] > 0) / n()) * 2 - 1
    )
}

popyr_summary = model_select$data %>%
  filter(taxon_capture %in% c("rana_muscosa", "rana_sierrae")) %>%
  group_by(population, year) %>%
  summarize(adult_n = sum(life_stage_simple == "adult"),
            subadult_n = sum(life_stage_simple == "subadult"),
            .groups = "drop") %>%
  filter(adult_n > 0 & subadult_n > 0) %>%
  group_by(population) %>%
  summarize(start_year = min(year),
            end_year = max(year),
            n_year = n())

pop_oi = popyr_summary %>%
  filter(n_year >= 10) %>%
  pull(population)
  

nd_pop_env = model_select$data |>
  group_by(taxon_capture, life_stage_simple) |>
  summarise(precip_mm_10_z = median(precip_mm_10_z),
            temp_c_30_z = median(temp_c_30_z),
            kg_code_2 = mmode(kg_code_2),
            .groups = "drop") |>
  filter(life_stage_simple == "subadult") %>%
  select(-life_stage_simple)

nd_popyr = model_select$data |>
  filter(population %in% pop_oi,
         taxon_capture %in% c("rana_muscosa", "rana_sierrae")) |>
  group_by(taxon_capture, life_stage_simple, population, year, `population:year`, `population:taxon_capture:life_stage_simple`) |>
  summarize(n_adult = sum(life_stage_simple == "adult"),
            n_subadult = sum(life_stage_simple == "subadult"),
            .groups = "drop") |>
  left_join(nd_pop_env, by = "taxon_capture")

hu_draws_popyr_spp_lss = model_select |>
  linpred_draws(
    newdata    = nd_popyr,
    dpar       = "hu",
    re_formula = ~ (1 | population:year) +
                  (1 | population:taxon_capture:life_stage_simple),
    allow_new_levels = FALSE
  )

hu_popyr_contrast_draws = hu_draws_popyr_spp_lss |>
  pivot_wider(
    id_cols     = c(population, year, taxon_capture, .draw),
    names_from  = life_stage_simple,
    values_from = hu
  ) |>
  mutate(hu_contrast = subadult - adult,
         contrast = "subadult - adult")

hu_popyr_contrast_agg = f_draw_agg(hu_popyr_contrast_draws %>%
                                   filter(!is.na(hu_contrast)), c("taxon_capture", "population", "year", "contrast"), "hu_contrast")
hu_popyr_adult_agg = f_draw_agg(hu_popyr_contrast_draws %>%
                                filter(!is.na(adult)) %>%
                                dplyr::rename(hu_adult = adult), c("taxon_capture", "population", "year", "contrast"), "hu_adult")
hu_popyr_subadult_agg = f_draw_agg(hu_popyr_contrast_draws %>%
                                   filter(!is.na(subadult)) %>%
                                dplyr::rename(hu_subadult = subadult), c("taxon_capture", "population", "year", "contrast"), "hu_subadult")

# mu
mu_draws_popyr_spp_lss = model_select |>
  linpred_draws(
    newdata    = nd_popyr,
    dpar       = "mu",
    re_formula = NULL
  )

mu_popyr_contrast_draws = mu_draws_popyr_spp_lss |>
  pivot_wider(
    id_cols     = c(population, year, taxon_capture, .draw),
    names_from  = life_stage_simple,
    values_from = mu
  ) |>
  mutate(mu_contrast = subadult - adult,
         contrast = "subadult - adult")

mu_popyr_contrast_agg = f_draw_agg(mu_popyr_contrast_draws %>%
                                   filter(!is.na(mu_contrast)), c("taxon_capture", "population", "year", "contrast"), "mu_contrast")
mu_popyr_adult_agg = f_draw_agg(mu_popyr_contrast_draws %>%
                                filter(!is.na(adult)) %>%
                                dplyr::rename(mu_adult = adult), c("taxon_capture", "population", "year", "contrast"), "mu_adult")
mu_popyr_subadult_agg = f_draw_agg(mu_popyr_contrast_draws %>%
                                   filter(!is.na(subadult)) %>%
                                dplyr::rename(mu_subadult = subadult), c("taxon_capture", "population", "year", "contrast"), "mu_subadult")

popyr_agg_combined = hu_popyr_contrast_agg %>%
  full_join(hu_popyr_adult_agg, by = c("taxon_capture", "population", "year", "contrast")) %>%
  full_join(hu_popyr_subadult_agg, by = c("taxon_capture", "population", "year", "contrast")) %>%
  left_join(mu_popyr_contrast_agg, by = c("taxon_capture", "population", "year", "contrast")) %>%
  left_join(mu_popyr_adult_agg, by = c("taxon_capture", "population", "year", "contrast")) %>%
  left_join(mu_popyr_subadult_agg, by = c("taxon_capture", "population", "year", "contrast"))


p_scatter_hu_life_stage_popyr_sn = ggplot(popyr_agg_combined, aes(x = -hu_adult_median, y = -hu_subadult_median, color = population)) +
  geom_abline(intercept = 0,
              slope = 1,
              alpha = 0.5) + 
  geom_point() +
  geom_line() +
  geom_text_repel(aes(label = year),
                  color = "#333",
                  size = 3) +
  theme_minimal()


p_scatter_mu_life_stage_linpred_pop_sn = ggplot(pop_agg_combined %>%
         filter(taxon_capture %in% c("rana_muscosa", "rana_sierrae")) %>%
         arrange(desc(taxon_capture)), aes(x = mu_adult_median, y = mu_subadult_median, color = species)) +
  geom_abline(intercept = 0,
              slope = 1,
              alpha = 0.5) + 
  geom_point() +
  geom_text_repel(aes(label = population),
                  color = "#333",
                  size = 3) +
  labs(title = "Adult v. Subadult Load by Population",
       x = "Adult median log(Bd load)",
       y = "Subadult median log(Bd load)") +
  theme_minimal()
```

## map
```{r}
library(leaflet)
library(RColorBrewer)
library(mapview)

pop_agg_contrast = pop_agg_combined %>%
  filter(!is.na(hu_contrast_median),
         taxon_capture %in% c("rana_muscosa", "rana_sierrae")) %>%
  mutate(species = get_species_name(taxon_capture))

hu_abs_max = max(abs(pop_agg_contrast$hu_contrast_median))
mu_abs_max = max(abs(pop_agg_contrast$mu_contrast_median))

pal_hu = colorNumeric("RdYlBu", domain = c(-hu_abs_max, hu_abs_max), reverse = TRUE)
hu_map = leaflet(pop_agg_contrast) %>%
  addTiles() %>%
  addCircleMarkers(lng = ~longitude, lat = ~latitude, color = ~pal_hu(-hu_contrast_median),
                   radius = 5, fillOpacity = 0.9,
                   popup = ~paste("Population:", population,"<br>Species:", species,"<br>log(detection odds) contrast:", -hu_contrast_median)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addLegend(pal = pal_hu, 
            values = ~(-hu_contrast_median),
            title = "Median log(Bd detection odds)<br>Contrast (subadult - adult)",
            position = "bottomright")

pal_mu = colorNumeric("RdYlBu", domain =  c(-mu_abs_max, mu_abs_max), reverse = TRUE)
mu_map = leaflet(pop_agg_contrast) %>%
  addTiles() %>%
  addCircleMarkers(lng = ~longitude, lat = ~latitude, color = ~pal_mu(mu_contrast_median),
                   radius = 5, fillOpacity = 0.9,
                   popup = ~paste("Population:", population,"<br>Species:", species, "<br>log(load) contrast:", mu_contrast_median)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addLegend(pal = pal_mu, 
            values = ~mu_contrast_median,
            title = "Median log(Bd load)<br>Contrast (subadult - adult)",
            position = "bottomright")

saveWidget(hu_map, here("size_bd_unified", "unified_figures", "maps", "map_prev_contrast_sn.html"))
saveWidget(mu_map, here("size_bd_unified", "unified_figures", "maps", "map_load_contrast_sn.html"))

hu_map
mu_map

```

## population density
```{r}

librarian::shelf(RIBBiTR-BII/ribbitrrr, dbplyr, RPostgres, DBI)
dbcon = hopToDB("ribbitr")

db_site = tbl(dbcon, Id("survey_data", "site"))

sn_sites = db_site %>%
  filter(site %in% brms_data$site) %>%
  select(site,
         area_sqr_m) %>%
  collect()

pop_dens = brms_data %>%
  left_join(sn_sites, by = "site") %>%
  group_by(taxon_capture, population) %>%
  mutate(pop_area_sqr_km = sum(area_sqr_m)/100000,
         n_visits = n_distinct(date),
         obs_pop_dens = n() / (n_visits * pop_area_sqr_km))

ggplot(pop_dens %>%
         filter(taxon_capture %in% c("rana_muscosa", "rana_sierrae")), aes(x = obs_pop_dens, y = bd_load_co_density, color = taxon_capture)) +
  geom_point(alpha = 0.2) +
  scale_y_log10() +
  scale_x_log10()

n_pops = brms_data %>%
  group_by(taxon_capture) %>%
  count

```

# Combined rank plot
```{r}
#| fig-width: 20
#| fig-height: 20

x_range <- ggplot_build(p_cascade_hu_contrast_linpred)$layout$panel_scales_x[[1]]$range$range
y_range <- ggplot_build(p_cascade_mu_contrast_linpred)$layout$panel_scales_x[[1]]$range$range

p_combined = wrap_plots(p_median_effect_scatter +
                          scale_x_continuous(position = "top",
                                             limits = x_range
                                             ) +
                          ylim(y_range) +
                          labs(title = NULL,
                               xlab = NULL,
                               ylab = NULL),
                        p_cascade_mu_contrast_linpred +
                          scale_y_discrete(position = "right",
                                           limits = rev) +
                          coord_flip() +
                          theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1)) +
                          labs(title = NULL,
                               xlab = NULL,
                               ylab = NULL),
                        p_cascade_hu_contrast_linpred +
                          scale_y_discrete(limits = rev) +
                          labs(title = NULL,
                               xlab = NULL,
                               ylab = NULL),
                        p_phylo_heatmap +
                          scale_y_discrete(position = "right",
                                           limits = rev) +
                          labs(title = NULL,
                               xlab = NULL,
                               ylab = NULL),
                        nrow = 2)

p_combined

```

# Export Essential Data
```{r}
# define manuscript destination dir
repo_dir = gsub("size_bd_load_analysis", "", here())
manuscript_data_dir = paste0(repo_dir, "size-bd-dynamics-manuscript/resources/01_essential_data/")

# save essential data
write_csv(bd_model_env_stats, paste0(manuscript_data_dir, "unified_03_a_species_data_stats.csv"))
write_parquet(spp_lss_draws, paste0(manuscript_data_dir, "unified_03_b_spp_lss_draws.parquet"))
write_csv(combined_agg, paste0(manuscript_data_dir, "unified_03_c_combined_effects_agg.csv"))
write_parquet(agg_precip, paste0(manuscript_data_dir, "unified_03_d_precipitation_agg.parquet"))
write_parquet(agg_temp, paste0(manuscript_data_dir, "unified_03_e_temperature_agg.parquet"))
write_parquet(pop_agg_combined, paste0(manuscript_data_dir, "unified_03_f_population_agg.parquet"))


write.tree(jp_pruned, file = here(manuscript_data_dir, "jp_tree_pruned.nwk"))

```